### 2014-10-02 manuscript/manuscript.tex
starting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% American Geophysical Union (AGU)
% LaTeX Template
% Version 1.0 (3/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% The AGUTeX class and agu-ps referencing style were created and are owned 
% by AGU: http://publications.agu.org/author-resource-center/author-guide/latex-formatting-toolkit/
%
% This template has been modified from the blank AGU template to include
% examples of how to insert content and drastically change commenting. The
% structural integrity is maintained as in the original blank template.
%
% Important notes: 
% This template retains extensive commenting from the AGU template. It is heavily 
% advised you read these comments and follow them in order to insure a speedy 
% submission process.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified March 2013
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Step 1: Set the \documentclass

% There are two options for article format: two column (default) and draft.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.

% Choose the journal abbreviation for the journal you are submitting to:

% jgrga	JOURNAL OF GEOPHYSICAL RESEARCH
% gbc	GLOBAL BIOCHEMICAL CYCLES
% grl		GEOPHYSICAL RESEARCH LETTERS
% pal	PALEOCEANOGRAPHY
% ras	RADIO SCIENCE
% rog	REVIEWS OF GEOPHYSICS
% tec	TECTONICS
% wrr	WATER RESOURCES RESEARCH
% gc		GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw	SPACE WEATHER
% ms	JAMES
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft,jgrga]{AGUTeX}

% To create numbered lines:

% If you don't already have lineno.sty, you can download it from http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/ (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN). Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno} and \linenumbers*[1]command, below:

\usepackage{lineno}
\linenumbers*[1]

%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables

% DO NOT USE \psfrag or \subfigure commands.

%  Figures and tables should be placed AT THE END OF THE ARTICLE, after the references.

%  Uncomment the following command to include .eps files (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
\usepackage{graphicx}

% Substitute one of the following for [dvips] above if you are using a different driver program and want to proof your illustrations on your machine:
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]

%  Uncomment the following command to allow illustrations to print when using Draft:
\setkeys{Gin}{draft=false}

% See how to enter figures and tables at the end of the article, after references.

%----------------------------------------------------------------------------------------
%	RUNNING HEAD AND CORRESPONDING AUTHOR
%----------------------------------------------------------------------------------------

% Author names in capital letters:
\authorrunninghead{ASSUMPCAO M.}

%------------------------------------------------

% Shorter version of title entered in capital letters:
\titlerunninghead{SHORT TITLE}

%------------------------------------------------

% Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: Jane Smith, Department of Geography, Ohio State University, Columbus, Ohio, USA. (j.smith@ohio.edu)}

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\title{manuscript agu}

%----------------------------------------------------------------------------------------
%	AUTHORS AND AFFILIATIONS
%----------------------------------------------------------------------------------------

% Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote; matching \altaffiltext will appear at bottom of page.

\authors{John Smith,\altaffilmark{1}
James Smith,\altaffilmark{1,2}, and
Jane Smith\altaffilmark{2}}

\altaffiltext{1}{Department of Hydrology and Water Resources, University of Arizona, Tucson, Arizona, USA.}

\altaffiltext{2}{Department of Geography, Ohio State University, Columbus, Ohio, USA.}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% Do NOT include any \begin...\end commands within the body of the abstract.

\begin{abstract}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed vehicula metus sapien. Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu. Vestibulum ultrices eros sed odio tincidunt bibendum. Pellentesque fermentum ante vel nulla commodo fermentum. Vestibulum in augue sit amet libero viverra accumsan eu at magna. Sed at ligula quis nibh pharetra facilisis non eu libero. Suspendisse non quam sit amet massa luctus interdum sit amet in purus. Integer id orci elit, vitae sollicitudin lectus.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENT
%----------------------------------------------------------------------------------------

% The body of the article must start with a \begin{article} command
% \end{article} must follow the references section, before the figures and tables.

\begin{article}

\section{Introduction}

Nam fermentum sapien at enim varius consectetur. Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae. Integer lacinia purus vel metus tempus suscipit. Curabitur ac sapien quis mauris euismod commodo. Sed pharetra sem elit. Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor. Donec convallis ligula vitae sem viverra fermentum. Mauris in ullamcorper erat. Donec ultrices tempus nibh quis vestibulum. This statement requires citation \citep{AtkinsonSloan}. This one is an in-text citation because the authors of \citet{ColtonKress1} are specifically mentioned.

Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus. Aliquam a tellus nec felis sagittis consequat. Quisque convallis imperdiet neque a tempor. Nulla non erat urna. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Suspendisse viverra tortor pretium orci lacinia eleifend. Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem. Vestibulum pharetra malesuada egestas. Sed non orci sit amet erat suscipit fringilla in et diam. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla. Aenean porta rhoncus suscipit.

%------------------------------------------------

\section{Materials and Methods}

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Mauris eu sapien nunc, sit amet accumsan dui. Nulla ac diam ut nunc placerat semper eget et libero. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Cras hendrerit ullamcorper sapien vitae luctus. Quisque vel diam massa. Vestibulum dui nibh, facilisis vel vestibulum eu, viverra in quam.

\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%------------------------------------------------

\section{Results}

Referencing equation \ref{eq:emc}. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Referencing Table \ref{sampletable}. Referencing Figure \ref{placeholder}.

\subsection{Simulations}

\subsubsection{Simulation 1}

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

\subsubsection{Simulation 2}

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

\subsection{Real Data}

Aliquam interdum pellentesque scelerisque. Sed tincidunt suscipit purus, id aliquet nulla vehicula quis. Duis sed nisl lorem. Vivamus erat ante, dignissim et aliquam vel, adipiscing vitae magna. Cras id dapibus metus. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Proin ut lectus ut nisi congue ullamcorper. Ut ac turpis ligula. Sed faucibus bibendum nunc eget gravida.

%------------------------------------------------

\section{Discussion}

Nam fermentum sapien at enim varius consectetur. Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae. Integer lacinia purus vel metus tempus suscipit. Curabitur ac sapien quis mauris euismod commodo. Sed pharetra sem elit. Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor. Donec convallis ligula vitae sem viverra fermentum. Mauris in ullamcorper erat. Donec ultrices tempus nibh quis vestibulum.

Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus. Aliquam a tellus nec felis sagittis consequat. Quisque convallis imperdiet neque a tempor. Nulla non erat urna. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Suspendisse viverra tortor pretium orci lacinia eleifend. Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem. Vestibulum pharetra malesuada egestas. Sed non orci sit amet erat suscipit fringilla in et diam. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla. Aenean porta rhoncus suscipit.

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

%----------------------------------------------------------------------------------------
%	APPENDICES (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here

% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing  \appendix

% \section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title

\appendix

\section{Appendix Title}

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

%----------------------------------------------------------------------------------------
%	GLOSSARY OR NOTATION (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgments}
This work was partially supported by a grant from the Spanish Ministry of Science and Technology.
\end{acknowledgments}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}

% Or,

% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.

% Follow these steps:
% 1. Run LaTeX on your LaTeX file.

% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.

% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.

% 4. Comment out the old \bibliographystyle and \bibliography commands.

% 5. Run LaTeX on your new file before submitting.

% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

\begin{thebibliography}{}

\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
  Comp.}, \textit{56}(193), 119--139.

\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
  Scattering Theory}, John Wiley, New York.

\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
  Wendland}}]{StephanHsiao}
Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
  Appl. Math.}, \textit{34}(1), 1--19.

\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
  components and line integrals of currents in modified edge representation,
  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

\end{thebibliography}

% Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

\end{article}

%----------------------------------------------------------------------------------------
%	FIGURES AND TABLES
%----------------------------------------------------------------------------------------

%% Enter Figures and Tables here:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information should be placed in footnotes below the table.
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}
% \noindent\includegraphics[width=20pc]{samplefigure.eps}
% \caption{Caption text here}
% \label{figure_label}
% \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% See below for how to make sideways figures or tables.

\newpage

\begin{figure}
\includegraphics[width=0.4\linewidth]{placeholder.jpg}
\caption{Figure caption}\label{placeholder}
\end{figure}

\begin{table}
\caption{Table caption}\label{sampletable}
\begin{tabular}{l l l}
\hline
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\hline
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\hline
\end{tabular}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with the first character following the equal sign on the first line.

% Use an \hspace{} command to insert horizontal space into your equation if necessary. Place an appropriate unit of measure between the curly braces, e.g. \hspace{1in}; you may have to experiment to achieve the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}

### 2014-10-02 manuscript.tex
1st steps... title, author, affiliation, etc...

--- 2014-10-02 manuscript/manuscript.tex
@@
 
-\documentclass[draft,jgrga]{AGUTeX}
+\documentclass[draft,grl]{AGUTeX}
 
@@
 % Author names in capital letters:
-\authorrunninghead{ASSUMPCAO M.}
+\authorrunninghead{PIRCHINER M.}
 
@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{SHORT TITLE}
+\titlerunninghead{Brazilian Smoothing Seismicity and PSHA}
 
@@
 % Corresponding author mailing address and e-mail address:
-\authoraddr{Corresponding author: Jane Smith, Department of Geography, Ohio State University, Columbus, Ohio, USA. (j.smith@ohio.edu)}
+\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, São Paulo, Brazil.
+(marlon.pirchiner@iag.usp.br)}
 
@@
 
-\title{manuscript agu}
+\title{Smoothing seismicity techniques applied to 
+seismic sources characterization and 
+probabilistic seismic hazard analysis in Brazil}
 
@@
 
-\authors{John Smith,\altaffilmark{1}
-James Smith,\altaffilmark{1,2}, and
-Jane Smith\altaffilmark{2}}
+\authors{Marlon Pirchiner\altaffilmark{1,2}}
 
-\altaffiltext{1}{Department of Hydrology and Water Resources, University of Arizona, Tucson, Arizona, USA.}
+\altaffiltext{1}{Department of Geophysics, 
+Institute of Astronomy, Geophysics and Atmospherical Sciences, 
+University of São Paulo, São Paulo, Brazil.}
 
-\altaffiltext{2}{Department of Geography, Ohio State University, Columbus, Ohio, USA.}
+\altaffiltext{2}{School of Applied Math, Getúlio Vargas Foundation, Rio de Janeiro, Brazil.}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed vehicula metus sapien.
-Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu.
-Vestibulum ultrices eros sed odio tincidunt bibendum.
-Pellentesque fermentum ante vel nulla commodo fermentum.
-Vestibulum in augue sit amet libero viverra accumsan eu at magna.
-Sed at ligula quis nibh pharetra facilisis non eu libero.
-Suspendisse non quam sit amet massa luctus interdum sit amet in purus.
-Integer id orci elit, vitae sollicitudin lectus.
+abrastcasrtasdlkjh alksjd flkajhsd flkjahds flkjahsdlkfja slkdjfh alskjdfh laksjdhf lakjshdfl ajhds lfkja lksdjflaksjdf
+lakjsf akjshdfl akjdsfljkahds lfkjahds lkjahsdlfkjas ldkjfh laskjdhf laksjf lasdfh
+asdflkajf lkajshdf lajshdfliuhp 9qerjhtgnv.k, zmxcnpq;awe. ksmdnf va.,fxcnvpq ;.jars,mndf vq;asgdv
+qolehfbd nvqlas,dbnv qalsd.vjhbq oalsk.ezjrbnfc;wq asdl.Zxgjyvqehratnpg oq;ve.krasj gbeq;adr.fxngb
 

### 2014-10-07 manuscript.tex
abstract

@@
 
-abrastcasrtasdlkjh alksjd flkajhsd flkjahds flkjahsdlkfja slkdjfh alskjdfh laksjdhf lakjshdfl ajhds lfkja lksdjflaksjdf
-lakjsf akjshdfl akjdsfljkahds lfkjahds lkjahsdlfkjas ldkjfh laskjdhf laksjf lasdfh
-asdflkajf lkajshdf lajshdfliuhp 9qerjhtgnv.k, zmxcnpq;awe. ksmdnf va.,fxcnvpq ;.jars,mndf vq;asgdv
-qolehfbd nvqlas,dbnv qalsd.vjhbq oalsk.ezjrbnfc;wq asdl.Zxgjyvqehratnpg oq;ve.krasj gbeq;adr.fxngb
+Intraplate seismicity is a hard problem to solve in tectonics, main during the lack of observed seismicity in
+many areas where seismic hazard assesment are also difficult and the methods applied
+are often close correlated to the geometries defined by the experts, of course, with valuable criteria.
+
+Three smoothing seismicity techniques are reviewed in order to investigate their appliability to the 
+Brazilian seismicity and seismic hazard assesment case. 
+All these methods are \emph{catalog-based} in contrast to standard zoning \emph{expert-based} methods from Cornell and
+McGuire. Of course it does not mean that the expert there is no value and many parameters needs to be configured.
+
+All the methods are based on a kernel density estimation, but the main diffence between them is the
+\emph{bandwidth selection} process. The Frankel method use a \emph{fixed} bandwidth in space, Woo method is based on
+\emph{magnitude} also and the Helmstetter is a \emph{local-adaptive} in space and time bandwidth selection. This last
+mathod was proposed originally to care the background seismicity rate on long-term earthquake forecast but the one year
+forecast is exactly the annual seismic rate used to define a seismic source under a time-independent assumption.
+
+In the conclusions the results shown that all methods gives good results and could be used in the Brazilian
+seismicity case, all of them, improoving the spatial resolution of the actual state-of-art GSHAP model. 
+The Last two methods was implemented and are available as a free software on HMTK from GEM Foundation. All the hazard
+calculations was performed on OpenQuake open source software.
 

### 2014-10-07 manuscript.tex
intro

@@
 
-Nam fermentum sapien at enim varius consectetur.
-Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae.
-Integer lacinia purus vel metus tempus suscipit.
-Curabitur ac sapien quis mauris euismod commodo.
-Sed pharetra sem elit.
-Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor.
-Donec convallis ligula vitae sem viverra fermentum.
-Mauris in ullamcorper erat.
-Donec ultrices tempus nibh quis vestibulum.
-This statement requires citation \citep{AtkinsonSloan}.
-This one is an in-text citation because the authors of \citet{ColtonKress1} are specifically mentioned.
+Brazil have no recent seismic hazard study at national scale despite some important and localized recent studies
+focused mainly on specific buildings.
 
-Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus.
-Aliquam a tellus nec felis sagittis consequat.
-Quisque convallis imperdiet neque a tempor.
-Nulla non erat urna.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Suspendisse viverra tortor pretium orci lacinia eleifend.
-Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem.
-Vestibulum pharetra malesuada egestas.
-Sed non orci sit amet erat suscipit fringilla in et diam.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla.
-Aenean porta rhoncus suscipit.
+The Brazilain seismic hazard building code was made based on GSHAP model which is the well accepted state of art model
+for the whole region.
+
+A significant number of earthquakes has been recorded in the last years, mainly after the national seismic network
+(RSBR) operation, and an actualized hazard map whould be computed.
+
+Although the standard Cornell Mcguire zoning methodolgy for the classical hazard seismic assesment, the goal is
+review a few smoothing seismicity methods based on kernel density estimation. This estimation allow a seismic point
+sources grid definition. The seismic rate (a-value) associated to each source is definided by the smoothed seismicity
+methods, but the more parameters like b-values, minimum and maximum magnitudes, rupture and depth
+distributions must to be assigned in other ways.
+
+In the kernel density estimation theory two things there always be handled. The kernel shape and bandwidth.
+Almost kernels used are gaussian or power-laws in one of their variants. The mainly distinction between methods decribed
+is the proposed bandwidth selection process.
+
+Frankel smoothing method applies a gaussian Nadaraya-Watson smoothing method with a \emph{fixed}
+smoothing-distance bandwidth to estimate seismicity rate. Gordon Woo proposes an \emph{magnitude-dependent} kernel
+bandwidth based on the nearest neighbour mean distance in magnitude bins. Helmstetter proposes a \emph{local-adaptive}
+bandwidth, based on kernel estimations on both space and time dimentions, estimation of the background seismicity rate
+for long-term forecasts. Using one-year forecast and assume time-independence, the same assumption of PSHA, of
+seismicity rate is possible to use these values on seismic sources characterization.
+
+To perform the hazard calculation, the OpenQuake suite was choosed. The modeller environment used was the HMTK also
+available by GEM scientifical board. The Frankel method was already implemented on the toolkit and the other two was
+implemented in the context of this paper and they are free available on a public repository.
+
+The earthquake catalog data comes from the Brazilian seismological research authorities.
+
+The Brazilian seismicity data is presented including some overview plots. Next the method's theory is presented follow
+by the specific decisions and optimizations performed on each method modelling. And last, the results, conclusions and
+further considerations are discussed.
 

### 2015-01-08 manuscript.tex
last changes at home

@@
 
-\documentclass[draft,grl]{AGUTeX}
+\documentclass[draft,grl]{agutex}
 
@@
 
-Intraplate seismicity is a hard problem to solve in tectonics, main during the lack of observed seismicity in
-many areas where seismic hazard assesment are also difficult and the methods applied
-are often close correlated to the geometries defined by the experts, of course, with valuable criteria.
+Intraplate seismicity is a hard problem to solve in tectonics, 
+main during the lack of observed seismicity in many areas where seismic 
+hazard assesment are also difficult and the methods applied are often 
+close correlated to the geometries defined by the experts with valuable 
+criteria.
 
-Three smoothing seismicity techniques are reviewed in order to investigate their appliability to the 
-Brazilian seismicity and seismic hazard assesment case. 
-All these methods are \emph{catalog-based} in contrast to standard zoning \emph{expert-based} methods from Cornell and
-McGuire. Of course it does not mean that the expert there is no value and many parameters needs to be configured.
+Three smoothing seismicity techniques are reviewed in order 
+to investigate their results when applied over the Brazilian seismicity 
+and suggesting a seismic hazard assesment.
 
-All the methods are based on a kernel density estimation, but the main diffence between them is the
-\emph{bandwidth selection} process. The Frankel method use a \emph{fixed} bandwidth in space, Woo method is based on
-\emph{magnitude} also and the Helmstetter is a \emph{local-adaptive} in space and time bandwidth selection. This last
-mathod was proposed originally to care the background seismicity rate on long-term earthquake forecast but the one year
-forecast is exactly the annual seismic rate used to define a seismic source under a time-independent assumption.
+All the methods are based on a kernel density estimation. The main 
+difference between them is the \emph{bandwidth selection} process.
 
-In the conclusions the results shown that all methods gives good results and could be used in the Brazilian
-seismicity case, all of them, improoving the spatial resolution of the actual state-of-art GSHAP model. 
-The Last two methods was implemented and are available as a free software on HMTK from GEM Foundation. All the hazard
-calculations was performed on OpenQuake open source software.
+The Frankel method use a \emph{fixed} bandwidth in space. Woo method
+suggest a bandwidth as function of the \emph{magnitude}.
+Helmstetter and Werner proposes a \emph{local-adaptive} 
+in space and time bandwidth, chosen for each event. 
+It was proposed initialy to take care of the background seismicity rate 
+on long-term earthquake forecast context, but the one year forecasted 
+seismic rate could be a grid of points seismic sources under a 
+time-independent assumption.
+
+In the conclusions the results shown that
+
+all methods gives good results and 
+could be used in the Brazilian seismicity case,
+
+all of them, 
+% ###### improoving #########
+the spatial resolution of the actual state-of-art GSHAP model.
+
+The Last two methods was implemented and are available as a free software on HMTK from GEM Foundation. 
+All the hazard calculations was performed on OpenQuake open source software.
 
@@
 \section{Introduction}
+
+\subsection{Intraplate Seismicity}
+\subsection{Previous Brazilian Studies}
+\subsection{Previous Smoothing Seismicity (Frankel,Woo,Helmstetter\ldots)}
+\subsection{Specifical Purpose}
 
@@
 \section{Materials and Methods}
+
+\subsection{Brazilian Seismic Catalog}
+Boletim sismico brasileiro.
+
+História e versão com Mw.
+\subsection{Brazilian Seismic Catalog}
 

### 2015-01-13 c01-introducao.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Introdução}
\label{cap:introducao}

Um elemento primordial na análise
de \emph{risco} sísmico é a análise da \emph{ameaça} sísmica,
onde a identificação e caracterização das fontes sismogênicas (causadoras de
movimento do chão, fundamentalmente tremores de terra) é a primeira das etapas.

Considera-se nessa fase, principalmente as falhas
geológicas, o acúmulo de tensão medido através do movimento relativo da crosta
terrestre, a neotecnônica da crosta, o possível acoplamento entre placas, os tremores
(rupturas e falhamentos) já registrados anteriormente, enfim, todo conhecimento geológico
disponível, para caracterizar (a) a geometria espacial da feição geológica e provável fonte
sísmica e (b) o número de ocorrência - taxa - dos tremores conforme a
proporção de energia liberada por cada um - magnitude.

No Brasil, onde a ocorrência de tremores não é desprezível mas menor do que a de
outras partes do planeta, o processo de identificação das fontes sísmicas é
executado geralmente através da opinão de especialistas que fazem o zoneamento
sísmico segundo as informações, técnicas e a experiência que dispõem.

Para cada uma dessas zonas sísmicas, que serão consideradas como tendo atividade
sísmica uniforme, é determinada a distribuição da ocorrência de tremores em função
da magnitude de cada tremor.

Existem também outras propostas metodológicas envolvendo técnicas de suavização
que permitem estimativas da taxa de sismicidade, por exemplo por funções de núcleo.
As propostas de \citet{frankel_1995}, a de \citet{woo_1996} e a de
\citet{helmstetter_2012} serão discutidas com maior detalhe.

O que todas elas possuem em comum é o objetivo de caracterizar a taxa de
sismicidade (ocorrência de tremores) em uma malha sobre a região de interesse
através da soma da contribuição de funções de núcleo (gaussianas, leis de
potência, etc) em cada nó dessa malha.

O pressuposto central dessa
idéia é que os grandes sismos (com menos evidências, pois
aconteceram poucos fenômenos observáveis desse tipo) 
tendem a ocorrer no entorno de
onde já ocorreram antes outros tremores (menores e mais frequentes).

Fundamentalmente o que os diferencia é a forma de escolher a largura para essas
funções de núcleo associadas a cada tremor do catálogo.

O que se pretende é observar um pouco mais detalhadamente o comportamento
desses diferentes métodos num ambiente com baixa e esparsa sismicidade.

Perifericamente, aproveitou-se a oportunidade para avaliar um recente
conjunto de programas de computador, com código livre, implementado para esse tipo de análise.

%% ------------------------------------------------------------------------- %%
\section{Objetivos}
\label{sec:objetivo}

O principal objetivo pretendido é a avaliação da
aplicabilidade de algumas técnicas de suavização para a
caracterização da ocorrência de sismos no Brasil.

Secundariamente, aproveita-se a oportunidade para testar o uso de um conjunto
recente de programas de computador disponível livremente pela e para a comunidade
científica, o Openquake\footnote{\url{http://www.globalquakemodel.org/openquake}}.

Perifericamente, será possível fazer alguma contribuição ao Openquake
ou a alguma outra biblioteca associada destinada a modelagem dos parâmetros de entrada
para o cálculo de ameaça.

%% ------------------------------------------------------------------------- %%
\section{Contribuições}
\label{sec:contribucoes}

As principais contribuições deste trabalho são:

\begin{itemize}
  \item Discorrer sobre métodos alternativos ao zoneamento para a caracterização de fontes
  sismogênicas, a primeira das etapas da análise probabilística de ameaça 
  sísmica.

  \item Compreender e assimilar como usar Openquake, um conjunto de programas de
  computador desenvolvido recentemente e oferecido com código livre, contribuindo
  para oferecer uma ferramenta a mais para o cálculo da ameaça sísmica à comunidade
  de sismologia e engenharia sísmica brasileira.

\end{itemize}

%% ------------------------------------------------------------------------- %%
\section{Organização do Trabalho}
\label{sec:organizacao_trabalho}

No capítulo \ref{cap:conceitos}, são apresentados os conceitos mais elementares
de sismologia e de estatística relevantes para uma melhor compreensão do tema.
Em seguida, no capítulo \ref{cap:regiao_de_estudo} é apresentado com maior detalhe
a região de estudo sob o ponto de vista geológico e tectônico. No capítulo \ref{cap:teoria}
é apresentada e formalizada a teoria e os fundamentos dos métodos discutidos. O capítulo \ref{cap:processamento}
discorre sobre as etapas de processamento propriamente ditas, enquanto o capítulo \ref{cap:resultados}
foca apenas nos resultados de cada método e em resultados anteriores. As discussões relevantes a partir da análise
dos resultados é apresentada no capítulo \ref{cap:conclusoes} finalmente.

### 2015-01-13 c04-metodologia.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Contexto Teórico}
\label{cap:teoria}

%% ------------------------------------------------------------------------- %%
\section{Apresentação}
%\index{área do trabalho!fundamentos}
\label{sec:c04_apresentacao}

Este capítulo apresenta a formalização das teorias aplicadas na fase de
processamento.

Trata-se essencialmente das \gls{smoothing} que, em geral, permitem extrair 
feições importantes do conjunto de dados.

Quando aplicadas à caracterização das \glspl{seismic_source} em \gls{psha}
tornam possível gerar um conjunto regular de \glspl{point_source} 
singularmente definidas pela suavização das \glspl{seismic_rate} 
nas células de uma malha regular.

%% ------------------------------------------------------------------------- %%
\section{\Gls{smoothing}}
\index{suavização!fundamentos, metodologia}
\label{sec:04_smoothing_general}

A ideia, no fundo, é estimar a distribuição espacial da taxa anual de sismicidade $R$
ou sua \gls{pdf}.
O método mais simples conhecido para essa estimativa seria o histograma.

\subsection{Histograma 2D: uma possível \glsdesc{pdf} para a taxa de sismicidade}

Numa malha regular a taxa anual de sismicidade em cada célula seria calculada 
contando, à partir de um catálogo com tempo de observação conhecido,
\begin{equation}
\scriptsize
	\ensuremath{
	\frac{\text{o número observado de sismos na célula}}
		 {\text{área/volume da célula} \times 
		  \text{número total de sismos observados}}
	/
	\text{tempo de observação em anos}
	}
\normalsize.
\label{eq:rate_count}
\end{equation}

Isso seria equivalente a preparar um histograma normalizado 
dos tremores em duas (ou três, considerando a profundidade) dimensões.

O que se busca, em geral, por essas técnicas é suavizar justamente essas contagens ou esse histograma
normalizado que representa uma estimativa da função de densidade de probabilidade da taxa de ocorrência espacial de tremores.

%\newtheorem{prop}{Proposição}
%\begin{prop}

\subsection{Regressão e Suavizadores}

Para os $n$ pares (célula, taxa de sismicidade) $(x_1, R_1), (x_2, R_2), \cdots, (x_n, R_n)$
obtidos pela contagem anterior, considere um modelo para a taxa de sismicidade $R$ 
em uma determinada célula $x_i$ a partir dessa amostra dado por
\begin{equation}
	\ensuremath{
		R(x_i) = \lambda(x_i) + \epsilon(x_i),\;\;\; i=1,\dots,n
	}
\label{eq:rate_model}
\end{equation}

onde os $\epsilon_i$ são \gls{va} não-correlacionadas que representam os erros 
tais que $E(\epsilon_i \arrowvert X = x_i) = 0$ 
e a  $Var(\epsilon_i \arrowvert X = x_i) = \sigma^2(x_i)$. A função  
$\lambda(x_i) = E(R_i \arrowvert X = x_i)$ é uma função de regressão.
%\end{prop}

É possível definir um estimador ou suavizador linear $\hat{\lambda}$ para $\lambda$, 
se para todo $x \in \mathbb{R}$ existe uma sequência de pesos $w_1(x), w_2(x),\cdots,w_n(x)$ tais que
$\sum_{i=1}^{n}w_i(x) = 1$, como sendo

\begin{equation}
	\ensuremath{
		\hat{\lambda}(x) = \sum_{i=1}^{n}w_i(x)\,R_i.
	}
\label{eq:rate_estim}
\end{equation}

A questão passa a ser então como encontrar essa sequência de pesos $w_i$.

\subsection{Função de Núcleo e Estimadores de Nadaraya-Watson}
\label{sec:nadaraya}
Uma função de núcleo $K$ (\emph{Kernel}) é qualquer função par, contínua e limitada que satisfaz as seguintes
propriedades:

%\begin{enumerate}[(i)]
%	\item $\int \lvert K(\boldsymbol{u})\rvert \,\mathrm{d}\boldsymbol{u} < \infty $
%	\item $\underset{ \lvert\boldsymbol{u} \rvert \to \infty }{\lim} \lvert \boldsymbol{u} K(\boldsymbol{u})\rvert = 0$
%	\item $\int \! K(\boldsymbol{u}) \,\mathrm{d}\boldsymbol{u} = 1 $
%\end{enumerate}
\begin{center}
(i) $\int \lvert K(\boldsymbol{r})\rvert \,\mathrm{d}\boldsymbol{r} < \infty $
\;\;\;\;\;(ii) $\underset{ \lvert\boldsymbol{r} \rvert \to \infty }{\lim} 
				\lvert \boldsymbol{r} \, K(\boldsymbol{r})\rvert =0$ 
\;\;\;\;\;(iii) $\int \! K(\boldsymbol{r})	\,\mathrm{d}\boldsymbol{r} = 1 $.
\end{center}

Uma das possíveis maneiras de se encontrar os pesos $w_i$ é o 
Estimador de Nadaraya-Watson \citep{nadaraya_1965}.

Seja $h \in \mathbb{R}, h > 0$ e $K$ uma função de núcleo. 
Nadaraya e Watson propõem, para a estimativa \eqref{eq:rate_estim}, os pesos
\begin{equation}
	\ensuremath{
		w_i(x) = \frac{ K\left( \frac{x - x_i}{h} \right)}
					  {\sum_{j=1}^{n} K\left( \frac{x - x_j}{h} \right) },
	}
\label{eq:rate_wi}
\end{equation}
onde $h$ é conhecida como largura de banda ou \emph{bandwidth}.

\subsection{Formas das funções de núcleo}

Dentre as possíveis expressões para as funções de núcleo é relevante destacar duas \citep{kagan_2000}.

A equação \eqref{eq:kernel_gs} apresenta a expressão da função de núcleo gaussiana:

\begin{equation}
	\ensuremath{
		K_{gs}(\gls{sym:r}\arrowvert h) = \eta_1(h)
			e^{- \frac{\|\gls{sym:r}\|^2}
 				 	  {2 h^2 }},
 	}
\label{eq:kernel_gs}
\end{equation}
onde $h$ é a largura de banda definida para a função de núcleo e $\eta_1(h)$ um fator de normalização
para que sua integral seja igual à unidade.

Outra forma possível para a função de núcleo é uma Lei de Potência (\emph{power-law}), que decai com o 
inverso do cubo da distância, como na equação
\eqref{eq:kernel_pl} a seguir:

\begin{equation}
	\ensuremath{
		K_{pl}(\gls{sym:r}\arrowvert h) = 
			\frac{\eta_2(h)}
 				 {\left( \|\gls{sym:r}\|^2 + h^2 \right)^{\frac{3}{2}} },
 	}
\label{eq:kernel_pl}
\end{equation}
onde $h$ é a largura de banda definida para a função de núcleo e $\eta_2(h)$ um fator de normalização
para que sua integral também seja igual à unidade.

Quaisquer dessas duas funções podem ser usadas como função de núcleo 
para os estimadores de Nadaraya-Watson apresentados anteriormente.

\subsection{Contribuição de uma função de núcleo bidimensional}

É importante notar que há uma outra abordagem possível para a estimativa da taxa de sismicidade.

Se, em vez de suavizar o histograma do catálogo sísmico numa malha regular, a proposta for
avaliar a contribuição em probabilidade pela \gls{pdf} de cada tremor em uma determinada célula da malha, 
é possível considerar uma função de núcleo como a \gls{pdf} da ocorrência de cada tremor, e,
essa contribuição numa célula $j$, devido ao tremor $i$ em $\boldsymbol{r}_i$, dada pela integral da
função de núcleo na área/volume da célula \citep{zechar_2010-2}: 
\begin{equation}
	\ensuremath{
		R_{j}(\boldsymbol{r}_i \arrowvert h) = \int\limits_{\mathrm{cell}\;j}\!K( \gls{sym:r} -
		\gls{sym:ri}\,\arrowvert\,h)\,\mathrm{d}\gls{sym:r} },
\label{eq:kernel_int}
\end{equation}
onde $h$ é a largura de banda da suavização.

Em todos os casos apresentados, é fácil notar que, a determinação da largura de banda influencia diretamente
a estimativa da taxa de sismicidade. Os métodos estudados variam essencialmente 
na forma com que a escolha da largura de banda é concebida e proposta.

%% ------------------------------------------------------------------------- %%
\section{Frankel, 1995}
\index{Frankel, 1995}
\label{sec:frankel}

A proposta de Arthur \citet{frankel_1995} foi usar uma largura de banda fixa, 
nomeada distância de correlação $d_F$, e aplicar o estimador de Nadaraya-Watson (\ref{sec:nadaraya})
para suavizar o histograma 2D da sismicidade utilizando uma função de núcleo gaussiana:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
onde $\tilde{n}_j$ é a taxa de sismicidade (número de sismos com magnitude $m$ maior que a mínima magnitude
\gls{sym:Md} do catálogo) suavizada
na célula $j$, $n_i$ é o número de sismos em cada outra célula $i$ e
	\gls{sym:dij} é \glsdesc{sym:dij}.

\citet{zechar_2010-2} propuseram uma variação da abordagem de Frankel avaliando a contribuição
de cada tremor $i$ na célula $j$ em vez de suavizar o histograma 2D, mas para isso
é preciso avaliar essa contribuição como na equação \eqref{eq:kernel_int}.

%% ------------------------------------------------------------------------- %%
\section{Woo, 1996}
\index{Woo, 1996}
\label{sec:woo}

Já Gordon \citet{woo_1996} propôs avaliar a contribuição de uma função de núcleo 
de cada sismo $i$ ocorrido em $\boldsymbol{r}_i$ 
na célula centrada em $\boldsymbol{r}$ que dependa também de sua
magnitude $m$:
\begin{equation}
	\ensuremath{
		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
											{T({\gls{sym:ri})}},
	}
	\label{eq:Rrm}
\end{equation}
onde $N$ é o número de tremores $i$ no catálogo 
e $T(\gls{sym:ri})$ é o período em que todo sismo de magnitude acima de $m$ é completamente observado 
em \gls{sym:ri}.

A função de núcleo utilizada por Woo foi a proposta por \citet{kagan_knopoff_1980}
para um domínio espacial infinito:
\begin{equation}
	\ensuremath{
		K_{KJ}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
							\left( 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right)^{-\gls{sym:aW}},
	}
	\label{eq:k_kj}
\end{equation}
onde \gls{sym:aW} é, segundo \citet{verejones_1992}, \glsdesc{sym:aW}.

Uma variante limitada para a função de núcleo foi proposta por \citet{verejones_1992}:\begin{equation}
	\ensuremath{
		K_{VJ}(\gls{sym:r}, m \arrowvert \gls{sym:DW}) = 
		\begin{cases}
			\frac{\gls{sym:DW}}{2\pi\,\gls{sym:hm}} 
			\left( \frac{\gls{sym:hm}}{\gls{sym:r}} \right)^{2 - \gls{sym:DW}} 
			  & \gls{sym:r} \leq \gls{sym:hm} \\
			0 & \gls{sym:r} > \gls{sym:hm}
		\end{cases},
	}
	\label{eq:k_vj}
\end{equation}
onde \gls{sym:DW} é \glsdesc{sym:DW}.

Para a largura de banda \gls{sym:hm}, nos dois casos, Gordon Woo propôs utilizar a relação
\begin{equation}
	\ensuremath{
		h(m\arrowvert \gls{sym:a0}, \gls{sym:a1}) = \gls{sym:a0}e^{\gls{sym:a1}m},
	}
	\label{eq:hm}
\end{equation}
em que $a_0$ e $a_1$ são determinados pela regressão entre a 
distância média $h$ de cada tremor ao vizinho mais próximo em cada faixa de magnitude $m \pm \mathrm{d}m$.
É uma largura de banda dependente da magnitude.

%% ------------------------------------------------------------------------- %%
\section{Helmstetter, 2012}
\index{hemlstetter, 2012}
\label{sec:helmstetter}

O trabalho de  Agnès Helmstetter e Maximilian Werner \citep{helmstetter_2012} é focado principalmente em 
projeção de longo-prazo para ocorrência de tremores/rupturas. 
O principal pressuposto dessa técnica (seção \ref{sec:projecao}) é que nesse período a taxa
de sismicidade se mantenha invariante, que também é o pressuposto da \gls{psha}.

\subsection{Taxa de sismicidade}
Para eles a taxa de sismicidade em uma localização
$\boldsymbol{r}$ e em um instante $t$ 
em função dos tempos $T_i$ e localizações $\mathbf{r}_i$ dos tremores $i$ em um catálogo 
poderia ser expresso por
\begin{equation}
	\ensuremath{\gls{sym:R} = \sum_{i=1}^{N}{ \frac{1}{h_i\,{d_i}^2} \gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms01}
\end{equation}
onde \gls{sym:R} é \glsdesc{sym:R}, 
	  $K_t$ é a \glsdesc{sym:Kt}, 
	  $K_r$ é a \glsdesc{sym:Kr}.

Como seu interesse é em projeção, restringiram os tempos do catálogo aos $t_i
< t$ e ponderaram segundo um peso $w$ da seguinte maneira:
\begin{equation}
\ensuremath{\gls{sym:R} = \gls{sym:Rmin} + \sum_{t_i < t}{ 
	\frac{2\,w(\boldsymbol{r}_i,t_i)}{h_i\,{d_i}^2}
			\gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms02}
\end{equation}
onde \gls{sym:Rmin} é a \glsdesc{sym:Rmin}, positiva, permitindo que ocorram eventos-surpresa 
onde não se tem registro de sismicidade.

Os pesos $w(\boldsymbol{r}_i,t_i)$, calculados em cada tremor $i$, propostos por Helmstetter são a transformação de
Gutemberg-Richter. Consideraram uma contribuição maior para a taxa de sismicidade daqueles sismos 
que ocorreram onde e quando a magnitude de completude $M_c$ era maior que 
a mínima magnitude $M_d$ do catálogo 
em locais onde a magnitude de completude e o \emph{valor-b} variam. 
Note que o modelo de pesos acomoda bem variações espaciais e
temporais tanto da magnitude de completude como do \emph{valor-b}. A expressão é a seguinte:
\begin{equation}
	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
	\right] } },
	\label{eq:helms_wi}
\end{equation}
onde  \gls{sym:wi} é o \glsdesc{sym:wi} na localização $\boldsymbol{r}$ e no instante $t$, 
	  \gls{sym:b}$(\boldsymbol{r},t)$ é o \glsdesc{sym:b}, 
	  \gls{sym:Mc_rt} é a \glsdesc{sym:Mc_rt}, 
	  \gls{sym:Md} é a \glsdesc{sym:Md}.

\subsection{Método acoplado dos vizinhos mais próximos}

Em sismologia, as funções de núcleo no espaço e no tempo são usualmente estimadas
pela distância entre cada tremor e seu $k^{simo}$ vizinho mais próximo.

A definição de cada $h_i$ e $d_i$ é feita por otimização, 
pelo método acoplado do $k^{ésimo}$ vizinho mais próximo
proposto por \citet{choi_1999}
com uma modificação para a função de núcleo temporal assimétrica.

Para cada tremor $i$, as larguras de banda  $h_i$ e $d_i$ são escolhidas para
minimizar a soma $h_i + a_{cnn}\,d_i$ sob a condição de que hajam ao menos $k_{cnn}$
eventos a uma distância espacial menor que $d_i$ e, simultaneamente, estejam
no intervalo de tempo $[t_i - h_i, t_i[$. O parâmetro $k_{cnn}$ controla a suavização 
geral do modelo, e $a_{cnn}$ controla a importância relativa entre o tempo e o espaço.
Valores altos de $a_{cnn}$ apresentam alta resolução espacial e, ao mesmo tempo, 
são mais suaves no domínio do tempo. Em outras palavras:

\begin{equation}
	\ensuremath{
%		h_i, d_i = \underset{d_i \ge \gls{sym:dk}, h_i \ge \gls{sym:hk}}{\argmin} 
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           } 
		\left[ s \left(h_i,d_i 
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right) 
			   := h_i + \gls{sym:a_cnn}d_i 
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
onde \gls{sym:k_cnn} é o \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} é o \glsdesc{sym:a_cnn},
	 \gls{sym:dk} é o \glsdesc{sym:dk} e 
	 \gls{sym:hk} é o \glsdesc{sym:hk}.

As larguras de bandas são escolhidas localmente, podendo ser menor, aumentando a resolução 
onde há maior densidade de informação, ou crescer onde há escassez de informação e menor resolução.

\subsection{Taxa de sismicidade estacionária}
Mesmo havendo modelado a dependência com o tempo para distribuição da taxa de sismicidade,
a taxa estacionária $\bar{R}$ que se preserva no longo-prazo em uma determinada localização  
$\boldsymbol{r}_0$ é a que mais interessa à \gls{psha} e,
segundo o método proposto por \citet{helmstetter_2012}, será definida
pelo valor da mediana da taxa fornecida pelo modelo no
período considerado:

\begin{equation}
	\ensuremath{
		\bar{R}(\boldsymbol{r}_0) = \text{Mediana}\left[R(\boldsymbol{r}_0, t)\right].
	}
	\label{eq:helms_mediana}
\end{equation}

O fato de se tomar a mediana faz com que variações na taxa de sismicidade devido a pré e pós-eventos
não interfiram significativamente no valor da taxa estacionária. A decorrência mais importante desse fato
é que o método dispensa a remoção de agrupamentos (seção \ref{sec:declustering}) do catálogo, um método paramétrico e
quase sempre controverso.

\subsection{Verossimilhança}

Para otimizar os parâmetros do modelo de Helmstetter, é preciso dividir o catálogo de sismos.
Uma parte serve para o aprendizado do modelo e a outra para avaliação.

Se a ocorrência de tremores pode ser modelada por um processo de Poisson com taxa $N_p$, 
então a probabilidade de se observar exatamente $n$ eventos no período de tempo considerado é dada por

\begin{equation}
	\ensuremath{
		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
	}
	\label{eq:loglik}
\end{equation}

E o logarítmo, a ser maximizado, da verossimilhança entre 
o que o modelo predisse e o que foi observado 
pode ser escrito como

\begin{equation}
	\ensuremath{
		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
	}
	\label{eq:loglik}
\end{equation}
onde \gls{sym:Np} é \glsdesc{sym:Np},
	\gls{sym:nxy} é \glsdesc{sym:nxy}.

Os parâmetros $R_{min}$, $a_{cnn}$ e $k_{cnn}$ do modelo são otimizados 
pela maximização da verossimilhança $L$.

\subsection{Ganho}

Uma distribuição de Poisson espacialmente uniforme para a sismicidade com taxa $N_u$
teria sua verossimilhança $L_u$ expressa por
\begin{equation}
	\ensuremath{
		\gls{sym:Lu} = -N_t + 
		\sum_{i_x = 1}^{N_x}\sum_{i_y=1}^{N_y}
		\gls{sym:nxy}\log N_u - \log \left[ \gls{sym:nxy}! \right]
	}
	\label{eq:lu}
\end{equation}
onde $N_u = N_t/N_c$ com $N_t$ o número de sismos no catálogo-alvo e $N_c$ o número de células,

O ganho de probabilidade $G$ por tremor predito no catálogo-alvo (de teste)
sobre o que seria predito por uma distribuição espacialmente uniforme 
foi definido por \citet{kagan_knopoff_1977} como
\begin{equation}
	\ensuremath{
		\gls{sym:G} = e^{ \frac{\gls{sym:L} - \gls{sym:Lu}}{\gls{sym:Nt}}   }.
	}
	\label{eq:gain}
\end{equation}

Quando se compara dois modelos frente o mesmo catálogo-alvo, com a mesma quantidade de tremores $N_t$,
a equação \eqref{eq:gain} pode ser simplificada, pois 
\begin{equation}
	\ensuremath{
	\begin{align}
		L - L_u & = \sum_{i_x = 1}^{N_x}\sum_{i_y=1}^{N_y}
				  \gls{sym:nxy}\log \left[ \frac{\gls{sym:Np}}{\gls{sym:Nu}} \right] \\
				& = \sum_{i = 1}^{N_t}\log \left[\frac{\gls{sym:Npi}}{\gls{sym:Nu}} \right]
	\end{align}}
	\label{eq:llu}
\end{equation}

e o ganho $G$ do modelo passa a ser o mesmo que média geométrica da taxa prevista pelo modelo sobre 
a taxa uniforme em todas as células:
\begin{equation}
	\ensuremath{
	\begin{align}
		G & = e^{\sum_{i = 1}^{N_t}
					\frac{\log \left[  \gls{sym:Npi} / \gls{sym:Nu}  \right]}
						 {\gls{sym:Nt}}
			  } \\
		  & = {\langle  \gls{sym:Npi} / \gls{sym:Nu}  \rangle}_{geom}
	\end{align}}
	\label{eq:G}
\end{equation}
onde $\langle\cdot\rangle_{geom}$ significa média geométrica.

\subsection{Testes}

\citet{helmstetter_2012} usam os testes propostos por \citet{schorlemmer_2007}
e \citet{rhoades_2011} para comparar diferentes modelos.

O teste-T avalia quando o ganho de informação de um modelo é significativamente diferente
de um outro.

Para isso as taxas preditas pelos modelos são re-normalizadas para que o valor da taxa de sismicidade
prevista pelos modelos seja igual à taxa observada.

Sendo \gls{sym:NAi} a \glsdesc{sym:NAi} e \gls{sym:NBi} a \glsdesc{sym:NBi}, o ganho de informação $I_{inf}$
por tremor se reduz à 
\begin{equation}
	\ensuremath{
		\gls{sym:I} = \frac{1}{\gls{sym:Nt}} 
					  \sum_{i = 1}^{N_t}\log\left[ \frac{\gls{sym:NAi}}
					  								  {\gls{sym:NBi}}  \right].
	}
	\label{eq:gain_info}
\end{equation}

O ganho de informação $I_{inf}$ se relaciona com o ganho dos modelos (eq. \eqref{eq:gain}) por
\begin{equation}
	\ensuremath{
		I_{inf} = \log\left(\frac{G_A}{G_B}\right).
	}
	\label{eq:gain_info_G}
\end{equation}

A variância da amostra de Rhoades apresentada por Helmstetter é

\begin{equation}
	\ensuremath{
		\sigma^2 = 	\frac{1}{N_t - 1}
					{\left(
					\sum_{i-1}^{N_t}
						{x_i}^2 
					\right)}
					- 
					\frac{1}{{N_t}^2 - N_t}
					{\left(
						\sum_{i=1}^{N_t}{x_i}
					\right)}^2,
	}
	\label{eq:var}
\end{equation}
onde $x_i = \log\left[ \gls{sym:NAi} / \gls{sym:NBi}  \right]$.

Seja o valor $T_s$ também definido por Rhoades e apresentado por Helmstetter

\begin{equation}
	\ensuremath{
		\gls{sym:Ts} = \frac{I_{inf}\sqrt{N_t}}{\sigma}.
	}
	\label{eq:T}
\end{equation}

Se $x_i$ é independente e tem distribuição normal,
então $T_s$ deve ter distribuição de Student com $N_t - 1$ graus de liberdade. Para grandes valores de $N_t$
a distribuição converge rapidamente para a normal.

O valor do ganho de informação é considerado significante se $T_s > 2$ no intervalo de 95\% de confiança.

O teste de Wilcoxon é usado no caso de que $x_i$ não tenha distribuição normal
mas permanece simétrica e independente \citep{rhoades_2011}.
Esse teste avalia o quanto a mediana de $x_i$ difere significativamente de zero.
Isso é o mesmo que avaliar o quanto a taxa prevista por um modelo é significativamente diferente da prevista por 
outro modelo.

A probabilidade $p_W$ de se observar um valor maior que a mediana de $x_i$ é retornada pelo teste.
Valores de $p_W < 0.05$ indicam que a mediana de $x_i$ é significativamente diferente de 0.

### 2015-01-13 c05-processamento.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Processamento}
\label{cap:processamento}

Esse capítulo apresenta como as teorias do capítulo anterior foram utilizadas para obtenção 
dos resultados que serão discutidos no próximo capítulo.

%% ------------------------------------------------------------------------- %%
\section{Conjunto de Dados}
\index{Conjunto de Dados}
\label{sec:dados}

Os dados utilizados foram os dados do catálogo \gls{iscgem} \citep{storchak_2013} 
para a América do Sul e 
os dados do \glsdesc{bsb2013} \citep{bsb_2013} para o Brasil.

O dados são texto e formatados em arquivos de \gls{csv}.

A figura \ref{fig:eq_record} ilustra o número de registros de tremores por ano nos dois catálogos.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.48\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_sa3_rate}
			\subcaption{Número de tremores registrados por ano, \gls{iscgem}}
			\label{fig:sa_eq_record}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.48\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_bsb2013_rate}
			\subcaption{Número de tremores registrados por ano, \gls{bsb2013}}
			\label{fig:br_eq_record}
    \end{subfigure}%
	\caption{Número de tremores registrados por ano após 1900}
	\label{fig:eq_record}
\end{figure}

Em ambos é possível notar um incremento do número de registros de sismos com o passar dos anos, 
além de algumas variações bem destacadas em períodos reduzidos de poucos anos.

%% ------------------------------------------------------------------------- %%
\subsection{Catálogo ISC-GEM}
\index{Conjunto de Dados!ISC-GEM}
\label{sec:data_source}

O \glsdesc{iscgem} versão 1.04 possui licença CC-BY-SA e é fruto da redeterminação de parâmetros
dos terremotos com os dados disponíveis no ISC-GEM e de um detalhado estudo para que ao menos um valor de
magnitude $\gls{sym:Mw}$ estivesse disponível com incertezas.

Possui pouco mais de 110 000 registros.

\subsection{Boletim Sísmico Brasileiro}
\index{Conjunto de Dados!BSB}
\label{sec:data_source2}

O \glsdesc{bsb} \citep{bsb_2013}, versão 2013.08 possui licença CC-BY e é fruto do esforço de compilação de
dados e determinação de epicentros e magnitudes que contou com a colaboração de várias instituições
como o \gls{obsis} da \gls{unb}, o \gls{ipt}, a \gls{unesp}, a \gls{ufrn} liderados pelo \gls{iag} da \gls{usp}.

Possui algo em torno de 900 registros.

%% ------------------------------------------------------------------------- %%

\section{Ferramentas}
\index{Ferramentas}
\label{sec:ferramentas}

As ferramentas utilizadas para as análises e cálculos apresentados nesse texto
foram de código aberto.
Utilizou-se o Latex para a edição do texto, o software de processamento de dados
geoespacial QGIS, o console do Linux BASH, a IDE Eclipse, o git/GitHub para o controle de versao,

O iPython como console interativo, a MatplotLib e Basemap para o gráficos, 
a SciPy para um conjunto de funções científicas e estatísticas, a shapely e gdal
para lidar com objetos com atributos geométricos.

%% ------------------------------------------------------------------------- %%
\subsection{Linguagens de Programação para o Cálculo da Taxa de Sismicidade}
\index{Linguagens de Programação}
\label{sec:linguagens}

A principal linguagem de programação utilizada foi Python.

Os códigos-fonte dos modelos de Woo e de Helmstetter foram providos pelos autores
e estão em Fortran. O método de Frankel estava disponível em Python.

%% ------------------------------------------------------------------------- %%
\subsection{Programas de Computador para Cálculo de Ameaça Sísmica}
\index{Ferramentas!programas}\index{software}
\label{sec:software}

O programa mais comumente utilizado para análise de ameaça sísmica é o CRISIS \citep{crisis_2007} nas suas
mais diversas versões. Como nesse trabalho a opção foi por código aberto, poderia ter se optado pelo 
\gls{opensha}\footnote{\url{http://www.opensha.org}} que é escrito em Java.
A intenção no entanto foi provar o conjunto de tecnologias mantido pela fundação \gls{gem}
para o cálculo de ameaça e rísco sísmico que é o \gls{oq} \citep{pagani_2014}.

O código do \gls{oq} está abertamente disponível para consulta e clonagem no GitHub
\footnote{\url{https://github.com/GEM}} e seu \emph{ecossistema} é ilustrado na figura a seguir:

\begin{figure}[!h]
  \centering
  \includegraphics[width=.80\textwidth]{oq_ecosystem} 
  \caption{Ecossistema de módulos, bibliotecas e utilitários do OpenQuake}
  \label{fig:oq} 
\end{figure}

O destaque em vermelho na figura \ref{fig:oq} fica para o \gls{oqe} que executa as simulações de cada tronco da
arvore-lógica e as usa para calcular o cenário de risco e armazenar o resultado.
Faz isso distribuindo o processamento em tarefas e permitindo escalar o cálculo.

A \gls{oqp} permite interagir com os dados armazenados e gerenciados na nuvem, como catálogos, 
falhas, dados de esforços, fontes sismogênicas e suas geometrias usados em cálculos de ameaça/risco.
Permite também interagir com o resultado do cálculo, visualizando mapas de ameaça e curvas de intensidade.

%% ------------------------------------------------------------------------- %%
\subsection{Bibliotecas de Funções}
\index{ferramentas!bibliotecas}
\label{sec:bibliotecas}

Em amarelo na figura \ref{fig:oq} está a \gls{hl} que contém toda lógica e ciência para o cálculo de ameaça,
como os tipos de fontes sísmogênicas, as \gls{mfd}, as \glspl{gmpe}, etc.

Está também a \gls{rl}
que contém os modelos de vulnerabilidade e exposição para a análise de risco.

E por último,
a \gls{nrml} que é a sintaxe da linguagem de representação das árvores-lógicas, fontes sísmicas e
resultados, como mapas de risco, espectro de alguma medida de intensidade, etc.

%% ------------------------------------------------------------------------- %%
\subsection{Implementações e Novos Códigos}
\index{Implementações e novos códigos }
\label{sec:implementacao}

Incluído no suporte da fundação \gls{gem} está a manutenção de um Comitê Científico
que desenvolve ferramentas auxiliares adicionais como conversores, ferramentas para trabalho com 
o catálogo, ferramentas gráficas, entre outras para interação com o
\gls{oq} como pode ser visto na figura \ref{fig:oq} em azul, cinza e bege.

Uma dessas ferramentas em especial é o \gls{hmtk} \citep{weatherill_2012, weatherill_2014-1} que facilita todo o
processo de modelagem da \gls{psha} como a remoção de agrupamentos, a caracterização de zonas sísmicas,  
a visualização da evolução da taxa de sismicidade, a análise da magnitude de completude e
estimativa do \emph{valor-b}.

O \gls{hmtk} já trazia um módulo para trabalhar com a sismicidade que usava as \gls{smoothing}
implementando o método de \citet{frankel_1995}. E é na \gls{hmtk} que se pretende
contribuir com a implementação dos métodos de \citet{woo_1996} e de \citet{helmstetter_2012}.

%% ------------------------------------------------------------------------- %%
\section{Pré-Processamento}
\index{pré-processamento}
\label{sec:pre_processamento}

Para aplicar as \gls{smoothing} no conjunto de dados é necessário alguns primeiros procedimendos.

%% ------------------------------------------------------------------------- %%
\subsection{Controle de Qualidade}
\index{pre-processamento!controle de qualidade}
\label{sec:qualicontrol}

A primeira coisa a fazer no conjunto de dados é uma checagem geral.

Nessa hora é preciso observar se não há pontos com coordenadas erradas, invertidas, faltando valores de dias ou horas.

É recomendado também fazer uma varredura em busca de vieses no catálogo \citep{van_stiphout_2010}.
Isso é feito com, por exemplo, um histograma do dia da semana da ocorrência dos tremores em busca
de algum descréscimo em fins de semana que seriam um provável indicativo de contaminação do catálogo 
por atividade humana, por exemplo explosões em pedreiras. 
Também é possível fazer um histograma para observar a
distribuição do horário de ocorrência durante o dia, 
ou mesmo da profundidade para estimar a resolução do catálogo
nesse quesito. A figura \ref{fig:qc_histograms} apresenta alguns dos histogramas sugeridos para a análise
exploratória dos dados dos catálogos.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_sa3_weekday}
			\subcaption{Distribuição dos tremores nos dias da semana, \gls{iscgem}}
			\label{fig:sa_week_hist}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.45\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_bsb2013_weekday}
			\subcaption{Distribuição dos tremores nos dias da semana, \gls{bsb2013}}
			\label{fig:br_week_hist}
    \end{subfigure}%
        %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)

 	\begin{subfigure}[b]{0.45\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_sa3_hour}
			\subcaption{Distribuição do horário (GMT) de ocorrência dos tremores, \gls{iscgem}}
			\label{fig:sa_hour_hist}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.45\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_bsb2013_hour}
			\subcaption{Distribuição do horário (GMT) de ocorrência dos tremores, \gls{bsb2013}}
			\label{fig:br_hour_hist}
    \end{subfigure}%
        %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)

	\begin{subfigure}[b]{0.45\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{dep_sa_hist}
			\subcaption{Distribuição da profundidade dos tremores, \gls{iscgem}}
			\label{fig:sa_dep_hist}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.4\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{dep_br_hist}
			\subcaption{Distribuição da profundidade dos tremores, \gls{bsb2013}}
			\label{fig:br_dep_hist}
        \end{subfigure}%
        %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.

  \caption{Checagem de qualidade.}
  \label{fig:qc_histograms} 
\end{figure}

Nas figuras \ref{fig:sa_week_hist} e \ref{fig:br_week_hist} não se observam grandes vieses aparentes 
em relação aos dias da semana, nem aos horários de ocorrência durante o dia 
(figuras \ref{fig:sa_week_hist} e \ref{fig:br_week_hist}) nos dois catálogos.

No que diz respeito à distribuição de profundidade, o catálogo do ISC tem as
profundidades mais distribuidas, mesmo com grande maioria sendo eventos rasos 
também se observa um número razoável de sismos a mais de 600km de profundidade.
O BSB por sua vez é composto praticamente todo de sismos rasos, com menos de 10km
de profundidade, ou na prática, com profundidade desconhecida.

%% ------------------------------------------------------------------------- %%
\subsection{Conversão de Magnitudes}
\index{magnitudes!conversão}
\label{sec:mag_conv}

Para que a magnitude dos tremores possam ser analizadas em termos do momento sísmico e do 
tamanho da ruptura que geram é preciso que os sismos do catálogo apresentem pelo menos um valor
\gls{sym:MW} para a magnitude.

No catálogo do ISC-GEM isso é assunto resolvido pois as magnitudes \gls{sym:MW} já foram computadas \citep{storchak_2013}, mas no
caso do \gls{bsb} a maior parte das magnitudes são regionais $m_R$ que se assemelham às magnitudes $m_b$ 
ou magnitudes estimadas a partir de dados macrossísmicos. Nos dois casos é preciso avaliar funções que relacionem os
vários valores de magnitude e \gls{sym:MW}.

Para o catálogo \gls{bsb} as relações conhecidas\footnote{por Assumpção,M. e Drouet, S. (não-publicado)} para serem
usadas são as de equivalência entre $m_R$ e $m_b$:
\begin{equation}
	\ensuremath{
		m_R \equiv m_b,
	}
\label{eq:mrmb}
\end{equation}
as de conversão de $m_b$ em $\gls{sym:MW}$
\begin{equation}
	\ensuremath{
		M_W(m_b) = 1.12 m_b - 0.76,
	}
\label{eq:mwmb}
\end{equation}
as de conversão entre $A_f$ e $M_W$
\begin{equation}
	\ensuremath{
		M_W(A_f) = 0.6 + 0.8\log A_f,
	}
\label{eq:mwaf}
\end{equation}
onde \gls{sym:Af} é \glsdesc{sym:Af}; e por fim as estimativas de $m_b$ a partir da \glsdesc{sym:I0} (\gls{sym:I0})
presentes no catálogo.

NOTA: Essa etapa de pré-processamento, por simplicidade, não foi feita no escopo desse trabalho mas é aqui apresentada 
como referência para trabalhos futuros.

%% ------------------------------------------------------------------------- %%
\subsection{Remoção de agrupamentos}
\index{remoção de agrupamentos}\index{declustering}
\label{sec:declustering}

A maior parte dos modelos de sismicidade estudados assumem que a ocorrência de
tremores segue um processo de Poisson.

Mas sabe-se que os predecessores e sucessores de tremores principais não são independentes.
O aumento do número de sismos pouco antes e pouco depois de um grande tremor contamina
o catálogo com mais sismos do que seriam realmente esperados se os tremores fossem realmente independentes.
Por exemplo, logo após um sismo de magnitude 8, geralmente ocorrem durante um curto 
espaço de tempo mais alguns sismos de magnitude próximo a 7 e cerca de dezenas de sismos 
de magnitude menores que 5, que não ocorreriam com essa mesma frequência na ausência do sismo
principal, de magnitude 8.

Existem outros modelos que usam exatamente essa variação das taxas de sismicidade com o tempo
para fazerem projeções de curto-prazo, mas que não estão no escopo desse trabalho.

Os métodos mais difundidos para a remoção desses agrupamentos não-Poissonianos são métodos que consistem em definir
janelas de tempo e espaço em função da magnitude \citep{gardner_1974}, dentro das quais, os sismos são considerados como pertencentes ao mesmo
agrupamento.

A seguir estão três formulações implementadas no HMTK para a remoção de agrupamentos.
A primeira de \citet{gardner_1974} é 
\begin{equation}\begin{split} 
\mbox{d(m)} = &10^{0.1238 m + 0.983}\\
\mbox{t(m)} = & 
\begin{cases} 10^{0.032 m + 2.7389} & \text{se $m \geq 6.5$} \\ 
              10^{0.5409 m - 0.547} & \mbox{caso contrário}  \end{cases}\end{split}
\end{equation}

Uma formulação alternativa foi proposta por Gr\"unthal \citep{marsan_david_2012} :

\begin{equation}\begin{split} 
\mbox{d(m)} = & e^{1.77 + \left( {0.037 + 1.02 m} \right)^2} \\ 
   \mbox{t(m)} = & \begin{cases}   |e^{-3.95+ \left( {0.62 + 17.32 m}
    \right)^2}|    & \text{se $m \geq 6.5$ } \\ 10^{2.8 + 0.024 m} & 
    \text{caso contrário}  \end{cases}\end{split}
\end{equation}

E outra sugerida por \citet{uhrhammer_1986}
%
\begin{equation}
\mbox{d(m)} = e^{-1.024 + 0.804 m} \quad \mbox{t(m)} = 
    e^{-2.87 + 1.235 m}
\end{equation}

Enquanto os métodos acima definem explicitamente a janela de tempo, \citet{musson_1999}, propôs que as janelas de tempo
fossem janelas móveis em vez de fixas. Sua teoria é consistente com a de \citet{gardner_1974}.

A figura \ref{fig:eq_decluster_cum} apresenta os resultados de distintos métodos e janelas aplicados aos 
catálogos.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.48\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{decluster_sa}
			\subcaption{Número cumulativo de tremores registrados por ano para o \gls{iscgem}
			original e para diferentes métodos/janelas de remoção de agrupamentos.}
			\label{fig:sa_eq_record}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.48\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{decluster_br}
			\subcaption{Número cumulativo de tremores registrados por ano para o \gls{bsb2013}
			original e para diferentes métodos/janelas de remoção de agrupamentos.}
			\label{fig:br_eq_record}
    \end{subfigure}%
	\caption{Número cumulativo de tremores registrados por ano após 1900. A magnitude usada foi a do catálogo. 
	As diferentes respostas vêm das diferentes janelas utilizadas por cada método.}
	\label{fig:eq_decluster_cum}
\end{figure}

No caso do \gls{bsb2013} os resultados são relativamente equivalentes quando comparados aos resultados no caso do 
\glsdesc{iscgem}, onde é possível notar claramente as diferenças nos resultados de cada método de remoção de
agrupamento. Isso demonstra que uma discussão mais aprofundada sobre os métodos de 
remoção de agrupamentos é necessária mas está além dos propósitos desse texto.

A figura \ref{fig:eq_decluster} ilustra os agrupamentos de sismos não-Poissonianos encontrados 
pelo método de AFTERAN \citep{musson_2000} e janelas de Gr\"uenthal quando aplicados
aos catálogos \gls{iscgem} e \gls{bsb2013}. As sequencia de cores indica o número de cada agrupamento, e sua
distribuição espacial não apresenta qualquer padrão.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.46\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_sa3_pp_decluster}
			\subcaption{Número de tremores registrados por ano, \gls{iscgem}}
			\label{fig:sa_decluster}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[t]{0.50\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hmtk_bsb2013_pp_decluster}
			\subcaption{Número de tremores registrados por ano, \gls{bsb2013}}
			\label{fig:br_decluster}
    \end{subfigure}%
	\caption{Número de tremores registrados por ano após 1900}
	\label{fig:eq_decluster}
\end{figure}

Os catálogos com os agrupamentos removidos passam a ser os catálogos utilizados nas etapas posteriores.

%% ------------------------------------------------------------------------- %%
\subsection{Análise da Magnitude de Completude}
\index{Magnitude de Completude}
\label{sec:completeness}

Uma evidência da necessidade de se avaliar a magnitude de completude \gls{sym:Mc}
pode ser percebida ao contar a quantidade de sismos registrados ao longo do tempo
para cada intervalo de magnitude. A figura \ref{fig:qc_time_mag_count} apresenta
esses gráficos para os catálogos considerados.

\begin{figure}[H]
	  \centering
	  \begin{subfigure}[b]{0.7\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{time_mag_count_sa}
			\caption{Catálogo \gls{iscgem} 1900-2012}
			\label{fig:tmf_sa}
        \end{subfigure}%
        %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)

	  \begin{subfigure}[b]{0.7\textwidth}
		  	\centering
  			\includegraphics[width=1.00\textwidth]{time_mag_count_br}
			\caption{Catálogo \gls{bsb2013} 1900-2012}
			\label{fig:tmf_br}
       \end{subfigure}%

	   \begin{subfigure}[b]{0.7\textwidth}
		  	\centering
  			\includegraphics[width=1.00\textwidth]{time_mag_count_br_1960}
			\caption{Catálogo \gls{bsb2013} no período de 1960-2012}
			\label{fig:tmf_br_1960}
       \end{subfigure}%

  \caption{Contagem de sismos em tempo e em magnitude.}
  \label{fig:qc_time_mag_count} 
\end{figure}

Na figura \ref{fig:qc_time_mag_count} é possível perceber, nos dois catálogos,
que a quantidade de sismos no mesmo intervalo de magnitude varia bastante ao longo do tempo.
Isso se deve a uma série de fatores, mas principalmente à capacidade de registro
da rede sismográfica, implantação de redes globais e 
à modernização e aumento de sensibilidade de equipamentos.

A magnitude de completude é importante \citep{woessner_2005} por sua influência direta no cálculo dos parâmetros $a$ e
$b$ de Gutemberg-Richter.
Ela pode variar ao longo do tempo e do espaço (variação da densidade de estações sismográficas).

\begin{p}
\textbf{Stepp Test}
\end{p}

O método de \citet{stepp_1971} permite obter a variação temporal da magnitude
de completude \gls{sym:Mc} diretamente a partir do catálogo com um teste simples.

Esse procedimento analítico foi um dos primeiros a serem propostos para esse fim.
Sejam $\lambda_1, \lambda_2, \cdots, \lambda_n$ o número de sismos por unidade de tempo.
Assumindo que os tremores de certa faixa de magnitude têm distribuição de Poisson, a estimativa da taxa de
média de sismicidade $\lambda$ por unidade de tempo é expressa por

\begin{equation}
	\ensuremath{
		\lambda =  \frac{1}{n}\sum_{i=1}^{n} \lambda_i.
	}
\label{eq:mwaf}
\end{equation}
Sua variância é $\sigma_{\lambda}^2 =  \lambda/n$, onde $n$ o é número de intervalos.

Com intervalos de 1 ano, $n$ intervalos é o período $T$ de observação com desvio padrão
\begin{equation}
	\ensuremath{
		 \sigma_{\lambda} = \frac{\sqrt{\lambda}}{\sqrt{T}}.
	}
\label{eq:mwaf}
\end{equation}

A taxa de sismicidade sendo estacionária, seu desvio padrão deve se comportar como $1/\sqrt{T}$ em 
tempos de observação crescentes $T$.

Os diagramas abaixo tomam intervalos de observação crescentes com passo de 2.5 anos,
em intervalos de 0.5 de magnitude.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{stepp_sa}
			\subcaption{Diagrama de Stepp para o \gls{iscgem} (\emph{declustered})}
			\label{fig:sa_stepp}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.47\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{stepp_br}
			\subcaption{Diagrama de Stepp para o \gls{bsb2013} (\emph{declustered})}
			\label{fig:br_stepp}
    \end{subfigure}%
	\caption{Diagrama de Stepp para análise da magnitude de completude \gls{sym:Mc}}
	\label{fig:eq_stepp}
\end{figure}

Com esses parâmetros, as magnitudes de completude \gls{sym:Mc}, determinadas pelo método
de Stepp, estão listadas nas tabelas \ref{tab:mc_sa} e \ref{tab:mc_br}.

	\begin{table}[h]
	  	\centering
		\begin{tabular}{l|*{11}{c}}
		$M_c$ & 3.0  & 3.5  & 4.0  & 4.5  & 5.0  & 5.5  & 6.0  & 6.5  & 7.0  & 7.5  & 8 \\  \hline
		Ano   & 1986 & 1986 & 1986 & 1960 & 1958 & 1958 & 1927 & 1898 & 1885 & 1885 & 1885   \\
		\end{tabular}
		\caption{Magnitude de completude, \gls{iscgem}}
		\label{tab:mc_sa}
	\end{table}

	\begin{table}[h]
	  	\centering
		\begin{tabular}{l|*{7}{c}}
		$M_c$ & 3.0  & 3.5  & 4.0  & 4.5  & 5.0  & 5.5  & 6.0  \\  \hline
		Ano   & 1980 & 1975 & 1975 & 1965 & 1965 & 1860 & 1860 \\
		\end{tabular}
		\caption{Magnitude de completude, Cat.BSB2013}
		\label{tab:mc_br}
	\end{table}

Os parâmetros a serem utilizados podem ser controversos, inclusive a opção de forçar com que 
os anos de completude de magnitudes maiores sejam ao menos iguais aos de magnitude menores, 
mas novamente, a discussão sobre os melhores parâmetros para o método de Stepp, sua validade, ou 
mesmo outros métodos estão além da proposta desse texto.

Os valores de \gls{sym:Mc} das tabelas \ref{tab:mc_sa} e \ref{tab:mc_br} são os valores utilizados
no cálculo da recorrência no resto do trabalho.

%% ------------------------------------------------------------------------- %%
\section{Frankel, 1995}
\index{Frankel, 1995!processamento}
\label{sec:proc_frankel}

Tanto no método de Frankel como nos demais métodos avaliados,
a malha para a estimativa de taxa de sismicidade para o Brasil 
foi de $1^o \times 1^o$, o \emph{valor-b} foi fixado em 1.

Para o processamento pelo método de Frankel, utilizou-se o catálogo com os agrupamentos removidos,
e a tabela de magnitude de completude mencionada. A distância de correlação utilizada apenas como exemplo
foi de 150km a 68\% (totalizando uma influência total da ordem de 500km). A zona de influência foi limitada em 3 vezes a
distância de correlação.

%% ------------------------------------------------------------------------- %%
\section{Woo, 1996}
\index{Woo, 1996!processamento}
\label{sec:proc_woo}

O processamento pelo método de Woo também foi feito com a remoção dos agrupamentos no catálogo.

A largura de banda dependente da magnitude é ajustada pelo próprio método.
O ajuste da função de largura de banda é apresentado na figura~\ref{fig:woo_b} para o \gls{bsb2013}:

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{woo_bandwidth} 
  \caption{Ajuste da largura de banda para o método de Woo1996.
  	Os valores de 1.38 e 1.18 para $a_0$ e $a_1$ foram obtidos por mínimos quadrados.}
  \label{fig:woo_b} 
\end{figure}

Apenas como comparação estão apresentados também os ajustes citados por \citet{beauval_2003} para a Espanha e Noruega.
Como é possível observar o ajuste brasileiro apresenta distâncias médias sistematicamente maiores, provavelmente
devido à escala continental do Brasil ou ao baixo volume de registros.

A função de núcleo utilizada foi a proposta pela equação \eqref{eq:k_kj}.

%% ------------------------------------------------------------------------- %%
\section{Helmstetter, 2012}
\index{Helmstetter!processamento}
\label{sec:proc_helmstetter}

Utilizou-se, para a projeção da taxa de sismicidade, como catálogo de teste
o período 1950-2007. Os sismos de 2007-2012 juntamente
com os sismos ocorridos antes de 1950 foram colocados no catálogo-alvo.

A escolha por colocar os sismos anteriores a 1950 no catálogo se deve ao baixo número
de sismos no Brasil e de pouca capacidade histórica de observação, de forma que esses eventos trazem informações
importantes sobre essas fontes sismogênicas que não aparecem no catálogo no periodo escolhido para a aprendizagem.

Para a aprendizagem foram utilizados sismos com magnitudes acima de 3.8 buscando projeções para
sismos com magnitudes acima de 4.0. O pressuposto é de que sismos com magnitudes grandes
ocorrem onde geralmente ocorrem sismos de magnitude inferior.

A figura \ref{fig:h_catalogue} apresenta os catálogos de treino e de teste usados na otimização.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{helmstetter_catalogues} 
  \caption{Catálogos de aprendizado e de teste para o método de \citet{helmstetter_2012}}
  \label{fig:h_catalogue} 
\end{figure}

As magnitudes de completude, embora haja espaço no modelo para sua representação espacial e temporal
e, teoricamente, pudessem ser calculadas para o instante e posição de cada sismo para que fossem
considerados com os pesos adequados, não foram consideradas dessa forma. Em vez disso a magnitude
de completude utilizada foi a mínima magnitude no catálogo de teste ($M_c = M_d = 3.8$) 
uniformemente para todo o catálogo.

Os cálculos da largura de banda $h_i$ e $d_i$ de cada sismo é ilustrado no caso de um evento particular na figura
\ref{fig:h_hidi}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{helmstetter_hidi} 
  \caption{Exemplo da largura de banda para um determinado evento para o método de Helmstetter, com $k_{cnn} = 5$ e
  $a_{cnn} = 100$ arbitrários, para um determinado tremor $i$. A linha pontilhada vermelha marca a distância temporal
  $h_i$ e a linha pontilhada azul representa a distância espacial $d_i$}
  \label{fig:h_hidi} 
\end{figure}

Nota-se que as larguras de banda são escolhidas de forma a deixar $k_{cnn}$ eventos entre $h_i$ e $d_i$.
O mais importante é notar que como $k_{cnn} = 5$ existem exatamente 5 outros eventos (bolinhas) no retângulo
definido pelo estremo inferior esquerdo (0,0) e o superior direito ($h_i$,$d_i$). Fica evidente
que os valores determinados para $h_i$ e $d_i$ são os que, preservando ao menos 5 eventos, minimiza a soma
($h_i + a_{cnn}d_i$) mas que seriam outros valores caso a $a_{cnn}$ apontasse para outra proporção espaço-temporal.

Em seguida, na figura \ref{fig:hidi_hist} são apresentadas as distribuições de $h_i$ e $d_i$ calculadas
para os parâmetros já otimizados do modelo.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{hi_histogram}
			\subcaption{Distribuição das larguras de banda temporais}
			\label{fig:hi_hist}
	\end{subfigure}%
	\quad %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
	\begin{subfigure}[b]{0.47\textwidth}
		  	\centering
			\includegraphics[width=1.00\textwidth]{di_histogram}
			\subcaption{Distribuição das larguras de banda espaciais}
			\label{fig:di_hist}
    \end{subfigure}%
	\caption{Distribuição das larguras de banda calculadas com os parâmetros otimizados do modelo}
	\label{fig:hidi_hist}
\end{figure}

A taxa estacionária em cada célula é calculada a partir da mediana da taxa prevista pelo modelo em cada célula,
como na figura \ref{fig:h_stationary}.
\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{helmstetter_stationary_a} 
  \caption{Taxa de sismicidade estacionaria calculada a partir da mediana da taxa de sismicidade
  modelada pelo método de Helmstetter para uma determinada célula $r_0$}
  \label{fig:h_stationary} 
\end{figure}
A célula em questão fica próxima à região de Mato Grosso e o pico de sismicidade em torno de 1955,
ocorre justamente na época do grande tremor da região.

A função de núcleo utilizada na dimensão do espaço foi a gaussiana \eqref{eq:kernel_gs} multivariada, 
e univariada na dimensão do tempo.

%% ------------------------------------------------------------------------- %%
\section{Pós-Processamento}
\index{pós-processamento}
\label{sec:pos_proc}

Como o objetivo desse trabalho é a caracterização das fontes sísmogênicas,
o cálculo da ameaça pelo método clássico propriamente dito
acaba figurando como pós-processamento.

A figura \ref{fig:classical_psha} apresenta o fluxograma de trabalho para uma análise clássica de \gls{psha} conforme
implementado no Openquake \citep{pagani_2010, weatherill_2012}.

\begin{figure}[H]
	\centering
	\begin{tabular}{l}
	\includegraphics[width=0.70\textwidth]{classical_psha_workflow}
	\end{tabular}
	\caption{Fluxo de trabalho clássico para a \gls{psha} \citep{crowley_2013}.}
\label{fig:classical_psha}
\end{figure}

### 2015-01-13 c07-conclusoes.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Discussão}
\label{cap:conclusoes}

Apesar de possuirem implementações distintas de um mesmo fundamento matemático e estatítstico
que é a estimativa de densidade por funções de núcleo, os métodos, no geral, conseguiram
estimar das características da sismicidade a partir da sismicidade histórica catalogada.

Alguns distribuiram mais a taxa de sismicidade e a ameaça, uns destacando certas feições, outros outras.
Pode-se perceber que a caracterização da distribuição de frequência e magnitude, o uso da sua forma discreta 
ou truncada, pode ter influência no valor calculado para a ameaça e merece atenção.

Não parece haver um modelo mais correto, mesmo
porque todas as formulações são coerentes ao que se propuseram. Mesmo assim
os resultados, considerando suas diferentes proposições variam bastante.
O modelo de projeção de ocorrência de rupturas de Helmstetter talvez nem
fosse aplicável num universo de poucas amostras distribuídas em uma área tão vasta.

Essa caracterização suavizada da taxa de sismicidade, mesmo que executada de forma
distinta por critérios distintos, certamente terá relevãncia e deverá ser considerada
de alguma forma pelos sismólogos ou engenheiros modeladores de ameaça sísmica,
mesmo que apenas para cumprir certo percentual de uma árvore lógica.

Ficou claro também, mesmo não tendo sido o foco específico desse trabalho, 
não ser possível negligenciar a ocorrência de sismos profundos da zona de 
subducção para a ameaça sísmica do extremo oeste do país. Mesmo os sismos muito profundos (da ordem de centenas de km)
não provocando grandes impactos nas estruturas, sismos maiores e um pouco mais distantes podem ter razoáveis amplitudes em determinados perídos do espectro de aceleração.

O openquake como calculador de ameaças foi positivo do ponto de vista metodológico,
e o material bem documentado serviu de apoio e esclarecimento do cálculo.

O resultado discrepante entre os valores apresentados pelo Crisis-2007 e o Openquake,
precisam ser investigados com maior detalhe, mas os resultados do openquake foram compatíveis com o
modelo global, mesmo os valores calculados com o mesmo modelo de fontes sísmicas.

O HMTK se mostrou uma ferramenta essencial para a modelagem da ameaça. 
Muitas funcionalidades permitem facilmente a implementação dos principais fluxos de trabalho.

Por fim, talvez seja possível dizer que os objetivos propostos foram cumpridos.

%------------------------------------------------------
\section{Considerações Finais}

Foi possível de certa forma aplicar as técnicas de suavização, estimadoras das taxas suavizadas de sismicidade, 
gerar uma grade regular de fontes sísmicas pontuais, e então calcular a ameaça sísmica obtendo 
valores razoáveis, sem com isso definir zonas sísmicas, é o que se conhece também como métodos de \emph{zoneless}.

As diferenças nas formas de escolher a largura de banda das funções de núcleo de cada método,
podem talvez render ao método de Hemlstetter alguma vantagem, por ser localmente adaptável,
quando define claramente feições no nordeste. as outros métodos, com escolhas mais rígidas,
também o fizeram, privilegiando outras regiões.

O ajuste do método de Woo para o Brasil forneceu enormes larguras de banda, com cerca de 1500km para
as maiores magnitudes (são poucas e o método se baseia em vizinhos mais próximos), isso dilui
a influência dos grandes sismos nas taxas de sismicidade.

No caso do Brasil, essas primeiras estimativas sugerem que estudos de maior detalhe 
sejam feitos nas regiões de maior destaque.

%------------------------------------------------------
\section{Sugestões para Pesquisas Futuras}

Ainda há muito o que explorar.

A variação temporal e espacial da magnitude de completude no Brasil seria a principal fonte de informação 
para melhoria significativa da resposta dos métodos que dependem dessas correções para darem bons resultados.
Seu conhecimento permite também um mapeamento dos \emph{valores-b}. \citet{vorobieva_2013} aponta caminhos muito
interessantes.

Consideração de modelos que levem em consideração a distribuição da perda de tensão na forma de momento sísmico,
e mesmo a suavização da distribuição de momento acumulado espacialmente pela placa também podem ser outra alternativa.

Na modelagem da ameaça sísmica, um passo importante, seria trabalhar daqui por diante também com uma metodologia
para seleção das relações de atenuação.

Nas regiões de maior sismicidade, talvez seja possível estudar modelos onde a taxa de sismicidade varia com o tempo
como os modelos de sequência epidêmica de pós-abalos, que podem dar melhor resposta com pouco volume de dados.

Coletar incertezas e elaborar um cenário de ameaça,
detalhando a árvore lógica de possibilidades e variações
tanto nos modelos de fonte sísmica, incluíndo eventos históricos característicos, 
como nos modelos de atenuação.

Estudos de desagregação \citep{pagani_2007} também precisarão ser feitos no futuro.

### 2015-01-13 dissertacao.tex
adding images

% Arquivo LaTeX de exemplo de dissertação/tese a ser apresentados à CPG do IME-USP
% 
% Versão 5: Sex Mar  9 18:05:40 BRT 2012
%
% Criação: Jesús P. Mena-Chalco
% Revisão: Fabio Kon e Paulo Feofiloff
%  
% Obs: Leia previamente o texto do arquivo README.txt

\documentclass[12pt,twoside,a4paper]{book}

% ---------------------------------------------------------------------------- %
% Pacotes 
\usepackage[T1]{fontenc}
\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}           % usamos arquivos pdf/png como figuras
\usepackage{setspace}                   % espaçamento flexível
\usepackage{indentfirst}                % indentação do primeiro parágrafo
\usepackage{makeidx}                    % índice remissivo
\usepackage[nottoc]{tocbibind}          % acrescentamos a bibliografia/indice/conteudo no Table of Contents
\usepackage{courier}                    % usa o Adobe Courier no lugar de Computer Modern Typewriter
\usepackage{type1cm}                    % fontes realmente escaláveis
\usepackage{listings}                   % para formatar código-fonte (ex. em Java)
\usepackage{titletoc}
\usepackage{enumerate}
\usepackage{float}
%\usepackage[bf,small,compact]{titlesec} % cabeçalhos dos títulos: menores e compactos
\usepackage[fixlanguage]{babelbib}
\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\usepackage[font=footnotesize]{subcaption}
%\usepackage{subfigure}
\usepackage[usenames,svgnames,dvipsnames]{xcolor}
\usepackage[a4paper,top=2.54cm,bottom=2.0cm,left=2.0cm,right=2.54cm]{geometry} % margens
%\usepackage[pdftex,plainpages=false,pdfpagelabels,pagebackref,colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,filecolor=black,bookmarksopen=true]{hyperref} % links em preto
%\usepackage[pdftex,plainpages=false,pdfpagelabels,pagebackref,colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=green,bookmarksopen=true]{hyperref}
%hyperref
\usepackage[pdftex,plainpages=false,pdfpagelabels,pagebackref,colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=Green,bookmarksopen=true]{hyperref}
% links coloridos

\captionsetup[subcaption]{font=footnotesize}

% ---------------------------------------------------------------------------- %
% fontes e símbolos
\usepackage{amsfonts}
\usepackage{textcomp}

% multicols wrapfig
\usepackage{multicol}
\usepackage{wrapfig}

% ---------------------------------------------------------------------------- %
% Teoremas
\usepackage{amsthm}
\newtheorem{theorem}{Teorema}[section]
\newtheorem{lemma}[theorem]{\bf{Lema}}

% ---------------------------------------------------------------------------- %
% Equações Numeradas
\usepackage{amsmath}
\numberwithin{equation}{section}
%\numberwithin{equation}{subsection}

% % links coloridos
\usepackage[all]{hypcap}                    % soluciona o problema com o hyperref e capitulos
\usepackage[round,sort,nonamebreak]{natbib} % citação bibliográfica textual(plainnat-ime.bst)
\bibpunct{(}{)}{;}{a}{\hspace{-0.7ex},}{,} % estilo de citação. Veja alguns exemplos em http://merkel.zoneo.net/Latex/natbib.php

\fontsize{60}{62}\usefont{OT1}{cmr}{m}{n}{\selectfont}

% ---------------------------------------------------------------------------- %
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\erf}{erf}
% ---------------------------------------------------------------------------- %

% ---------------------------------------------------------------------------- %
% Glossarios, Acronimos e Simbolos, 
%\usepackage{hyperref}                    % 
\usepackage[sanitize=none,acronym,toc]{glossaries}
%\usepackage[acronym,toc]{glossaries}
% Define a new glossary type
\newglossary[slg,toc]{symbols}{sym}{sbl}{Lista de S\'imbolos}
\newglossary[elg]{equations}{eqn}{eql}{Equations}
\makeglossaries

%\usepackage{imakeidx}
\usepackage{imakeidx} % para o índice remissivo
\makeindex[intoc]

% ---------------------------------------------------------------------------- %
% Dicionario de Termos:
\input{d01-terms}       
\input{d02-symbols}       
\input{d03-acronyms}       
\input{d04-equations}       
% ---------------------------------------------------------------------------- %

% ---------------------------------------------------------------------------- %
% Cabeçalhos similares ao TAOCP de Donald E. Knuth
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{#1}}{}}
\renewcommand{\sectionmark}[1]{\markright{\MakeUppercase{#1}}{}}
\renewcommand{\headrulewidth}{0pt}

% ---------------------------------------------------------------------------- %
\graphicspath{{./images/}}             % caminho das figuras (recomendável)
\frenchspacing                          % arruma o espaço: id est (i.e.) e exempli gratia (e.g.) 
\urlstyle{same}                         % URL com o mesmo estilo do texto e não mono-spaced
\raggedbottom                           % para não permitir espaços extra no texto
\fontsize{60}{62}\usefont{OT1}{cmr}{m}{n}{\selectfont}
\cleardoublepage
\normalsize

% ---------------------------------------------------------------------------- %
% Opções de listing usados para o código fonte
% Ref: http://en.wikibooks.org/wiki/LaTeX/Packages/Listings
\lstset{ %
language=Python,                % choose the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=single,	                % adds a frame around the code
framerule=0.6pt,
tabsize=2,	                    % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)},         % if you want to add a comment within your code
backgroundcolor=\color[rgb]{1.0,1.0,1.0}, % choose the background color.
rulecolor=\color[rgb]{0.8,0.8,0.8},
extendedchars=true,
xleftmargin=10pt,
xrightmargin=10pt,
framexleftmargin=10pt,
framexrightmargin=10pt
}

% ---------------------------------------------------------------------------- %
% Corpo do texto
\begin{document}
\frontmatter 
% cabeçalho para as páginas das seções anteriores ao capítulo 1 (frontmatter)
\fancyhead[RO]{{\footnotesize\rightmark}\hspace{2em}\thepage}
\setcounter{tocdepth}{2}
\fancyhead[LE]{\thepage\hspace{2em}\footnotesize{\leftmark}}
\fancyhead[RE,LO]{}
\fancyhead[RO]{{\footnotesize\rightmark}\hspace{2em}\thepage}

\onehalfspacing  % espaçamento

% ---------------------------------------------------------------------------- %
% CAPA
% Nota: O título para as dissertações/teses do IME-USP devem caber em um 
% orifício de 10,7cm de largura x 6,0cm de altura que há na capa fornecida pela SPG.
\thispagestyle{empty}
\begin{center}
    \vspace*{2.3cm}
    \textbf{\Large{\glsdesc*{titulo_da_dissertacao}}}\\

    \vspace*{1.2cm}
    \Large{Marlon Pirchiner}

    \vskip 2cm
    \textsc{
    Dissertação apresentada à\\[-0.25cm] 
    Escola de Matemática Aplicada da\\[-0.25cm]
    Fundação Getúlio Vargas-RJ\\[-0.25cm]
    para obtenção do título de \\[-0.25cm]
    Mestre em Ciências}

    \vskip 1.5cm
    Programa: Modelagem Matemática de Informação\\
    Orientador: Prof. Dr. Vincent Guigues\\
    %Coorientador: Prof. Dr. Stephane Drouet

   	\vskip 1cm
    \normalsize{}

    \vskip 0.5cm
    \normalsize{Rio de Janeiro, agosto de 2014}
\end{center}

% ---------------------------------------------------------------------------- %
% Página de rosto (SÓ PARA A VERSÃO DEPOSITADA - ANTES DA DEFESA)
% Resolução CoPGr 5890 (20/12/2010)
%
% IMPORTANTE:
%   Coloque um '%' em todas as linhas
%   desta página antes de compilar a versão
%   final, corrigida, do trabalho
%
%\newpage
%\thispagestyle{empty}
%    \begin{center}
%        \vspace*{2.3 cm}
%	    \textbf{\Large{\glsdesc*{titulo_da_dissertacao}}}\\
%        \vspace*{2 cm}
%    \end{center}

%    \vskip 2cm

%    \begin{flushright}
%	Esta é a versão original da dissertação elaborada pelo\\
%	candidato Marlon Pirchiner, tal como \\
%	submetida à Comissão Julgadora.
%   \end{flushright}

%\pagebreak

% ---------------------------------------------------------------------------- %
% Página de rosto (SÓ PARA A VERSÃO CORRIGIDA - APÓS DEFESA)
% Resolução CoPGr 5890 (20/12/2010)
%
% Nota: O título para as dissertações/teses do IME-USP devem caber em um 
% orifício de 10,7cm de largura x 6,0cm de altura que há na capa fornecida pela SPG.
%
% IMPORTANTE:
%   Coloque um '%' em todas as linhas desta
%   página antes de compilar a versão do trabalho que será entregue
%   à Comissão Julgadora antes da defesa
%
%
\newpage
\thispagestyle{empty}
     \begin{center}
         \vspace*{2.3 cm}
         \textbf{\Large{\glsdesc*{titulo_da_dissertacao}}}\\
         \vspace*{2 cm}
     \end{center}
% 
     \vskip 2cm
% 
     \begin{flushright}
 	Esta versão da dissertação contém as correções e alterações sugeridas\\
 	pela Comissão Julgadora durante a defesa da versão original do trabalho,\\
 	realizada em 15/07/2014. 
     \vskip 2cm
% 
     \end{flushright}
     \vskip 4.2cm
% 
     \begin{quote}
     \noindent Comissão Julgadora:
%     
     \begin{itemize}
 		\item Vincent Guigues (orientador) - EMAp/FGV-RJ 
 		\item Flávio Codeço Coelho - EMAp/FGV-RJ 
 		\item Marcelo Assumpção - IAG-USP 
 		\item Moacyr Alvim Horta Barbosa da Silva - EMAp/FGV-RJ 
 		\item Stéphane Drouet - ON 
     \end{itemize}
%       
     \end{quote}
 \pagebreak

% CATALOGRAPHIC
\newpage
\thispagestyle{empty}

	\begin{figure}[H]
	  \centering
	  \includegraphics[width=\textwidth]{_catalographic} 
	\end{figure}

\pagebreak

% SIGNATURES
\newpage
\thispagestyle{empty}

	\begin{figure}[H]
	  \centering
	  \includegraphics[width=\textwidth]{_signatures} 
	\end{figure}

\pagebreak

\pagenumbering{roman}     % começamos a numerar

% ---------------------------------------------------------------------------- %
% Agradecimentos:
% Se o candidato não quer fazer agradecimentos, deve simplesmente eliminar esta página 
\chapter*{Agradecimentos}
Aos meus familiares e amigos pela benevolência de sempre.

À Didi.

A todos do Grupo de Sismologia (e também a todo pessoal) do Instituto de
Astronomia Geofísica e Ciências Atmosféricas (IAG) da Universidade de São Paulo (USP) 
por todo apoio e suporte de sempre e durante o tempo em que
estive entre o curso de mestrado e o trabalho.

Aos companheiros e professores pelas conversas e discussões ao longo do curso.

% ---------------------------------------------------------------------------- %
% Resumo
\chapter*{Resumo}

A avaliação de risco sísmico, fundamental para as decisões sobre as estruturas de obras de engenharia
e mitigação de perdas, envolve fundamentalmente a análise de ameaça sísmica. 
Calcular a ameaça sísmica é o mesmo que calcular a probabilidade de que certo nível de 
determinada medida de intensidade em certo local durante um certo tempo seja excedido.

Dependendo da complexidade da atividade geológica essas estimativas podem ser bastante sofisticadas.
Em locais com baixa sismicidade, como é o caso do Brasil, o pouco tempo (geológico) de observação
e a pouca quantidade de informação são fontes de muitas incertezas e dificuldade de análise pelos métodos mais clássicos
e conhecidos que geralmente consideram, através de opiniões de especialistas, determinadas zonas sísmicas.

Serão discutidas algumas técnicas de suavização e seus fundamentos como métodos alternativos ao zoneamento,
em seguida se exemplifica suas aplicações no caso brasileiro.

\noindent \textbf{Palavras-chave:} terremotos, ameaça sísmica, suavização.

% ---------------------------------------------------------------------------- %
% Abstract
\chapter*{Abstract}
\noindent PIRCHINER, M. \textbf{Smoothing seismicity techniques applied to probabilistic seismic
hazard analysis in Brazil}.
2014.
Master thesis - Applied Math School,
Getúlio Vargas Foundation, Rio de Janeiro, 2014.
\\

Seismic risk assesment is crucial to make better decisions about engineering structures and loss mitigation.
It involves, mainly, the evaluation of seismic hazard.
Seismic hazard assesment is the computation of probability which the level of some ground
motion intensity measure, in a given site, which, in some time window, will be exceeded.

Depending on geological and tectonic complexity, the seismic hazard evaluation becomes more sofisticated.
At sites with low seismicity, which is the brazilian case, the relative (geologically) low observation time
and the lack of earthquake and tectonic information, increases the uncertainties and makes more difficult
the standard analysis, which in general, consider expert opinions, to characterize the
seismicity into disjoint zones.

This text discusses some smoothing seismicity techniques and their theoretical foundations as alternative
methods for seismicity characterization. Next, the methods are exemplified in the brazilian context.

\noindent \textbf{Keywords:} earthquakes, seismic hazard, smoothing.

% ---------------------------------------------------------------------------- %
% Sumário
\tableofcontents    % imprime o sumário

% Lista de Abreviaturas
\printglossary[type=\acronymtype,title=Lista de Abreviaturas]

% Lista de Simbolos
\printglossary[type=symbols,title=Lista de S\'imbolos]

%\printglossaries
%\printglossaries

% ---------------------------------------------------------------------------- %
%\chapter{Lista de Abreviaturas}
%\begin{tabular}{ll}
%         CFT         & Transformada contínua de Fourier (\emph{Continuous Fourier Transform})\\
%         DFT         & Transformada discreta de Fourier (\emph{Discrete Fourier Transform})\\
%        EIIP         & Potencial de interação elétron-íon (\emph{Electron-Ion Interaction Potentials})\\
%        STFT         & Tranformada de Fourier de tempo reduzido (\emph{Short-Time Fourier Transform})\\
%\end{tabular}

% ---------------------------------------------------------------------------- %
%\chapter{Lista de Símbolos}
%\begin{tabular}{ll}
%        $\lambda$   & Seismic Rate\\
%        $\psi$      & Função de análise \emph{wavelet}\\
%        $\Psi$      & Transformada de Fourier de $\psi$\\
%\end{tabular}

% ---------------------------------------------------------------------------- %
% Listas de figuras e tabelas criadas automaticamente
\listoffigures            
\listoftables

% ---------------------------------------------------------------------------- %
% Capítulos do trabalho
\mainmatter

% cabeçalho para as páginas de todos os capítulos
\fancyhead[RE,LO]{\thesection}

\singlespacing              % espaçamento simples
%\onehalfspacing            % espaçamento um e meio

% ============================================================================ %

\input{c01-introducao}        % associado ao arquivo: 'cap-introducao.tex'
\input{c02-conceitos}         % associado ao arquivo: 'cap-conceitos.tex'
\input{c03-regiao}        	  % associado ao arquivo: 'cap-conclusoes.tex'
\input{c04-metodologia}       % associado ao arquivo: 'cap-introducao.tex'
\input{c05-processamento}     % associado ao arquivo: 'cap-conceitos.tex'
\input{c06-resultados}        % associado ao arquivo: 'cap-conclusoes.tex'
\input{c07-conclusoes}        % associado ao arquivo: 'cap-conclusoes.tex'

% ============================================================================ %

% cabeçalho para os apêndices
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\appendixname\ \thechapter}} {\MakeUppercase{#1}} }
\fancyhead[RE,LO]{}
\appendix

% ============================================================================ %
%\include{a01-conjuntos}      % associado ao arquivo: 'ape-conjuntos.tex'
% ============================================================================ %

% ---------------------------------------------------------------------------- %
% Bibliografia
\backmatter \singlespacing   				% espaçamento simples
\bibliographystyle{styles/plainnat-ime} 	% citação bibliográfica textual
\bibliography{bib/bibliografia}  			% associado ao arquivo: 'bibliografia.bib'

% ---------------------------------------------------------------------------- %
% Índice remissivo
%\index{TBP|see{periodicidade região codificante}}
%\index{DSP|see{processamento digital de sinais}}
%\index{STFT|see{transformada de Fourier de tempo reduzido}}
%\index{DFT|see{transformada discreta de Fourier}}
%\index{Fourier!transformada|see{transformada de Fourier}}

% Glossário
\printglossary[title=Gloss{á}rio,toctitle=Gloss{á}rio]

% índice remissivo no documento 
\printindex

\end{document}

### 2015-01-13 c02-conceitos.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Conceitos}
\label{cap:conceitos}

Este capítulo apresenta, um a um, os conceitos mais elementares
e tenta harmonizar a terminologia empregada no decorrer do texto.

%% ------------------------------------------------------------------------- %%
\section{\Gls{tectonic}}
\index{\gls{tectonic}}
\label{sec:02_tectonica}

A \gls{tectonic} é \glsdesc*{tectonic}.

Uma das principais evidências das transformações geológicas do planeta
são os \glspl{equake}. A figura \ref{f:global_epicenters} \citep{lowman_jr_1998}
é um mapa global com a ocorrência geográfica dos tremores. Nele é possivel notar que
os sismos não são distribuídos uniformemente pelo globo.

\begin{figure}[H]
   \centering
   \includegraphics[width=0.80\textwidth]{global_pde_mag_all}
   \caption{Mapa Mundial de Epicentros}
   \label{f:global_epicenters}
\end{figure}

O padrão apresentado pela \gls{seismic_activity} global foi essencial
para o desenvolvimento posterior da \gls*{tectonic_plate_theory}.

%% ------------------------------------------------------------------------- %%
\subsection{\Gls{tectonic_plate_theory}}
\index{\Gls{tectonic}!\Gls{tectonic_plate_theory}}
\label{sec:02_placas}

A \gls*{tectonic_plate_theory}, desenvolvida na segunda metade do século XX,
cartografava na superfície do globo as \glspl{litho_plate}.

\begin{figure}[H]
   \centering
   \includegraphics[width=0.80\textwidth]{litho_plates_overview}
   \caption[Cartografia das placas litosféricas]
   		   {Cartografia das placas litosféricas}
   \label{f:plates_overview}
\end{figure}

As \glspl{litho_plate}, como pode ser visto na figura \ref{f:plates_overview} \citep{usgs_plates_1996},
e o conceito de \gls{astenosphere} (\glsdesc{astenosphere})
surgem para conformar uma teoria capaz de explicar
uma série de fenômenos tectônicos observados e ainda não bem explicados naquela época.

%% ------------------------------------------------------------------------- %%
\subsubsection{Bordas}
\index{\Gls{tectonic_plate_theory}!bordas}
\label{sec:02_bordas}

Nas bordas das \glspl{litho_plate} a tectônica é mais intensa
provocando uma enorme diversidade de fenômenos geológicos de acordo
com o tipo de interação. Algumas delas estão esquematizadas na
figura \ref{f:plate_boundaries} \citet{vigil_1997}.

\begin{figure}[H]
   \centering
   \includegraphics[width=0.95\textwidth]{plate_boundaries}
   \caption[Diferentes tipos de interações entre \glspl{litho_plate} em suas bordas]
   		   {Diferentes tipos de interações entre \glspl{litho_plate} em suas bordas}
   \label{f:plate_boundaries}
\end{figure}

Na figura \ref{f:plate_boundaries} estão ilustrados os diferentes tipos de interação
entre as \glspl{litho_plate} nas suas bordas, que causam, como já se sabe, a maior
parte dos \glspl{equake} e vulcanismo.

Só na borda das placas é liberada cerca de 95\% da quantidade total da energia
disseminada na forma de \glspl{equake} no globo.

%% ------------------------------------------------------------------------- %%
\subsubsection{Interior}
\index{\Gls{tectonic_plate_theory}!interior}
\label{sec:02_interior}

A dificuldade é explicar, com maior detalhe, porque e como são liberados os outros
5\% do total de energia em \glspl{equake}, mais raros, no interior das \glspl{litho_plate}.

Não há pleno consenso nem um modelo geral para a explicação do mecanismo de ocorrência dos
sismos no interior das placas \citep{talwani_2014} embora sejam conhecidas diversas zonas sísmicas em regiões no
interior de placas que apresentam sismicidade importante com sismos cujas
magnitudes em alguns casos foram superior a 7, como em Nova Madrid, nos Estados Unidos
e em locais da China e da Austrália para citar alguns outros.

%% ------------------------------------------------------------------------- %%
\subsection{Sismotectônica}
\index{\gls*{seismotectonic}}
\label{sec:sismotectonica}

A \gls{seismotectonic} é \glsdesc*{seismotectonic}.

Na prática consiste por um lado, num esforço de compreensão dos
processos geológicos através da observação dos tremores e analogamente, compreender os tremores através da observação
de processos geológicos mensuráveis.

É fácil notar a contribuição dessa disciplina para a análise de sismicidade e vice-versa.

\section{Probabilidade}
\index{probabilidade}
\label{sec:probabilidade}

\subsection{\Glsdesc{pdf}}
\index{\glsdesc{pdf}}
\label{sec:pdf}

A \gls{pdf} de uma \gls{va}
descreve a probabilidade de que essa \gls{va}
assuma, entre todas as realizações possíveis, uma em especial.

Seja $X$, uma \gls{va} unidimensional. A \gls{pdf}
$f_X(x)$ de $X$ é definida por

\begin{equation}
	P \left\{ X \in [x_0,x_1[\, \right\} = \int_{x_0}^{x_1}\!f_X(x)\,\mathrm{d}x, \; \forall x_0 \leq x_1 \in \mathbb{R}.
	\label{eq:pdf}
\end{equation}

Para que uma função possa assumir o papel de \gls{pdf} é necessário que ela
possua as seguites propriedades:
\begin{enumerate}[(i)]
	\item $f_X(x) \ge 0\;\forall x$  (a função $f_X$ deve ser sempre positiva), e
	\item $\int_{-\infty}^{+\infty} f_X(x) \mathrm{d}x = 1$ (e deve somar, sobre todos os valores possíveis, a unidade).
\end{enumerate}

\subsection{\Glsdesc{pmf}}
\index{\glsdesc{pmf}}
\label{sec:pmf}

Outro conceito importante e diretamente relacionado à \gls{pdf} é a
\gls{pmf}.

No caso da \gls{va} $X$, sua \gls{pmf} $F_X(x)$ é definida como

\begin{equation}
	P \left\{ X \leq x\right\} = F_X(x) = \int_{-\infty}^{x}\!f_X(u)\,\mathrm{d}u.
	\label{eq:pmf}
\end{equation}

\subsection{Histograma}
\index{histograma}
\label{sec:histogram}

Quando a \gls{pdf} de uma \gls{va} não é conhecida e se deseja estudar seu comportamento
é preciso estimá-la e para isso o histograma é uma das técnicas mais antigas e amplamente utilizadas.

O histograma divide o universo das observações, possíveis realizações $X_1, X_2,\cdots, X_n$
da \gls{va}, em compartimentos (\emph{bins}).

Dados uma origem arbitrária $x_0$ e uma largura $h$ de cada um, os compartimentos
são definidos como os intervalos $[x_0 + (j -1)h,\; x_0 + jh[$
com $j\in\mathbb{Z}$, um identificador para cada um deles.

Considere um determinado intervalo $[-h/2, h/2[$.
A probabilidade (\eqref{eq:pdf}) de que uma observação qualquer venha a pertencer a esse intervalo é

\begin{equation}
	P \left\{ X \in [-h/2,h/2[ \; \right\} = \int_{-h/2}^{h/2}\!f_X(x)\,\mathrm{d}x.
	\label{eq:hist01}
\end{equation}

E um estimador natural $\hat{f}_X$ para a densidade $f_X$ seria contar o número de observações

\begin{equation}
	P \left\{ X \in [-h/2,h/2[ \; \right\} \approx \frac{\# \left\{ X_i \in [-h/2,h/2[ \; \right\}}{n} =
	\int_{-h/2}^{h/2}\!\hat{f}_X(x)\,\mathrm{d}x ,
	\label{eq:hist02}
\end{equation}
de onde
\begin{equation}
	\hat{f}_X(x) = \frac{\# \left\{ X_i \in [-h/2,h/2[ \; \right\}}{nh},
	\label{eq:hist_03}
\end{equation}
para todo $x \in [-h/2,h/2[$.

De modo geral, sejam $X_1, \cdots, X_n$ observações \gls{iid} da \gls{va} $X$ com densidade desconhecida $f$.
Considere $N_I$ intervalos de comprimento $h$ e o conjunto de compartimentos $C_j = [x_0 + (j -1)h,\; x_0 + jh[,\;
j=1..N_I$.
Defina
\[
	I_A(x) := \begin{cases}
		1 & \text{se } x \in A \\
		0 & \text{caso contrário}
	\end{cases}
\]
e
\[	n_j := \sum_{i=1}^{n}I_{C_j}(X_i)\; \text{tal que} \;  \sum_{j=1}^{N_I}n_j = n. \]

Dessa forma a estimativa $\hat{f}$ parametrizada pela largura $h$ para a densidade $f$ seria
\begin{equation}
	\hat{f}(x \arrowvert\, h) = \frac{1}{nh} \sum_{j=1}^{N_I}n_jI_{C_j}(x)
	\label{eq:hist_04}
\end{equation}
para toda realização possível $x$ de $X$.

\section{Sismicidade}
\index{sismicidade}
\label{sec:sismicidade}

A sismicidade é a ocorrência dos tremores de terra. Como, quando, onde, de que tamanho?

É sabido que pequenos abalos são mais frequentes que os tremores de terra
muito grandes e catastróficos cujos registros são extremamente raros.

A figura \ref{f:m9} apresenta os sismos de magnitude acima de nove conhecidos.

\begin{figure}[H]
   \centering
   \includegraphics[width=0.5\textwidth]{M9}
   \caption[Sismos com magnitude acima de 9,0.]
   		   {Sismos com magnitude acima de 9,0. Fonte: \gls{isc}}
   \label{f:m9}
\end{figure}

Tremores de terra, abalos, \glspl{equake}, sismos são a ocorrência de
fenômenos geológicos de ruptura, instantânea, por certo mecanismo, de certa dimensão, na
crosta terrestre.

\subsection{Ocorrência}
\index{\gls{equake}!ocorrência}
\label{sec:ocorrencia}

Os tremores acontecem por uma ruptura geológica (figura \ref{f:rupture})
num instante \gls{sym:t}, num lugar \gls{sym:r} e cada um
com sua magnitude \gls{sym:m} associada.

\begin{figure}[H]
   \centering
   \includegraphics[width=0.50\textwidth]{rupture}
   \caption[Ilustração da área de ruptura em um falhamento geológico]
   		   {Ilustração da área de ruptura em um falhamento geológico\footnotemark}
   \label{f:rupture}
\end{figure}
\footnotetext{\citet{opensha_team_2010}}

O local em que se iniciou a ruptura que deu origem ao tremor é um \gls{hypocenter},
enquanto sua projeção na superfície, desconsiderando-se a profundidade, é o \gls{epicenter}.

\subsection[Magnitude]{Magnitude (da ruptura)}
\index{magnitude}
\label{sec:magnitude}

A magnitude de um tremor de terra é um valor medido numa escala que versa sobre a energia liberada pelo sismo.
Essa energia é proporcional à área rompida e ao deslocamento geológico relativo entre os blocos de rocha na superficie
de ruptura.

O desenvolvimento experimental de escalas de magnitude, para medir o tamanho dos tremores,
é marcado pelo trabalho do sismólogo Charles \citet{richter_1935}. Existem, entretanto,
uma série de diferentes escalas de magnitude, baseadas em diversos tipos de medidas.

A escolha de qual usar fica a critério
de cada sismólogo (ou analista) e de cada rede sismográfica.
Geralmente usam escalas diferentes
para avaliar a magnitude dos tremores ou até mesmo divulgam mais de um tipo de magnitude para
um mesmo evento.

As escalas são calibradas para fornecerem valores similares, de acordo com
o intervalo de utilidade para o qual foram desenvolvidas, mas apresentam diferenças consideráveis para um mesmo evento.
Isso pode comprometer as análises estatísticas baseadas em um catálogo
cujas magnitudes não tenham sido calculadas de maneira uniforme.

\subsubsection{Magnitude Richter}
\index{magnitude!Richter}
\label{sec:magnitude_richter}

As escalas de magnitude mais comuns são as que derivam da definição de \citet{richter_1935}
fundada na relação empírica entre o logarítmo da amplitude do registro das ondas sísmicas e a distância onde foram
registradas. Em 1935 Richter notou a seguinte proporcionalidade

\begin{equation}
	\gls{eqn:richter},
	\label{eq:richter}
\end{equation}
\glsdesc*{eqn:richter}.

A amplitude máxima de sua
escala foi definida pela amplitude máxima observada em um sismômetro Wood-Anderson, com período de 0.8s, registrando a 100km
do tremor.

Algumas correções poderiam ser cogitadas, principalmente pelo fato da escala estar intimamente
relacionada a um determinado equipamento, hoje obsoleto, e porque sismos locais (a menos de 100km) têm sua magnitude
melhor calculada usando frequências mais altas que as registráveis pelo sismômetro da época.

Outras escalas foram desenvolvidas a partir da medida da amplitude máxima de determinadas fases
de diferentes tipos de onda sísmica e apresentam bons resultados para a maior parte dos sismos,
mas não refletem, com precisão,
o tamanho dos maiores e mais destrutivos eventos, com magnitude acima de 7 ou 8 porque geralmente saturam.

\subsubsection{Magnitude de Momento Sísmico}
\index{magnitude!momento sísmico}
\label{sec:risco_sismico}

O evento de natureza sismológica ocorre num
instante \gls{sym:t} liberando uma certa quantidade de energia na forma de \glsdesc{sym:M_0}
\gls{sym:M_0}. A \glsdesc{sym:MW} \gls{sym:MW} desse evento é proporcional ao logarítmo dessa energia gls{sym:M_0}.

O \glsdesc{sym:M_0} é apresentado na equação \eqref{eq:M_0}:

\begin{equation}
	\gls{eqn:M_0}
	\label{eq:M_0}
\end{equation}
\glsdesc*{eqn:M_0}.

O momento sísmico é estimado geralmente pela inversão duplamente acoplada de um tensor de momento aos registros em
forma de onda do movimento do chão causado pelo terremoto. Ou, em casos de tremores muito bem registrados, ele pode
ser estimado a partir de algum modelo numérico para a ruptura.

A \glsdesc{sym:MW} \gls{sym:MW} \citep{hanks_1979} é baseada no
logarítmo do \glsdesc{sym:M_0} \gls{sym:M_0}, e não se satura no caso de grandes eventos.
Sua definição é dada pela equação \eqref{eq:M_W}

\begin{equation}
	\gls{eqn:M_W}
	\label{eq:M_W}
\end{equation}
\glsdesc*{eqn:M_W}.

\subsubsection{Intensidade Macrossísmica}
\index{instensidade macrossísmica}
\label{sec:intensidade}

A intensidade macrossísmica é uma escala para medir, não a energia proporcional
à ruptura que originou o tremor de terra, mas para retratar a percepção humana, e os efeitos sobre
construções e objetos, do movimento do chão onde quer este tenha produzido seus efeitos.

Uma das mais difundidas é a escala Modificada de Mercalli \citep{richter_1958} apresentada em sua versão simplificada
na tabela \ref{tab:mercalli}:

\begin{table}[H]
\begin{center}
\begin{scriptsize}
\noindent\begin{tabular}{c|c|p{12cm}}
\hline
Categoria  	& Sensação & Efeitos \\
\hline
I 			&	Imperceptível 	&	Não sentido. Apenas registado pelos sismógrafos.
\\II 		&	Muito fraco 	&	Sentido por um muito reduzido número de pessoas em
								repouso, em especial pelas que habitam em andares
								elevados.
\\III 		&	Fraco 			&	Sentido por um pequeno número de pessoas. Bem sentido nos andares elevados.
\\IV 		&	Moderado 		&	Sentido dentro das habitações, podendo despertar do sono um pequeno número de pessoas.
								Nota-se a vibração de
								portas e janelas e das loiças dentro dos armários.
\\V 		&	Forte 			&	Praticamente sentido por toda a população, fazendo acordar muita gente.
								Há queda de alguns objectos menos estáveis e param os pêndulos dos relógios.
								Abrem-se pequenas fendas nos estuques das paredes.
\\VI 		&	Bastante forte 	&	Provoca início de pânico nas populações. Produzem-se leves danos nas habitações,
								caindo algumas chaminés. O mobiliário menos pesado é deslocado.
\\VII 		&	Muito forte 	&	Caem muitas chaminés. Há estragos limitados em edifícios de boa construção,
								mas importantes e generalizados nas construções mais frágeis.
								Facilmente perceptível pelos condutores de veículos automóveis em trânsito.
								Desencadeia pânico geral nas populações.
\\VIII 		&	Ruinoso  		&	Danos acentuados em construções sólidas. Os edifícios de muito boa construção
								sofrem alguns danos. Caem campanários e chaminés de fábricas.
\\IX 		&	Desastroso 		&	Desmoronamento de alguns edifícios. Há danos consideráveis em construções muito sólidas.
\\X 		&	Destruidor 		&	Abrem-se fendas no solo. Há cortes nas canalizações, torção nas vias de caminho
								de ferro e empolamentos e fissuração nas estradas.
\\XI 		&	Catastrófico 	&	Destruição da quase totalidade dos edifícios, mesmo os mais sólidos.
								Caem pontes, diques e barragens. Destruição das redes de canalização e das vias de comunicação.
								Formam-se grandes fendas no terreno, acompanhadas de desligamento. Há grandes escorregamentos de terrenos.
\\XII 		&	Cataclismo 		&	Destruição total. Modificação da topografia. Nunca foi presenciado no período histórico. \\
\hline
\end{tabular}
\caption{Escala simplificada de intensidade sísmica,
modificada em 1956 derivada da escala original de Giuseppe Mercalli de 1902.}
\label{tab:mercalli}
\end{scriptsize}
\end{center}
\end{table}

Existem estudos \citep{bakun_1999} que propõem a inferência sobre o tamanho da ruptura, e sua
magnitude, a partir de observações macrossísmicas, ou dos efeitos relatados pela escala de intenside, georreferenciados.

%% ------------------------------------------------------------------------- %%
\subsection{Catálogos}
\index{catálogos}
\label{sec:catalogos}

Os catálogos podem ser vistos como uma coleção de parâmetros sobre os tremores.
Podem ser classificados em três categorias \citep{woessner_2010} enumeradas a seguir:

\begin{itemize}\setlength{\itemsep}{0em}
	\item Pré-históricos: baseados na coleta de dados feitas por
	geólogos estruturais em trincheiras ou campos de subsidência. Podem conter registros de tremores que ocorreram
	há milhares de anos.
	\item Históricos: catálogos formados a partir de relatos históricos e inferência de valores de intensidade
	(seção \ref{sec:intensidade}), de análises de forma de onda com instrumentos antigos (registros em papel), eventualmente
	digitalizados.
	Cobrem o período das primeiras descrições humanas até os catálogos intrumentais.
	\item Instrumentais: são os catálogos de sismicidade definidos por dados produzidos por uma rede sismográfica bem estabelecida
	 gerando localizações continuamente (que começam a existir a partir de 1960).
\end{itemize}

Os catálogos instrumentais são uma listagem onde se epera encontrar, para cada evento, as
seguintes informações:

\begin{itemize}\setlength{\itemsep}{0em}
	\item algum identificador,
	\item a localização (\gls{hypocenter}) do evento (longitude, latitude, profundidade) em algum sistema de referência,
	\item o tempo de origem (data e hora) com precisão de pelo menos décimos de segundo e
	\item uma ou várias informações de \glsdesc{sym:m}.
\end{itemize}

Adicionalmente, embora não seja muito frequente, podem ser fornecidas informações adicionais obtidas pela análise das
formas de onda, como:

\begin{itemize}\setlength{\itemsep}{0em}
	\item incertezas sobre as magnitudes,
	\item incertezas sobre a localização (erro padrão, elipses de erro, cobertura dos sismogramas em diversas distâncias, cobertura
	dos sismogramas em vários ângulos azimutais, acurácia do modelo de velocidades utilizado, para enumerar alguns),
	\item intensidade máxima,
	\item intensidade no epicentro,
	\item número de, e as vezes as próprias, informações usadas para a determinação do hipocentro e hora de origem,
	\item sobre o mecanismo (alinhamento, mergulho e sentido do deslocamento na falha geológica) focal, entre outras.
\end{itemize}

É importante salientar \citep{woessner_2010} que cada um dos parâmetros determinados é fruto de uma série de decisões
e etapas de processamento. Começam pela escolha dos sismômetros e pelos locais onde serão instalados para registrar as
formas de onda. Sinais acima do nível de ruído são associados à chegadas de fases quando registradas em mais de uma estação.
A localização e o tempo de origem são determinados juntando-se os tempos de chegadas das fases a um modelo de velocidade
das ondas ao longo de camadas da crosta (ao qual a localização é extremamente dependente). As magnitudes, por fim, são
computadas a partir das amplitudes e/ou da duração do sinal, dependendo profundamente da calibração dos instrumentos.

\subsection{\glsdesc{mfd}}
\index{MFD}
\label{sec:mfd}

Observa-se que os sismos menores são muito mais frequentes.
Entretanto, os maiores e mais raros são os que trazem a maior ameaça e os que causam as maiores perdas.
Em virtude disso, uma análise conveniente seria explorar como se distribuem as magnitudes.

\subsubsection{\glsdesc*{mfd} (\gls{mfd}) de Gutenberg-Richter}
\index{Gutenberg-Richter MFD}
\label{sec:grmfd}

\citet{gutenberg_1944} observaram empiricamente que a distribuição da frequência de ocorrência dos tremores e
das magnitudes seguiam uma distribuição (de Pareto) cuja versão clássica é apresentada na equação \eqref{eq:gr_mfd} a
seguir:

\begin{equation}
	\gls{eqn:gr_mfd}
	\label{eq:gr_mfd}
\end{equation}
\glsdesc*{eqn:gr_mfd}.

Com uma simples transformação de variáveis ($\alpha = 10^a$ e $\beta = b\ln{10}$), observa-se que o número de sismos que ocorrem
com magnitudes dentro de um pequeno intervalo $[m, m+\mathrm{d}m]$ é

\begin{equation}
	\begin{split}
		\gls{sym:N_m} &= 10^{\gls{sym:a} - \gls{sym:b}\gls{sym:m}} \\
					  &= \alpha e^{-\beta m}
	\end{split}
	\label{eq:gr_exp}
\end{equation}

A distribuição cumulativa, ou seja, o número de eventos com magnitude maior que um certo valor $m_{min}$, é apresentada
na equação \eqref{eq:gr_cum}:

\begin{equation}
	\begin{split}
		N(m > m_{min}) &= \alpha \int\limits_{m_{min}}^{\infty}e^{-\beta m}\mathrm{d}m \\
					   &= \frac{\alpha}{\beta} e^{-\beta m} \\
					   &= \alpha_{cum} e^{-\beta m}.
	\end{split}
	\label{eq:gr_cum}
\end{equation}
onde $\alpha_{cum} = \alpha / \beta $ é o valor cumulativo da atividade sísmica.

Entretanto, a distribuição clássica de \gls{gr} não impunha restrições sobre um limite inferior $m_{min}$
ou superior $m_{max}$ à validade da distribuição.

\subsubsection{MFD Truncada}
\index{MFD Truncada}
\label{sec:TMFD}

Variações da distribuição clássica de \gls{gr} foram propostas em vista de melhor representar as MFD estudadas à partir de
catálogos de diversas regiões.

A equação \eqref{eq:gr_max} versão incremental truncada com um limite superior $m_{max}$:

\begin{equation}
		\gls*{sym:N_m} = \frac{e^{-\beta m}}{1 -e^{-\beta m_{max}} }, m \leq m_{max}
	\label{eq:gr_max}
\end{equation}

Na equação \eqref{eq:gr_dtr} versão incremental duplamente truncada com um limite inferior $m_{min}$ e superior $m_{max}$ :

\begin{equation}
		\gls*{sym:N_m} = \frac{e^{-\beta (m - m_{min})}}{1 -e^{-\beta (m_{max} - m_{min}) } } , m_{min} \leq m \leq m_{max}
	\label{eq:gr_dtr}
\end{equation}

A figura \ref{f:mfd} ilustra essas distribuições.

\subsubsection{MFD Limitada}
\index{MFD Limitada}
\label{sec:BMFD}

Outra possibilidade, é limitar suavemente a parte final da curva (ver figura \ref{f:mfd}). A equação \eqref{eq:gr_bounded} apresenta
a distribuição:

\begin{equation}
		\gls*{sym:N_m} = \alpha [ e^{-\beta (m - m_{min})} - e^{-\beta (m_{max} - m_{min}) } ], m_{min} \leq m \leq m_{max}
	\label{eq:gr_bounded}
\end{equation}

\subsubsection{MFD com decaimento exponencial}
\index{MFD com decaimento exponencial}
\label{sec:KMFD}

Yan Kagan \citep{kagan_2002} propôs uma distribuição de magnitude mais adequada e acoplada à energia liberada pelos sismos, que
pode ser descrita como na equação \eqref{eq:gr_tapered}:

\begin{equation}\ensuremath{
		\gls*{sym:N_m} = [\gls*{sym:beta_p} + \frac{m}{m_{min}}]
				m_{min}^{\gls*{sym:beta_p}}
				\gls*{sym:m_corner}^{-1 -\gls*{sym:beta_p}}
				e^{\frac{m_{min} - m}{\gls*{sym:m_corner}}},
				m_{min} \leq m < \infty
		}
	\label{eq:gr_tapered}
\end{equation}
onde \glsdesc*{sym:beta_p} e \gls*{sym:m_corner} \glsdesc*{sym:m_corner}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.95\textwidth]{mfd}
   \caption[Distribuições de frequência e magnitude]
   		   {Distribuições de frequência e magnitude}
   \label{f:mfd}
\end{figure}

A figura \ref{f:mfd} apresenta um comparativo de algumas distribuições. Para ilustração,
há também na figura um histograma de um catálogo de
uma pequena região do norte do Chile, onde se pode observar que tanto a porção inferior (em torno de $m=5$), como a porção
posterior ($m > 7$) do histograma não seguem perfeitamente a distribuição. Há essencialmente duas zonas críticas em que é preciso
estar atento à física do problema:
(i) na parte inferior, muitos sismos de magnitude pequena não são registrados, seja por não terem energia suficiente
para sensibilizar um conjunto razoável de estações que permitam determinar suas localizações, seja porque o número de
estações é insuficiente na região onde os pequenos tremores ocorrem; (ii) a parte superior, por sua vez, é crítica
por se acoplar diretamente aos limites físicos do tamanho da maior ruptura possível, relacionada diretamente ao limite de liberação de energia na forma de momento sísmico $M_0$.

Nas distribuições de magnitude e frequência é importante que se possa reconhecer claramente alguns parâmetros
fundamentais.

\subsection{Valor-b}
\index{valor-b}
\label{sec:b_value}

O \emph{valor-b} foi apresentado na seção \ref{sec:grmfd} como sendo a inclinação da reta que representa a parte linear
descrescente da distribuição. Representa a proporção entre sismos pequenos e catastróficos que uma determinada fonte
sísmica é capaz de produzir (figura \ref{f:occurrence}).

\begin{figure}[H]
   \centering
   \includegraphics[width=0.95\textwidth]{occurrence}
   \caption[Distribuição incremental e cumulativa de frequência e magnitude dos sismos presentes no catálogo ISC-GEM
   para a América do Sul unido com o \gls{bsb2013}]
   {Distribuição incremental e cumulativa de frequência e magnitude dos sismos presentes no catálogo ISC-GEM
   para a América do Sul unido com o \gls{bsb2013}}
   \label{f:occurrence}
\end{figure}
%\footnotetext{\citet{img_opensha_rupture}}

\subsection{Taxa de Sismicidade}
\index{taxa de sismicidade}
\label{sec:seismic_rate}

A taxa de sismicidade é a medida da ocorrência dos tremores por uma determinada unidade de tempo (geralmente anos).
Representa para cada magnitude, a frequência média de ocorrência de sismos.

\subsection{Valor-a}
\index{valor-a}
\label{sec:a_value}

O \emph{valor-a} é a interseção da \gls{mfd} no eixo das frequências e representa o nível geral de sismos
que as fontes observadas pelo catálogo são capazes de produzir.

Costuma ser confundido pela forma de representação adotadas para a distribuição (incremental e/ou
cumulativa) e pelos truncamentos onde por vezes se apresenta o valor-a como a taxa de sismicidade da
magnitude mínima ou de completude do catálogo.

No presente trabalho o \emph{valor-a} significará sempre o da distribuição cumulativa de sismos por unidade de
tempo com magnitudes positivas.

\subsection{Magnitude de Completude $M_c$}
\index{magnitude de completude}
\label{sec:completeness}

A magnitude de completude é o valor mínimo para o qual a distribuição é capaz de observar completamente o conjunto
de sismos. Em outras palavras representa o limite de observação completa do catálogo.

Sua identificação é bem simples quando se observa a distribuição incremental de magnitudes. É facilmente notado o valor
de magnitude na porção inferior na qual o número de sismos registrados começa a divergir da tendência geral da
distribuição.

Seu mapeamento é importante uma vez que os métodos de ajuste e determinação dos parâmetros da distribuição baseados na
máxima verossimilhança \citep{aki_1965, weichert_1980} dependem fundamentalmente desse valor mínimo.

\section{Risco Sísmico}
\index{Risco Sísmico}
\label{sec:risco_sismico}

A redução do risco sísmico é um problema complexo que envolve geralmente muitas pessoas, informações, decisões e ações.

A palavra risco, ao pé da letra, significa a exposição à possibilidade de injúria ou perda. E geralmente é usada como
sinônimo de ameaça. Na literatura acerca do tema risco, inclusive, as palavras risco e ameaça são usadas com certa
confusão.

No glossário da \gls{eeri} (EERI Committee on Seismic Risk, 1984) a definição de risco sísmico é a probabilidade de
que perdas sociais ou econômicas aconteçam como decorrência de tremores por superarem limiares estabelecidos para
determinado local ou região durante um certo período de exposição.

A ameaça sísmica, por outro lado, é qualquer fenômeno físico (oscilação, falhamento) associado à terremotos que possam
produzir efeitos adversos às atividades humanas. Na prática são avaliados por dadas probabilidades de ocorrência.

Pode-se deduzir que o risco sísmico é então uma combinação da ameaça sísmica com outros fatores:

\begin{equation}
		\text{Risco Sísmico} = \text{Ameaça Sísmica} \ast \text{Vulnerabilidade} \ast \text{Valor Exposto},
	\label{eq:risk}
\end{equation}

onde a vulnerabilidade é a quantidade de danos induzidos por um dado grau de ameaça e expressa como uma fração do
valor exposto ao dano e varia de acordo com o modelo proposto.

Frequentemente, o fator vulnerabilidade advém das análises das (ii) respostas das estruturas edificadas ao espectro de
acelerações produzidos pela (i) provável ameaça sísmica e da análise de possíveis (iii) danos estruturais à edificação.

A decisão de alterar ou não o desenho estrutural das edificações é feito a partir da análise dos (iv) prejuízos
(quantidade de moeda, mortes, tempo inoperante) causados caso as estruturas sejam danificadas conforme as análises anteriores.

\section{Ameaça Sísmica}
\index{ameaça sísmica}
\label{sec:ameaca_sismica}

A ameaça sísmica poderia ser definida de modo geral como a possibilidade de ocorrerem efeitos potencialmente destrutivos
de um terremoto em uma particular localização. Com exceção de \textit{tsunamis} ou falhamentos geológicos superficiais,
todos os efeitos destrutivos de um tremor de terra estão diretamente relacionados ao movimento do chão induzido pela
passagem das ondas sísmicas. Existem, entretanto diferenças de abordagem para a avaliação da ameaça sísmica.

A \glsdesc{psha} foi introduzida por \citet{cornell_1968}, aprimorada por \citet{mcguire_1976} e se tornou técnica mais
amplamente utilizada para a avaliação da ameaça sísmica. Também é possível fazer essa avaliação deterministicamente,
através de cenários definidos pelo espectro de movimento forte do chão que pode ser causado pela ocorrência de um
determinado tremor de terra em certa localização e de certa magnitude. O possível espectro de movimento no local de
interesse é avaliado através de relações de atenuação ou \glspl{gmpe}.

Os mecanismos da \gls{psha} \citep{bazzurro_1999, abrahamson_2006} são menos óbvios do que os da \gls{dsha}
\citep{reiter_1991, kramer_1996}, e em essência significam identificar todos os possíveis tremores que podem afetar o local de interesse,
incluíndo todas as possíveis combinações de distâncias e caracterizar a frequência de ocorrência das diferentes magnitudes através de relações de recorrência. As equações de
atenuação são utilizadas para calcular os parâmetros do movimento do chão no local de interesse devido a esses tremores
e consequentemente a taxa com que diferentes níveis de movimento do chão ocorram no local de interesse.

Seus resultados também apresentam certa distinção. Se por um lado a \gls{psha} traz consigo o aspecto
temporal, ou a taxa com que diferentes níveis de aceleração excederão determinado limiar em determinado local de interesse,
por outro, a \gls{dsha} apresenta o movimento do chão esperado quando ocorra determinado evento de controle.

\section{Projeção da Ocorrência de Rupturas}
\index{projeção de ocorrência de rupturas}
\label{sec:projecao}

As projeções (\textit{forecasting}) são feitas para se estimar a ocorrência de futuros tremores
\citep{kagan_2000,marzocchi_2011}, principalmente dos maiores, com grandes chances de provocar perdas.

Nas de curto prazo, estimam-se os próximos tremores
numa escala de dias ou horas considerando uma taxa de sismicidade variável
com o tempo como no caso dos pré e pós-abalos, ou de quando
acontece um enxame sísmico, período de maior atividade numa região.
Sua principal aplicação é auxiliar a tomada de decisões de curto período,
como evacuação de edifícios.

Nas de longo prazo, foco desse texto, a principal consideração feita é de que a
\gls{seismic_rate} não varie ao longo do tempo, servindo para estimar as acelerações
provocadas por tremores, mesmo que possam ocorrer
muito raramente, de grandes proporções.

São geralmente aplicadas quando
se deseja saber o nível de segurança e resistência estrutural que devem ser impostos
às edificações em geral, ou quando se deseja estimar o valor de um contrato de resseguro
de algum outro grande investimento.

\section{Análise Probabilística de Ameaça Sísmica }
\index{PSHA, Análise Probabilística de Ameaça Sísmica}
\label{sec:psha}

Na \gls{psha} são considerados todos os possíveis tremores, as rupturas que os originaram e os movimentos do chão
resultantes conjuntamente com suas probabilidades de ocorrência associadas de modo a encontrar o nível de movimento do
chão que será excedido, numa janela de tempo, com uma pré-definida baixa tolerância \citep{baker_2008-1}.

Uma das formas de se enxergar o resultado de uma análise de ameaça sísmica
é como uma estimativa da pequena probabilidade $\xi$, em uma janela de tempo dada,
com que determinada medida de intensidade $I$ é raramente excedida.

Na análise probabilística de ameaça sísmica o nível de confiança $\xi$
e a janela de tempo são fixadas (por exemplo, $t$ anos). A tarefa é então estimar
o valor da intensidade $I$, em um determinado local $S$, de forma que a probabilidade
do evento
\begin{equation} \label{eventEt}
\begin{array}{lll}
 E_t(I, S)&=&\{\mbox{Haja pelo menos um evento causando intensidade }  \\
 &&  \;\;\mbox{maior que $I$ em }S  \mbox{ nos próximos }t \mbox{ anos}\} \\
\end{array}
\end{equation}
seja $\xi$.

\subsection{Método de Zoneamento}
\index{PSHA!zoneamento}
\label{sec:zoneamento}

Na ausência de informações geológicas mais precisas, o método introduzido por \citet{cornell_1968} e
\citet{mcguire_1976} para modelar e resolver essse problema consiste
primeiramente em identificar quais zonas sísmicas (que em geral não se sobrepõem)
podem ter impacto sobre o valor de intensidade em $S$.

O número de tremores que povocam intensidade em $S$ maior que $I$ em $t$ anos depende da frequência
de tremores em cada zona. O valor da intensidade $I$ provocada por cada tremor depende da magnitude,
aleatória, desses tremores e de sua localização, também aleatória.

Para considerar esses fatores Cornell \& McGuire propõem que
\begin{itemize}
\item[(i)] em cada zona $i$, o processo de ocorrência de tremores seja
modelado como um processo de Poisson com taxa $\lambda_i$, assumindo que os tremores em diferentes zonas são
independentes.
\item[(ii)] Numa zona $i$, a magnitude dos tremores é modelada como uma \gls{va}
com densidade $f_{M_i}(\cdot)$.
\item[(iii)] A distância entre cada tremor da zona $i$ e o local $S$ é modelada como uma \gls{va}
com densidade $f_{D_i}(\cdot)$.
\item[(iv)] O modelo de predição do movimento do chão (\gls{gmpe}) é expresso pela regressão da medida de intensidade
em magnitude, distância, condições geológicas do local $S$ e outros fatores.
\end{itemize}

A habilidade de se calcular a probabilidade do evento
\eqref{eventEt} para qualquer $I$ faz com que seja possível a estimativa, por dicotomia, de uma intensidade $I$ que satisfaça
$P\left\{ E_t(I, S) \right\} \geq \xi$.

\subsection{Identificação das Fontes Sísmicas}
Para identificar fontes sísmicas são utilizados desde registros históricos de sismicidade à evidências geológicas de
falhamentos/rupturas datados com deslocamento e magnitudes inferidos e busca-se aproveitar de toda informação relevante
disponível, como a medida secular de deslocamento relativo entre observações geodésicas contínuas ou mesmo da
sismicidade recente.

Quando se identifica uma fonte sísmica é comum representá-la por uma forma geométrica simples mas consistente com o
conjunto das observações disponíveis para descrever as possíveis rupturas \citep{crowley_2013}.

\subsubsection{Ponto}
\index{fonte sísmica!ponto}
\label{sec:point_source}

Se toda informação disponível é uma localização isolada de um tremor antigo, com magnitude e com mecanismo de
falhamento conhecido, é possível representá-lo como uma fonte sísmica de tipo pontual.
Nesse tipo de fonte são definidos os limites superior e inferior da ruptura,
sua orientação e tipo de falhamento (quando disponível)
e o hipocentro era definido como o centro de cada ruptura. Recentemente porém é possível definir uma distribuição de
progundidades hipocentrais.

\subsubsection{Área}
\index{fonte sísmica!área}
\label{sec:area_source}

Quando o conhecimento sobre a geologia, a tectônica, ou mesmo a correlação espacial dos tremores no catálogo permitam
o delineamento de zonas ou áreas com características sísmicas comuns se costuma representar por um polígono na
superfície.

Essas áreas, para efeito de cálculo, são discretizadas como um conjunto de fontes sísmicas de
característica pontual distribuídas uniformemente por toda área.

\subsubsection{Falha Simples}
\index{fonte sísmica!falha simples}
\label{sec:simple_fault_source}

Muitas vezes os parâmentros de um falhamento ativo são claramente conhecidos e monitorados. Isso permite uma maior
especificidade na representação da fonte sísmica, restringindo mais, por exemplo, as flutuações na orientação das
rupturas.
Nesse caso a geometria da falha se caracteriza pela projeção do traço de falha na superfície e pelos limites superior e
inferior da ruptura no plano de mergulho (ver figura \ref{f:rupture}).

\subsubsection{Falha Complexa}
\index{fonte sísmica!falha complexa}
\label{sec:complex_fault_source}

Casos de sismicidade em zonas de subducção ou encontro de placas, de contexto geológico mais complexo
geralmente apresentam variações laterais, de mergulho, de acúmulo de esforços, de orientação, etc. Fontes sísmicas em
situações como essa são modeladas por um poliedro unido de forma suave.

\subsection{Caracterização da \glsdesc*{mfd}}
\index{caracterização da \glsdesc*{mfd}}
\label{sec:psha_mfd}

Conhecida a fonte sísmica e sua representação geométrica, é preciso caracterizar sua capacidade sismogênica determinando
uma (ou mais) possíveis \gls{mfd}s que se ajustam às observações. Isso inclui a espressão matemática da distribuição, a
taxa geral de sismicidade (\emph{valor-a}) e frequentemente as magnitudes mínima e máxima.

A densidade $f_{M_i}(\cdot)$ usada para a distribuição de magnitude dos tremores na zona $i$ depende do
histórico de magnitudes na mesma zona, e possui uma das formas funcionais da seção \ref{sec:mfd},
como por exemplo a distribuição duplamente truncada,
com $M_i$ sendo o intervalo entre as magnitudes mínima e máxima $[M_{\min}(i), M_{\max}(i)]$
em cada zona $i$.

\subsection{Caracterização da Distribuição de Distâncias}
\index{caracterização da distribuição de distâncias}
\label{sec:psha_distances}

Dados um local de interesse e uma provável ruptura em uma fonte sísmica é necessário calcular a distribuição das
distâncias da fonte, isto é, das possíveis rupturas, ao local em que se deseja avaliar a ameaça.

Em cada zona as rupturas são consideradas como tendo distribuição espacial uniforme $f_{D_i}(\cdot)$.
A distribuição das distâncias, usadas pelas equações de atenuação, podem então ser calculadas
analiticamente ou por aproximação.

\subsection{Predição do Movimento do Chão}
\index{predição do movimento do chão}
\label{sec:gmpe}

Para se estimar os possíveis níveis de movimento do chão causados por eventos de uma determinada magnitude à uma certa
distância do local de interesse são utilizadas as equações de predição de movimento do chão (\glspl{gmpe}).

As GMPEs são modelos (equações e coeficientes) de regressão representando certo valor de
intensidade $I$ induzida por um tremor de magnitude $M$ a uma distância $D$ do epicentro (ruptura, hipocentro, etc, dependendo da
modelagem da GMPE) mas que depende também de outros fatores $\theta$ que considerem as condições
do local de estudo e do tipo de falhamento por exemplo. Geralmente as GMPEs tem a forma

\begin{equation} \label{pgamodel}
\ln I = \overline{\ln I}(M, D, \theta) + \sigma(M, D, \theta) \varepsilon.
\end{equation}

Nessa relação, $\overline{\ln I}(M, D, \theta)$ (e respectivamente $\sigma(M, D, \theta)$)
é a média condicional (e desvio padrão) de $\ln I$ para certa magnitude $M$, distância $D$
e condições $\theta$, enquanto $\varepsilon$ é uma \gls{va} gaussiana padrão.

Essa média $\overline{\ln I}(M, D, \theta)$ deve crescer com $M$ já que quanto maior a magnitude,
maior a intensidade provocada, e, por outro lado, decrescer com a distância $D$ pois quanto maior a distância, menor
a intensidade, como é possível observar nas figuras \ref{fig:gmpe_magnitude} e \ref{fig:gmpe_distance}.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.47\textwidth}
		\centering
		\includegraphics[width=1.0\textwidth]{gmpe_magnitude}
		\caption{Variação da intensidade \gls{pga} com a magnitude.}
		\label{fig:gmpe_magnitude}
	\end{subfigure}
	\quad
	\begin{subfigure}[t]{0.47\textwidth}
		\centering
		\includegraphics[width=1.0\textwidth]{gmpe_distance}
		\caption{Variação da intensidade \gls{pga}) com a distância de Joyner-Boore.
		  para um valor de magnitude de 5.05 }
		\label{fig:gmpe_distance}
	\end{subfigure}
	\caption{Variação das intensidades com a magnitude e distância, para diferentes GMPEs}
	\label{fig:gmpe}
\end{figure}

A distância de \citet{joyner_1981} é uma das várias medidas de distância utilizadas e corresponde especificamente à
menor distância entre o local de avaliação da ameaça $S$ à projeção da ruptura na superfície da topografia.

Apenas como exemplo, a forma funcional que \citet{toro_1997} dão à GMPE é

\begin{equation}
\begin{array}{lcl}
\ln I(M, R_M, \theta)
&=& \theta_1 + \theta_2(M-6) + \theta_3(M-6)^2  \\
& & - \theta_4\ln (R_M) -(\theta_5 - \theta_4)\max\left[ \ln\left( \frac{R_M}{100} \right), 0 \right] -\theta_6 R_M \\
& & + \varepsilon_e + \varepsilon_a, \\
\end{array}
\end{equation}
com
$$
R_M =  \sqrt{ R_{jb}^2 + \theta_7^2 }
$$
onde $R_{jb}$ é a distância de Joyner-Boore, $M$ é a magnitude de momento,
$\theta_i$ são as componentes do vetor de coeficientes e $\varepsilon_e$ + $\varepsilon_a$
são as incertezas epistêmica e aleatória do modelo.

Nesse exemplo
\begin{equation}
\begin{array}{lcl}
\overline{\ln I}(M, D, \theta) & = & \theta_1 + \theta_2(M-6) + \theta_3(M-6)^2  \\
& & - \theta_4\ln (R_M)
-(\theta_5 - \theta_4)\max\left[ \ln\left( R_M / 100 \right), 0 \right]
-\theta_6 R_M
\end{array}
\end{equation}
e
\begin{equation}
	\sigma(M, D, \theta)=\varepsilon_e(M,D) + \varepsilon_a(M,D).
\end{equation}

\subsection{Combinação de Incertezas e Avaliação da Ameaça Sísmica}
\index{cálculo da ameaça}
\label{sec:hazard}

A distribuição do número de tremores $N_{t i}$, na zona $i$, na janela de tempo $t$, por Poisson,
é dada por
$$
P(N_{t i}=k)=e^{-\lambda_i t} \frac{(\lambda_i t)^k}{k!},\;k \in \mathbb{N},
$$
onde a taxa $\lambda_i$ representa o número médio de tremores na zona
$i$ por unidade de tempo, por exemplo, por ano.

Fixado um valor de intensidade $I$, o evento
\begin{equation} \label{formulapi}
\begin{array}{lll}
E(I, S, i)  & =  & \{ \mbox{Um tremor da zona }i \mbox{ gera uma intensidade} \\
&  &  \;\;\mbox{maior ou igual a} I \mbox{ em }S \}
\end{array}
\end{equation}
tem probabilidade $p_i=P \left\{ E(I, S, i)\right\}$.

Para cada tremor na zona $i$, o evento $E(I, S, i)$ ocorre, ou não.
Como resultado, é possível definir dois novos processos de contagem
em cada zona $i$: o processo $\tilde N_{t i}$, contando os tremores que
causam intensidade maior ou igual a $I$ em $S$ e os que causam intensidade menor que $I$ em $S$.

\begin{figure}[H]
	\centering
	\begin{tabular}{l}
	\includegraphics[width=0.80\textwidth]{poisson}
	\end{tabular}
	\caption{Separação do processo de chegada dos tremores da zona $i$
	em um processo causando, em $S$, intensidade maior que $I$ (bolas pretas) e intensidade
	menor ou igual a $I$ (bolas brancas).}
\label{fig:poisson}
\end{figure}

É preciso utilizar o seguinte
\begin{lemma}
Considere um processo de Poisson $N_t$ com taxa de ocorrência $\lambda$.
Assumindo que as chegadas são de dois tipos: tipo 1 com probabilidade $p$,
e tipo 2 com probabilidade  $1 - p$ e que elas são independentes,
então o processo $\tilde N_t$ do tipo 1 é um proceso de Poisson com taxa
$\lambda p$.
\end{lemma}
\begin{proof} Para cada $k \in \mathbb{N}$, calcule
$$
\begin{array}{lll}
P\Big(\tilde N_{t} =k \Big) & = & \displaystyle \sum_{j=k}^{+\infty} P\Big(\tilde N_{t} =k | N_{t}
=j\Big)  P \Big( N_{t} =j  \Big)\;\;\;\mbox{[Teorema da Probabilidade Total]}\\
&=& \displaystyle \sum_{j=k}^{+\infty} C_j^k p^k (1-p)^{j-k} e^{-\lambda t} \frac{(\lambda t)^j}{j!} \\
&=&e^{-\lambda t} \frac{(\lambda p t)^k}{k!} \displaystyle \sum_{j=0}^{+\infty} \frac{[\lambda (1-p)]^{j}}{j!} = e^{-\lambda p t} \frac{(\lambda p t)^k}{k!},
\end{array}
$$
o que mostra que $\tilde N_{t}$, usando a independência dos tipos de chegada em
intervalos disjuntos, é uma \gls{va} de Poisson com parâmetro
$\lambda p t$.\hfill
\end{proof}

Esse lema mostra que o processo $(\tilde N_{t i})_t$ é um processo de Poisson com taxa $\lambda_i p_i$.

Sendo $\mathcal{N}$ o número de zonas sísmicas, segue que a probabilidade de que haja $k$ sismos causando
intensidade maior que $I$ em $S$ na janela de tempo $t$ é
$$
\displaystyle
\begin{array}{lll}
P\Big(\displaystyle \sum_{i=1}^{\mathcal{N}} \tilde N_{t i} = k\Big)
&= &\displaystyle \sum_{x_1+\ldots +x_{\mathcal{N}}=k} P\Big(\tilde N_{t 1} = x_1; \ldots; \tilde N_{t \mathcal{N}} = x_{\mathcal{N}} \Big)\\
&= &\displaystyle{\sum_{x_1+\ldots +x_{\mathcal{N}}=k} \prod_{i=1}^{\mathcal{N}}}P\Big(\tilde N_{t i} = x_i \Big)\\
&= &\displaystyle \sum_{x_1+\ldots +x_{\mathcal{N}}=k} \prod_{i=1}^{\mathcal{N}} e^{-\lambda_i p_i t} \frac{(\lambda_i p_i t)^{x_i}}{x_{i}!}
\end{array}
$$
em que para a segunda igualdade foi utilizada a independência de  ${\tilde N}_{t 1}, \ldots, {\tilde N}_{t
\mathcal{N}}$.

Para $k=0$ na relação acima obtêm-se que
\begin{equation} \label{probinterest}
1-P( E_t(I, S) )= P(\overline{E_t(I, S)})=e^{-(\sum_{i=1}^{\mathcal{N}} \lambda_i p_i) t}.
\end{equation}

Com $\tilde N_t=\sum_{i=1}^{\mathcal{N}} \tilde N_{t i}$,
o valor esperado $\tilde N_t$ é o número médio de tremores causando intensidade maior que $I$ em $S$
 nos próximos $t$ anos e pode ser espresso por
\begin{equation} \label{formlambdatA}
\lambda_t(I, S)=\mathbb{E}\Big[ \tilde N_t \Big] =\sum_{i=1}^{\mathcal{N}} \mathbb{E}\Big[ \tilde N_{t i}\Big] =
\left( \sum_{i=1}^{\mathcal{N}} \lambda_i p_i \right) t.
\end{equation}

Usando essa relação e a equação \eqref{probinterest}, a probabilidade do evento $E_t(I, S)$
pode ser reescrita
\begin{equation}
P \left\{   E_t(I, S) \right\} =1-e^{-\lambda_t(I, S)}
\end{equation}
com $\lambda_t(I, S)$ dado por \eqref{formlambdatA}.

\subsubsection{Combinando todos os elementos}

Assumindo que as variáveis aleatórias da distância $D_i$ entre o tremor na zona $i$ e o local $S$ e magnitude $M_i$
são independentes e usando o Teorema da Probabilidade Total, obtém-se que
\begin{equation}
p_i= \displaystyle \int_{m_i =M_{\min}(i)}^{M_{\max}(i)}
\int_{x_i =0}^{\infty}  P \left\{ \mbox{\small intensidade \normalsize }>I| M_i = m_i; D_i = x_i \right\}f_{M_i}(m_i) f_{D_i}( x_i )
\mathrm{d}m_i \mathrm{d}x_i
\end{equation}
onde $P \left\{ \mbox{\small intensidade \normalsize }>I| M_i = m_i; D_i = x_i \right\}$
é dada pelo de predição do movimento do chão \eqref{pgamodel}.

Para implementação, a integral acima é geralmente estimada discretizando as distribuições contínuas de
magnitude $M_i, i=1,\ldots, \mathcal{N}$, e distância $D_i, i=1,\ldots, \mathcal{N}$.

### 2015-01-13 c03-regiao.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Região de Estudo}
\label{cap:regiao_de_estudo}

Esse capítulo apresenta a região de estudo sob o ponto de vista tectônico e sismológico.

%% ------------------------------------------------------------------------- %%
\section{Contexto Geológico e Tectônico Sul-Americano}
\index{\gls{tectonic}!América do Sul}
\label{sec:03_america_do_sul}

A placa Sul-Americana, como mostra a figura \ref{fig:sa_plate} \citet{bizzi_2003}, tem ao norte a placa do Caribe e a
placa Norte-Americana.
Ao sul estão a placa de Scotia e a placa Antártica. Todas elas se deslocam 
majoritariamente tangencialmente
à placa Sul-Americana.

\begin{figure}[H]
  \centering
  \includegraphics[width=.95\textwidth]{placas_sa} 
  \caption{Placa Sul-Americana em seu contexto global}
  \label{fig:sa_plate} 
\end{figure}

Na divisa com placa Africana à leste está a Dorsal Meso-Atlântica que é resultado
do processo de abertura dos oceanos e separação dos continentes. A abertura dos Atlântico na 
dorsal é responsável por um considerável esforço de compressão horizontal na placa Sul-Americana.

E há também a subducção da placa de Nazca sob a placa Sul-Americana, à oeste,
responsável, entre outros processos, pelo surgimento das Fossas Oceânicas e da
cordilheira dos Andes com suas altitudes e vulcanismo.

Olhando um pouco mais de perto para a parte continental da placa Sul-Americana
(figura \ref{fig:sa_tec}) é interessante destacar três grupos principais de rochas: 
(i) o Embasamento Pré-Cambriano, (ii) as Coberturas Fanerozóicas e (iii) a 
Cadeia Andina.

\begin{figure}[H]
  \centering
  \includegraphics[width=.85\textwidth]{lithology_sa} 
  \caption{Mapa geológico da América do Sul. Fonte \url{http://onegeology.org}}
  \label{fig:sa_tec} 
\end{figure}

As rochas do embasamento pré-cambriano se originaram a mais de 500Ma. Por serem mais antigas
são mais estáveis. A cobertura Fanerozóica é resuldado da sedimentação ocorrida a menos de 250Ma. Formam as bacias
sedimentares. A Cadeia Andina com 30Ma é resultado da subducção e embora exponha rochas pré-cambrianas
em algumas partes é a parte mais ativa e interessante tectonicamente.

\subsubsection{Sismicidade Sul Americana}

A sismicidade Sul-Americana é marcada fortemente pela subducção à oeste e pela 
separação dos oceanos à leste. América Central, Caribe e a parte Antartica ao sul
são placas menores e seus movimentos merecem estudo de maior detalhe.

\begin{figure}[H]
  \centering
  \includegraphics[width=.70\textwidth]{seismicity_sa} 
  \caption{Sismicidade da América do Sul, Catálogo \gls*{iscgem}. 
  		   Geologia: \gls*{cgmw} via OneGeology.
		   Sismos mais profundos foram registrados no interior da placa, inclusive sobre o Acre.
  		   }
  \label{fig:sa_seis} 
\end{figure}

A figura \ref{fig:sa_seis} apresenta a sismicidade da América do Sul pelo catálogo \gls{iscgem}
(seção \ref{sec:data_source}). É possivel notar claramente a subducção, ou seja, do mergulho, da placa de Nazca sob a placa
Sul-Americana. Fica mais claro quando se observa que os sismos com profundidade variando cerca de centenas de
quilômetros e que vão se tornando mais profundos para interior da placa Sul-Americana.

Isso acontece porque parte da quantidade de rocha fria, oceânica e continental, está afundando sob o manto
e lentamente se incorporando à ele. Mas ainda existem atrito, compressões e processos de ruptura nessas
profundidades e que ocorrem majoritariamente na interface entre essas placas pelo acúmulo de tensão e deslocamentos 
mínimos durante milhares de anos e que são liberados instantaneamente na ruptura. 
Esse processo é também o responsável pelo soerguimento da 
cordilheira dos Andes e de boa parte do vulcanismo na região.

Sismos profundos, de 70 e 700km, geralmente provocam acelerações de baixa intensidade em seus epicentros 
devido à atenuação das ondas.

Também é latente a constatação da diferença de distribuição da sismicidade sul-americana nas bordas de placa e do
resto do continente como um todo. A maior parte do Brasil é praticamente inativa, não desprezível, no histórico
comparado.

%% ------------------------------------------------------------------------- %%
\section{Contexto Geológico e Tectônico Brasileiro}
\index{\gls{tectonic}!Brasil}
\label{sec:geotec_bras}

Tomando-se como referência 500Ma, destacam-se dois grandes grupos de rochas na figura \ref{fig:br_tec}
\citet{bizzi_2003} a seguir.

\begin{figure}[H]
  \centering
  \includegraphics[width=.70\textwidth]{tectonico_brasil} 
  \caption{Mapa Geológico do Brasil em escala 1:1.000.000}.
  \label{fig:br_tec} 
\end{figure}

O embasamento, mais antigo, exposto sob a Amazônia e em porções menores e mais recentes também expostos 
no sudeste e nordeste, cedem espaço à um segundo grupo de rochas mais jovens, fruto da sedimentação e metamorfismos
mais recentes \citep{bizzi_2003}.

%% ------------------------------------------------------------------------- %%
\section{Sismicidade do Brasil}
\index{Sismicidade do Brasil}
\label{sec:sismicidade_brasil}

No Brasil não há terremotos. Não ao menos em quantidade proporcional a 10\% dos sismos sul-americanos.
Mesmo assim, ou por isso mesmo, a ocorrência de um sismo é ainda mais ameaçadora. Onde se espera, 
já se está preparado. Por outro lado onde nunca se espera é sempre uma surpresa.

É fato que o Brasil por estar numa área continental, mais no interior da placa \citep{talwani_2014} e geologia
com uma formação antiga e estável tem um número reduzido, não desprezível, de tremores. 
A figura \ref{fig:br_seis} mostra o detalhe da sismicidade brasileira com a litologia ao fundo.

\begin{figure}[H]
  \centering
  \includegraphics[width=.90\textwidth]{seismicity_br} 
  \caption{Sismicidade do Brasil. Catálogo \gls{bsb2013} (seção \ref{sec:data_source2}).}
  \label{fig:br_seis} 
\end{figure}

É importante notar que já houve registro de sismos com magnitudes pouco acima de 6,
e que sismos de magnitude acima de 4, rasos, em área urbana e em um continente estável,
com baixa atenuação das amplitudes das ondas sísmicas, podem ser danosos.

%% ------------------------------------------------------------------------- %%
\subsection{Sul, Sudeste e Litoral Leste}
\index{Sismicidade do Brasil!sul, sudeste e litoral leste}
\label{sec:z_se}

A sismicidade do sudeste e seu litoral possui características diferentes.
Enquanto no litoral, a principal sismiciade ocorre na área do talude continental
(porção dos fundos marinhos com declive muito pronunciado que fica entre a plataforma continental e
a margem continental e onde começam as planícies abissais), com destaque nessa sismicidade para um dos maiores sismos
que se tem registro no Brasil (figura \ref{fig:z_se}).

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\textwidth]{z_se} 
  \caption{Zona sísmica do SE. \citet{dourado_2014}}
  \label{fig:z_se} 
\end{figure}

O continente é marcado por uma sismicidade difusa na área do cráton que se extende pelo norte de Minas
Gerais até quase o sul da Bahia e outra parte a nordeste da bacia do Paraná. 
Há também uma pequena sismicidade acompanhando a costa. Para maiores referencias, veja 
\citet{assumpcao_2004}.

É nessa região o único registro no Brasil de vítimas fatais decorrentes de tremores de terra, em Itacarambi, MG.

%% ------------------------------------------------------------------------- %%
\subsection{Centro-Norte}
\index{Sismicidade do Brasil!centro-norte}
\label{sec:z_cn}

É uma área com sismicidade peculiar. Na figura \ref{fig:z_cn}, de norte a sul, a sismicidade acompanha grosseiramente o
contato entre o cráton e parte da bacia do Parnaíba. O mesmo ocorre ambém no que seria a 
área central próxima à Chapada dos Veadeiros em uma outra formação cratônica.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.45\textwidth]{z_co} 
  \caption{Zona sísmica do Centro-Norte. \citet{dourado_2014}}
  \label{fig:z_cn} 
\end{figure}

A última porção, ao sul, a sismicidade ocorre na área sedimentar da bacia do Pantanal.

%% ------------------------------------------------------------------------- %%
\subsection{Mato-Grosso}
\index{Sismicidade do Brasil!centro-norte}
\label{sec:z_mt}

A sismicidade do Mato-Grosso, mais precisamente de Porto-de-Gaúchos (\ref{fig:z_mt}), é emblemática para o Brasil.
Poderia ser considerada com características similares a Nova Madri, EUA.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.45\textwidth]{z_mt} 
  \caption{Zona sísmica do Centro-Norte. \citet{dourado_2014}}
  \label{fig:z_mt} 
\end{figure}

A região sofreu um dos maiores sismos já registrados no Brasil, com magnitude pouco acima de 6.
Não há registros de falhas geológicas neo-tectônicas e mais complexa de ser explicada \citep{barros_2009}.

%% ------------------------------------------------------------------------- %%
\subsection{Extremo Oeste e Acre}
\index{Sismicidade do Brasil!extremo-oeste}
\label{sec:z_ac}

No extremo oeste do Brasil, na região do Acre, a sismicidade tem uma característica distinta das outras.
É possível reparar, na figura \ref{fig:z_ac}, primeiramente na quantidade de sismos, e em
seguida perceber a influência dos sismos sul americanos, desde os mais profundos aos intermediários
e relativamente rasos de 70km.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{z_ac} 
  \caption{Zona sísmica do Acre. \citet{dourado_2014}, Catálogo \gls*{iscgem}.}
  \label{fig:z_ac} 
\end{figure}

Note que a escala de profundidade da figura \ref{fig:z_ac} é diferente das demais que seguem a
mesma escala da figura \ref{fig:br_seis}.

%% ------------------------------------------------------------------------- %%
\subsection{Amazonas}
\index{Sismicidade do Brasil!Amazonas}
\label{sec:z_am}

É a região com menor quantidade de conhecimento e informação disponível.
Ainda assim, na figura \ref{fig:z_am} é possível observar a ocorrência de sismos
na direção norte-sul.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{z_am} 
  \caption{Zona sísmica de Manaus. \citet{dourado_2014}.}
  \label{fig:z_am} 
\end{figure}

O registro de um sismo de magnitude 5 determinada por dados macrossísmicos,
é o indício mais marcante na região.

%% ------------------------------------------------------------------------- %%
\subsection{Nordeste}
\index{Sismicidade do Brasil!nordeste}
\label{sec:z_ne}

A regisão nordeste brasileira é sismicamente a mais ativa (figura \ref{fig:z_ne}).

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\textwidth]{z_ne} 
  \caption{Zona sísmica do NE. \citet{dourado_2014}}
  \label{fig:z_ne} 
\end{figure}

Destaca-se a sismicidade na região de João Câmara no Rio Grande do Norte,
com um enxame sísmico em meados da década de 1980 \citep{takeya_1989, bezerra_1998}. 
Além disso são bem conhecidas as atividades sísmicas na região de Sobral-CE,
na região de Pernambuco e um pouco mais ao sul já na Bahia.

### 2015-01-13 c06-resultados.tex
adding images

%% ------------------------------------------------------------------------- %%
\chapter{Resultados}
\label{cap:resultados}

Os resultados obtidos através do processamento descrito no capítulo anterior
serão enumerados a seguir.

%% ------------------------------------------------------------------------- %%
\section{Resultados Anteriores}
\index{resultados anteriores}
\label{sec:old_results}

Antes dos resultados obtidos pelos modelos discutidos, 
serão apresentados alguns já conhecidos que poderão servir como referência ou ponto de partida.

%% ------------------------------------------------------------------------- %%
\subsection{GSHAP}
\index{GSHAP}
\label{sec:gshap}

Um dos resultados mais conhecidos de avaliação de ameaça sísmica global é o GSHAP \citep{giardini_1999}.

A figura \ref{fig:gshap} apresenta os resultados do GSHAP para América do Sul.
E escala está limitada para que sejam claras as variações no território brasileiro.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_gshap} 
  \caption{Resultado do GSHAP para o Brasil: \gls{PGA} (10\%/50anos) em unidades de $g$}
  \label{fig:gshap} 
\end{figure}

É possível observar no GSHAP a grande influência de sismos provenientes da zona de subdução nos Andes 
sob o território do Acre. Também é destacada a sismicidade no Nordeste, com um pico de 0.16g para a PGA, e por
último, chama a atenção o resto do país por haver um vazio de informação.

%% ------------------------------------------------------------------------- %%
\subsection{Zoneamento Sísmico}
\index{zoneamento}
\label{sec:zonning}

Existem também esforços \citep{dourado_2014} para se criar um modelo para a ameaça sísmica brasileira
usando a metodologia clássica de Cornell \& McGuire com zoneamento sísmico.

\citet{dourado_2014}, aplicou critérios de especialista para definir e caracterizar as zonas sísmicas
apresentadas na figura \ref{fig:a_dourado}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{a_dourado} 
  \caption{Zoneamento sísmico e caracterização das zonas sísmicas por \citep{dourado_2014}.
  Os valores para a magnitude mínima foram de 3.0, e os \emph{valores-a} correspondem a magnitude maior que 0.}
  \label{fig:a_dourado} 
\end{figure}

As zonas sísmicas foram apresentadas em detalhe de sismicidade e geologia no capítulo \ref{cap:regiao_de_estudo}.

O cálculo feito por Dourado com o programa Crisis \citep{crisis_2007} gerou os resultados da figura \ref{fig:pga_dourado}
para o mapa de ameaça sísmica (PGA, poe 10\% em 50 anos). Os valores originais \emp{Gal} $[cm/s^2]$ foram convertidos
para unidades de \emph{g} $[m/s^2]$.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_dourado_crisis} 
  \caption{Resultado de \citet{dourado_2014} para o Brasil: PGA (10\%/50anos) em unidades de $g$ calculadas com o programa
  Crisis-2007}
  \label{fig:pga_dourado} 
\end{figure}

As mesmas fontes sísmicas e suas respectivas características foram utilizadas para o cálculo no
Openquake e os resultados estão apresentados na figura \ref{fig:pga_dourado_oq}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_dourado_oq} 
  \caption{Mapa de ameaça sísmica, PGA(poe 0.1, 50y)[Dourado, 20014] OpenQuake-Engine }
  \label{fig:pga_dourado_oq} 
\end{figure}

Os resultados preservam as feições mas apresentam quantidade significativamente diferentes.
Provavelmente devido ao fator de escala entre a magnitude e a área de ruptura utilizada,
que sobrevalora a área de ruptura para magnitudes pequenas (menores que 5). Uma sugestão
para o futuro é separar, na definição das fonte, a \gls{mfd} em duas partes: na porção inferior
usar um fator de escala que considere toda a ruptura ocorrendo no hipocentro. Na porção superior,
a partir de uma certa magnitude, toma-se a ruptura tendo como geometria mais realista.

%% ------------------------------------------------------------------------- %%
\section{Suavização da Sismicidade}
\index{suavização da sismicidade}
\label{sec:suavizacao_resultados}

Os métodos avaliados resultaram nos mapas de 
\emph{valor-a} e de ameaça sísmica
que serão apresentados nessa seção.

%% ------------------------------------------------------------------------- %%
\subsection{Frankel, 1995}
\index{Frankel,1995!resultados}
\label{sec:frankel_resultados}

Como resultado do processamento apresentado na seção \ref{sec:proc_frankel},
com distância de correlação de 150km e utilizando-se do catálogo BSB,
a taxa de sismicidade suavizada gerada pelo modelo pode ser vista na figura \ref{fig:a_fran_br}.
\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{a_frankel_br} 
  \caption{Mapa do valor-a $[\log\left($\#tremores$/$ano$)/$área$\right)]$, usando o catálogo \gls{bsb2013} calculado
  pelo método de Frankel, 1995 }
  \label{fig:a_fran_br} 
\end{figure}

Apenas para comparação, o mesmo método foi empregado com o catálogo \gls{iscgem}, como pode ser visto na figura 
\ref{fig:a_fran_sa}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{a_frankel_sa} 
  \caption{Mapa do valor-a, catálogo \gls{iscgem} calculado pelo método de Frankel, 1995}
  \label{fig:a_fran_sa} 
\end{figure}

Nos dois casos é possível perceber a compatibilidade entre os resultados na maior parte do Brasil, mas com
diferenças significativas na região do Acre, Rondônia e na porção norte na Dorsal Atlântica. 
Note que pela limitação da escala, os valores em vermelho são quaisquer valores acima de 2.5

Os valores da ameaça calculados conforme a seção \ref{sec:hazard} estão na figura \ref{fig:pga_fran}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_frankel} 
  \caption{Mapa de ameaça sísmica, PGA (poe 10\%, 50y) [Frankel, 1995] }
  \label{fig:pga_fran} 
\end{figure}

Como técnica de suavização, é importante notar os principais altos de ameaça no nordeste e no centro do país.
A ameaça é recuperada razoávelmente na plataforma continental.

%% ------------------------------------------------------------------------- %%
\subsection{Woo, 1996}
\index{Woo, 1996!resultados}
\label{sec:woo_resultados}

O ajuste dos parâmetros da função de largura de banda resultou em $h(m)=1.39e^{1.18\,m}$km.

Aplicando-se o método de Woo ao catálogo brasileiro desagrupado, a taxa de sismicidade suavizada 
pode ser avaliada pela figura \ref{fig:a_woo}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{a_woo} 
  \caption{Mapa do valor-a, usando o catálogo \gls{bsb2013} calculado pelo método de Woo, 1996 }
  \label{fig:a_woo} 
\end{figure}

A sismicidade de regiões como o nordeste, Goiás/Tocantins, sudeste e plataforma continental 
foram bem recuperadas, enquanto partes da região norte e centro-oeste não tiveram tanto destaque.

O cálculo da ameaça (figura \ref{fig:pga_woo_inc}) foi feito com o resultado direto do método, utilizando
diretamente uma \gls{mfd} discreta.
\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_woo_inc} 
  \caption{Mapa de ameaça sísmica, PGA (poe 10\%, 50y) 
  		   calculado com o Openquake a partir das fontes sísmicas
  		   determinas pelo método de Woo, 1996, usando uma \gls{mfd}
  		   discreta e incremental.
  }
  \label{fig:pga_woo_inc} 
\end{figure}

Coerentemente com a taxa de sismicidade modelada. Não houve local em que a taxa
da figura \ref{fig:a_woo} fosse baixa e que tenha apresentado grande ameaça.
As regiões norte, oeste e sul do país quase não figuraram 
enquanto parte do sudeste, plataforma continental, Goiás e nordeste foram
destacadas. Merece ser notado o maior valor de ameaça localizado em Pernambuco, dentre as três principais regiões mais ameaçadas do nordeste.

Para quantificar as diferenças nos cálculos de ameaça em decorrência da \gls{mfd} utilizada, em seguida é apresentado o
mapa de ameaça (figura \ref{fig:pga_woo_cum}) calculado a partir das taxas geradas pelo modelo, mas utilizando-se do
acumulado nos valores de magnitude acima do mínimo do catálogo e aplicando a transformação de Gutemberg-Richter 
para obter o \emph{valor-a}. Nesse procedimento, a partir de certa magnitude minima de referencia, toma-se o cumulativo
das taxas para magnitudes acima da referência, e, se projeta (por \gls{GR}) o \emph{valor-a}.

Essa projeção acumula incertezas que não são inerentes ao método. A proposta de Woo é fornecer
diretamente a distribuição de magnitudes em cada ponto da malha enquanto essa projeção é feita
assumindo um \emph{valor-b} arbitrário.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.47\textwidth}
		\centering
		\includegraphics[width=1.0\textwidth]{pga_woo_cum} 
		\subcaption{Mapa de ameaça sísmica, PGA (poe 10\%, 50y), 
  		   calculado com o Openquake a partir das fontes sísmicas
  		   determinas pelo método de Woo, 1996, usando uma \gls{mfd}
  		   truncada usando o valor-a como o valor cumulativo
  		   contado a partir da \gls{sym:m_min}.
		}
		\label{fig:pga_woo_cum} 
	\end{subfigure}
	\quad
	\begin{subfigure}[t]{0.47\textwidth}
		\centering
		\includegraphics[width=1.0\textwidth]{pga_woo_dif} 
		\subcaption{Mapa diferencial de ameaça, PGA (poe 10\%, 50y), 
		   entre os modelos usando \gls{mfd} truncada e \gls{mfd}
		   discreta. Diferença entre os mapas das figuras 
		   \ref{fig:pga_woo_inc} e \ref{fig:pga_woo_cum}.
		   }
		\label{fig:pga_woo_dif} 
	\end{subfigure}
	\caption{Variação do resultado da ameaça em função do uso de diferentes 
			\glspl{mfd} no OpenQuake.}
	\label{fig:pga_woo} 
\end{figure}

A figura \ref{fig:pga_woo_dif} quantifica a diferença nos valores de ameaça obtitos quando 
utilizadas as diferentes distribuições de frequência e magnitude. Nota-se que enquanto
o modelo com \gls{mfd} incremental (método original) destaca principalmente o nordeste, o modelo utilizando o
\emph{valor-a} cumulativo (projetado) e uma \gls{mfd} truncada, destacou principalmente as ameaças no centro-oeste e
sudeste.

%% ------------------------------------------------------------------------- %%
\subsection{Helmstetter, 2012}
\index{Helmstetter, 2012!resultados}
\label{sec:helmstetter_resultados}

A otimização dos parâmetros do modelo de Helmstetter conforme descrito na seção \ref{sec:proc_helmstetter}
resultou, com ganho\footnote{equação \eqref{eq:gain}} $G = 2.43$ sobre uma distribuição uniforme, nos valores
apresentados na tabela \ref{tab:hemlstetter}.

\begin{table}[H]
	\centering
	\begin{tabular}{c|c}
		Parâmetro & Valor \\ \hline
		$R_{min}$ & $0.1\times10^{-13}$ \\
		$a_{cnn}$ & 325 \\
		$k_{cnn}$ & 1 \\
	\end{tabular}
	\caption{Parâmetros otimizados (método simplex) para o modelo de Helmstetter a partir do catálogo \gls{bsb2013}.}
	\label{tab:hemlstetter}
\end{table}

Usando os parâmetros otimizados, os valores calculados para a taxa de sismicidade
de longo-prazo pelo método proposto por Helmstetter pode ser visto na figura \ref{fig:helm_r}.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{a_helmstetter} 
  \caption{Mapa do valor-a, usando o catálogo \gls{bsb2013} calculado pelo método de Helmstetter, 2012.}
  \label{fig:helm_r} 
\end{figure}

Observa-se que a sismicidade de regiões como amazônia, Mato-Grosso e Pará foram destacadas,
enquanto no sudeste houve maior diluição. A sismicidade do nordeste aparece evidente, principalmente
nas regiões do Ceará, Rio Grande do Norte e Pernambuco.

É fato também que a sismicidade do centro-oeste rumo ao norte pelo estado de Goiás praticamente desapareceu.
Isso se deve provavelmente ao fato de que grande parte dos sismos dessa região possuem magnitudes não 
muito elevadas, tendo sido em grande parte removidas pela magnitude de completude como se pode notar nos catálogos
de aprendizado e testes na figura \ref{fig:h_catalogues}.

Na figura~\ref{fig:helm_h} estão os valores calculados para o mapa de ameaça 
com as taxas de sismicidade suavizadas pelo método.

\begin{figure}[H]
  \centering
  \includegraphics[width=.80\textwidth]{pga_helmstetter} 
  \caption{Mapa de ameaça sísmica, PGA (poe 10\%, 50y), 
  		   calculado com o OpenQuake a partir das fontes sísmicas
  		   determinas pelo método de Helmstetter, 2012.}
  \label{fig:helm_h} 
\end{figure}

Nota-se que ao passo em que houve um grande realçe da ameaça em algumas regiões como amazônia e Pará (próximo a Belém),
em regiões como a faixa sísmica no norte de Goiás e Tocantins não apresentaram o mesmo contraste.

### 2015-01-13 d01-terms.tex
adding images

%-----------------------------------------
% terms
%-----------------------------------------
\newglossaryentry{equake}
{
	name={terremoto},
	description={ruptura de alguma estrutura geológica},
	plural={terremotos}
}

\newglossaryentry{seismicity}
{
	name={sismicidade},
	description={ocorrência dos tremores},
}

\newglossaryentry{hypocenter}
{
	name={hipocentro},
	description={representação geométrica do ponto no espaço, onde 
		se iniciou o \gls{rupture_process} da \gls{crust}},
	plural={hipocentros}
}

\newglossaryentry{epicenter}
{
	name={epicentro},
	description={projeção ortogonal, sobre a superfície, do \gls{hypocenter}},
	plural={epicentros}
}

\newglossaryentry{seismotectonic}
{
	name={sismotectônica},
	description={o estudo das relações entre os \glspl{equake} e a \gls{tectonic} recente de uma região.
				 Procura compreender exatamente quais são os mecanismos que levam à uma ruptura geológica 
				 e são responsáveis pela
				 \gls{seismic_activity} em uma certa área. Isso é feito analisando-se de forma combinada 
				 registros recentes de tectonismo global e regional, 
				 considerando também evidências históricas e geomorfológicas},
}

\newglossaryentry{rupture_process}
{
	name={processo de ruptura},
	description={processo que envolve o rompimento de uma região da crosta,
			o deslocamento relativo entre essas regiões, e consequantemente,
			a liberação de uma grande quantidade de energia, de forma praticamente
			instantânea, tomando-se como referência o \gls{geologic_time}},
	plural={processos de ruptura}
}

\newglossaryentry{geologic_time}
{
	name={tempo geológico},
	description={escala de tempo que vai desde a formação do universo até os tempos atuais,
				englobando a formação do planeta e as transformações ocorridas desde então},
}

\newglossaryentry{tectonic}
{
	name={tect\^onica},
	description={disciplina científica focada nos processos respons\'aveis 
				 pela cria\c{c}\~ao e transforma\c{c}\~ao das estruturas geológicas da Terra e de outros planetas.},
	plural={tect\^onicas}
}

%\newglossaryentry{oq}
%{
%	name={OpenQuake},
%	description={programa de código aberto para o calculo de risco sísmico mantido pela Fundação \gls{gem}},
%	plural={OpenQuake}
%}

\newglossaryentry{crust}
{
	name={crosta terrestre},
	description={parte superficial, rígida e mais externa do planeta Terra},
}

\newglossaryentry{mantle}
{
	name={manto terrestre},
	description={material da por{ç}{ã}o intermediária do planeta, 
		fluido em tempo geológico},
}

\newglossaryentry{core}
{
	name={n{ú}cleo terrestre},
	description={por{ç}{ã}o mais central do planeta, com predomin{â}ncia de compostos metálicos},
}

\newglossaryentry{tectonic_plate_theory}
{
	name={teoria tect{ô}nica das placas},
	description={foi uma teoria revolucionária para a \gls{tectonic},
				propondo que a \gls{crust} terrestre estivesse dividida 
				em placas {à} deriva sobre o \gls{mantle}},
}

\newglossaryentry{litho_plate}
{
	name={placa litosf{é}rica},
	plural={placas litosf{é}ricas},
	description={placa de material da \gls{lithosphere}},
}

\newglossaryentry{lithosphere}
{
	name={litosfera},
	description={região rúptil, mais externa do planeta, formada pela \gls{crust} 
		(continental e ocêanica) e parte do \gls{mantle} superior, com aproximadamente 
		60\gls*{sym:km} de profundidade},
}

\newglossaryentry{astenosphere}
{
	name={astenosfera},
	description={região dúctil entre a \gls{lithosphere} e o \gls{mantle},
				com profundidades que variam de 60 a 700km},
}

\newglossaryentry{smoothing}
{
	name={técnicas de suavização},
	description={consiste em capturar importantes feições do conjunto de dados,
				 eliminando ruídos e outras estruturas de curto comprimento de onda
				 presentes nos dados},
}

\newglossaryentry{kernel_function}
{
	name={função de núcleo},
	description={funções n-dimensionais, cuja integral em todo o domínio resulta em 1,
				 podendo ser usadas como estimativas para 
				 funções de densidade de probabilidade},
	plural={funções de kernel},
}

\newglossaryentry{seismic_rate}
{
	name={taxa de sismicidade},
	description={taxa com que terremotos são produzidos por determinada \gls{seismic_source}},
	plural={taxas de sismicidade},
}

\newglossaryentry{seismic_activity}
{
	name={atividade sísmica},
	description={frequ{ê}cia de ocorr{ê}ncia de \glspl{equake}},
}

\newglossaryentry{poisson_process}
{
	name={processo de Poisson},
	description={uma sequencia de intervalos discretos com um experimento de Bernoulli em cada},
}

\newglossaryentry{seismic_source}
{
	name={fonte sísmica},
	description={estrutura geológica capaz de produzir tremores de terra},
	plural={fontes sísmicas}
}

\newglossaryentry{point_source}
{
	name={fonte sísmica pontual},
	description={representação geométrica por um ponto, de uma fonte sísmica},
	plural={fontes sísmicas pontuais},
}

\newglossaryentry{gmpe}
{
	name={GMPE},
	description={Equação de predição do movimento do chão},
	plural={GMPEs},
}

\newglossaryentry{area_source}
{
	name={fonte sísmica poligonal},
	description={representação geométrica por um polígono em superfície, 
				 de uma fonte sísmica},
	plural={fontes sísmicas poligonais},
}

\newglossaryentry{titulo_da_dissertacao}
{
	name={titulo_da_dissertacao},
	description={Técnicas de suavização aplicadas
					à caracterização de fontes sísmicas e 
					à análise probabilística de ameaça sísmica},
}

\newglossaryentry{isocista}
{
	name={isocista},
	description={curva que une valores de mesma intensidade sísmica},
}

### 2015-01-13 d03-acronyms.tex
adding images

%-----------------------------------------
% acronyms
%-----------------------------------------

\newacronym{psha}{PSHA}{Análise Probabilística de Ameaça Sísmica}
\newacronym{dsha}{DSHA}{Análise Determinística de Ameaça Sísmica}
\newacronym{GMPE}{\gls{gmpe}}{\glsdesc{gmpe}}
\newacronym{mfd}{MFD}{Distribuição de Frequência e Magnitude}
\newacronym{gr}{GR}{Gutenberg-Richter}
\newacronym{cnn}{CNN}{Acoplamento dos Vizinhos mais Próximos}
\newacronym{va}{v.a.}{variável aleatória}
\newacronym{iid}{i.i.d.}{independentes e identicamente distribuidas}
\newacronym{pdf}{densidade}{função de densidade de probabilidade}
\newacronym{pmf}{pmf}{função de distribuição acumulada de probabilidade}

\newacronym{pga}{PGA}{máxima aceleração do chão}

\newacronym{isc}{ISC}{International Seismological Centre}
\newacronym{gem}{GEM}{Global Earthquake Model}
\newacronym{opensha}{openSHA}{Open Source Seismic Hazard Analysis}
\newacronym{oq}{OQ}{Openquake}
\newacronym{eeri}{EERI}{Earthquake Engineering Research Institute}
\newacronym{cgmw}{CGMW}{Comission for Geological Map of the World}

\newacronym{hl}{HazardLib}{Biblioteca de Ameaça}
\newacronym{rl}{RiskLib}{Biblioteca de Risco}
\newacronym{nrml}{NRML}{Natural-hazard and Risk Markup Language}
\newacronym{hmtk}{HMTK}{Hazard Modeller's Toolkit}
\newacronym{oqp}{oq-platform}{Plataforma web de Interação com o OQ}
\newacronym{oqe}{oq-engine}{Motor de Cálculo do OQ}

\newacronym{obsis}{ObSis}{Observatório Sismológico}
\newacronym{unb}{UnB}{Universidade de Brasilia}
\newacronym{iag}{IAG}{Instituto de Astronomia, Geofísica e Ciências Atmosféricas}
\newacronym{usp}{USP}{Universidade de São Paulo}
\newacronym{ipt}{IPT}{Instituto de Pesquisas Tecnológicas}
\newacronym{unesp}{UNESP}{Universidade Estatual Paulista}
\newacronym{ufrn}{UFRN}{Universidade Federal do Rio Grande do Norte}

\newacronym{bsb}{BSB}{Boletim Sísmico Brasileiro}
\newacronym{csv}{CSV}{Valores Separados por Vírgulas}

\newacronym{bsb2013}{{BSB-2013.08}}{Brazilian Seismic Bulletin version 2013.08}
\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{Catálogo ISC-GEM para a América do
Sul}

### 2015-01-13 manuscript.tex
adding images

@@
+%!TEX program = pdflatex
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % See how to enter figures and tables at the end of the article, after references.
+\usepackage{float}
+%\usepackage{caption}
+\usepackage{subcaption}
+
+%----------------------------------------------------------------------------------------
+%	Glossaries
+%----------------------------------------------------------------------------------------
+
+% Glossaries
+\usepackage[sanitize=none,acronym,toc]{glossaries}
+%\usepackage[acronym,toc]{glossaries}
+% Define a new glossary type
+\newglossary[slg,toc]{symbols}{sym}{sbl}{Lista de Símbolos}
+\newglossary[elg]{equations}{eqn}{eql}{Equations}
+\makeglossaries
+
+% ---------------------------------------------------------------------------- %
+% Dicionario de Termos:
+\input{d01-terms}
+\input{d02-symbols}
+\input{d03-acronyms}
+\input{d04-equations}
+% ---------------------------------------------------------------------------- %
+
+% ---------------------------------------------------------------------------- %
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\erf}{erf}
+% ---------------------------------------------------------------------------- %
 
@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Brazilian Smoothing Seismicity and PSHA}
+\titlerunninghead{Smoothed seismicity Brazilian PSHA}
 
@@
 % Corresponding author mailing address and e-mail address:
-\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, São Paulo, Brazil.
+\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
 (marlon.pirchiner@iag.usp.br)}
@@
 Intraplate seismicity is a hard problem to solve in tectonics,
-main during the lack of observed seismicity in many areas where seismic 
+mainly during the lack of observed seismicity in many areas where seismic
 hazard assesment are also difficult and the methods applied are often
@@
 
-All the methods are based on a kernel density estimation. The main 
-difference between them is the \emph{bandwidth selection} process.
+All this methods are based on a kernel density estimation. The main
+difference between them is the bandwidth selection process.
 
-The Frankel method use a \emph{fixed} bandwidth in space. Woo method
+Frankel in their method uses a \emph{fixed} bandwidth in space. Woo's method
 suggest a bandwidth as function of the \emph{magnitude}.
 Helmstetter and Werner proposes a \emph{local-adaptive}
-in space and time bandwidth, chosen for each event. 
-It was proposed initialy to take care of the background seismicity rate 
-on long-term earthquake forecast context, but the one year forecasted 
-seismic rate could be a grid of points seismic sources under a 
-time-independent assumption.
+in space and time kernel bandwidths, chosen for each event.
+This last one was proposed initialy to take care of the background seismicity rate
+on long-term earthquake forecast context and the time-independent assumption
+about the annual background seismic rate could be used
+(not indiscriminately REFERENCE) as a grid of points seismic sources.
 
-In the conclusions the results shown that
+As conclusions the results shown that all methods gives defines better
+some known seismic regions and could be used in the Brazilian seismicity case,
+all of them, improoving the spatial resolution of the GSHAP model, the last one
+made on the region.
 
-all methods gives good results and 
-could be used in the Brazilian seismicity case,
+The Last two methods was implemented and are available as a free software on
+the Hazard Modeling Toolkit (HMTK) from GEM Foundation.
 
-all of them, 
-% ###### improoving #########
-the spatial resolution of the actual state-of-art GSHAP model.
-
-The Last two methods was implemented and are available as a free software on HMTK from GEM Foundation. 
 All the hazard calculations was performed on OpenQuake open source software.
@@
 
-\subsection{Intraplate Seismicity}
-\subsection{Previous Brazilian Studies}
-\subsection{Previous Smoothing Seismicity (Frankel,Woo,Helmstetter\ldots)}
-\subsection{Specifical Purpose}
-
 Brazil have no recent seismic hazard study at national scale despite some important and localized recent studies
@@
 
-The Brazilain seismic hazard building code was made based on GSHAP model which is the well accepted state of art model
+The Brazilian seismic hazard building code was made based on GSHAP model which is the well accepted state of art model
 for the whole region.
@@
 
+\subsection{Intraplate Seismicity}
+
+Onononono on on ono on on on on ononon on on on on on ono non on ono on on
+
+\subsection{Previous Brazilian Studies}
+
+Onononono on on ono on on on on ononon on on on on on ono non on ono on on
+
+\subsection{Previous Smoothing Seismicity (Frankel, Woo, Helmstetter\ldots)}
+
+Onononono on on ono on on on on ononon on on on on on ono non on ono on on
+
+\subsection{Specifical Purpose}
+
 %------------------------------------------------
+\section{Data}
 
-\section{Materials and Methods}
+%------------------------------------------------
+\section{Methods}
 
-\subsection{Brazilian Seismic Catalog}
-Boletim sismico brasileiro.
+\subsection{Frankel}
 
-História e versão com Mw.
-\subsection{Brazilian Seismic Catalog}
+A proposta de Arthur \citet{frankel_1995} foi usar uma largura de banda fixa,
+nomeada distância de correlação $d_F$, e aplicar o estimador de Nadaraya-Watson (\ref{sec:nadaraya})
+para suavizar o histograma 2D da sismicidade utilizando uma função de núcleo gaussiana:
+\begin{equation}
+	\ensuremath{
+		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
+						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
+	}
+	\label{eq:ni}
+\end{equation}
+onde $\tilde{n}_j$ é a taxa de sismicidade (número de sismos com magnitude $m$ maior que a mínima magnitude
+\gls{sym:Md} do catálogo) suavizada
+na célula $j$, $n_i$ é o número de sismos em cada outra célula $i$ e
+	\gls{sym:dij} é \glsdesc{sym:dij}.
 
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
+\subsection{Woo}
+
+Já Gordon \citet{woo_1996} propôs avaliar a contribuição de uma função de núcleo
+de cada sismo $i$ ocorrido em $\boldsymbol{r}_i$
+na célula centrada em $\boldsymbol{r}$ que dependa também de sua
+magnitude $m$:
+\begin{equation}
+	\ensuremath{
+		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
+											{T({\gls{sym:ri})}},
+	}
+	\label{eq:Rrm}
+\end{equation}
+onde $N$ é o número de tremores $i$ no catálogo
+e $T(\gls{sym:ri})$ é o período em que todo sismo de magnitude acima de $m$ é completamente observado
+em \gls{sym:ri}.
+
+\subsection{Helmstetter}
 
 \begin{equation}
-\label{eq:emc}
-e = mc^2
+	\ensuremath{\gls{sym:R} = \sum_{i=1}^{N}{ \frac{1}{h_i\,{d_i}^2} \gls{sym:Kt}\gls{sym:Kr} }},
+	\label{eq:helms01}
+\end{equation}
+onde \gls{sym:R} é \glsdesc{sym:R},
+	  $K_t$ é a \glsdesc{sym:Kt},
+	  $K_r$ é a \glsdesc{sym:Kr}.
+
+Como seu interesse é em projeção, restringiram os tempos do catálogo aos $t_i
+< t$ e ponderaram segundo um peso $w$ da seguinte maneira:
+\begin{equation}
+\ensuremath{\gls{sym:R} = \gls{sym:Rmin} + \sum_{t_i < t}{
+	\frac{2\,w(\boldsymbol{r}_i,t_i)}{h_i\,{d_i}^2}
+			\gls{sym:Kt}\gls{sym:Kr} }},
+	\label{eq:helms02}
+\end{equation}
+onde \gls{sym:Rmin} é a \glsdesc{sym:Rmin}, positiva, permitindo que ocorram eventos-surpresa
+onde não se tem registro de sismicidade.
+
+Os pesos $w(\boldsymbol{r}_i,t_i)$, calculados em cada tremor $i$, propostos por Helmstetter são a transformação de
+Gutemberg-Richter. Consideraram uma contribuição maior para a taxa de sismicidade daqueles sismos
+que ocorreram onde e quando a magnitude de completude $M_c$ era maior que
+a mínima magnitude $M_d$ do catálogo
+em locais onde a magnitude de completude e o \emph{valor-b} variam.
+Note que o modelo de pesos acomoda bem variações espaciais e
+temporais tanto da magnitude de completude como do \emph{valor-b}. A expressão é a seguinte:
+\begin{equation}
+	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
+	\right] } },
+	\label{eq:helms_wi}
+\end{equation}
+onde  \gls{sym:wi} é o \glsdesc{sym:wi} na localização $\boldsymbol{r}$ e no instante $t$,
+	  \gls{sym:b}$(\boldsymbol{r},t)$ é o \glsdesc{sym:b},
+	  \gls{sym:Mc_rt} é a \glsdesc{sym:Mc_rt},
+	  \gls{sym:Md} é a \glsdesc{sym:Md}.
+
+\subsubsection{Bandwidth computation}
+
+\begin{equation}
+	\ensuremath{
+%		h_i, d_i = \underset{d_i \ge \gls{sym:dk}, h_i \ge \gls{sym:hk}}{\argmin}
+		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
+						              d_i \ge \gls{sym:dk}}
+				           }
+		\left[ s \left(h_i,d_i
+			 		  \arrowvert
+					  \gls{sym:k_cnn},\gls{sym:a_cnn}
+			     \right)
+			   := h_i + \gls{sym:a_cnn}d_i
+	    \right],
+	}
+	\label{eq:helms_cnn}
+\end{equation}
+onde \gls{sym:k_cnn} é o \glsdesc{sym:k_cnn},
+	 \gls{sym:a_cnn} é o \glsdesc{sym:a_cnn},
+	 \gls{sym:dk} é o \glsdesc{sym:dk} e
+	 \gls{sym:hk} é o \glsdesc{sym:hk}.
+
+As larguras de bandas são escolhidas localmente, podendo ser menor, aumentando a resolução
+onde há maior densidade de informação, ou crescer onde há escassez de informação e menor resolução.
+
+\subsubsection{Taxa de sismicidade estacionária}
+Mesmo havendo modelado a dependência com o tempo para distribuição da taxa de sismicidade,
+a taxa estacionária $\bar{R}$ que se preserva no longo-prazo em uma determinada localização
+$\boldsymbol{r}_0$ é a que mais interessa à \gls{psha} e,
+segundo o método proposto por \citet{helmstetter_2012}, será definida
+pelo valor da mediana da taxa fornecida pelo modelo no
+período considerado:
+
+\begin{equation}
+	\ensuremath{
+		\bar{R}(\boldsymbol{r}_0) = \text{Mediana}\left[R(\boldsymbol{r}_0, t)\right].
+	}
+	\label{eq:stationary_rate}
 \end{equation}
 
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Mauris eu sapien nunc, sit amet accumsan dui.
-Nulla ac diam ut nunc placerat semper eget et libero.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Cras hendrerit ullamcorper sapien vitae luctus.
-Quisque vel diam massa.
-Vestibulum dui nibh, facilisis vel vestibulum eu, viverra in quam.
+O fato de se tomar a mediana faz com que variações na taxa de sismicidade devido a pré e pós-eventos
+não interfiram significativamente no valor da taxa estacionária. A decorrência mais importante desse fato
+é que o método dispensa a remoção de agrupamentos (seção \ref{sec:declustering}) do catálogo, um método paramétrico e
+quase sempre controverso.
 
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
+\subsubsection{Verossimilhança}
+
+Para otimizar os parâmetros do modelo de Helmstetter, é preciso dividir o catálogo de sismos.
+Uma parte serve para o aprendizado do modelo e a outra para avaliação.
+
+Se a ocorrência de tremores pode ser modelada por um processo de Poisson com taxa $N_p$,
+então a probabilidade de se observar exatamente $n$ eventos no período de tempo considerado é dada por
+
+\begin{equation}
+	\ensuremath{
+		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
+	}
+	\label{eq:poisson_probability}
+\end{equation}
+
+E o logarítmo, a ser maximizado, da verossimilhança entre
+o que o modelo predisse e o que foi observado
+pode ser escrito como
+
+\begin{equation}
+	\ensuremath{
+		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
+	}
+	\label{eq:loglik}
+\end{equation}
+onde \gls{sym:Np} é \glsdesc{sym:Np},
+	\gls{sym:nxy} é \glsdesc{sym:nxy}.
+
+Os parâmetros $R_{min}$, $a_{cnn}$ e $k_{cnn}$ do modelo são otimizados
+pela maximização da verossimilhança $L$.
 
@@
 Morbi vel felis mi.
-Referencing Table \ref{sampletable}.
+Referencing Table \ref{fig_seismicity}.
 Referencing Figure \ref{placeholder}.
@@
 
+% =========
+% intro
+% =========
+
 \begin{figure}
-\includegraphics[width=0.4\linewidth]{placeholder.jpg}
-\caption{Figure caption}\label{placeholder}
+	\centering
+	\begin{subfigure}[b]{0.3\textwidth}
+		\includegraphics[width=0.4\textwidth]{placeholder}
+		\caption{caption A}
+		\label{placeholder_a}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.3\textwidth}
+		\includegraphics[width=0.4\textwidth]{placeholder}
+		\caption{caption B}
+		\label{placeholder_b}
+	\end{subfigure}
+	\caption{main caption}
+	\label{placeholder}
 \end{figure}
 
-\begin{table}
-\caption{Table caption}\label{sampletable}
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_seismicity_br}
+	\caption{Brazilian earthquake catalog.}
+	\label{fig_seismicity}
+\end{figure}
+
+% =========
+% data
+% =========
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_hmtk_bsb2013_rate}
+	\caption{Earthquake records by year}
+	\label{fig_seismic_rate}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_time_mag_count_br_1960}
+	\caption{Magnitude count by year.}
+	\label{fig_time_mag_count}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_decluster_br}
+	\caption{Catalog declustering.}
+	\label{fig_decluster}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_occurrence}
+	\caption{Annual earthquake recurrence.}
+	\label{fig_occurrence}
+\end{figure}
+
+% =========
+% Methods
+% =========
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_a_frankel_br}
+	\caption{Smoothed seismicity: Frankel's method.}
+	\label{fig_a_frankel}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
+	\caption{Magnitude dependence bandwidth.}
+	\label{fig_woo_bandwidth}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_a_woo}
+	\caption{Smoothed seismicity: Woo's method.}
+	\label{fig_a_woo}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
+	\caption{Local bandwidth example.}
+	\label{fig_helmstetter_hidi}
+\end{figure}
+
+% TODO add HI_DI histograms
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
+	\caption{Stationary seismic rate}
+	\label{fig_helmstetter_stationary_rate}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
+	\caption{Learning and target catalogues.}
+	\label{fig_helmstetter_catalogues}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_a_helmstetter}
+	\caption{Smoothed seismicity: Helmstetter's method.}
+	\label{fig_a_helmstetter}
+\end{figure}
+
+% =========
+% Results
+% =========
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_pga_gshap}
+	\caption{GSHAP results: PGA (poe 10\%/50y)}
+	\label{fig_pga_gshap}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_pga_frankel}
+	\caption{Frankel's results: PGA (poe 10\%/50y)}
+	\label{fig_pga_frankel}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_pga_woo}
+	\caption{Woo's results: PGA (poe 10\%/50y)}
+	\label{fig_pga_woo}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_pga_helmstetter}
+	\caption{Helmstetter's results: PGA (poe 10\%/50y)}
+	\label{fig_pga_helmstetter}
+\end{figure}
+
+% TABLES
+
+\begin{table}[H]
+	\caption{Table caption}
 	\begin{tabular}{l l l}
@@
 	\end{tabular}
+	\label{sampletable}
 \end{table}
@@
  will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
+ \[ x^2=y^2 + z^2 \]
 

### 2015-01-14 gji_style/gjilguid2e.tex
gji + plain subfigure + partial translation + nadaraya ref correction

% gjilguid2e.tex
% V2.0 released 1998 December 18
% V2.1 released 2003 October 7 -- Gregor Hutton, updated the web address for the style files.

\documentclass{gji}
\usepackage{timet}

\title[Smoothing seismicity applied to Brazilian PSHA]
  {Smoothing seismicity applied to Brazilian PSHA}
\author[M. Pirchiner]
  {Marlon Pirchiner$^{1,2}$\thanks{Marcelo Assump\c{c}\~ao, Vincent Guigues and Stephane Drouet} \\
  $^1$ Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG), University of S\~ao Paulo (USP) \\
  $^2$ Applied Math School (EMAp), Get\'ulio Vargas Foundation (FGV), Rio de Janeiro.}
\date{original form January, 2015}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{XX}
\pubyear{2015}

%\def\LaTeX{L\kern-.36em\raise.3ex\hbox{{\small A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
%\def\LATeX{L\kern-.36em\raise.3ex\hbox{{\Large A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
% Authors with AMS fonts and mssymb.tex can comment out the following
% line to get the correct symbol for Geophysical Journal International.
\let\leqslant=\leq

\newtheorem{theorem}{Theorem}[section]

\begin{document}

\label{firstpage}

\maketitle

\begin{summary}
 This guide is for authors who are preparing papers for
 \textit{Geophysical Journal International} using the
\LaTeXe\ document preparation system and the GJI class file.
\end{summary}

\begin{keywords}
 \LaTeXe\ -- class files: \verb"gji.cls"\ -- sample text -- user guide.
\end{keywords}

\section{Introduction}

In addition to the standard submission of hardcopy from authors,
\textit{Geophysical Journal International} accepts machine-readable forms
of  papers in \LaTeX.
The layout design for \textit{Geophysical Journal International} has been
implemented as a \LaTeXe\ class file derived from the MN style file for
Monthly Notices of the Royal Astronomical Society. The GJI classfile is
based on the \verb"ARTICLE" style as discussed in the \LaTeX\ manual
\cite {la}. Commands which differ from the standard \LaTeX\ interface, or
which are provided in addition to the standard interface, are explained
in this guide. This guide is not a substitute for the \LaTeX\ manual
itself. Authors planning to submit their papers in \LaTeX\ are advised to
use \verb"gji.cls" as  early as possible in the creation of their
files.  This guide is modified from that produced by Woollatt et al (1994)
to describe the features of the MN style.

A very accessible guide to the features of \LaTeXe and the differences
from the earlier version is provided by Kopka \& Daly \shortcite{kd}.  This
reference provides in chapter 9 a summary of \LaTeX\ error messages and also
a full description of standard \LaTeX\ commands in Appendix F.

\subsection{The GJI document classes}

The use of \LaTeX\ document classes allows a simple change of class
(or class option) to transform the appearance of your document. The
GJI class file preserves the standard \LaTeX\ interface such that any
document which can be produced using the standard \LaTeX\ \verb"ARTICLE"
class can also be produced with the GJI class. However, the measure (or
width of text) is narrower than the default for \verb"ARTICLE", therefore
line breaks will change and long equations may need re-setting.

\subsection{General style issues}

For general style issues, authors are referred to the `Instructions for
Authors' on the inside back cover of \textit{Geophysical Journal
International}. Authors
who are interested in the details of style are referred to Butcher
\shortcite {bu} and The Chicago Manual \shortcite {cm}. The language of
the journal is British English and spelling should conform to this.

Use should be made of symbolic references (\verb"\ref") in order to
protect against late changes of order, etc.

\subsection{Submission of \LaTeX\ articles to the journal}

Papers should initially be submitted in the usual way to:
The Executive Secretary, Royal Astronomical Society, {\em or\/}
the EGS Editor,  {\em or\/} the DGG editor,  {\em or\/}
the American Editor,  {\em or\/} the Pacific Region Editor,
as set out in the Instructions for Authors on the inside back cover of
each issue of Geophysical Journal International.
Four hard copies should be supplied including figures, normally using the
\verb"[referee]" option, for papers with a  high mathematical content the
\verb"[mreferee]" option is recommended. In each case a separate page of
figure captions is preferred.
One of the copies should  be single-sided,
while the other two should  be weight-reduced, by being either
single-spaced or double-sided.   Copies of figures should
also be supplied. Authors should  ensure that their figures are suitable
(in terms of lettering size, etc.)  for the reductions they intend; they
should not attempt to include their figures inside a \TeX\ or \LaTeX\
file by using \verb"\special" or one of the style files for figure
handling.

Note that articles, or revised versions thereof, may not currently be
submitted by electronic mail. However when the article is accepted for
publication the \LaTeX\ file can be sent to the publisher by \verb"ftp"
together with appropriate forms of figures.  Instructions will be provided
following acceptance.

\section{Using the GJI class file}

If the file \verb"gji.cls" is not already in the appropriate
system directory for \LaTeX\ files, either arrange for it to be
put there, or copy it to your working directory. The class file
and related material, such as this guide, can be accessed via the
journal web-site  at
http://www.blackwellpublishing.com/journals/gji under {\em Author
Guidelines}.

The GJI document class is implemented as a complete document class, {\em
not\/} a document class option. In   order to use the GJI style, replace
\verb"article" by
\verb"gji" in the
\verb"\documentclass" command at the beginning of your document:
\begin{verbatim}
\documentclass{article}
\end{verbatim}
is replaced by
\begin{verbatim}
\documentclass{gji}
\end{verbatim}
In general, the following standard document class options should {\em
not\/} be used with the GJI style:
\begin{enumerate}
  \item \texttt{10pt}, \texttt{11pt}, \texttt{12pt} -- unavailable;
  \item \texttt{twoside} (no associated style file) --
     \texttt{twoside} is the default;
  \item \texttt{fleqn}, \texttt{leqno}, \texttt{titlepage} --
        should not be used (\verb"fleqn" is already incorporated into
        the GJI style);
  \item \texttt{twocolumn} -- is not necessary as it is the default style.
\end{enumerate}

In \LaTeX2e the use of postscript fonts and the inclusion of non-standard
options is carried out through the \verb"\usepackage" command, rather than
as options as in earlier versions.  Thus the Times font can be used for
text by including
\begin{verbatim}
\usepackage{times}
\end{verbatim}
on the line immediately after the \verb"\documentclass". If necessary,
\texttt{ifthen} and \texttt{bezier} can be included as packages.

The GJI class file has been designed to operate with the standard
version of \verb"lfonts.tex" that is distributed as part of \LaTeX
. If you have access to the source file for this guide,
\verb"gjilguid2e.tex", attempt to typeset it.  If you find font
problems you might investigate whether a non-standard version of
\verb"lfonts.tex" has been installed in your system.

\subsection{Additional document class options}\label{classoptions}

The following additional class options are available with the GJI style:
\begin{description}
  \item \texttt{onecolumn} -- to be used \textit{only} when two-column output
        is unable to accommodate long equations;
  \item \texttt{landscape} -- for producing wide figures and tables which
        need to be included in landscape format (i.e.\ sideways) rather
        than portrait (i.e.\ upright). This option is described below.
  \item \texttt{doublespacing} -- this will double-space your
        article by setting \verb"\baselinestretch" to 2.
  \item \texttt{referee} -- 12/20pt text size, single column,
        designed for submission of papers.
  \item \texttt{mreferee} -- 11/17pt text size, single column
        designed for submission of papers with mathematical content.
  \item \texttt{camera} -- designed for use with computer modern fonts to
        produce a closer representation of GJI style for camera
        ready material.
  \item \texttt{galley} -- no running heads, no attempt to align
        the bottom of columns.
\end{description}

\subsection{Landscape pages}

If a table or illustration is too wide to fit the standard measure, it
must be turned, with its caption, through 90 degrees anticlockwise.
Landscape illustrations and/or tables cannot be produced directly
using the GJI style file because \TeX\ itself cannot turn the
page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\begin{enumerate}
  \item Use the \verb"table*" or \verb"figure*" environments in your
        document to create the space for your table or figure on the
        appropriate page of your document. Include an empty
        caption in this environment to ensure the correct
        numbering of subsequent tables and figures. For instance, the
        following code prints a page with the running head, a message
        half way down and the figure number towards the bottom. If you
        are including a plate, the running headline is different, and you
        need to key in the three lines which are marked with \verb"% **",
        with an appropriate headline.
\begin{verbatim}
% ** \clearpage
% ** \thispagestyle{plate}
% ** \plate{Opposite p.~812, GJI, \textbf{135}}
\begin{figure*}
  \vbox to220mm{\vfil Landscape figure to
                go here. \vfil}
  \caption{}
  \label{landfig}
\end{figure*}
\end{verbatim}
\item Create a separate document with the corresponding document style
      but also with the \verb"landscape" document style option, and
      include the \verb"\pagestyle" command, as follows:
\begin{verbatim}
\documentclass[landscape]{gji}
\pagestyle{empty}
\end{verbatim}
  \item Include your complete tables and illustrations (or space for
        these) with captions using the \verb"table*" and \verb"figure*"
        environments.
  \item Before each float environment, use the
        \verb"\setcounter" command to ensure the correct numbering of
        the caption. For example,
\begin{verbatim}
\setcounter{table}{0}
\begin{table*}
 \begin{minipage}{115mm}
 \caption{Images of global seismic tomography.}
 \label{tab1}
 \begin{tabular}{@{}llllcll}
   :
 \end{tabular}
 \end{minipage}
\end{table*}
\end{verbatim}
The corresponding example for a figure would be:
\begin{verbatim}
\clearpage
\setcounter{figure}{12}
\begin{figure*}
 \vspace{144mm}
 \caption{Travel times for regional model.}
 \label{fig13}
\end{figure*}
\end{verbatim}
\end{enumerate}

\section{Additional facilities}

In addition to all the standard \LaTeX\ design elements, the GJI style
includes the following features.
\begin{enumerate}
  \item Extended commands for specifying a short version of the title and
        author(s) for the running headlines;
  \item A \verb"summary" environment to produce a suitably indented
        Summary
  \item An \verb"abstract" environment which produces the GJI style of
        Summary
  \item A \verb"keywords" environment and a \verb"\nokeywords" command;
  \item Use of the \verb"description" environment for unnumbered lists.
  \item A starred version of the \verb"\caption" command to produce
        captions for continued figures or tables.
 \end{enumerate}
 In general, once you have used the additional \verb"gji.cls" facilities
in your document, do not process it with a standard \LaTeX\ style file.

\subsection{Titles and author's name}

In the GJI style, the title of the article and the author's name (or
authors' names) are used both at the beginning of the article for the
main title and throughout the article as running headlines at the top
of every page. The title is used on odd-numbered pages (rectos) and the
author's name appears on even-numbered pages (versos). Although the
main heading can run to several lines of text, the running headline
must be a single line ($\leqslant 45$ characters). Moreover, the main
heading can also incorporate new line commands (e.g. \verb"\\") but
these are not acceptable in a running headline. To enable you to
specify an alternative short title and an alternative short author's
name, the standard \verb"\title" and \verb"\author" commands have been
extended to take an optional argument to be used as the running
headline. The running headlines for this guide were produced using the
following code:
\begin{verbatim}
\title[Geophys.\ J.\ Int.:
       \LaTeXe\ Guide for Authors]
  {Geophysical Journal International:
   \LaTeXe\ style guide for authors}
\end{verbatim}
and
\begin{verbatim}
\author[B.L.N. Kennett]
   {B.L.N. Kennett$^1$
  \thanks{Pacific Region Office, GJI} \\
  $^{1}$Research School of Earth Sciences,
    Australian National University,
    Canberra ACT \emph{0200}, Australia
  }
\end{verbatim}
The \verb"\thanks" note produces a footnote to the title or author.

\subsection{Key words and Summary}

At the beginning of your article, the title should be generated in the
usual way using the \verb"\maketitle" command. Immediately following
the title you should include a Summary followed by a list of key
words. The summary should be enclosed within an \verb"summary"
environment, followed immediately by the key words enclosed in a
\verb"keywords" environment. For example, the titles for this guide
were produced by the following source:
\begin{verbatim}
\maketitle
\begin{summary}
 This guide is for authors who are preparing
 papers for \textit{Geophysical Journal
 International} using the \LaTeXe\ document
 preparation system and the GJI style file.
\end{summary}
\begin{keywords}
 \LaTeXe\ -- class files: \verb"gji.cls"\ --
 sample text -- user guide.
\end{keywords}

\section{Introduction}
  :
\end{verbatim}
The heading `\textbf{Key words}' is included automatically and the key
words are followed by vertical space. If, for any reason, there are no
key words, you should insert the \verb"\nokeywords" command immediately
after the end of the \verb"summary" or \verb"abstract" environment. This
ensures that the   vertical space after the abstract and/or title is
correct and that any
\verb"thanks" acknowledgments are correctly included at the bottom of
the first column. For example,
\begin{verbatim}
\maketitle
\begin{abstract}
  :
\end{abstract}
\nokeywords

\section{Introduction}
  :
\end{verbatim}

Note that the \verb"summary" and \verb"abstract" environments have the same
effect for the documentclass \verb"gji.cls"

\subsection{Lists}

The GJI style provides numbered lists using the \verb"enumerate"
environment and unnumbered lists using the \verb"description"
environment with an empty label. Bulleted lists are not part of the GJI
style and the \verb"itemize" environment should not be used.

The enumerated list numbers each list item with roman numerals:
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
Alternative numbering styles can be achieved by inserting a
redefinition of the number labelling command after the
\verb"\begin{enumerate}". For example, the list
\begin{enumerate}
\renewcommand{\theenumi}{(\arabic{enumi})}
  \item first item
  \item second item
  \item etc\ldots
\end{enumerate}
was produced by:
\begin{verbatim}
\begin{enumerate}
 \renewcommand{\theenumi}{(\arabic{enumi})}
  \item first item
       :
\end{enumerate}
\end{verbatim}
Unnumbered lists are provided using the \verb"description" environment.
For example,
\begin{description}
  \item First unnumbered item which has no label and is indented from
        the left margin.
  \item Second unnumbered item.
  \item Third unnumbered item.
\end{description}
was produced by,
\begin{verbatim}
\begin{description}
 \item First unnumbered item...
 \item Second unnumbered item.
 \item Third unnumbered item.
\end{description}
\end{verbatim}

\subsection{Captions for continued figures and tables}

The \verb"\caption*" command may be used to produce a caption with the
same number as the previous caption (for the corresponding type of
float). For instance, if a very large table does not fit on one page,
it must be split into two floats; the second float should use the
\verb"caption*" command with a suitable caption:
\begin{verbatim}
\begin{table}
 \caption*{-- \textit{continued}}
  \begin{tabular}{@{}lccll}
  :
  \end{tabular}
\end{table}
\end{verbatim}

 \begin{figure}
     \vspace{5.5cm}
     \caption{An example figure in which space has been
              left for the artwork.}
     \label{sample-figure}
  \end{figure}

\section[]{Some guidelines for using\\* standard facilities}

The following notes may help you achieve the best effects with the GJI
style file.

\subsection{Sections}

\LaTeX\ provides five levels of section headings and they are all
defined in the GJI style file:
\begin{description}
  \item \verb"\section"
  \item \verb"\subsection"
  \item \verb"\subsubsection"
  \item \verb"\paragraph"
  \item \verb"\subparagraph"
\end{description}
Section numbers are given for section, subsection, subsubsection
and paragraph headings.  Section headings are automatically converted to
upper case; if you need any other style, see the example in section~\ref{headings}.

If you find your section/subsection (etc.)\ headings are wrapping round,
you must use the \verb"\\*" to end individual lines and include the
optional argument \verb"[]" in the section command. This ensures that
the turnover is flushleft.

\subsection{Illustrations (or figures)}

\begin{figure*}
 \vspace{5.5cm}
     \caption{An example figure spanning two-columns
             in which space has been left for the artwork.}
     \label{twocol-figure}
\end{figure*}

The GJI style will cope with positioning of your illustrations and
you should not use the positional qualifiers on the
\verb"figure" environment which would override these decisions. See
`Instructions for Authors' in {\em Geophysical Journal International\/}
for submission of
artwork. Figure captions should be below the figure itself, therefore
the \verb"\caption" command should appear after the figure or space
left for an illustration. For example, Fig.~\ref{sample-figure} is
produced using the following commands:
\begin{verbatim}
\begin{figure}
 \vspace{5.5cm}
 \caption{An example figure in which space has
          been left for the artwork.}
 \label{sample-figure}
\end{figure}
\end{verbatim}

Where a figure needs to span two-columns the \verb"figure*" environment
should be used as in  Fig.~\ref{twocol-figure} using the following commands
\begin{verbatim}
\begin{figure*}
 \vspace{5.5cm}
   \caption{An example figure spanning two-columns
     in which space has been left for the artwork.}
   \label{twocol-figure}
\end{figure*}
\end{verbatim}

\subsection{Tables}

The GJI style will cope with positioning of your tables and you
should not use the positional qualifiers on the
\verb"table" environment which would override these decisions. Table
captions should be at the top, therefore the \verb"\caption" command
should appear before the body of the table.

The \verb"tabular" environment can be used to produce tables with
single horizontal rules, which are allowed, if desired, at the head and
foot only. This environment has been modified for the GJI style in the
following ways:
\begin{enumerate}
  \item additional vertical space is inserted on either side of a rule;
  \item vertical lines are not produced.
\end{enumerate}
Commands to redefine quantities such as \verb"\arraystretch" should be
omitted. For example, Table~\ref{symbols} is produced using the
following commands.
\begin{table}
 \caption{Seismic velocities at major discontinuities.}
 \label{symbols}
 \begin{tabular}{@{}lcccccc}
  Class & depth & radius
        & $\alpha _{-}$ & $\alpha _{+}$
        & $\beta _{-}$ & $\beta _{+}$ \\
  ICB & 5154 & 1217 & 11.091 & 10.258
        & 3.438 &  0. \\
  CMB & 2889 & 3482 & 8.009 & 13.691
        & 0. & 7.301 \\
 \end{tabular}

 \medskip
 The ICB represents the boundary between the inner and outer cores and
the CMB the boundary between the core and the mantle.  Velocities with
subscript $-$ are evaluated just below the discontinuity and
those with subscript $+$ are evaluated just above the discontinuity.
\end{table}
\begin{verbatim}
\begin{table}
 \caption{Seismic velocities at major
          discontinuities.}
 \label{symbols}
 \begin{tabular}{@{}lcccccc}
  Class & depth & radius
        & $\alpha _{-}$ & $\alpha _{+}$
        & $\beta _{-}$ & $\beta _{+}$ \\
  ICB & 5154 & 1217 & 11.091 & 10.258
        & 3.438 &  0. \\
  CMB & 2889 & 3482 & 8.009 & 13.691
        & 0. & 7.301 \\
 \end{tabular}

 \medskip
 The ICB represents the boundary ...
... evaluated just above the discontinuity.

\end{table}
\end{verbatim}

If you have a table that is to extend over two columns, you need to use
\verb"table*" in a minipage environment, i.e., you can say
\begin{verbatim}
\begin{table*}
\begin{minipage}{80mm}
 \caption{Caption which will wrap round to the
          width of the minipage environment.}
 \begin{tabular}{%
      :
 \end{tabular}
\end{minipage}
\end{table*}
\end{verbatim}
The width of the minipage should more or less be the width of your table,
so you can only guess on a value on the first pass. The value will have to
be adjusted when your article is finally typeset, so don't worry
about making it the exact size.

\subsection{Running headlines}

As described above, the title of the article and the author's name (or
authors' names) are used as running headlines at the top of every page.
The headline on right pages can list up to three names; for more than
three use et~al. The \verb"\pagestyle" and \verb"\thispagestyle"
commands should {\em not\/} be used. Similarly, the commands
\verb"\markright" and \verb"\markboth" should not be necessary.

\subsection{Typesetting mathematics}

\subsubsection{Displayed mathematics}

The GJI style will set displayed mathematics flush with the left margin,
provided that you use the \LaTeX\ standard of open and closed square brackets
as delimiters. The equation
\[
 \sum_{i=1}^p \lambda_i =
{\mathrm{trace}}(\mathbf{S})
\]
was typeset in the GJI style using the commands
\begin{verbatim}
\[
 \sum_{i=1}^p \lambda_i =
{\mathrm{trace}}(\mathbf{S})
\]
\end{verbatim}
This correct positioning should be compared with that for
the following centred equation,
$$ \alpha_{j+1} > \bar{\alpha}+ks_{\alpha} $$
which was (wrongly) typeset using double dollars as follows:
\begin{verbatim}
$$ \alpha_{j+1} > \bar{\alpha}+ks_{\alpha} $$
\end{verbatim}
Note that \verb"\mathrm" will produce a roman character
in math mode.

For numbered equations use the \verb"equation" and \verb"eqnarray"
environments which will give the correct positioning.
If equation numbering by section is required the command
\verb"\eqsecnum" should appear after \verb"begin{document}"
at the head of the file.

\subsubsection{Bold math italic}\label{boldmathitalic}

The class file provides a font \verb"\mitbf" defined as:
\begin{verbatim}
\newcommand{\mitbf}[1]{
  \hbox{\mathversion{bold}$#1$}}
\end{verbatim}
Which can be used as follows, to typset the equation
\begin{equation}
  d(\mitbf{{s_{t_u}}}) = \langle [RM(\mitbf{{x_y}}
  + \mitbf{{s_t}}) - RM(\mitbf{{x_y}})]^2 \rangle
\end{equation}
the input should be
\begin{verbatim}
\begin{equation}
  d(\mitbf{{s_{t_u}}}) = \langle [RM(\mitbf{{x_y}}
  + \mitbf{{s_t}}) - RM(\mitbf{{x_y}})]^2 \rangle
\end{equation}
\end{verbatim}

If you are using version 1 of the New Font Selection Scheme, you may
have some messages in your log file that read something like ``Warning:
Font/shape `cmm/b/it' in size~\hbox{$< \!\! 9 \!\! >$} not available
on input line 649. Warning: Using external font `cmmi9' instead on input
line 649.'' If you have such messages, your system will have substituted
math italic characters where you wanted bold math italic ones: you
are advised to upgrade to version 2.

\subsubsection{Bold Greek}\label{boldgreek}

To get bold Greek you use the same
method as for bold math italic. Thus you can input
\begin{verbatim}
\[ \mitbf{{\alpha_{\mu}}} =
\mitbf{\Theta} \alpha. \]
\end{verbatim}
to typeset the equation
\[ \mitbf{{\alpha_{\mu}}} = \mitbf{\Theta} \alpha . \]

\subsection{Points to note in formatting text}\label{formtext}

A number of text characters require special attention
so that \LaTeX\ can properly format a file.

The following characters must be preceded by a
backslash or \LaTeX\ will interpret them as commands:
\begin{quote}
~~~~~~~~~\$~~~\&~~~\%~~~\#~~~\_~~~\{~~~and~~~\}
\end{quote}
must be typed
\begin{center}
\begin{quote}
~~~~~~\verb"\$"~~~\verb"\&"~~~\verb"\%"~~~\verb"\#"
~~~\verb"\_"~~~\verb"\{"~~~and~~~\verb"\}".
\end{quote}
\end{center}

\LaTeX\ interprets all double quotes as closing quotes.
Therefore quotation marks must be typed as pairs of
opening and closing single quotes, for example,
\texttt{ ``quoted text.''}

Note that \LaTeX\ will not recognize greater than or
less than symbols unless they are typed within math
commands (\verb"$>$" or \verb"$<$").

\subsubsection{Special symbols}

The macros for the special symbols in Tables~\ref{mathmode}
and~\ref{anymode}
have been taken from the Springer Verlag `Astronomy and Astrophysics'
design, with their permission. They are directly compatible and use the
same macro names.
These symbols will work in all text sizes, but are only guaranteed to work
in text and displaystyles. Some of the symbols will not get any smaller
when they are used in sub- or superscripts, and will therefore be
displayed at the wrong size. Don't worry about this as the typesetter
will be able to sort this out.
%
\begin{table*}
\begin{minipage}{106mm}
\caption{Special symbols which can only be used in math mode.}
\label{mathmode}
\begin{tabular}{@{}llllll}
Input & Explanation & Output & Input & Explanation & Output\\
\hline
\verb"\la"     & less or approx       & $\la$     &
  \verb"\ga"     & greater or approx    & $\ga$\\[2pt]
\verb"\getsto" & gets over to         & $\getsto$ &
  \verb"\cor"    & corresponds to       & $\cor$\\[2pt]
\verb"\lid"    & less or equal        & $\lid$    &
  \verb"\gid"    & greater or equal     & $\gid$\\[2pt]
\verb"\sol"    & similar over less    & $\sol$    &
  \verb"\sog"    & similar over greater & $\sog$\\[2pt]
\verb"\lse"    & less over simeq      & $\lse$    &
  \verb"\gse"    & greater over simeq   & $\gse$\\[2pt]
\verb"\grole"  & greater over less    & $\grole$  &
  \verb"\leogr"  & less over greater    & $\leogr$\\[2pt]
\verb"\loa"    & less over approx     & $\loa$    &
  \verb"\goa"    & greater over approx  & $\goa$\\
\hline
\end{tabular}
\end{minipage}
\end{table*}
%
\begin{table*}
\begin{minipage}{115mm}
\caption{Special symbols which don't have to be
used in math mode.}
\label{anymode}
\begin{tabular}{@{}llllll}
Input & Explanation & Output & Input & Explanation & Output\\
\hline
\verb"\sun"      & sun symbol            & $\sun$     &
  \verb"\earth"     & earth symbol         & $\earth$   \\[2pt]
\verb"\degr"     & degree                &$\degr$     &
  \verb"\micron"   & \micron               & \micron    \\[2pt]
\verb"\diameter" & diameter              & \diameter  &
  \verb"\sq"       & square                & \squareforqed\\[2pt]
\verb"\fd"       & fraction of day       & \fd        &
  \verb"\fh"       & fraction of hour      & \fh\\[2pt]
\verb"\fm"       & fraction of minute    & \fm        &
  \verb"\fs"       & fraction of second    & \fs\\[2pt]
\verb"\fdg"      & fraction of degree    & \fdg       &
  \verb"\fp"       & fraction of period    & \fp\\[2pt]
\verb"\farcs"    & fraction of arcsecond & \farcs     &
  \verb"\farcm"    & fraction of arcmin    & \farcm\\[2pt]
\verb"\arcsec"   & arcsecond             & \arcsec    &
  \verb"\arcmin"   & arcminute             & \arcmin\\

\hline
\end{tabular}
\end{minipage}
\end{table*}

The command \verb"\chemical" is provided to set chemical species with
an even level for subscripts (not produced in standard mathematics mode).
Thus \verb"\chemical{Fe_{2}^{2+}Cr_{2}O_{4}}" will produce
\chemical{Fe_{2}^{2+}Cr_{2}O_{4}}.

\subsection{Bibliography}

Two methods are provided for managing citations and
references.   The first approach uses the
\verb"\begin{thebibliography}{}"
and \verb"\end{thebibliography}{}" commands.

The second approach uses a simplified scheme using
\verb"\begin{references}" and \verb"\end{references}" commands.

References to published literature should be quoted in text by author
and date; e.g. Draine (1978) or (Begelman, Blandford \& Rees 1984).
Where more than one reference is cited having the same author(s) and date,
the letters a,b,c, \ldots\ should follow the date; e.g.\ Smith (1988a),
Smith (1988b), etc.
The first time you introduce a three-author paper, you should list all
three authors at the first citation, and thereafter, use et al.

\subsubsection{Biblography References in the text}

References in the text are given by author and date, and, whichever
method is used to produce the bibliography, the references in the text
are done in the same way. Each bibliographical entry has a key, which
is assigned by the author and used to refer to that entry in the text.
There is one form of citation -- \verb"\cite{key}" -- to produce the
author and date, and another form -- \verb"\shortcite{key}" -- which
produces the date only. Thus, Rutherford \& Hawker \shortcite{rh} is
produced by
\begin{verbatim}
Rutherford \& Hawker \shortcite{rh}
\end{verbatim}
while \cite{hi} is produced by
\begin{verbatim}
\cite{hi}
\end{verbatim}

\subsubsection{The bibliography}

The following listing shows some references prepared in the style of
the journal; the code produces the references at the end of this guide.
The following rules apply for the ordering of your references:
\begin{enumerate}
  \item if an author has written several papers, some with other authors,
        the rule is that the single-author papers precede the two-author
        papers, which, in turn, precede the multi-author papers;
  \item within the two-author paper citations, the order is determined
        by the second author's surname, regardless of date;
  \item within the multi-author paper citiations, the order is
        chronological, regardless of author's surnames.
\end{enumerate}
%
\begin{verbatim}
\begin{thebibliography}{}
  \bibitem[\protect\citename{Butcher }1992]{bu}
    Butcher J. 1992. \textit{Copy-editing: The
    Cambridge Handbook}, 3rd edn, Cambridge
    Univ. Press, Cambridge.
  \bibitem[\protect\citename{The Chicago Manual }%
    1982]{cm} \textit{The Chicago Manual of Style},
    Univ. Chicago Press, Chicago, 1982.
  \bibitem[\protect\citename{Chao }1985]{ch}
    Chao, B. F., 1985. Normal mode study of the
    Earth's rigid body motions,
    \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
  \bibitem[\protect\citename{Hinderer }1986]{hi}
    Hinderer, J., 1986. Resonance effects of the
    earth's fluid core in earth rotation,
    in \textit{Solved and Unsolved Problems},
    pp. 277-296, ed. Cazenave A., Reidel,
    Dordrecht.
  \bibitem[\protect\citename{Lamport }1986]{la}
    Lamport L., 1986,  \LaTeX: \textit{A Document
    Preparation System}, Addison--Wesley, New York
  \bibitem[\protect\citename{Lindberg }1986]{li}
    Lindberg, C., 1986.  Multiple taper harmonic
    analysis of terrestrial free oscillations,
    \textit{PhD thesis}, University of California.
  \bibitem[\protect\citename{Maupin }1992]{ma}
    Maupin, V., 1992. Modelling of laterally
    trapped surface waves with application to
    Rayleigh waves in the Hawaiian swell,
    \textit{Geophys. J. Int.}, \textbf{110}, 553-570.
  \bibitem[\protect\citename{Rutherford
    \& Hawker }1981]{rh} Rutherford, S. R.
    \& Hawker, K. E.,  1981, Consistent coupled
    mode theory of sound propagation for a
    class of non-separable problems,
    \textit{J. acoust. Soc. Am.}, \textbf{71},
    554-564
\end{thebibliography}
\end{verbatim}
Each entry takes the form
\begin{verbatim}
\bibitem[\protect\citename{Author(s), }%
  Date]{tag} Bibliography entry
\end{verbatim}
where \verb"Author(s)" should be the author names as they are cited in
the text, \verb"Date" is the date to be cited in the text, and
\verb"tag" is the tag that is to be used as an argument for the
\verb"\cite{}" and \verb"\shortcite{}" commands. \verb"Bibliography entry"
should be the material that is to appear in the bibliography,
suitably formatted.

\subsubsection{Simplified References and Citations}

The second approach to referencing is taken with permission from
the American Geophysical Union Latex macros

The reference section is started using a
\verb"\begin{references}" command which will
automatically produce a correctly formatted
``Reference'' head.  Each reference is then
preceded by a \verb"\reference"
command.  It is the author's
responsibility to place bibliographic reference
information in the proper order with correct
punctuation.  After the last reference in your
reference section, type an \verb"\end{references}"
command.

Authors may enter properly formatted citations directly
in the manuscript text and enclose those citations in
\verb"\markcite{}" commands.  This approach
marks all citations in your manuscript, but there
is no interaction between the \verb"\markcite"
commands and the reference section.

To create in-text citations, enclose each citation
within a \verb"\markcite" command.
There are two ways to include in-text citations,
depending on the way you phrase your sentence.
You may either include an entire reference within
brackets \markcite{(Merritt et al., 1996)} or you
may mention the author as part of your sentence and
include only the year in brackets, as in \markcite{Ono (1996)}.

As an example
\begin{verbatim}
\begin{references}
\reference
Azimi, Sh.A., Kalinin, A.Y., Kalinin, V.B.,
\& Pivovarov, B.L., 1968.
Impulse and transient characteristics of media
with linear and quadratic absorption laws,
\textit{Izv. Earth Phys.} (English Transl.),
\textbf{2}, 88--93.
\reference
Dahlen, F.A., \& Smith, M.L., 1975.
The influence of rotation on the free
oscillations of the Earth,
\textit{Phil. Trans. R. Soc. London Ser. A},
\textbf{279}, 143--167.
\reference
Durek, J.J., Ritzwoller, M.H.,
\& Woodhouse, J.H., 1993.
Constraining upper mantle anelasticity
using surface wave amplitude anomalies,
\gji, \textbf{114}, 249--272.
\end{references}
\end{verbatim}
produces the reference list
\begin{references}
 \reference
Azimi, Sh.A., Kalinin, A.Y., Kalinin, V.B.,
\& Pivovarov, B.L., 1968.
Impulse and transient characteristics of media
with linear and  quadratic absorption laws,
\textit{Izv. Earth Phys.} (English Transl.),
\textbf{2}, 88--93.
\reference
Dahlen, F.A., \& Smith, M.L., 1975.
The influence of rotation on the free oscillations of the Earth,
\textit{Phil. Trans. R. Soc. London Ser. A}, \textbf{279}, 143--167.
\reference
Durek, J.J., Ritzwoller, M.H., \& Woodhouse, J.H., 1993.
Constraining upper mantle anelasticity
using surface wave amplitude anomalies,
\gji \textbf{114}, 249--272.
\end{references}

\subsubsection{Common Journals}

The following abbreviations are provided for commonly cited
journals and can be used directly in the bibliography.

In the following table the abbreviation and the form of the
associated entry are presented
\newline
\begin{tabular}{ll}
\verb"\areps" & \areps \\
\verb"\bssa"  & \bssa \\
\verb"\eos"   & \eos  \\
\verb"\eps"   & \eps \\
\verb"\epsl"  & \epsl \\
\verb"\gca"   & \gca \\
\verb"\geo"   & \geo \\
\verb"\geop"  & \geop \\
\verb"\gji"   & \gji \\
\verb"\gjras" & \gjras \\
\verb"\grl"   & \grl \\
\verb"\gsab"  & \gsab \\
\verb"\gs"    & \gs \\
\verb"\jgr"   & \jgr \\
\verb"\jseis" & \jseis \\
\verb"\mnras" & \mnras \\
\verb"\pag"   & \pag \\
\verb"\pepi"  & \pepi \\
\verb"\rg"    & \rg \\
\verb"\tecto" & \tecto \\
\end{tabular}
%
% The following two tables have been moved back in the text to
% improve page layout
%
\begin{table*}
\begin{minipage}{130mm}
\caption{Authors' notes.}
\label{authors}
\begin{tabular}{@{}ll}
\verb"\title[optional short title]{long title}"
                    & short title used in running head\\
\verb"\author[optional short author(s)]{long author(s)}"
                    & short author(s) used in running head\\
\verb"\begin{abstract}...\end{abstract}"& for summary on
titlepage\\
\verb"\begin{summary}...\end{summary}"& for abstract on
titlepage\\
\verb"\begin{keywords}...\end{keywords}"& for keywords on titlepage\\
\verb"\nokeywords"  & if there are no keywords on titlepage\\
\verb"\begin{figure*}...\end{figure*}" & for a double spanning figure in two-column mode\\
\verb"\begin{table*}...\end{table*}" & for a double spanning table in
                                       two-column mode\\
\verb"\caption*"    & for continuation figure captions\\
\verb"\resetfigno" & resets figures numbers after an appendix\\
\verb"[referee]" & documentclass option for 12/20pt, single col,
                   for manuscript submission\\
\verb"[mreferee]" & documentclass option for 11/17pt, single col,
                   for submission of papers with extensive mathematics\\
\end{tabular}
\end{minipage}
\end{table*}
%
\begin{table*}
\begin{minipage}{130mm}
\caption{Editors' notes.}
\label{editors}
\begin{tabular}{@{}lp{270pt}}
\verb"\pagerange{000--000}"& for catchline, note use of en-rule\\
\verb"\pagerange{L00--L00}"& for letters option, used in catchline\\
\verb"\volume{000}" & volume number, for catchline\\
\verb"\pubyear{0000}" & publication year, for catchline\\
\verb"\microfiche{GJI000/0}" & for articles accompanied by microfiche\\
\verb"\journal" & replace the whole catchline at one go\\
\verb"[doublespacing]" & documentclass option for doublespacing\\
\verb"[galley]" & documentclass option for running to galley\\
\verb"[landscape]" & documentclass option for landscape illustrations\\
\verb"[fasttrack]" & documentclass option, for rapid short communications
                   (adds F to folios)\\
\verb"[onecolumn]" & documentclass option for one-column \\
\verb"\bsp"& typesets the final phrase `This paper has been produced
 using the Blackwell Publishing GJI \LaTeX2e\ class file.'\\
\end{tabular}
\end{minipage}
\end{table*}

\subsection{Appendices}

The appendices in this guide were generated by typing:
\begin{verbatim}
\appendix
\section{For authors}
     :
\section{For editors}
\end{verbatim}
You only need to type \verb"\appendix" once. Thereafter, every
\verb"\section" command will generate a new appendix which will be
numbered A, B, etc.

If figure captions are to provided after an appendix the figure number can
be reset to avoid extraneous labelling using the command
\verb"\resetfigno".

\section[]{Example of section heading with\\*
  {\mdseries \textsc{S}\lowercase{\textsc{mall}}
  \textsc{C}\lowercase{\textsc{aps}}},
  \lowercase{lowercase},
  \textit{ italic}, and bold\\* Greek such as
  $\mitbf{{\mu^\kappa}}$}\label{headings}

This can be built up using text commands and the \verb"mitbf"
command introduced above

\begin{verbatim}
\section[]{Example of section heading with\\*
  {\mdseries \textsc{S}\lowercase{\textsc{mall}}
  \textsc{C}\lowercase{\textsc{aps}}},
  \lowercase{lowercase},
  \textit{ italic}, and bold\\* Greek such as
  $\mitbf{{\mu^\kappa}}$}\label{headings}
\end{verbatim}

\subsection{Acknowledgments}
Acknowledgments after the main text and before the appendices can be
included with the
\texttt{acknowledgments} environment, as
\begin{verbatim}
\begin{acknowledgments}
We wish to thank ...
\end{acknowledgments}
\end{verbatim}
There is also a corresponding \texttt{acknowledgment} environment for a
single acknowledgment.

\begin{acknowledgments}
A number of colleagues have helped with suggestions for the
improvement of this material and I would particularly like to thank
Bob Geller, University of Tokyo for his criticisms and corrections.
\end{acknowledgments}

\begin{thebibliography}{}
  \bibitem[\protect\citename{Butcher }1992]{bu}
    Butcher J. 1992. \textit{Copy-editing: The Cambridge
    Handbook}, 3rd edn, Cambridge Univ. Press, Cambridge.
  \bibitem[\protect\citename{The Chicago Manual }%
    1982]{cm} \textit{The Chicago Manual of Style}, Univ.
    Chicago Press, Chicago, 1982.
  \bibitem[\protect\citename{Chao }1985]{ch}
    Chao, B. F., 1985. Normal mode study of the Earth's rigid
     body motions, \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
  \bibitem[\protect\citename{Hinderer }1986]{hi}
    Hinderer, J., 1986. Resonance effects of the earth's fluid
    core in earth rotation, in \textit{Solved and Unsolved
    Problems}, pp. 277-296, ed. Cazenave A., Reidel,
    Dordrecht.
  \bibitem[\protect\citename{Kopka \& Daly}1995]{kd}
    Kopka H. \& Daly P.W., 1995, \textit{A guide to} \LaTeX2e,
    Addison--Wesley, New York
  \bibitem[\protect\citename{Lamport }1986]{la}
    Lamport L., 1986,  \LaTeX: \textit{A Document
    Preparation System}, Addison--Wesley, New York
  \bibitem[\protect\citename{Lindberg }1986]{bl}
    Lindberg, C., 1986.  Multiple taper harmonic analysis of
    terrestrial free oscillations, \textit{PhD thesis},
    University of California.
  \bibitem[\protect\citename{Maupin }1992]{ma}
    Maupin, V., 1992. Modelling of laterally trapped surface
    waves with application to Rayleigh waves in the Hawaiian
    swell, \textit{Geophys. J. Int.}. \textbf{110}, 553-570.
  \bibitem[\protect\citename{Rutherford \& Hawker }%
    1981]{rh}  Rutherford, S. R. \& Hawker, K. E., 1981,
    Consistent coupled mode theory of sound propagation for a
    class of non-separable problems,
   \textit{J. acoust. Soc. Am.}, \textbf{71}, 554-564
\end{thebibliography}

\appendix
\section{For authors}

Table~\ref{authors} is a list of design macros which are unique to GJI. The
list displays each macro's name and description.

\section{For editors}

The additional features shown in Table~\ref{editors} may be used for
production purposes.

\bsp % ``This paper has been produced using the Blackwell
     %   Publishing GJI \LaTeXe\ class file.''

\label{lastpage}

\end{document}

### 2015-01-14 manuscript.tex
gji + plain subfigure + partial translation + nadaraya ref correction

@@
 
-\documentclass[draft,grl]{agutex}
+\documentclass[draft, grl]{agutex}
 
@@
 
-\usepackage{lineno}
-\linenumbers*[1]
+% \usepackage{lineno}
+% \linenumbers*[1]
 
@@
 % See how to enter figures and tables at the end of the article, after references.
+\usepackage[utf8]{inputenc}
 \usepackage{float}
 %\usepackage{caption}
-\usepackage{subcaption}
-
-%----------------------------------------------------------------------------------------
+%\usepackage{subcaption}
+
+% \usepackage[round,sort,nonamebreak]{natbib} % citação bibliográfica textual(plainnat-ime.bst)
+% \bibpunct{(}{)}{;}{a}{\hspace{-0.7ex},}{,} % estilo de citação. Veja alguns exemplos em http://merkel.zoneo.net/Latex/natbib.php
+
+\usepackage{amsfonts}
+\usepackage{textcomp}
+
+%-----------------------------------------------------------------------------
+%	colored references
+%-----------------------------------------------------------------------------
+% \usepackage[usenames,svgnames,dvipsnames]{xcolor}
+\usepackage[a4paper,top=2.5cm,bottom=3.0cm,left=2.0cm,right=2.0cm]{geometry}
+% \usepackage[colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=Green,bookmarksopen=true]{hyperref}
+% \usepackage[all]{hypcap}      % soluciona o problema com o hyperref
+%-----------------------------------------------------------------------------
 %	Glossaries
-%----------------------------------------------------------------------------------------
+%-----------------------------------------------------------------------------
 
@@
 \altaffiltext{1}{Department of Geophysics,
-Institute of Astronomy, Geophysics and Atmospherical Sciences,
-University of São Paulo, São Paulo, Brazil.}
-
-\altaffiltext{2}{School of Applied Math, Getúlio Vargas Foundation, Rio de Janeiro, Brazil.}
+Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG),
+University of São Paulo (USP), São Paulo, Brazil.}
+
+\altaffiltext{2}{Applied Math School (EMAp), Getúlio Vargas Foundation (FGV), Rio de Janeiro, Brazil.}
 
@@
 
-A proposta de Arthur \citet{frankel_1995} foi usar uma largura de banda fixa,
-nomeada distância de correlação $d_F$, e aplicar o estimador de Nadaraya-Watson (\ref{sec:nadaraya})
-para suavizar o histograma 2D da sismicidade utilizando uma função de núcleo gaussiana:
+Arthur \citet{frankel_1995} smoothing seismicity proposal consists originally in use a called correlation distance $d_F$ as the kernel \textit{fixed} bandwidth and next to apply the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
 \begin{equation}
@@
 \end{equation}
-onde $\tilde{n}_j$ é a taxa de sismicidade (número de sismos com magnitude $m$ maior que a mínima magnitude
-\gls{sym:Md} do catálogo) suavizada
-na célula $j$, $n_i$ é o número de sismos em cada outra célula $i$ e
-	\gls{sym:dij} é \glsdesc{sym:dij}.
+where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$.
+$n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.
 
@@
 
-Já Gordon \citet{woo_1996} propôs avaliar a contribuição de uma função de núcleo
-de cada sismo $i$ ocorrido em $\boldsymbol{r}_i$
-na célula centrada em $\boldsymbol{r}$ que dependa também de sua
-magnitude $m$:
+Gordon \citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, far $\boldsymbol{r}_i$ from its kernel, on the cell centered on $\boldsymbol{r}$ in dependence of their magnitude $m$:
 \begin{equation}
@@
 \end{equation}
-onde $N$ é o número de tremores $i$ no catálogo
-e $T(\gls{sym:ri})$ é o período em que todo sismo de magnitude acima de $m$ é completamente observado
-em \gls{sym:ri}.
+where $N$ is the number of earthquakes $i$ on the catalogue
+and $T(\gls{sym:ri})$ is the timeframe which all earthquakes with magnitude above $m$ is completely observed on \gls{sym:ri}.
+
+Any kernel could be applied on that definition. In practice Woo used a \citet{kagan_knopoff_1980} for a infinity spatial domain:
+\begin{equation}
+	\ensuremath{
+		K_{KJ}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
+							\left[ 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right]^{-\gls{sym:aW}},
+	}
+	\label{eq:k_kj}
+\end{equation}
+where \gls{sym:aW} is, accordingly \citet{verejones_1992}, \glsdesc{sym:aW}.
+
+To compute the magnitude-dependent bandwidth function \gls{sym:hm}, Gordon Woo suggested the follow relation
+\begin{equation}
+	\ensuremath{
+		h(m\arrowvert \gls{sym:a0}, \gls{sym:a1}) = \gls{sym:a0}e^{\gls{sym:a1}m},
+	}
+	\label{eq:hm}
+\end{equation}
+where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$.
 
@@
 
-\subsubsection{Bandwidth computation}
-
+\subsubsection{Local bandwidth computation}
+
+The method implemented to compute the space and time bandwidths for each earthquake was the Couppled-Nearest-Neighbour \citep{helmstetter_2012}:
 \begin{equation}
 	\ensuremath{
-%		h_i, d_i = \underset{d_i \ge \gls{sym:dk}, h_i \ge \gls{sym:hk}}{\argmin}
 		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
@@
 \end{equation}
-onde \gls{sym:k_cnn} é o \glsdesc{sym:k_cnn},
-	 \gls{sym:a_cnn} é o \glsdesc{sym:a_cnn},
-	 \gls{sym:dk} é o \glsdesc{sym:dk} e
-	 \gls{sym:hk} é o \glsdesc{sym:hk}.
-
-As larguras de bandas são escolhidas localmente, podendo ser menor, aumentando a resolução
-onde há maior densidade de informação, ou crescer onde há escassez de informação e menor resolução.
+where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
+	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
+	 \gls{sym:dk} is the \glsdesc{sym:dk} e
+	 \gls{sym:hk} is the \glsdesc{sym:hk}.
+
+The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.
 
@@
 \begin{acknowledgments}
-This work was partially supported by a grant from the Spanish Ministry of Science and Technology.
+The author thanks specially to Marcelo Assumpção for the support and critical discussion as well all IAG-USP seismological team.
+It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
 \end{acknowledgments}
@@
 
-\begin{thebibliography}{}
-
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-
-\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
-Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
-  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
-  Comp.}, \textit{56}(193), 119--139.
-
-\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
-Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
-  Scattering Theory}, John Wiley, New York.
-
-\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
-  Wendland}}]{StephanHsiao}
-Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
-  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
-  Appl. Math.}, \textit{34}(1), 1--19.
-
-\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
-Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
-  components and line integrals of currents in modified edge representation,
-  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.
-
-\end{thebibliography}
+% ---------------------------------------------------------------------------- %
+% Bibliografia
+\bibliographystyle{BibTeX/agufull08} 	% citação bibliográfica textual
+\bibliography{bib/bibliografia}  		% associado ao arquivo:
+
+% \begin{thebibliography}{}
+
+% \providecommand{\natexlab}[1]{#1}
+% \expandafter\ifx\csname urlstyle\endcsname\relax
+%   \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+%   \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+%   \urlstyle{rm}\Url}\fi
+
+% \bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
+% Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
+%   logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
+%   Comp.}, \textit{56}(193), 119--139.
+
+% \bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
+% Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
+%   Scattering Theory}, John Wiley, New York.
+
+% \bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
+%   Wendland}}]{StephanHsiao}
+% Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
+%   problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
+%   Appl. Math.}, \textit{34}(1), 1--19.
+
+% \bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
+% Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
+%   components and line integrals of currents in modified edge representation,
+%   \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.
+
+% \end{thebibliography}
 
@@
 \begin{figure}
-	\centering
-	\begin{subfigure}[b]{0.3\textwidth}
-		\includegraphics[width=0.4\textwidth]{placeholder}
-		\caption{caption A}
-		\label{placeholder_a}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.3\textwidth}
-		\includegraphics[width=0.4\textwidth]{placeholder}
-		\caption{caption B}
-		\label{placeholder_b}
-	\end{subfigure}
-	\caption{main caption}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\footnotesize(a)
+		& 
+		\footnotesize(b)
+		\\		
+		\includegraphics[width=0.45\textwidth]{placeholder} &
+		\includegraphics[width=0.45\textwidth]{placeholder}
+		\end{tabular}
+	\end{table}
+	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
 	\label{placeholder}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.33\textwidth} 
+						 p{0.33\textwidth} 
+						 p{0.33\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\				
+		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
+		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
+		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
+	\label{placeholder2}
+	\end{center}
 \end{figure}

### 2015-01-14 manuscript.tex
methodology

@@
 \documentclass[draft, grl]{agutex}
+%\documentclass[grl]{agutex}
 
@@
 
-% \usepackage{lineno}
-% \linenumbers*[1]
+\usepackage{lineno}
+\linenumbers*[1]
 
@@
 % Glossaries
-\usepackage[sanitize=none,acronym,toc]{glossaries}
+%\usepackage[sanitize=none,acronym,toc]{glossaries}
+\usepackage[acronym,toc]{glossaries}
 %\usepackage[acronym,toc]{glossaries}
@@
 
-Intraplate seismicity is a hard problem to solve in tectonics,
-mainly during the lack of observed seismicity in many areas where seismic
-hazard assesment are also difficult and the methods applied are often
-close correlated to the geometries defined by the experts with valuable
-criteria.
-
-Three smoothing seismicity techniques are reviewed in order
-to investigate their results when applied over the Brazilian seismicity
-and suggesting a seismic hazard assesment.
-
-All this methods are based on a kernel density estimation. The main
-difference between them is the bandwidth selection process.
-
-Frankel in their method uses a \emph{fixed} bandwidth in space. Woo's method
-suggest a bandwidth as function of the \emph{magnitude}.
-Helmstetter and Werner proposes a \emph{local-adaptive}
-in space and time kernel bandwidths, chosen for each event.
-This last one was proposed initialy to take care of the background seismicity rate
-on long-term earthquake forecast context and the time-independent assumption
-about the annual background seismic rate could be used
+Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assesment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+
+Three smoothing seismicity techniques are reviewed in order to investigate their results when applied over the Brazilian seismicity and suggesting a seismic hazard assesment.
+
+All this methods are based on a kernel density estimation. The main difference between them is the bandwidth selection process.
+
+Frankel in their method uses a \emph{fixed} bandwidth in space.
+Woo's method suggest a bandwidth as function of the \emph{magnitude}.
+Helmstetter and Werner proposes a \emph{local-adaptive} in space and time kernel bandwidths, chosen for each event.
+This last one was proposed initialy to take care of the background seismicity rateon long-term earthquake forecast context and the time-independent assumption about the annual background seismic rate could be used
 (not indiscriminately REFERENCE) as a grid of points seismic sources.
 
-As conclusions the results shown that all methods gives defines better
-some known seismic regions and could be used in the Brazilian seismicity case,
-all of them, improoving the spatial resolution of the GSHAP model, the last one
-made on the region.
-
-The Last two methods was implemented and are available as a free software on
-the Hazard Modeling Toolkit (HMTK) from GEM Foundation.
+As conclusions the results shown that all methods gives defines better some known seismic regions and could be used in the Brazilian seismicity case, all of them, improoving the spatial resolution of the GSHAP model, the last one made on the region.
+
+The Last two methods was implemented and are available as a free software on the Hazard Modeling Toolkit (HMTK) from GEM Foundation.
 
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 %	ARTICLE CONTENT
-%----------------------------------------------------------------------------------------
-
+%------------------------------------------------------------------------------
 % The body of the article must start with a \begin{article} command
@@
 
-Brazil have no recent seismic hazard study at national scale despite some important and localized recent studies
-focused mainly on specific buildings.
-
-The Brazilian seismic hazard building code was made based on GSHAP model which is the well accepted state of art model
-for the whole region.
-
-A significant number of earthquakes has been recorded in the last years, mainly after the national seismic network
-(RSBR) operation, and an actualized hazard map whould be computed.
-
-Although the standard Cornell Mcguire zoning methodolgy for the classical hazard seismic assesment, the goal is
-review a few smoothing seismicity methods based on kernel density estimation. This estimation allow a seismic point
-sources grid definition. The seismic rate (a-value) associated to each source is definided by the smoothed seismicity
-methods, but the more parameters like b-values, minimum and maximum magnitudes, rupture and depth
-distributions must to be assigned in other ways.
-
-In the kernel density estimation theory two things there always be handled. The kernel shape and bandwidth.
-Almost kernels used are gaussian or power-laws in one of their variants. The mainly distinction between methods decribed
-is the proposed bandwidth selection process.
-
-Frankel smoothing method applies a gaussian Nadaraya-Watson smoothing method with a \emph{fixed}
-smoothing-distance bandwidth to estimate seismicity rate. Gordon Woo proposes an \emph{magnitude-dependent} kernel
-bandwidth based on the nearest neighbour mean distance in magnitude bins. Helmstetter proposes a \emph{local-adaptive}
-bandwidth, based on kernel estimations on both space and time dimentions, estimation of the background seismicity rate
-for long-term forecasts. Using one-year forecast and assume time-independence, the same assumption of PSHA, of
-seismicity rate is possible to use these values on seismic sources characterization.
-
-To perform the hazard calculation, the OpenQuake suite was choosed. The modeller environment used was the HMTK also
-available by GEM scientifical board. The Frankel method was already implemented on the toolkit and the other two was
-implemented in the context of this paper and they are free available on a public repository.
+Brazil have no recent seismic hazard study at national scale despite some important and localized recent studies focused mainly on specific buildings.
+
+The Brazilian seismic hazard building code was made based on GSHAP model which is the well accepted state of art model for the whole region.
+
+A significant number of earthquakes has been recorded in the last years, mainly after the national seismic network (RSBR) operation, and an actualized hazard map whould be computed.
+
+Although the standard Cornell Mcguire zoning methodolgy for the classical hazard seismic assesment, the goal is review a few smoothing seismicity methods based on kernel density estimation.
+This estimation allow a seismic point sources grid definition.
+The seismic rate (\emph{a}-value) associated to each source is definided by the smoothed seismicity methods, but the more parameters like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions must to be assigned in other ways.
+
+In the kernel density estimation theory two things there always be handled.
+The kernel shape and bandwidth.
+Almost kernels used are gaussian or power-laws in one of their variants.
+The mainly distinction between methods decribed is the proposed bandwidth selection process.
+
+Frankel smoothing method applies a gaussian Nadaraya-Watson smoothing method with a \emph{fixed} smoothing-distance bandwidth to estimate seismicity rate.
+Gordon Woo proposes an \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbour mean distance in magnitude bins.
+Helmstetter proposes a \emph{local-adaptive} bandwidth, based on kernel estimations on both space and time dimentions, estimation of the background seismicity rate for long-term forecasts.
+Using one-year forecast and assume time-independence, the same assumption of PSHA, of seismicity rate is possible to use these values on seismic sources characterization.
+
+To perform the hazard calculation, the OpenQuake suite was choosed.
+The modeller environment used was the HMTK also available by GEM scientifical board.
+The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.
 
@@
 
-The Brazilian seismicity data is presented including some overview plots. Next the method's theory is presented follow
-by the specific decisions and optimizations performed on each method modelling. And last, the results, conclusions and
-further considerations are discussed.
+The Brazilian seismicity data is presented including some overview plots.
+Next the method's theory is presented follow by the specific decisions and optimizations performed on each method modelling.
+And last, the results, conclusions and further considerations are discussed.
 
@@
 %------------------------------------------------
-\section{Methods}
+\section{Smoothing methods}
+Under this work, three smoothing methods was considered. On this section brief explanation of each one will be done.
 
@@
 \end{equation}
-where $N$ is the number of earthquakes $i$ on the catalogue
+where $N$ is the number of earthquakes $i$ on the catalog
 and $T(\gls{sym:ri})$ is the timeframe which all earthquakes with magnitude above $m$ is completely observed on \gls{sym:ri}.
@@
 	\ensuremath{
-		K_{KJ}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
+		K_{KK}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
 							\left[ 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right]^{-\gls{sym:aW}},
@@
 
-\subsection{Helmstetter}
-
+\subsection{Helmstetter and Werner}
+
+Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to charachterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.
+
+\citet{helmstetter_2012} proposes a seismicity model space and time dependent using kernels for space and time independently:
 \begin{equation}
@@
 \end{equation}
-onde \gls{sym:R} é \glsdesc{sym:R},
-	  $K_t$ é a \glsdesc{sym:Kt},
-	  $K_r$ é a \glsdesc{sym:Kr}.
-
-Como seu interesse é em projeção, restringiram os tempos do catálogo aos $t_i
-< t$ e ponderaram segundo um peso $w$ da seguinte maneira:
+where \gls{sym:R} is the \glsdesc{sym:R},
+	  $K_t$ is the \glsdesc{sym:Kt},
+	  $K_r$ is the \glsdesc{sym:Kr}.
+
+As they are interested in forecast, just past time $t_i < t$ need to be considered. Also the observation completeness is taken in account by a set of weights $w$ in this follow way:
 \begin{equation}
@@
 \end{equation}
-onde \gls{sym:Rmin} é a \glsdesc{sym:Rmin}, positiva, permitindo que ocorram eventos-surpresa
-onde não se tem registro de sismicidade.
-
-Os pesos $w(\boldsymbol{r}_i,t_i)$, calculados em cada tremor $i$, propostos por Helmstetter são a transformação de
-Gutemberg-Richter. Consideraram uma contribuição maior para a taxa de sismicidade daqueles sismos
-que ocorreram onde e quando a magnitude de completude $M_c$ era maior que
-a mínima magnitude $M_d$ do catálogo
-em locais onde a magnitude de completude e o \emph{valor-b} variam.
-Note que o modelo de pesos acomoda bem variações espaciais e
-temporais tanto da magnitude de completude como do \emph{valor-b}. A expressão é a seguinte:
+where \gls{sym:Rmin} is the \glsdesc{sym:Rmin}, positive, allowing earthquakes to occur where their never occurred yet.
+
+The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection.
+These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$.
+These weights could easily consider space and time completenes and \emph{b}-value fluctuations.
+The expression used to compute it:
 \begin{equation}
@@
 \end{equation}
-onde  \gls{sym:wi} é o \glsdesc{sym:wi} na localização $\boldsymbol{r}$ e no instante $t$,
-	  \gls{sym:b}$(\boldsymbol{r},t)$ é o \glsdesc{sym:b},
-	  \gls{sym:Mc_rt} é a \glsdesc{sym:Mc_rt},
-	  \gls{sym:Md} é a \glsdesc{sym:Md}.
+where \gls{sym:wi} is the \glsdesc{sym:wi} at location $\boldsymbol{r}$ on instant $t$,
+	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
+	  \gls{sym:Mc_rt} is the \glsdesc{sym:Mc_rt},
+	  \gls{sym:Md} is the \glsdesc{sym:Md}.
 
@@
 
-\subsubsection{Taxa de sismicidade estacionária}
-Mesmo havendo modelado a dependência com o tempo para distribuição da taxa de sismicidade,
-a taxa estacionária $\bar{R}$ que se preserva no longo-prazo em uma determinada localização
-$\boldsymbol{r}_0$ é a que mais interessa à \gls{psha} e,
-segundo o método proposto por \citet{helmstetter_2012}, será definida
-pelo valor da mediana da taxa fornecida pelo modelo no
-período considerado:
-
+\subsubsection{Stationary seismic rate}
+After compute the model parameters and completely define it, the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ have to be computed. 
+The way to do it proposed by \citet{helmstetter_2012} was to get the median
+on $\boldsymbol{r}_0$ over all considered time window:
 \begin{equation}
 	\ensuremath{
-		\bar{R}(\boldsymbol{r}_0) = \text{Mediana}\left[R(\boldsymbol{r}_0, t)\right].
+		\bar{R}(\boldsymbol{r}_0) = \text{Median}\left[R(\boldsymbol{r}_0, t)\right].
 	}
@@
 
-O fato de se tomar a mediana faz com que variações na taxa de sismicidade devido a pré e pós-eventos
-não interfiram significativamente no valor da taxa estacionária. A decorrência mais importante desse fato
-é que o método dispensa a remoção de agrupamentos (seção \ref{sec:declustering}) do catálogo, um método paramétrico e
-quase sempre controverso.
-
-\subsubsection{Verossimilhança}
-
-Para otimizar os parâmetros do modelo de Helmstetter, é preciso dividir o catálogo de sismos.
-Uma parte serve para o aprendizado do modelo e a outra para avaliação.
-
-Se a ocorrência de tremores pode ser modelada por um processo de Poisson com taxa $N_p$,
-então a probabilidade de se observar exatamente $n$ eventos no período de tempo considerado é dada por
-
+The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures.
+
+\subsubsection{Maximum likelyhood optimization}
+
+The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}.
+To optimize these parameters the catalog is divided in two parts: one for learning about the parameters and other to test them.
+
+If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the consider time frame is
 \begin{equation}
@@
 
-E o logarítmo, a ser maximizado, da verossimilhança entre
-o que o modelo predisse e o que foi observado
-pode ser escrito como
-
+Then, over all cells, the (log) likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed on testing catalog is written as
 \begin{equation}
@@
 \end{equation}
-onde \gls{sym:Np} é \glsdesc{sym:Np},
-	\gls{sym:nxy} é \glsdesc{sym:nxy}.
-
-Os parâmetros $R_{min}$, $a_{cnn}$ e $k_{cnn}$ do modelo são otimizados
-pela maximização da verossimilhança $L$.
+where \gls{sym:Np} is the \glsdesc{sym:Np} and
+	  \gls{sym:nxy} is the \glsdesc{sym:nxy}.
+
+The model parameters $R_{min}$, $a_{cnn}$ and $k_{cnn}$ should be optimized by the maximization of log-likelihood $L$.
 
@@
 
-Referencing equation \ref{eq:emc}.
+Referencing equation.
 Mauris vel lorem magna, tristique auctor ipsum.
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 %	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
+%------------------------------------------------------------------------------
 \begin{acknowledgments}
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 %	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 
@@
 
-% ---------------------------------------------------------------------------- %
-% Bibliografia
+% ---------------------------------------------------------------------------- % Bibliografy
+% ---------------------------------------------------------------------------- 
 \bibliographystyle{BibTeX/agufull08} 	% citação bibliográfica textual
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 %	FIGURES AND TABLES
-%----------------------------------------------------------------------------------------
+%------------------------------------------------------------------------------
 
@@
 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
-	\caption{Learning and target catalogues.}
+	\caption{Learning and target catalogs.}
 	\label{fig_helmstetter_catalogues}
@@
 %
-% Use the \subsubsubsection{} command to identify level 3 heads
+% Use the \subsubsubsection{} command to identify level 4 heads
 % An example:

### 2015-01-15 manuscript.tex
bsb2014.11 description and Mw guidelines

@@
 
-To perform the hazard calculation, the OpenQuake suite was choosed.
+To perform the hazard computation, the OpenQuake suite was choosed.
 The modeller environment used was the HMTK also available by GEM scientifical board.
@@
 %------------------------------------------------
-\section{Data}
+\section{Brazilian seismicity data}
+
+The main source of brazilian intraplate seismicity data is current the Brazilian Seismic Bulletin maintained mainly by the Seismological Certer, Institute of Astronomy, Geophysics and Atmospherical Sciences, University of São Paulo \citep{bsb_2014}.
+This catalog is a compilation from different sources.
+
+Until 1981, the catalog corresponds to \citet{berrocal_1984} which is a huge compilation effort from historic and instrumental earthquake data with participation of Univesidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON), Universidade Estadual de São Paulo (UNESP) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
+From 1982 to 1995 the information comes from Revista Brasileira de Geofísica.
+Since 1995, the Bulletin is maintained mainly by University of São Paulo under the same cooperation network.
+Earlier versions of this catalog was used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.
+The number of earthquakes on the catalog
+
+Today the catalog is distributed in two ways.
+One could be called \emph{raw} and contains all compiled information even events outside country which effects was felt in Brazil, and other could be called \emph{clean}, since that events with high error and low quality was removed.
+
+This catalog only cover crustal events.
+Andean deep subduction earthquakes with depht higher than 50km at west was discarded.
+This deep earthquakes contributions to the ground shaking hope to be consider in the future based on some new comprehensive south-american catalog.
+
+Almost magnitude values was computed as $mb$ type or equivalent $mR$ \citep{assumpcao_1983}.
+Part of them was computed from macroseismic felt area data \citep{berrocal_1984}.
+Since almost \gls{gmpe} are based on $M_W$ moment scale magnitude, on this present study, a new version of the catalog with $M_W$ values as delivered.
+\citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.
+On this context the conversion was made folowhing the guidelines on table \ref{tab:guidelines} from an \emph{ad-hoc} communication distributed with the catalog.
 
@@
 \begin{table}[H]
-	\caption{Table caption}
+	\caption{BSB 2014.11 $M_W$ proxies \emph{ad-hoc} guidelines.}
 	\begin{tabular}{l l l}
 		\hline
-		\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
+		\textbf{magnitude source} & \textbf{rule} & \textbf{uncertainty}\\
 		\hline
-		Treatment 1         & 0.0003262           & 0.562 \\
-		Treatment 2         & 0.0015681           & 0.910 \\
-		Treatment 3         & 0.0009271           & 0.296 \\
+		% mb or equival & $$ M_W = 1.21m - 0.76 $$  & 0.3 \\
+		% 		% magnitude from $A_f$  & $$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 8.4\times10^{-4}A_f^{1/2}& 0.4 \\
+		$mb$ or $mR$ 
+		&$ M_W(m) = 1.21 m - 0.76 $  
+		& 0.3 \\
+		Area felt $A_f$ 	
+		&$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 0.00084\sqrt{A_f}$& 0.4 \\
+		Maximum intensity $I_0$ 	
+		& $mb(I_0) = 1.21 + 0.45 I_0$ then $M_W(m)$
+		& 0.6 \\
+		$mb$ and $A_f$ 	
+		& $M_W(m, A_f) = 0.7 M_W(m) + 0.3 M_W(A_f)$
+		& 0.33 \\
 		\hline
 	\end{tabular}
-	\label{sampletable}
+	\label{tab:guidelines}
 \end{table}

### 2015-01-15 manuscript.tex
update some figures

@@
 %------------------------------------------------
+%
+% seismicity data
+% 
+%------------------------------------------------
+
 \section{Brazilian seismicity data}
 
-The main source of brazilian intraplate seismicity data is current the Brazilian Seismic Bulletin maintained mainly by the Seismological Certer, Institute of Astronomy, Geophysics and Atmospherical Sciences, University of São Paulo \citep{bsb_2014}.
+Brazilian intraplate seismicity data is current the Brazilian Seismic Bulletin maintained mainly by the Seismological Certer, Institute of Astronomy, Geophysics and Atmospherical Sciences, University of São Paulo \citep{bsb_2014}.
 This catalog is a compilation from different sources.
@@
 This deep earthquakes contributions to the ground shaking hope to be consider in the future based on some new comprehensive south-american catalog.
+The epicenters on the catalog is shown on the figure \ref{fig_seismicity}.
 
@@
 
+\subsection{Catalog overview}
+
+Figure \ref{fig_overview} represent some parameters distribution on the catalog.
+Is easy to note that depth are not well defined into the catalog, and here all earthquakes will considered as in the shallow crust, around 15km.
+From the figures \ref{fig_overview}a and \ref{fig_overview}b, is possible to note the behavior almost uniform on weekday distribution, and a slightly tendence to record earthquakes during the night.
+This could suggest some influence of daily noise level.
+
+\subsection{Declustering procedures}
+
+blsbalba
+
+\subsection{Catalog completeness}
+
+blsbalba
+
+%------------------------------------------------
+%
+% smooothing methods
+% 
 %------------------------------------------------
@@
 \begin{figure}
+	\includegraphics[width=0.5\linewidth]{z_img_seismicity}
+	\caption{Brazilian seismicity from 1767 to 2014. Source: \citep{bsb_2014}.}
+	\label{fig_seismicity}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.20\textwidth} 
+						 p{0.40\textwidth} 
+						 p{0.40\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\				
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
+		\end{tabular}
+	\end{table}
+	\caption{Catalogue overview.
+The histogram (a) shows the depth distribution.
+The histograms (b) and (c) represent the distributions of weekday and hour which earthquakes occur respectly.
+Dashed lines represent the mean value.}
+	\label{fig_overview}
+	\end{center}
+\end{figure}
+
+\begin{figure}
 	\begin{center}
@@
 
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_seismicity_br}
-	\caption{Brazilian earthquake catalog.}
-	\label{fig_seismicity}
-\end{figure}
-
 % =========
@@
 % =========
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_hmtk_bsb2013_rate}
-	\caption{Earthquake records by year}
-	\label{fig_seismic_rate}
-\end{figure}
 
@@
 		$mb$ or $mR$ 
-		&$ M_W(m) = 1.21 m - 0.76 $  
+		&$ M_W(m) = 1.121 m - 0.76 $  
 		& 0.3 \\

### 2015-01-16 manuscript.tex
improove data description and figures layout

@@
 
+\subsection{Hypocenters}
+
 Until 1981, the catalog corresponds to \citet{berrocal_1984} which is a huge compilation effort from historic and instrumental earthquake data with participation of Univesidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON), Universidade Estadual de São Paulo (UNESP) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
@@
 
+\subsection{Magnitudes}
+
 Almost magnitude values was computed as $mb$ type or equivalent $mR$ \citep{assumpcao_1983}.
@@
 
-\subsection{Catalog overview}
+\subsection{Catalog check}
 
@@
 
+The annual earthquake records since 1900 on the catalog is on the figure \ref{fig_records} where is evident the seismicity increase arround 1986 and 1998 (???) derived by João Câmara [REFERENCE] and BEBEDOURO, SOBRAL?!?
+activity.
+Also is possible to note the increasing number of records after 1960 due the operation of global, regional and local stations since that.
+
 \subsection{Declustering procedures}
 
-blsbalba
-
-\subsection{Catalog completeness}
-
-blsbalba
+To preserve the Poisson assumption about the independece of the events on the catalog, it was submited to some decluster algorithms and the results is shown on the figure \ref{fig_decluster}a.
+
+The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uental \citep{marsan_david_2012}.
+In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uental distance window.
+
+In the case of this catalog the methods did not presented strong differences at the final results (fig.
+\ref{fig_decluster}a) and the final choose of Gardner-Knopof/Uhrhammer was based on criteria to maximize the number of events to be analysed.
+
+The map on figure \ref{fig_decluster}b locate all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where in some time, earthquakes occur sistematically.
+
+\subsection{Frequency-magnitude distribution}
+
+Figure \ref{fig_occurrence} presents the magnitude and frequency distribution.
+A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.
+
+\subsection{Catalogue completeness}
+
+The Brazilian seismicity data completeness has evident time and space dependence.
+The number of stations and their accuracy increased in the time, and the spatial occupation respects the population density coverage concentrated in the east coast in almost cases with exception only in the central part of the country by the Brasilia Seismic Station (REFERENCE).
+
+The spatial completeness dependence was not considered on this work. History of almost regional stations was not easy available and blocks methodologies, e.g. \citet{mignan_2011} work.
+
+Time dependence was defined by expert criteria and it is shown on the figure
 
@@
 %\end{table}
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+% 		\footnotesize(a)
+% 		& 
+% 		\footnotesize(b)
+% 		\\		
+% 		\includegraphics[width=0.45\textwidth]{placeholder} &
+% 		\includegraphics[width=0.45\textwidth]{placeholder}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
+% 	\label{placeholder}
+% 	\end{center}
+% \end{figure}
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.33\textwidth} 
+% 						 p{0.33\textwidth} 
+% 						 p{0.33\textwidth} }
+% 		\footnotesize(a) &
+% 		\footnotesize(b) &
+% 		\footnotesize(c)
+% 		\\				
+% 		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
+% 		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
+% 		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
+% 	\label{placeholder2}
+% 	\end{center}
+% \end{figure}
 
@@
 	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
+		\\		
+		\footnotesize(b) 
+		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
+		\end{tabular}
+	\end{table}
+	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
+	\label{fig_records}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
+		\\		
+		\footnotesize(b) 
+		& \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2013_pp_decluster}
+		\end{tabular}
+	\end{table}
+	\caption{Decluster evaluation (a) and the clusters map (b).}
+	\label{fig_decluster}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=1.0\linewidth]{z_img_hmtk_bsb2014_11_completeness}
+	\caption{Annual earthquake recurrence.}
+	\label{fig_occurrence}
+\end{figure}
+
+% =========
+% Methods
+% =========
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
+	\caption{Magnitude dependence bandwidth.}
+	\label{fig_woo_bandwidth}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
+	\caption{Local bandwidth example.}
+	\label{fig_helmstetter_hidi}
+\end{figure}
+
+% TODO add HI_DI histograms
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
+	\caption{Stationary seismic rate}
+	\label{fig_helmstetter_stationary_rate}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
+	\caption{Learning and target catalogs.}
+	\label{fig_helmstetter_catalogues}
+\end{figure}
+
+% =========
+% Smoothed Rates
+% =========
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.33\textwidth} 
+						 p{0.34\textwidth} 
+						 p{0.33\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\				
+		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
+		\includegraphics[height=5cm]{z_img_a_woo} &
+		\includegraphics[height=5cm]{z_img_a_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
+	\label{fig_rates}
+	\end{center}
+\end{figure}
+
+% =========
+% Results
+% =========
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.33\textwidth} 
+						 p{0.34\textwidth} 
+						 p{0.33\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\				
+		\includegraphics[height=5cm]{z_img_pga_frankel}  &
+		\includegraphics[height=5cm]{z_img_pga_woo} &
+		\includegraphics[height=5cm]{z_img_pga_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{PGA (poe 10\%/50y) comparission for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
+	\label{fig_pga}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
@@
 		\\		
-		\includegraphics[width=0.45\textwidth]{placeholder} &
-		\includegraphics[width=0.45\textwidth]{placeholder}
+		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
+		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
 		\end{tabular}
 	\end{table}
-	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
+	\caption{Previous available Brazillian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
 	\label{placeholder}
 	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		\begin{tabular}{ p{0.33\textwidth} 
-						 p{0.33\textwidth} 
-						 p{0.33\textwidth} }
-		\footnotesize(a) &
-		\footnotesize(b) &
-		\footnotesize(c)
-		\\				
-		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
-		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
-		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
-		\end{tabular}
-	\end{table}
-	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
-	\label{placeholder2}
-	\end{center}
-\end{figure}
-
-% =========
-% data
-% =========
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_time_mag_count_br_1960}
-	\caption{Magnitude count by year.}
-	\label{fig_time_mag_count}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_decluster_br}
-	\caption{Catalog declustering.}
-	\label{fig_decluster}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_occurrence}
-	\caption{Annual earthquake recurrence.}
-	\label{fig_occurrence}
-\end{figure}
-
-% =========
-% Methods
-% =========
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_a_frankel_br}
-	\caption{Smoothed seismicity: Frankel's method.}
-	\label{fig_a_frankel}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
-	\caption{Magnitude dependence bandwidth.}
-	\label{fig_woo_bandwidth}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_a_woo}
-	\caption{Smoothed seismicity: Woo's method.}
-	\label{fig_a_woo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
-	\caption{Local bandwidth example.}
-	\label{fig_helmstetter_hidi}
-\end{figure}
-
-% TODO add HI_DI histograms
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
-	\caption{Stationary seismic rate}
-	\label{fig_helmstetter_stationary_rate}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
-	\caption{Learning and target catalogs.}
-	\label{fig_helmstetter_catalogues}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_a_helmstetter}
-	\caption{Smoothed seismicity: Helmstetter's method.}
-	\label{fig_a_helmstetter}
-\end{figure}
-
-% =========
-% Results
-% =========
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_pga_gshap}
-	\caption{GSHAP results: PGA (poe 10\%/50y)}
-	\label{fig_pga_gshap}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_pga_frankel}
-	\caption{Frankel's results: PGA (poe 10\%/50y)}
-	\label{fig_pga_frankel}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_pga_woo}
-	\caption{Woo's results: PGA (poe 10\%/50y)}
-	\label{fig_pga_woo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_pga_helmstetter}
-	\caption{Helmstetter's results: PGA (poe 10\%/50y)}
-	\label{fig_pga_helmstetter}
 \end{figure}

### 2015-01-19 manuscript.tex
add Assumpcao suggestions

@@
 \title{Smoothing seismicity techniques applied to
-seismic sources characterization and
+seismic source characterization and
 probabilistic seismic hazard analysis in Brazil}
@@
 
-Three smoothing seismicity techniques are reviewed in order to investigate their results when applied over the Brazilian seismicity and suggesting a seismic hazard assesment.
-
-All this methods are based on a kernel density estimation. The main difference between them is the bandwidth selection process.
-
-Frankel in their method uses a \emph{fixed} bandwidth in space.
-Woo's method suggest a bandwidth as function of the \emph{magnitude}.
-Helmstetter and Werner proposes a \emph{local-adaptive} in space and time kernel bandwidths, chosen for each event.
-This last one was proposed initialy to take care of the background seismicity rateon long-term earthquake forecast context and the time-independent assumption about the annual background seismic rate could be used
+Three smoothing seismicity techniques are reviewed and their application tothe Brazilian seismicity and a seismic hazard assesment are discussed.
+
+All these smoothing methods are based on a kernel density estimation. The main difference between them is the bandwidth selection process.
+
+The method of \citet{frankel_1995} uses a \emph{fixed} bandwidth in space.
+\citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}.
+\citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.
+This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
 (not indiscriminately REFERENCE) as a grid of points seismic sources.
 
-As conclusions the results shown that all methods gives defines better some known seismic regions and could be used in the Brazilian seismicity case, all of them, improoving the spatial resolution of the GSHAP model, the last one made on the region.
+The results shown that all methods clearly define some known seismic regions and could be used in the Brazilian seismicity.
+All methods improve the spatial resolution of the GSHAP model, the only previously available for the region.
 
@@
 
-Brazil have no recent seismic hazard study at national scale despite some important and localized recent studies focused mainly on specific buildings.
-
-The Brazilian seismic hazard building code was made based on GSHAP model which is the well accepted state of art model for the whole region.
-
-A significant number of earthquakes has been recorded in the last years, mainly after the national seismic network (RSBR) operation, and an actualized hazard map whould be computed.
-
-Although the standard Cornell Mcguire zoning methodolgy for the classical hazard seismic assesment, the goal is review a few smoothing seismicity methods based on kernel density estimation.
-This estimation allow a seismic point sources grid definition.
-The seismic rate (\emph{a}-value) associated to each source is definided by the smoothed seismicity methods, but the more parameters like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions must to be assigned in other ways.
-
-In the kernel density estimation theory two things there always be handled.
-The kernel shape and bandwidth.
-Almost kernels used are gaussian or power-laws in one of their variants.
-The mainly distinction between methods decribed is the proposed bandwidth selection process.
-
-Frankel smoothing method applies a gaussian Nadaraya-Watson smoothing method with a \emph{fixed} smoothing-distance bandwidth to estimate seismicity rate.
-Gordon Woo proposes an \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbour mean distance in magnitude bins.
-Helmstetter proposes a \emph{local-adaptive} bandwidth, based on kernel estimations on both space and time dimentions, estimation of the background seismicity rate for long-term forecasts.
-Using one-year forecast and assume time-independence, the same assumption of PSHA, of seismicity rate is possible to use these values on seismic sources characterization.
-
-To perform the hazard computation, the OpenQuake suite was choosed.
-The modeller environment used was the HMTK also available by GEM scientifical board.
+\subsection{Intraplate Seismicity}
+
+Onononono on on ono on on on on ononon on on on on on ono non on ono on on
+
+\subsection{Previous Brazilian Studies}
+
+Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984}, 2000, Assumpção in \citep{talwani_2014} (REFERENCE), Brazil has no seismic hazard map at national scale.
+
+The Brazilian seismic hazard building code (REFERENCE) [ABNT] was based on the GSHAP results for South America, which is the last published model for the whole region.
+Howerver, some significant seismic areas (Porto dos Gaúchos [REFERENCIA]), are not represented in the GPSHAP-based building codes.
+
+A significant number of earthquakes has been recorded in the last decades allowing a better definition of seismic zones. For this reason an updated hazard map is needed.
+
+\subsection{Previous Smoothing Seismicity (Frankel, Woo, Helmstetter\ldots)}
+
+Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on kernel density estimation which results in a grid of smoothed seismicity.
+
+This kernel density estimation or the smoothed seismicity grid allow to define a grid of seismic point sources.
+The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.
+
+In the kernel density estimation both its shape and its bandwidth need to be defined.
+Most kernels are gaussian or power-laws in one of their variants.
+The main distinction between the methods described here is the bandwidth selection process.
+
+\citet{frankel_1995} smoothing method applies a gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
+\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbour mean distance in magnitude bins.
+\citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth, based on kernel estimations for the background seismicity rate in long-term forecasts on both space and time dimensions.
+
+Using one-year forecast and assuming time-independence of the seismicity rate, the same assumption of PSHA, it is possible to use these seismicity values on seismic sources characterization.
+(REVER E COMENTAR)
+
+\subsection{Hazard computation}
+
+To perform the hazard computation, the OpenQuake suite was chosen.
+The modelling tool was the HMTK also available by GEM scientifical board.
 The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.
@@
 
-\subsection{Intraplate Seismicity}
-
-Onononono on on ono on on on on ononon on on on on on ono non on ono on on
-
-\subsection{Previous Brazilian Studies}
-
-Onononono on on ono on on on on ononon on on on on on ono non on ono on on
-
-\subsection{Previous Smoothing Seismicity (Frankel, Woo, Helmstetter\ldots)}
-
-Onononono on on ono on on on on ononon on on on on on ono non on ono on on
-
 \subsection{Specifical Purpose}
@@
 
-Brazilian intraplate seismicity data is current the Brazilian Seismic Bulletin maintained mainly by the Seismological Certer, Institute of Astronomy, Geophysics and Atmospherical Sciences, University of São Paulo \citep{bsb_2014}.
-This catalog is a compilation from different sources.
+Data on intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo \citep{bsb_2014}.
+This catalog is a joint compilation from different sources.
 
@@
 
-Until 1981, the catalog corresponds to \citet{berrocal_1984} which is a huge compilation effort from historic and instrumental earthquake data with participation of Univesidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON), Universidade Estadual de São Paulo (UNESP) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
-From 1982 to 1995 the information comes from Revista Brasileira de Geofísica.
-Since 1995, the Bulletin is maintained mainly by University of São Paulo under the same cooperation network.
-Earlier versions of this catalog was used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.
-The number of earthquakes on the catalog
+For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation of Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
+From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica.
+Since 1995, the Bulletin is maintained mainly by University of São Paulo under the same cooperation network in addition to Universidade Estadual Paulista (UNESP).
+Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.
 
 Today the catalog is distributed in two ways.
-One could be called \emph{raw} and contains all compiled information even events outside country which effects was felt in Brazil, and other could be called \emph{clean}, since that events with high error and low quality was removed.
-
-This catalog only cover crustal events.
-Andean deep subduction earthquakes with depht higher than 50km at west was discarded.
-This deep earthquakes contributions to the ground shaking hope to be consider in the future based on some new comprehensive south-american catalog.
-The epicenters on the catalog is shown on the figure \ref{fig_seismicity}.
+One, which could be called \emph{raw}, contains all compiled information, such as: events outside the country which were felt in Brazil, errors in previous versions and also non-seismic events (known quarry blasts, sonic boom, etc).
+The other way could be called \emph{clean}, where all non-seismic and low quality events were removed.
+This last one was used in this paper and the total number of entries is [CALCULAR O NUMERO DE EVENTOS NO CATALOGO].
+
+This catalog only cover crustal events. Andean subduction earthquakes deeper than 50km in the Brazil-Peru border was discarded. The epicenters of the catalog are shown in figure \ref{fig_seismicity}.
 
@@
 
-Almost magnitude values was computed as $mb$ type or equivalent $mR$ \citep{assumpcao_1983}.
-Part of them was computed from macroseismic felt area data \citep{berrocal_1984}.
-Since almost \gls{gmpe} are based on $M_W$ moment scale magnitude, on this present study, a new version of the catalog with $M_W$ values as delivered.
-\citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.
-On this context the conversion was made folowhing the guidelines on table \ref{tab:guidelines} from an \emph{ad-hoc} communication distributed with the catalog.
+Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude.
+For historical events magnitudes were estimated from macroseismic data (felt area and maximum intensity) \citep{berrocal_1984}.
+Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
+REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS..
+(Johnston)
+% Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.
 
@@
 
-Figure \ref{fig_overview} represent some parameters distribution on the catalog.
-Is easy to note that depth are not well defined into the catalog, and here all earthquakes will considered as in the shallow crust, around 15km.
-From the figures \ref{fig_overview}a and \ref{fig_overview}b, is possible to note the behavior almost uniform on weekday distribution, and a slightly tendence to record earthquakes during the night.
+Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour.
+Most events have no depth determination and the well determined depths (from pP or local networks) are mostly less than 10km.
+
+From the figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night.
 This could suggest some influence of daily noise level.
 
-The annual earthquake records since 1900 on the catalog is on the figure \ref{fig_records} where is evident the seismicity increase arround 1986 and 1998 (???) derived by João Câmara [REFERENCE] and BEBEDOURO, SOBRAL?!?
-activity.
-Also is possible to note the increasing number of records after 1960 due the operation of global, regional and local stations since that.
+The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara [REFERENCE, TAKEYA et al 1989] and Porto dos Gaúchos [REFERENCE, BARROS 2009].
+The increasing number of records after 1960 is due to the operation of global, regional and local stations.
 
@@
 
-To preserve the Poisson assumption about the independece of the events on the catalog, it was submited to some decluster algorithms and the results is shown on the figure \ref{fig_decluster}a.
-
-The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uental \citep{marsan_david_2012}.
+To preserve the Poisson assumption about the independence of the events several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a.
+The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uental \citep{marsan_david_2012}
+(REVER REFERENCIA DO PROPRIO gruental ou CORRIGIN POIS NAO EH MARSAN DAVID mas Thomas van Stiphout, Jiancang Zhuang, and David Marsan].
 In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uental distance window.
 
-In the case of this catalog the methods did not presented strong differences at the final results (fig.
-\ref{fig_decluster}a) and the final choose of Gardner-Knopof/Uhrhammer was based on criteria to maximize the number of events to be analysed.
-
-The map on figure \ref{fig_decluster}b locate all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where in some time, earthquakes occur sistematically.
+In the case of Brazilian catalog the methods did not present large differences at the final number of earthquakes (fig.
+\ref{fig_decluster}a) and the Gardner-Knopof/Uhrhammer combination was chosen to maximize the number of events.
+
+The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where long earthquake series tend to occur more frequently.
 
@@
 
-\subsection{Catalogue completeness}
-
-The Brazilian seismicity data completeness has evident time and space dependence.
-The number of stations and their accuracy increased in the time, and the spatial occupation respects the population density coverage concentrated in the east coast in almost cases with exception only in the central part of the country by the Brasilia Seismic Station (REFERENCE).
-
-The spatial completeness dependence was not considered on this work. History of almost regional stations was not easy available and blocks methodologies, e.g. \citet{mignan_2011} work.
-
-Time dependence was defined by expert criteria and it is shown on the figure
+\subsection{Catalog completeness}
+
+The completeness of Brazilian seismicity data has evident time and space dependence.
+
+The number and quality of stations increased over time. The population density, previously concentrated along the coast increased in other regions.
+
+% and the spatial occupation respects the population density coverage concentrated in the east coast in most cases with exception only in the central part of the country by the Brasilia Seismic Station (REFERENCE).
+
+The histories of most regional stations are not easy available which hinders application of methodologies, such as \citet{mignan_2011}.
+For this reason, spatial completeness was not considered and temporal completeness parameters for the whole country was defined by expert criteria (figure FIGURA COMPLETENESS )
+
+- completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
+- se conhecia metodos para lidar com completeza espacial
+- também se conhece o método multiescala (provavel bom solucionador)
+- método de Stepp
+- mas usa-se uma completeness temporal simples.
 
@@
 \section{Smoothing methods}
-Under this work, three smoothing methods was considered. On this section brief explanation of each one will be done.
-
-\subsection{Frankel}
-
-Arthur \citet{frankel_1995} smoothing seismicity proposal consists originally in use a called correlation distance $d_F$ as the kernel \textit{fixed} bandwidth and next to apply the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
+Three smoothing methods are now briefly explained.
+
+\subsection{Frankel: fixed bandwidth}
+
+Arthur \citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
 \begin{equation}
@@
 
-\subsection{Woo}
-
-Gordon \citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, far $\boldsymbol{r}_i$ from its kernel, on the cell centered on $\boldsymbol{r}$ in dependence of their magnitude $m$:
+\subsection{Woo: magnitude-dependent bandwidth}
+
+\citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, located at $\boldsymbol{r}_i$, to the seismicity rate $R$ (number of earthquakes per year and unity area) on the cell centered in $\boldsymbol{r}$, depending on the magnitude $m$:
 \begin{equation}
@@
 \end{equation}
-where $N$ is the number of earthquakes $i$ on the catalog
-and $T(\gls{sym:ri})$ is the timeframe which all earthquakes with magnitude above $m$ is completely observed on \gls{sym:ri}.
-
-Any kernel could be applied on that definition. In practice Woo used a \citet{kagan_knopoff_1980} for a infinity spatial domain:
+where $N$ is the number of earthquakes $i$ in the magnitude bin $m \pm \mathrm{d}m$,
+and $T(\gls{sym:ri})$ is the completess time of magnitude $m$ observed on \gls{sym:ri}.
+
+Any kernel could be applied on that definition. I used used the \citet{kagan_knopoff_1980} kernel for infinite spatial domain as described by \citet{woo_1996}:
 \begin{equation}
@@
 \end{equation}
-where \gls{sym:aW} is, accordingly \citet{verejones_1992}, \glsdesc{sym:aW}.
-
-To compute the magnitude-dependent bandwidth function \gls{sym:hm}, Gordon Woo suggested the follow relation
+where \gls{sym:aW} is \glsdesc{sym:aW} \citep{verejones_1992}.
+
+VERIFICAR SE É O CASO DE APRESENTAR OUTRAS FORMAS DE KERNEL. DAR PREFERENCIAS AOS REALMENTE UTILIZADOS.
+
+To compute the magnitude-dependent bandwidth function \gls{sym:hm}, \citet{woo_1996} suggested the follow relation
 \begin{equation}
@@
 \end{equation}
-where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$.
-
-\subsection{Helmstetter and Werner}
+where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$, as illustrated on figure \ref{fig_woo_bandwidth}.
+
+Just to compare, in figure \ref{fig_woo_bandwidth} it is shown the functions computed by \citet{beauval_2003} for Norway and Spain.
+The angular coeficient ($a_1$) is in the same range of the others but the general mean distance between events into the same magnitude bin is higher than other countries.
+It is in some way expected.
+Brazil is quite large than mentioned countries.
+
+\subsection{Helmstetter and Werner: locally-adaptive space and time bandwidths}
 

### 2015-01-27 manuscript.tex
changing some figures
minimal changes

@@
 % Corresponding author mailing address and e-mail address:
-\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
-(marlon.pirchiner@iag.usp.br)}
+\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil. (marlon.pirchiner@iag.usp.br)}
 
@@
 
-\title{Smoothing seismicity techniques applied to
-seismic source characterization and
-probabilistic seismic hazard analysis in Brazil}
+\title{Smoothing seismicity techniques applied to seismic source characterization and probabilistic seismic hazard analysis in Brazil}
 
@@
 
-\altaffiltext{1}{Department of Geophysics,
-Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG),
-University of São Paulo (USP), São Paulo, Brazil.}
+\altaffiltext{1}{Department of Geophysics, Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG), University of São Paulo (USP), São Paulo, Brazil.}
 
@@
 
-Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assesment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
-
-Three smoothing seismicity techniques are reviewed and their application tothe Brazilian seismicity and a seismic hazard assesment are discussed.
+Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+
+Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assessment are discussed.
 
@@
 \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.
-This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
+This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
 (not indiscriminately REFERENCE) as a grid of points seismic sources.
@@
 The Brazilian seismic hazard building code (REFERENCE) [ABNT] was based on the GSHAP results for South America, which is the last published model for the whole region.
-Howerver, some significant seismic areas (Porto dos Gaúchos [REFERENCIA]), are not represented in the GPSHAP-based building codes.
+However, some significant seismic areas (Porto dos Gaúchos [REFERENCIA]), are not represented in the GPSHAP-based building codes.
 
@@
 In the kernel density estimation both its shape and its bandwidth need to be defined.
-Most kernels are gaussian or power-laws in one of their variants.
+Most kernels are Gaussian or power-laws in one of their variants.
 The main distinction between the methods described here is the bandwidth selection process.
 
-\citet{frankel_1995} smoothing method applies a gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
-\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbour mean distance in magnitude bins.
+\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
+\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins.
 \citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth, based on kernel estimations for the background seismicity rate in long-term forecasts on both space and time dimensions.
@@
 To perform the hazard computation, the OpenQuake suite was chosen.
-The modelling tool was the HMTK also available by GEM scientifical board.
+The modelling tool was the HMTK also available by GEM scientific board.
 The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.
@@
 
-\subsection{Specifical Purpose}
+\subsection{Specific Purpose}
+
+asdfasdfa
 
@@
 
+%------------------------------------------------------------------------------
 \subsection{Catalog check}
+%------------------------------------------------------------------------------
 
@@
 
+%------------------------------------------------------------------------------
 \subsection{Declustering procedures}
+%------------------------------------------------------------------------------
 
@@
 
+%------------------------------------------------------------------------------
 \subsection{Frequency-magnitude distribution}
+%------------------------------------------------------------------------------
 
@@
 
+%------------------------------------------------------------------------------
 \subsection{Catalog completeness}
+%------------------------------------------------------------------------------
 
@@
 %------------------------------------------------
-%
-% smooothing methods
-%
+% M E T O D O L O G Y 
 %------------------------------------------------
 \section{Smoothing methods}
+
 Three smoothing methods are now briefly explained.
 
+%------------------------------------------------------------------------------
 \subsection{Frankel: fixed bandwidth}
-
-Arthur \citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
+%------------------------------------------------------------------------------
+
+Arthur \citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
 \begin{equation}
@@
 
+%------------------------------------------------------------------------------
 \subsection{Woo: magnitude-dependent bandwidth}
+%------------------------------------------------------------------------------
 
@@
 where $N$ is the number of earthquakes $i$ in the magnitude bin $m \pm \mathrm{d}m$,
-and $T(\gls{sym:ri})$ is the completess time of magnitude $m$ observed on \gls{sym:ri}.
+and $T(\gls{sym:ri})$ is the completeness time of magnitude $m$ observed on \gls{sym:ri}.
 
@@
 Just to compare, in figure \ref{fig_woo_bandwidth} it is shown the functions computed by \citet{beauval_2003} for Norway and Spain.
-The angular coeficient ($a_1$) is in the same range of the others but the general mean distance between events into the same magnitude bin is higher than other countries.
+The angular coefficient ($a_1$) is in the same range of the others but the general mean distance between events into the same magnitude bin is higher than other countries.
 It is in some way expected.
@@
 
+%------------------------------------------------------------------------------
+
 \subsection{Helmstetter and Werner: locally-adaptive space and time bandwidths}
-
-Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to charachterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.
+%------------------------------------------------------------------------------
+
+Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to characterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.
 
@@
 
+%------------------------------------------------------------------------------
+
 \subsubsection{Local bandwidth computation}
+%------------------------------------------------------------------------------
 
@@
 
-The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.
+The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarely or regions with information lack.
+
+%------------------------------------------------------------------------------
 
 \subsubsection{Stationary seismic rate}
+%------------------------------------------------------------------------------
+
 After compute the model parameters and completely define it, the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ have to be computed.
@@
 
-\subsubsection{Maximum likelyhood optimization}
+%------------------------------------------------------------------------------
+
+\subsubsection{Maximum likelihood optimization}
+%------------------------------------------------------------------------------
 
@@
 
-If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the consider time frame is
+If the earthquake occurrence could be modeled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the consider time frame is
 \begin{equation}
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
+%	R E S U L T S 
+%----------------------------------------------------------------------------------------
 
@@
 
-\subsubsection{Simulation 2}
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
-
 \subsection{Real Data}
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
+%	D I S C U S S I O N
+%----------------------------------------------------------------------------------------
 
@@
 
-Nam fermentum sapien at enim varius consectetur.
-Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae.
-Integer lacinia purus vel metus tempus suscipit.
-Curabitur ac sapien quis mauris euismod commodo.
-Sed pharetra sem elit.
-Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor.
-Donec convallis ligula vitae sem viverra fermentum.
-Mauris in ullamcorper erat.
-Donec ultrices tempus nibh quis vestibulum.
-
-Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus.
-Aliquam a tellus nec felis sagittis consequat.
-Quisque convallis imperdiet neque a tempor.
-Nulla non erat urna.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Suspendisse viverra tortor pretium orci lacinia eleifend.
-Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem.
-Vestibulum pharetra malesuada egestas.
-Sed non orci sit amet erat suscipit fringilla in et diam.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla.
-Aenean porta rhoncus suscipit.
-
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
+Conclusion comments
 
@@
 
-\appendix
-
-\section{Appendix Title}
-
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
+% \appendix
+
+% \section{If I'll append something}
+
+% something appended
 
@@
 
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		\begin{tabular}{ p{0.20\textwidth}
-						 p{0.40\textwidth}
-						 p{0.40\textwidth} }
-		\footnotesize(a) &
-		\footnotesize(b) &
-		\footnotesize(c)
-		\\
-		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
-		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
-		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
-		\end{tabular}
-	\end{table}
-	\caption{Catalogue overview.
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.20\textwidth}
+% 						 p{0.40\textwidth}
+% 						 p{0.40\textwidth} }
+% 		\footnotesize(a) &
+% 		\footnotesize(b) &
+% 		\footnotesize(c)
+% 		\\
+% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
+% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
+% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{Catalogue overview.
 The histogram (a) shows the depth distribution.
-The histograms (b) and (c) represent the distributions of weekday and hour which earthquakes occur respectly.
+The histograms (b) and (c) represent the distributions of weekday and hour which earthquakes occur respectively.
 Dashed lines represent the mean value.}
-	\label{fig_overview}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-		\begin{tabular}{ c c }
-		\footnotesize(a)
-		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
-		\\
-		\footnotesize(b)
-		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
-		\end{tabular}
-	\end{table}
-	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
-	\label{fig_records}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-		\begin{tabular}{ c c }
-		\footnotesize(a)
-		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
-		\\
-		\footnotesize(b)
-		& \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2013_pp_decluster}
-		\end{tabular}
-	\end{table}
-	\caption{Decluster evaluation (a) and the clusters map (b).}
-	\label{fig_decluster}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=1.0\linewidth]{z_img_hmtk_bsb2014_11_completeness}
-	\caption{Annual earthquake recurrence.}
-	\label{fig_occurrence}
-\end{figure}
-
-% =========
-% Methods
-% =========
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
-	\caption{Magnitude dependence bandwidth.}
-	\label{fig_woo_bandwidth}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
-	\caption{Local bandwidth example.}
-	\label{fig_helmstetter_hidi}
-\end{figure}
-
-% TODO add HI_DI histograms
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
-	\caption{Stationary seismic rate}
-	\label{fig_helmstetter_stationary_rate}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
-	\caption{Learning and target catalogs.}
-	\label{fig_helmstetter_catalogues}
-\end{figure}
-
-% =========
-% Smoothed Rates
-% =========
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		\begin{tabular}{ p{0.33\textwidth}
-						 p{0.34\textwidth}
-						 p{0.33\textwidth} }
-		\footnotesize(a) &
-		\footnotesize(b) &
-		\footnotesize(c)
-		\\
-		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
-		\includegraphics[height=5cm]{z_img_a_woo} &
-		\includegraphics[height=5cm]{z_img_a_helmstetter}
-		\end{tabular}
-	\end{table}
-	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
-	\label{fig_rates}
-	\end{center}
-\end{figure}
-
-% =========
-% Results
-% =========
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		\begin{tabular}{ p{0.33\textwidth}
-						 p{0.34\textwidth}
-						 p{0.33\textwidth} }
-		\footnotesize(a) &
-		\footnotesize(b) &
-		\footnotesize(c)
-		\\
-		\includegraphics[height=5cm]{z_img_pga_frankel}  &
-		\includegraphics[height=5cm]{z_img_pga_woo} &
-		\includegraphics[height=5cm]{z_img_pga_helmstetter}
-		\end{tabular}
-	\end{table}
-	\caption{PGA (poe 10\%/50y) comparission for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
-	\label{fig_pga}
-	\end{center}
-\end{figure}
-
-\begin{figure}
-	\begin{center}
-	\begin{table}[H]
-		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-		\footnotesize(a)
-		&
-		\footnotesize(b)
-		\\
-		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
-		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
-		\end{tabular}
-	\end{table}
-	\caption{Previous available Brazillian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
-	\label{placeholder}
-	\end{center}
-\end{figure}
+% 	\label{fig_overview}
+% 	\end{center}
+% \end{figure}
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+% 		\begin{tabular}{ c c }
+% 		\footnotesize(a)
+% 		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
+% 		\\
+% 		\footnotesize(b)
+% 		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
+% 	\label{fig_records}
+% 	\end{center}
+% \end{figure}
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+% 		\begin{tabular}{ c c }
+% 		\footnotesize(a)
+% 		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
+% 		\\
+% 		\footnotesize(b)
+%         & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{Decluster evaluation (a) and the clusters map (b).}
+% 	\label{fig_decluster}
+% 	\end{center}
+% \end{figure}
+
+% \begin{figure}
+%     \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
+%     \caption{Completeness}
+%     \label{fig_completeness}
+% \end{figure}
+
+% \begin{figure}
+%     \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
+%     \caption{Annual earthquake recurrence.}
+%     \label{fig_occurrence}
+% \end{figure}
+
+% % =========
+% % Methods
+% % =========
+
+% \begin{figure}
+% 	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
+% 	\caption{Magnitude dependence bandwidth.}
+% 	\label{fig_woo_bandwidth}
+% \end{figure}
+
+% \begin{figure}
+% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
+% 	\caption{Local bandwidth example.}
+% 	\label{fig_helmstetter_hidi}
+% \end{figure}
+
+% % % TODO add HI_DI histograms
+
+% \begin{figure}
+% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
+% 	\caption{Stationary seismic rate}
+% 	\label{fig_helmstetter_stationary_rate}
+% \end{figure}
+
+% \begin{figure}
+% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
+% 	\caption{Learning and target catalogs.}
+% 	\label{fig_helmstetter_catalogues}
+% \end{figure}
+
+% % =========
+% % Smoothed Rates
+% % =========
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.33\textwidth}
+% 						 p{0.34\textwidth}
+% 						 p{0.33\textwidth} }
+% 		\footnotesize(a) &
+% 		\footnotesize(b) &
+% 		\footnotesize(c)
+% 		\\
+% 		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
+% 		\includegraphics[height=5cm]{z_img_a_woo} &
+% 		\includegraphics[height=5cm]{z_img_a_helmstetter}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{Smoothed rates comparison: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectively.}
+% 	\label{fig_rates}
+% 	\end{center}
+% \end{figure}
+
+% % =========
+% % Results
+% % =========
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.33\textwidth}
+% 						 p{0.34\textwidth}
+% 						 p{0.33\textwidth} }
+% 		\footnotesize(a) &
+% 		\footnotesize(b) &
+% 		\footnotesize(c)
+% 		\\
+% 		\includegraphics[height=5cm]{z_img_pga_frankel}  &
+% 		\includegraphics[height=5cm]{z_img_pga_woo} &
+% 		\includegraphics[height=5cm]{z_img_pga_helmstetter}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{PGA (poe 10\%/50y) comparison for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
+% 	\label{fig_pga}
+% 	\end{center}
+% \end{figure}
+
+% \begin{figure}
+% 	\begin{center}
+% 	\begin{table}[H]
+% 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+% 		\footnotesize(a)
+% 		&
+% 		\footnotesize(b)
+% 		\\
+% 		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
+% 		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
+% 		\end{tabular}
+% 	\end{table}
+% 	\caption{Previous available Brazilian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
+% 	\label{placeholder}
+% 	\end{center}
+% \end{figure}
 

### 2015-01-27 manuscript.tex
histogram figure
completeness study...

@@
 
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		\begin{tabular}{ p{0.20\textwidth}
-% 						 p{0.40\textwidth}
-% 						 p{0.40\textwidth} }
-% 		\footnotesize(a) &
-% 		\footnotesize(b) &
-% 		\footnotesize(c)
-% 		\\
-% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
-% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
-% 		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{Catalogue overview.
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.20\textwidth}
+						 p{0.40\textwidth}
+						 p{0.40\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
+		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
+		\end{tabular}
+	\end{table}
+	\caption{Catalogue overview.
 The histogram (a) shows the depth distribution.
@@
 Dashed lines represent the mean value.}
-% 	\label{fig_overview}
-% 	\end{center}
-% \end{figure}
-
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-% 		\begin{tabular}{ c c }
-% 		\footnotesize(a)
-% 		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
-% 		\\
-% 		\footnotesize(b)
-% 		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
-% 	\label{fig_records}
-% 	\end{center}
-% \end{figure}
-
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-% 		\begin{tabular}{ c c }
-% 		\footnotesize(a)
-% 		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
-% 		\\
-% 		\footnotesize(b)
-%         & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{Decluster evaluation (a) and the clusters map (b).}
-% 	\label{fig_decluster}
-% 	\end{center}
-% \end{figure}
-
-% \begin{figure}
-%     \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
-%     \caption{Completeness}
-%     \label{fig_completeness}
-% \end{figure}
-
-% \begin{figure}
-%     \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
-%     \caption{Annual earthquake recurrence.}
-%     \label{fig_occurrence}
-% \end{figure}
+	\label{fig_overview}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
+		\\
+		\footnotesize(b)
+		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
+		\end{tabular}
+	\end{table}
+	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
+	\label{fig_records}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
+		\\
+		\footnotesize(b)
+        & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
+		\end{tabular}
+	\end{table}
+	\caption{Decluster evaluation (a) and the clusters map (b).}
+	\label{fig_decluster}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
+    \caption{Completeness}
+    \label{fig_completeness}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
+    \caption{Annual earthquake recurrence.}
+    \label{fig_occurrence}
+\end{figure}
 
@@
 
-% \begin{figure}
-% 	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
-% 	\caption{Magnitude dependence bandwidth.}
-% 	\label{fig_woo_bandwidth}
-% \end{figure}
-
-% \begin{figure}
-% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
-% 	\caption{Local bandwidth example.}
-% 	\label{fig_helmstetter_hidi}
-% \end{figure}
-
-% % % TODO add HI_DI histograms
-
-% \begin{figure}
-% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
-% 	\caption{Stationary seismic rate}
-% 	\label{fig_helmstetter_stationary_rate}
-% \end{figure}
-
-% \begin{figure}
-% 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
-% 	\caption{Learning and target catalogs.}
-% 	\label{fig_helmstetter_catalogues}
-% \end{figure}
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
+	\caption{Magnitude dependence bandwidth.}
+	\label{fig_woo_bandwidth}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
+	\caption{Local bandwidth example.}
+	\label{fig_helmstetter_hidi}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \begin{table}[H]
+        %\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+        \begin{tabular}{ c c }
+        \footnotesize(a)
+        & \includegraphics[width=0.45\textwidth]{z_img_hi_histogram}
+        \\
+        \footnotesize(b)
+        & \includegraphics[width=0.45\textwidth]{z_img_di_histogram}
+        \end{tabular}
+    \end{table}
+    \caption{(a) $h_i$ and (b) $d_i$ distributions for BSB2014.}
+    \label{fig_hidi_distributions}
+    \end{center}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
+	\caption{Stationary seismic rate}
+	\label{fig_helmstetter_stationary_rate}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
+	\caption{Learning and target catalogs.}
+	\label{fig_helmstetter_catalogues}
+\end{figure}
 
@@
 
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		\begin{tabular}{ p{0.33\textwidth}
-% 						 p{0.34\textwidth}
-% 						 p{0.33\textwidth} }
-% 		\footnotesize(a) &
-% 		\footnotesize(b) &
-% 		\footnotesize(c)
-% 		\\
-% 		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
-% 		\includegraphics[height=5cm]{z_img_a_woo} &
-% 		\includegraphics[height=5cm]{z_img_a_helmstetter}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{Smoothed rates comparison: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectively.}
-% 	\label{fig_rates}
-% 	\end{center}
-% \end{figure}
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.33\textwidth}
+						 p{0.34\textwidth}
+						 p{0.33\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\
+		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
+		\includegraphics[height=5cm]{z_img_a_woo} &
+		\includegraphics[height=5cm]{z_img_a_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{Smoothed rates comparison: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectively.}
+	\label{fig_rates}
+	\end{center}
+\end{figure}
 
@@
 
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		\begin{tabular}{ p{0.33\textwidth}
-% 						 p{0.34\textwidth}
-% 						 p{0.33\textwidth} }
-% 		\footnotesize(a) &
-% 		\footnotesize(b) &
-% 		\footnotesize(c)
-% 		\\
-% 		\includegraphics[height=5cm]{z_img_pga_frankel}  &
-% 		\includegraphics[height=5cm]{z_img_pga_woo} &
-% 		\includegraphics[height=5cm]{z_img_pga_helmstetter}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{PGA (poe 10\%/50y) comparison for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
-% 	\label{fig_pga}
-% 	\end{center}
-% \end{figure}
-
-% \begin{figure}
-% 	\begin{center}
-% 	\begin{table}[H]
-% 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
-% 		\footnotesize(a)
-% 		&
-% 		\footnotesize(b)
-% 		\\
-% 		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
-% 		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
-% 		\end{tabular}
-% 	\end{table}
-% 	\caption{Previous available Brazilian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
-% 	\label{placeholder}
-% 	\end{center}
-% \end{figure}
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.33\textwidth}
+						 p{0.34\textwidth}
+						 p{0.33\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c)
+		\\
+		\includegraphics[height=5cm]{z_img_pga_frankel}  &
+		\includegraphics[height=5cm]{z_img_pga_woo} &
+		\includegraphics[height=5cm]{z_img_pga_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{PGA (poe 10\%/50y) comparison for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
+	\label{fig_pga}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\footnotesize(a)
+		&
+		\footnotesize(b)
+		\\
+		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
+		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
+		\end{tabular}
+	\end{table}
+	\caption{Previous available Brazilian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
+	\label{placeholder}
+	\end{center}
+\end{figure}
 

### 2015-01-27 process/completeness.py

@@ -1,0 +1,129 @@
+# -*- coding: utf-8 -*-
+# Python Numerical and Plotting Libraries
+import numpy as np
+import matplotlib.pyplot as plt
+from copy import deepcopy
+
+# HMTK Catalogue Import/Export Libraries
+from hmtk.parsers.catalogue.csv_catalogue_parser import CsvCatalogueParser, CsvCatalogueWriter
+
+
+# HMTK Declustering Tools
+from hmtk.seismicity.declusterer.dec_afteran import Afteran
+from hmtk.seismicity.declusterer.dec_gardner_knopoff import GardnerKnopoffType1
+from hmtk.seismicity.declusterer.distance_time_windows import (GardnerKnopoffWindow,
+                                                               GruenthalWindow,
+                                                               UhrhammerWindow)
+
+# HMTK Completeness Tools
+from hmtk.seismicity.completeness.comp_stepp_1971 import Stepp1971
+
+from hmtk.plotting.seismicity.completeness.cumulative_rate_analysis import SimpleCumulativeRate
+
+
+# HMTK Plotting Tools
+from hmtk.plotting.mapping import HMTKBaseMap
+from hmtk.plotting.seismicity.catalogue_plots import (plot_depth_histogram,
+                                                      plot_magnitude_time_scatter,
+                                                      plot_magnitude_time_density,
+                                                      plot_magnitude_depth_density,
+                                                      plot_rate,
+                                                      plot_weekday_histogram,
+                                                      plot_hour_histogram,
+                                                      plot_observed_recurrence)
+
+
+print 'Imports OK!'
+
+project_home = "/Users/pirchiner/dev/pshab_source_models"
+_model = "background"
+
+config = {
+#    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
+    'catalog': project_home + "/data/catalogs/hmtk_bsb2014.11.csv",
+    'declustered_catalog': project_home + "/"+_model+"/catalogs/bsb2014.11_declustered_catalog.csv",
+    'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
+    'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
+    # BR_Geral
+    'fixed_completeness1' : np.array([[2014., 3.5],
[…]

### 2015-02-03 d01-terms.tex
minnor corrections, references and symbol translations

@@
 {
-	name={terremoto},
-	description={ruptura de alguma estrutura geológica},
-	plural={terremotos}
+	name={earthquake},
+	description={rupture in some geological strutcture},
+	plural={earthquakes}
 }
@@
 {
-	name={sismicidade},
-	description={ocorrência dos tremores},
+	name={seismicity},
+	description={earthquake occurrence},
 }
@@
 {
-	name={hipocentro},
-	description={representação geométrica do ponto no espaço, onde 
-		se iniciou o \gls{rupture_process} da \gls{crust}},
-	plural={hipocentros}
+	name={hypocenter},
+	description={geometrical representation of the starting point of the \gls{rupture_process} on the \gls{crust}},
+	plural={hypocenters}
 }
@@
 {
-	name={epicentro},
-	description={projeção ortogonal, sobre a superfície, do \gls{hypocenter}},
+	name={epicenter},
+	description={orthogonal projecton over Earth surface of \gls{hypocenter}},
 	plural={epicentros}
@@
 {
-	name={sismotectônica},
+	name={seismotectonic},
 	description={o estudo das relações entre os \glspl{equake} e a \gls{tectonic} recente de uma região.
@@
 {
-	name={tect\^onica},
+	name={tectonic},
 	description={disciplina científica focada nos processos respons\'aveis 
 				 pela cria\c{c}\~ao e transforma\c{c}\~ao das estruturas geológicas da Terra e de outros planetas.},
-	plural={tect\^onicas}
+	plural={tectonics}
 }
@@
 {
-	name={técnicas de suavização},
+	name={smoothing techniques},
 	description={consiste em capturar importantes feições do conjunto de dados,
@@
 {
-	name={função de núcleo},
-	description={funções n-dimensionais, cuja integral em todo o domínio resulta em 1,
-				 podendo ser usadas como estimativas para 
-				 funções de densidade de probabilidade},
-	plural={funções de kernel},
+	name={kernel function},
+	description={n-dimentional function, which integral over whole domain is equal one and could be used as a probability density estimation.},
+	plural={kernel functions},
 }
@@
 {
-	name={taxa de sismicidade},
-	description={taxa com que terremotos são produzidos por determinada \gls{seismic_source}},
-	plural={taxas de sismicidade},
+	name={seismic rate},
+	description={rate within earthquakes are generated in some \gls{seismic_source}},
+	plural={seismic rates},
 }
@@
 {
-	name={atividade sísmica},
-	description={frequ{ê}cia de ocorr{ê}ncia de \glspl{equake}},
+	name={seismic activity},
+	description={frequency of \glspl{equake} occurence},
 }
@@
 {
-	name={fonte sísmica},
-	description={estrutura geológica capaz de produzir tremores de terra},
-	plural={fontes sísmicas}
+	name={seismic source},
+	description={geological structure able to produce \glspl{equake}},
+	plural={seismic sources}
 }
@@
 {
-	name={fonte sísmica pontual},
-	description={representação geométrica por um ponto, de uma fonte sísmica},
-	plural={fontes sísmicas pontuais},
+	name={point seismic source},
+	description={geometrical representation as point of some seismogenic source},
+	plural={point seismic sources},
 }
@@
 	name={GMPE},
-	description={Equação de predição do movimento do chão},
+	description={ground motion prediction equation},
 	plural={GMPEs},
@@
 {
-	name={fonte sísmica poligonal},
+	name={area seismic source},
 	description={representação geométrica por um polígono em superfície, 
 				 de uma fonte sísmica},
-	plural={fontes sísmicas poligonais},
+	plural={farea seismic sources},
 }
@@
 {
-	name={isocista},
-	description={curva que une valores de mesma intensidade sísmica},
+	name={isocist},
+	description={border of a region with the same seismic intensity from the same \gls{equake}},
 }

### 2015-02-03 d03-acronyms.tex
minnor corrections, references and symbol translations

@@
 
-\newacronym{psha}{PSHA}{Análise Probabilística de Ameaça Sísmica}
-\newacronym{dsha}{DSHA}{Análise Determinística de Ameaça Sísmica}
+\newacronym{psha}{PSHA}{Probabilistic Seismic Hazard Analysis}
+\newacronym{dsha}{DSHA}{Deterministic Seismic Hazard Analysis}
 \newacronym{GMPE}{\gls{gmpe}}{\glsdesc{gmpe}}
-\newacronym{mfd}{MFD}{Distribuição de Frequência e Magnitude}
+\newacronym{mfd}{MFD}{Magnitude Frequency Distribution}
 \newacronym{gr}{GR}{Gutenberg-Richter}
-\newacronym{cnn}{CNN}{Acoplamento dos Vizinhos mais Próximos}
-\newacronym{va}{v.a.}{variável aleatória}
-\newacronym{iid}{i.i.d.}{independentes e identicamente distribuidas}
-\newacronym{pdf}{densidade}{função de densidade de probabilidade}
-\newacronym{pmf}{pmf}{função de distribuição acumulada de probabilidade}
+\newacronym{cnn}{CNN}{Coupled-Nearest-Neighbour}
+\newacronym{va}{r.v.}{random variable}
+\newacronym{iid}{i.i.d.}{independent and identicaly distributed}
+\newacronym{pdf}{density}{probability density function}
+\newacronym{pmf}{pmf}{probability mass function}
+\newacronym{kde}{KDE}{Kernel Density Estimation}
 
-\newacronym{pga}{PGA}{máxima aceleração do chão}
+\newacronym{pga}{PGA}{peak ground-motion acceleration}
 
@@
 
-\newacronym{hl}{HazardLib}{Biblioteca de Ameaça}
-\newacronym{rl}{RiskLib}{Biblioteca de Risco}
+\newacronym{hl}{HazardLib}{hazard library}
+\newacronym{rl}{RiskLib}{risk library}
 \newacronym{nrml}{NRML}{Natural-hazard and Risk Markup Language}
@@
 \newacronym{oqp}{oq-platform}{Plataforma web de Interação com o OQ}
-\newacronym{oqe}{oq-engine}{Motor de Cálculo do OQ}
+\newacronym{oqe}{oq-engine}{OQ hazard and risk calculator}
 
-\newacronym{obsis}{ObSis}{Observatório Sismológico}
-\newacronym{unb}{UnB}{Universidade de Brasilia}
-\newacronym{iag}{IAG}{Instituto de Astronomia, Geofísica e Ciências Atmosféricas}
-\newacronym{usp}{USP}{Universidade de São Paulo}
-\newacronym{ipt}{IPT}{Instituto de Pesquisas Tecnológicas}
-\newacronym{unesp}{UNESP}{Universidade Estatual Paulista}
-\newacronym{ufrn}{UFRN}{Universidade Federal do Rio Grande do Norte}
+\newacronym{obsis}{ObSis}{Seismological Observatory}
+\newacronym{unb}{UnB}{Brasilia University}
+\newacronym{iag}{IAG}{Institute of Astronomy, Geophysics and Atmosferical Sciences}
+\newacronym{usp}{USP}{São Paulo University}
+\newacronym{ipt}{IPT}{Institute of Technological Research}
+\newacronym{unesp}{UNESP}{São Paulo State University}
+\newacronym{ufrn}{UFRN}{Federal University of Rio Grande do Norte}
 
-\newacronym{bsb}{BSB}{Boletim Sísmico Brasileiro}
-\newacronym{csv}{CSV}{Valores Separados por Vírgulas}
+\newacronym{bsb}{BSB}{Brazilian Seismic Bulletin}
+\newacronym{csv}{CSV}{Common separated values}
 
 \newacronym{bsb2013}{{BSB-2013.08}}{Brazilian Seismic Bulletin version 2013.08}
-\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{Catálogo ISC-GEM para a América do
-Sul}
+\newacronym{bsb2014}{{BSB-2014.11}}{Brazilian Seismic Bulletin version 2011.11}
+\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{ISC-GEM catalog for South America}

### 2015-02-03 manuscript.tex
minnor corrections, references and symbol translations

--- 2015-01-19 manuscript.tex
@@
 % Author names in capital letters:
-\authorrunninghead{PIRCHINER M.}
+\authorrunninghead{PIRCHINER M. and ASSUMP\c{C}\~AO M.}
 
@@
 
-Three smoothing seismicity techniques are reviewed and their application tothe Brazilian seismicity and a seismic hazard assesment are discussed.
-
-All these smoothing methods are based on a kernel density estimation. The main difference between them is the bandwidth selection process.
+Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assesment are discussed.
+
+All these smoothing methods are based on a \gls{kde}. The main difference between them is the bandwidth selection process.
 
@@
 \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.
+
 This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
@@
 
-The Brazilian seismic hazard building code (REFERENCE) [ABNT] was based on the GSHAP results for South America, which is the last published model for the whole region.
-Howerver, some significant seismic areas (Porto dos Gaúchos [REFERENCIA]), are not represented in the GPSHAP-based building codes.
-
-A significant number of earthquakes has been recorded in the last decades allowing a better definition of seismic zones. For this reason an updated hazard map is needed.
+The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{shedlock_tanner_1999}, which is the last published model for the whole region.
+Howerver, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.
+
+A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better definition of seismic zones. For this reason an updated hazard map is needed.
 
@@
 
-Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on kernel density estimation which results in a grid of smoothed seismicity.
-
-This kernel density estimation or the smoothed seismicity grid allow to define a grid of seismic point sources.
+Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.
+
+The seismicity \gls{kde} or the smoothed seismicity grid allow to define a grid of seismic point sources.
 The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.
 
-In the kernel density estimation both its shape and its bandwidth need to be defined.
-Most kernels are gaussian or power-laws in one of their variants.
+To perform the \gls{kde}, both its shape and bandwidth needs to be defined.
+Most kernels are gaussian or power-laws or one of their variants.
 The main distinction between the methods described here is the bandwidth selection process.
@@
 From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica.
-Since 1995, the Bulletin is maintained mainly by University of São Paulo under the same cooperation network in addition to Universidade Estadual Paulista (UNESP).
+Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP).
 Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.
@@
 Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
-REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS..
+REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS.
 (Johnston)
@@
 
-\subsection{Catalog check}
+\subsection{Catalog checking}
 
 Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour.
-Most events have no depth determination and the well determined depths (from pP or local networks) are mostly less than 10km.
-
-From the figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night.
+Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10km.
+
+Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night.
 This could suggest some influence of daily noise level.
 
-The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara [REFERENCE, TAKEYA et al 1989] and Porto dos Gaúchos [REFERENCE, BARROS 2009].
+The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara \citep{takeya_1989} and Porto dos Gaúchos \citep{barros_2009}.
 The increasing number of records after 1960 is due to the operation of global, regional and local stations.
@@
 
-To preserve the Poisson assumption about the independence of the events several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a.
-The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uental \citep{marsan_david_2012}
-(REVER REFERENCIA DO PROPRIO gruental ou CORRIGIN POIS NAO EH MARSAN DAVID mas Thomas van Stiphout, Jiancang Zhuang, and David Marsan].
-In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uental distance window.
+To preserve the Poissonian assumption about the independence of the events, several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a.
+The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uenthal \citep{van_stiphout_2012}.
+In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window.
 
@@
 
-Figure \ref{fig_occurrence} presents the magnitude and frequency distribution.
+Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
 A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.
@@
 
-- completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
-- se conhecia metodos para lidar com completeza espacial
-- também se conhece o método multiescala (provavel bom solucionador)
-- método de Stepp
-- mas usa-se uma completeness temporal simples.
+\begin{itemize}
+	\item completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
+	\item se conhecia metodos para lidar com completeza espacial
+	\item também se conhece o método multiescala (provavel bom solucionador)
+	\item conhece-se o método de Stepp
+	\item mas usa-se uma completeness temporal simples
+\end{itemize}
 
@@
 
-Arthur \citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
+\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
 \begin{equation}
@@
 
-The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection.
-These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$.
-These weights could easily consider space and time completenes and \emph{b}-value fluctuations.
-The expression used to compute it:
+The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection. This is the expression used to compute it:
 \begin{equation}
@@
 \end{equation}
-where \gls{sym:wi} is the \glsdesc{sym:wi} at location $\boldsymbol{r}$ on instant $t$,
+where \gls{sym:wi} is the \glsdesc{sym:wi} factor computed on location $\boldsymbol{r}$ on instant $t$,
 	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
@@
 
+These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$.
+They also could easily take into account the space and time completenes and \emph{b}-value fluctuations.
+
 \subsubsection{Local bandwidth computation}
 
-The method implemented to compute the space and time bandwidths for each earthquake was the Couppled-Nearest-Neighbour \citep{helmstetter_2012}:
+The method implemented to compute the space and time bandwidths for each earthquake was the \gls{cnn} \citep{helmstetter_2012} expressed as:
 \begin{equation}
@@
 	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
-	 \gls{sym:dk} is the \glsdesc{sym:dk} e
+	 \gls{sym:dk} is the \glsdesc{sym:dk} and
 	 \gls{sym:hk} is the \glsdesc{sym:hk}.
 
-The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.
+The bandwidths are defined locally by this simple optimization process. It could be small on high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.
 
 \subsubsection{Stationary seismic rate}
-After compute the model parameters and completely define it, the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ have to be computed.
-The way to do it proposed by \citet{helmstetter_2012} was to get the median
-on $\boldsymbol{r}_0$ over all considered time window:
+After compute the model parameters it is completely defined, and the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ can be computed on the way proposed by \citet{helmstetter_2012} taking the median value for some location $\boldsymbol{r}_0$ over all considered time window:
 \begin{equation}
@@
 
-The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures.
+The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures. This is one of the most important achievement of this methodology.
 
@@
 To optimize these parameters the catalog is divided in two parts: one for learning about the parameters and other to test them.
-
-If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the consider time frame is
+The target catalog will able us to check the model prediction performance derived from its chosen parameters.
+The best model parameters will maximize the prediction model capacity likely the target catalog.
+
+If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
 \begin{equation}
@@
 
-Then, over all cells, the (log) likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed on testing catalog is written as
+Then, over all cells, the log-likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed (on testing catalog) is written as
 \begin{equation}
@@
 \begin{acknowledgments}
-The author thanks specially to Marcelo Assumpção for the support and critical discussion as well all IAG-USP seismological team.
-It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
+The author thanks specially to 
+%Marcelo Assumpção for the support and critical discussion as well 
+all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
 \end{acknowledgments}
@@
 		\footnotesize(b)
-		& \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2013_pp_decluster}
+        & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
 		\end{tabular}
@@
 \begin{figure}
-	\includegraphics[width=1.0\linewidth]{z_img_hmtk_bsb2014_11_completeness}
+    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
+    \caption{Completeness}
+    \label{fig_completeness}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
     \caption{Annual earthquake recurrence.}
@@
 % TODO add HI_DI histograms
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+		& \includegraphics[width=0.45\textwidth]{z_img_hi_histogram}
+		\\
+		\footnotesize(b)
+        & \includegraphics[width=0.45\textwidth]{z_img_di_histogram}
+		\end{tabular}
+	\end{table}
+	\caption{(a) $h_i$ and (b) $d_i$ distributions from \gls{bsb}.}
+	\label{fig_hidi_histogram}
+	\end{center}
+\end{figure}
 

### 2015-02-03 manuscript.tex
fixing typos and cleaning for now

@@
 % Define a new glossary type
-\newglossary[slg,toc]{symbols}{sym}{sbl}{Lista de Símbolos}
+\newglossary[slg,toc]{symbols}{sym}{sbl}{Symbols List}
 \newglossary[elg]{equations}{eqn}{eql}{Equations}
@@
 % Corresponding author mailing address and e-mail address:
-\authoraddr{Corresponding author: Marlon Pirchiner, Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
+\authoraddr{Corresponding author: 
+Marlon Pirchiner, 
+Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
 (marlon.pirchiner@iag.usp.br)}
@@
 
-\authors{Marlon Pirchiner\altaffilmark{1,2}}
+\authors{Marlon Pirchiner\altaffilmark{1,2} 
+		and Marcelo Assumpção\altaffilmark{1}
+		}
 
@@
 
-Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assesment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
-
-Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assesment are discussed.
+Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+
+Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assessment are discussed.
 
@@
 
-This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
-(not indiscriminately REFERENCE) as a grid of points seismic sources.
+This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
+(not indiscriminately REFERENCE PAGANI WEATHERILL) as a grid of points seismic sources.
 
@@
 
-All the hazard calculations was performed on OpenQuake open source software.
+All the hazard calculations was performed on OpenQuake \citep{pagani_2014} open source software.
 
@@
 The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{shedlock_tanner_1999}, which is the last published model for the whole region.
-Howerver, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.
+However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.
 
@@
 To perform the \gls{kde}, both its shape and bandwidth needs to be defined.
-Most kernels are gaussian or power-laws or one of their variants.
+Most kernels are Gaussian or power-laws or one of their variants.
 The main distinction between the methods described here is the bandwidth selection process.
 
-\citet{frankel_1995} smoothing method applies a gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
-\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbour mean distance in magnitude bins.
+\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
+\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins.
 \citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth, based on kernel estimations for the background seismicity rate in long-term forecasts on both space and time dimensions.
@@
 
-\subsection{Specifical Purpose}
+\subsection{Specific Purpose}
 
@@
 
-\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
+\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
 \begin{equation}
@@
 
-Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to charachterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.
+Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to characterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.
 
@@
 
-\subsubsection{Maximum likelyhood optimization}
+\subsubsection{Maximum likelihood optimization}
 
@@
 
-If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
+If the earthquake occurrence could be modeled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
 \begin{equation}
@@
 
-Referencing equation.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Referencing Table \ref{fig_seismicity}.
-Referencing Figure \ref{placeholder}.
+.
 
@@
 
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-\subsubsection{Simulation 2}
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
-
-\subsection{Real Data}
-
-Aliquam interdum pellentesque scelerisque.
-Sed tincidunt suscipit purus, id aliquet nulla vehicula quis.
-Duis sed nisl lorem.
-Vivamus erat ante, dignissim et aliquam vel, adipiscing vitae magna.
-Cras id dapibus metus.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Proin ut lectus ut nisi congue ullamcorper.
-Ut ac turpis ligula.
-Sed faucibus bibendum nunc eget gravida.
+.
 
@@
 
-Nam fermentum sapien at enim varius consectetur.
-Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae.
-Integer lacinia purus vel metus tempus suscipit.
-Curabitur ac sapien quis mauris euismod commodo.
-Sed pharetra sem elit.
-Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor.
-Donec convallis ligula vitae sem viverra fermentum.
-Mauris in ullamcorper erat.
-Donec ultrices tempus nibh quis vestibulum.
-
-Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus.
-Aliquam a tellus nec felis sagittis consequat.
-Quisque convallis imperdiet neque a tempor.
-Nulla non erat urna.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Suspendisse viverra tortor pretium orci lacinia eleifend.
-Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem.
-Vestibulum pharetra malesuada egestas.
-Sed non orci sit amet erat suscipit fringilla in et diam.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla.
-Aenean porta rhoncus suscipit.
-
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
+.
 
@@
 
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
+.
 

### 2015-02-04 process/completeness.py
completeness code

@@ -126,4 +126,4 @@ completeness_table = completeness_algorithm.completeness(catalogue_uh,
                                         _config,
                                         saveplot="cumulative_rate_plots")
 
-#print completeness_table
+print completeness_table

### 2015-02-04 process/figure_kernels.py
kernel picture

@@ -1,0 +1,75 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+#import matplotlib.ticker as mtick
+from matplotlib.ticker import FuncFormatter
+
+def to_percent(y, position):
+    # Ignore the passed in position. This has the effect of scaling the default
+    # tick locations.
+    s = "%.2f"%(100 * y)
+    # The percent symbol needs escaping in latex
+    if matplotlib.rcParams['text.usetex'] == True:
+        return s + r'$\%$'
+    else:
+        return s + '%'
+
+
+# kernel functions
+def h(m, p):
+	return p[0]*np.exp(m*p[1])
+
+
+def kernel_kagan_knopoff(r, m, p, aw=1.5):
+	h2 = h(m, p)
+	r2 = r*r
+
+	factor1 = (aw - 1)/np.pi/h2
+	factor2 = (1 + (r2/h2))**(-aw)
+	return factor1 * factor2
+
+
+def kernel_kagan_knopoff(r, m, p, aw=1.5):
+	h2 = h(m, p)
+	r2 = r*r
+
+	factor1 = (aw - 1)/np.pi/h2
+	factor2 = (1 + (r2/h2))**(-aw)
+	return factor1 * factor2
+
+
+
+# plot setup
+parameters = [1.39, 1.18]
+distances = np.linspace(-50, 50, 200)
+magnitudes = [3, 4, 5, 6]
+colors= ['g', 'b', 'r', 'orange']
+
+f = plt.figure(figsize=(10,12))
+
+for i, m in enumerate(magnitudes):
+	values = kernel_kagan_knopoff(
+		distances,
+		m,
+		parameters)
+
+	plt.plot(distances, values, c=colors[i],
+		linewidth=2, 
+		#marker=None,
+		#marker='none', 
+		label="mag=%.1f"%m)
+
+plt.xlim((-50,50))
+plt.xlabel("distances [km]")
+plt.ylabel("density")
+plt.title("Kagan-Knopoff kernels for Brazilian bandwidths")
+plt.legend()
+
+# format percentual axis
+# fmt = '%.2f%%' # Format you want the ticks, e.g. '40%'
+# ticks = mtick.FormatStrFormatter(fmt)
+formatter = FuncFormatter(to_percent)
+plt.gca().yaxis.set_major_formatter(formatter)
+
+
+plt.show()

### 2015-02-04 manuscript.tex
lastest version

@@
 
-The Last two methods was implemented and are available as a free software on the Hazard Modeling Toolkit (HMTK) from GEM Foundation.
-
-All the hazard calculations was performed on OpenQuake \citep{pagani_2014} open source software.
+The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.
+
+All the hazard calculations was performed on OpenQuake \citep{pagani_2014} open source software also from GEM Foundation.
 
@@
 In the case of Brazilian catalog the methods did not present large differences at the final number of earthquakes (fig.
-\ref{fig_decluster}a) and the Gardner-Knopof/Uhrhammer combination was chosen to maximize the number of events.
+\ref{fig_decluster}a) and the Gardner-Knopoff/Uhrhammer combination was chosen to maximize the number of events.
 
@@
 The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}.
-To optimize these parameters the catalog is divided in two parts: one for learning about the parameters and other to test them.
-The target catalog will able us to check the model prediction performance derived from its chosen parameters.
+To optimize these parameters the catalog is divided in two parts: one for \emph{learning} about the parameters and other to test them.
+The \emph{testing} catalog will able us to check the model prediction performance derived from its chosen parameters.
 The best model parameters will maximize the prediction model capacity likely the target catalog.
@@
 
-Then, over all cells, the log-likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed (on testing catalog) is written as
+Then, over all cells, the log-likelihood, to be maximized, between model prediction (from the learning catalog) and the observed earthquakes (on the testing catalog) is written as
 \begin{equation}
@@
 \begin{figure}
+	\includegraphics[width=0.4\linewidth]{z_img_woo_kernels}
+	\caption{\citet{kagan_knopoff_1980} kernel shapes for some magnitude values.}
+	\label{fig_woo_kernels}
+\end{figure}
+
+\begin{figure}
 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
@@
 	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
-	\caption{Learning and target catalogs.}
+	\caption{Learning and testing catalogs.}
 	\label{fig_helmstetter_catalogues}

### 2015-02-05 manuscript.tex
bibliography update

@@
 
-Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984}, 2000, Assumpção in \citep{talwani_2014} (REFERENCE), Brazil has no seismic hazard map at national scale.
+Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984, berrocal_1996, assumpcao_et_al_2014}, Brazil has no seismic hazard map at national scale.
 
@@
 
-\subsection{Previous Smoothing Seismicity (Frankel, Woo, Helmstetter\ldots)}
+\subsection{Smoothing Seismicity}
 

### 2015-02-05 process/completeness.py

@@ -42,7 +42,7 @@ config = {
 #    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
     'catalog': project_home + "/data/catalogs/hmtk_bsb2014.11.csv",
     'declustered_catalog': project_home + "/"+_model+"/catalogs/bsb2014.11_declustered_catalog.csv",
-    'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
+    #'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
     'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
     # BR_Geral
     'fixed_completeness1' : np.array([[2014., 3.5],

### 2015-02-06 manuscript.tex
adding completeness biblio, text and figures

@@
 
-The number and quality of stations increased over time. The population density, previously concentrated along the coast increased in other regions.
-
-% and the spatial occupation respects the population density coverage concentrated in the east coast in most cases with exception only in the central part of the country by the Brasilia Seismic Station (REFERENCE).
-
-The histories of most regional stations are not easy available which hinders application of methodologies, such as \citet{mignan_2011}.
-For this reason, spatial completeness was not considered and temporal completeness parameters for the whole country was defined by expert criteria (figure FIGURA COMPLETENESS )
-
-\begin{itemize}
-	\item completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
-	\item se conhecia metodos para lidar com completeza espacial
-	\item também se conhece o método multiescala (provavel bom solucionador)
-	\item conhece-se o método de Stepp
-	\item mas usa-se uma completeness temporal simples
-\end{itemize}
+Brazil just recently was installed its permanent seismic network \citep{pirchiner_2011}.
+The number and quality of stations increased over time.
+The population density, previously concentrated along the coast increased in other regions changing the historical detectability.
+
+There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveforms \citep{schorlemmer_2008}, but the time and the effort to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance, was not intended to be done on the scope of this paper.
+
+The waveforms from the earliest years was not in digital form.
+The histories of most regional (temporary) stations are not easy available.
+In addition, the low number of earthquakes in the catalog also difficults some high resolution statistical methods.
+All of this hinders the direct application of most that methodologies.
+
+For this reason, the spatial completeness was not considered and the temporal completeness parameters for the whole country was defined by the simplest expert criteria.
+
+Figure COMPLETENESS shows the magnitude time distribution and the used completeness table. We chosen just one time and magnitude cut.
 
@@
 \begin{figure}
-    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
-    \caption{Completeness}
+	\begin{center}
+	\begin{table}[H]
+		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
+		\begin{tabular}{ c c }
+		\footnotesize(a)
+        & \includegraphics[width=0.45\textwidth]{z_img_completeness_temporal_stepp}
+		\\
+		\footnotesize(b)
+		& \includegraphics[width=0.45\textwidth]{z_img_completeness_temporal_scatter}
+		\end{tabular}
+	\end{table}
+	\caption{Completeness evaluation. 
+	Magnitude-Time scatter plot (a) and the Stepp plot (b).}
 	\label{fig_completeness}
+	\end{center}
 \end{figure}
@@
 
+\begin{table}[H]
+	\caption{Temporal completeness comparision. For each magnitude interval,
+	the completess was computed by the \citet{stepp_1972} method and the respective assigned value.}
+	\begin{tabular}{c c c}
+		\hline
+	\textbf{magnitude} & \textbf{year (Stepp)} & \textbf{year (simplest)} \\
+		\hline
+		2.0 & 1982 & 1885 \\
+		2.5 & 1980 & 1965 \\
+		3.0 & 1976 &  \\
+		3.5 & 1969 & 1960 \\
+		4.0 & 1955 &  \\
+		4.5 &  & 1900 \\
+		5.0 & 1952 &  \\
+		5.5 & 1927 &  \\
+		6.0 &  &  \\
+		6.5 &  &  \\
+		\hline
+	\end{tabular}
+	\label{tab:completeness}
+\end{table}
+
 \end{document}

### 2015-02-06 process/completeness.py

@@ -1,129 +1,0 @@
-# -*- coding: utf-8 -*-
-# Python Numerical and Plotting Libraries
-import numpy as np
-import matplotlib.pyplot as plt
-from copy import deepcopy
-
-# HMTK Catalogue Import/Export Libraries
-from hmtk.parsers.catalogue.csv_catalogue_parser import CsvCatalogueParser, CsvCatalogueWriter
-
-
-# HMTK Declustering Tools
-from hmtk.seismicity.declusterer.dec_afteran import Afteran
-from hmtk.seismicity.declusterer.dec_gardner_knopoff import GardnerKnopoffType1
-from hmtk.seismicity.declusterer.distance_time_windows import (GardnerKnopoffWindow,
-                                                               GruenthalWindow,
-                                                               UhrhammerWindow)
-
-# HMTK Completeness Tools
-from hmtk.seismicity.completeness.comp_stepp_1971 import Stepp1971
-
-from hmtk.plotting.seismicity.completeness.cumulative_rate_analysis import SimpleCumulativeRate
-
-
-# HMTK Plotting Tools
-from hmtk.plotting.mapping import HMTKBaseMap
-from hmtk.plotting.seismicity.catalogue_plots import (plot_depth_histogram,
-                                                      plot_magnitude_time_scatter,
-                                                      plot_magnitude_time_density,
-                                                      plot_magnitude_depth_density,
-                                                      plot_rate,
-                                                      plot_weekday_histogram,
-                                                      plot_hour_histogram,
-                                                      plot_observed_recurrence)
-
-
-print 'Imports OK!'
-
-project_home = "/Users/pirchiner/dev/pshab_source_models"
-_model = "background"
-
-config = {
-#    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
-    'catalog': project_home + "/data/catalogs/hmtk_bsb2014.11.csv",
-    'declustered_catalog': project_home + "/"+_model+"/catalogs/bsb2014.11_declustered_catalog.csv",
-    #'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
-    'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
-    # BR_Geral
-    'fixed_completeness1' : np.array([[2014., 3.5],
[…]

### 2015-02-06 process/figure_completeness_temporal.py

@@ -1,0 +1,174 @@
+# -*- coding: utf-8 -*-
+# Python Numerical and Plotting Libraries
+import numpy as np
+import matplotlib.pyplot as plt
+from copy import deepcopy
+
+# HMTK Catalogue Import/Export Libraries
+from hmtk.parsers.catalogue.csv_catalogue_parser import CsvCatalogueParser, CsvCatalogueWriter
+
+
+# HMTK Declustering Tools
+from hmtk.seismicity.declusterer.dec_afteran import Afteran
+from hmtk.seismicity.declusterer.dec_gardner_knopoff import GardnerKnopoffType1
+from hmtk.seismicity.declusterer.distance_time_windows import (GardnerKnopoffWindow,
+                                                               GruenthalWindow,
+                                                               UhrhammerWindow)
+
+# HMTK Completeness Tools
+from hmtk.seismicity.completeness.comp_stepp_1971 import Stepp1971
+
+from hmtk.plotting.seismicity.completeness.cumulative_rate_analysis import SimpleCumulativeRate
+from hmtk.plotting.seismicity.completeness.plot_stepp_1972 import create_stepp_plot
+
+# HMTK Plotting Tools
+from hmtk.plotting.mapping import HMTKBaseMap
+from hmtk.plotting.seismicity.catalogue_plots import (plot_depth_histogram,
+                                                      plot_magnitude_time_scatter,
+                                                      plot_magnitude_time_density,
+                                                      plot_magnitude_depth_density,
+                                                      plot_rate,
+                                                      plot_weekday_histogram,
+                                                      plot_hour_histogram,
+                                                      plot_observed_recurrence)
+
+
+print 'Imports OK!'
+
+project_home = "/Users/pirchiner/dev/pshab_source_models"
+_model = "background"
+
+config = {
+#    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
+    'catalog': project_home + "/data/catalogs/hmtk_bsb2014.11.csv",
+    'declustered_catalog': project_home + "/"+_model+"/catalogs/bsb2014.11_declustered_catalog.csv",
+    #'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
[…]

### 2015-02-06 manuscript.tex
add dayplot ref
completeness and introduction almost finished... figure and bibliography update

@@
 
-\subsection{Intraplate Seismicity}
-
-Onononono on on ono on on on on ononon on on on on on ono non on ono on on
-
-\subsection{Previous Brazilian Studies}
+\subsection{Intraplate seismicity}
+
+COMENTAR CAPITULO DO LIVRO
+
+\subsection{Previous studies in Brazil}
 
@@
 
-The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{shedlock_tanner_1999}, which is the last published model for the whole region.
+The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region.
 However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.
@@
 
-\subsection{Smoothing Seismicity}
+\subsection{Smoothed seismicity}
 
@@
 
-\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate.
+\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate smoothing the 2D histogram counting.
+This version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1}.
+There is also a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which sum the contribution of each earthquake kernel on given cell.
 \citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins.
@@
 
-Using one-year forecast and assuming time-independence of the seismicity rate, the same assumption of PSHA, it is possible to use these seismicity values on seismic sources characterization.
-(REVER E COMENTAR)
+Using the background seismic rate computed to the forecast and assuming time-independence of that, the same assumption of PSHA, it is possible, preserving their differences \citep{marzocchi_2011}, to use these forecasted seismicity values on seismic sources characterization \citep{weatherill_pagani_2014}.
 
@@
 
-To perform the hazard computation, the OpenQuake suite was chosen.
-The modelling tool was the HMTK also available by GEM scientific board.
+To perform the hazard computation, the OpenQuake suite \citep{pagani_2014} was chosen.
+The modelling tool was the \gls{hmtk} also open available by \gls{gem} scientific board.
 The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.
@@
 
-The Brazilian seismicity data is presented including some overview plots.
+\subsection{Specific purposes}
+
+The specific purposes of this paper are: (i) review 3 distincts smoothing/zoneless methods to describe spatially the seismicity rate, (ii) update the seismic hazard studies at national scale by proposing a new hazard model,
+(iii) use the GEM-OpenQuake tools to perform hazard calculation and (iv) add and share with the community two more new smoothing methods into the \gls{hmtk}.
+
+To aim this, the Brazilian seismicity data is presented including some overview plots.
 Next the method's theory is presented follow by the specific decisions and optimizations performed on each method modelling.
 And last, the results, conclusions and further considerations are discussed.
-
-\subsection{Specific Purpose}
 
@@
 
-Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour.
+Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour following \citet{gulia_2012}.
 Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10km.
@@
 
-There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveforms \citep{schorlemmer_2008}, but the time and the effort to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance, was not intended to be done on the scope of this paper.
+There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, stucchi_2004, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveforms \citep{schorlemmer_2008}, but the time and the effort to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance, was not intended to be done on the scope of this paper.
 
@@
 
-For this reason, the spatial completeness was not considered and the temporal completeness parameters for the whole country was defined by the simplest expert criteria.
-
-Figure COMPLETENESS shows the magnitude time distribution and the used completeness table. We chosen just one time and magnitude cut.
+For this reason, the spatial completeness was not considered and the temporal completeness parameters for the whole country was defined by an easy cumulative rate criteria using historical information about the network evolution and the \citet{stepp_1972} plot (fig.
+\ref{fig_completeness}a).
+TALK ABOUT BRASILIA STATION???
+
+Figure \ref{fig_completeness}b shows the magnitude time distribution in a scatter plot as the cumulative number of events which is important to perform the \gls{tcef} \citep{nasir_2013}.
+The linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend.
+Figure \ref{fig_completeness}b also show three (from many) possible completeness tables to allow quick check their distinctions.
+We chosen the ``simplest'' one.
+
+IMPROVE THIS DECISION EXPLANATION
 
@@
 
-\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
+\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a ``correlation distance'' $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
 \begin{equation}
@@
 \begin{figure}
-	\includegraphics[width=0.5\linewidth]{z_img_seismicity}
+	\includegraphics[width=0.99\linewidth]{z_img_seismicity}
 	\caption{Brazilian seismicity from 1767 to 2014. Source: \citep{bsb_2014}.}
@@
 		\footnotesize(a)
-		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_rate}
+		& \includegraphics[width=0.99\textwidth]{z_img_hmtk_bsb2014_11_rate}
 		\\
 		\footnotesize(b)
-		& \includegraphics[width=0.40\textwidth]{z_img_time_mag_count_br_1960}
+		& \includegraphics[width=0.90\textwidth]{z_img_time_mag_count_br_1960}
 		\end{tabular}
@@
 		\footnotesize(a)
-		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
+		& \includegraphics[width=0.70\textwidth]{z_img_hmtk_bsb2014_11_decluster}
 		\\
 		\footnotesize(b)
-        & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
+        & \includegraphics[width=0.60\textwidth]{z_img_hmtk_bsb2014_11_clusters}
 		\end{tabular}
 	\end{table}
-	\caption{Decluster evaluation (a) and the clusters map (b).}
+	\caption{Decluster evaluation (a) using distinct algorithms and windows.
+There is no significant differences between final results.
+The map (b) of clusters from selected method shows places where earthquake sequences tends to occur more frequently.}
 	\label{fig_decluster}
@@
 		\footnotesize(a)
-        & \includegraphics[width=0.45\textwidth]{z_img_completeness_temporal_stepp}
+        & \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_stepp}
 		\\
 		\footnotesize(b)
-		& \includegraphics[width=0.45\textwidth]{z_img_completeness_temporal_scatter}
+		& \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_scatter}
 		\end{tabular}
@@
 	\caption{Completeness evaluation. 
-	Magnitude-Time scatter plot (a) and the Stepp plot (b).}
+	The Stepp plot (a) was made using $\Delta_{mag}=0.5$ and $\Delta_{time}=2$ as magnitude and time bins.
+The magnitude-time scatter plot (b) shows in addition, three diferent temporal completess models and some cumulative rates to better perform a visual analysis.}
 	\label{fig_completeness}
@@
 \begin{figure}
-    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
+    \includegraphics[width=.99\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
     \caption{Annual earthquake recurrence.}
@@
 \begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_woo_bandwidth}
+	\includegraphics[width=0.99\linewidth]{z_img_woo_bandwidth}
 	\caption{Magnitude dependence bandwidth.}
@@
 \begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_woo_kernels}
+	\includegraphics[width=0.99\linewidth]{z_img_woo_kernels}
 	\caption{\citet{kagan_knopoff_1980} kernel shapes for some magnitude values.}
@@
 \begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_hidi}
+	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_hidi}
 	\caption{Local bandwidth example.}
@@
 		\footnotesize(a)
-		& \includegraphics[width=0.45\textwidth]{z_img_hi_histogram}
+		& \includegraphics[width=0.99\textwidth]{z_img_hi_histogram}
 		\\
 		\footnotesize(b)
-        & \includegraphics[width=0.45\textwidth]{z_img_di_histogram}
+        & \includegraphics[width=0.99\textwidth]{z_img_di_histogram}
 		\end{tabular}
@@
 \begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_stationary_a}
+	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_stationary_a}
 	\caption{Stationary seismic rate}
@@
 \begin{figure}
-	\includegraphics[width=0.4\linewidth]{z_img_helmstetter_catalogues}
+	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_catalogues}
 	\caption{Learning and testing catalogs.}
@@
 		4.0 & 1955 	&  \\
-		4.5 &  & 1900 \\
+		4.5 &  		& 1935 \\
 		5.0 & 1952 	&  \\

### 2015-02-06 process/figure_completeness_temporal.py

@@ -46,16 +46,16 @@ config = {
     'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
     # BR_Geral
     'fixed_completeness1' : np.array([[2014., 3.5],
-                                   [1980., 4.5],
-                                   [1968., 5.0],
-                                   [1962., 6.0],
-                                   [1940., 7.0]]),
+                                     [1980., 4.5],
+                                     [1968., 5.0],
+                                     [1962., 6.0],
+                                     [1940., 6.5]]),
 
     # pirchiner_simplificado
     'fixed_completeness2' : np.array([[1985., 2.0],
                                       [1975., 3.0],
                                       [1960., 4.0],
-                                      [1900., 6.0]]),
+                                      [1935., 6.5]]),
 
 }
 
@@ -131,6 +131,7 @@ completeness_table = completeness_algorithm.completeness(catalogue_uh,
 completeness_table_stepp = completeness_table
 print completeness_table
 print config['fixed_completeness2']
+
 plot_magnitude_time_scatter(catalogue_uh,
   # completeness_table=config['fixed_completeness2'],
   # completeness_table=completeness_table,
@@ -143,22 +144,97 @@ plot_magnitude_time_scatter(catalogue_uh,
   figsize=(9, 4),
   overlay=True)
 
-completeness_table=config['fixed_completeness2']
-plt.step(completeness_table[:,0], completeness_table[:,1], 
+completeness_table = config['fixed_completeness2']
+plt.step(completeness_table[:, 0], completeness_table[:, 1], 
   where='post', 
   linewidth=4,
   color='Navy',
   label="Simplified")
 
+completeness_table = config['fixed_completeness1']
+plt.step(completeness_table[:, 0], completeness_table[:, 1],
+  where='post',
+  linewidth=2,
+  color='darkgreen',
+  label=u"Assumpção")
+
 completeness_table=completeness_table_stepp
-plt.step(completeness_table[:,0], completeness_table[:,1], 
[…]

### 2015-02-10 manuscript.tex
completeness figure and map + clean abstract and review intro

@@
 
-\documentclass[draft, grl]{agutex}
-%\documentclass[grl]{agutex}
+% \documentclass[draft, grl]{agutex}
+\documentclass[grl]{agutex}
 
@@
 \begin{abstract}
-
 Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+% Intraplate seismicity is a hard problem to solve in tectonics.
+In general, the earthquake occurrence is lower than others ``active'' regions.
+In the Brazilian case, there is a lack of observed seismicity due the short (in geological scale) observation timeframe.
+In addition, the populational density decreases towards to interior of the country and changes the reporting capabilities historicaly.
+
+% This diffused intraplate seismicity pattern is often modeled as polygons representing different seismic zones or area sources for \gls{psha} purposes.
+Their geometries are defined by expert opinions.
+
+% Other possible approach to deal this is develop zoneless or smoothed  models for the seismicity, conside
 
@@
 
-This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
-(not indiscriminately REFERENCE PAGANI WEATHERILL) as a grid of points seismic sources.
+This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast which time-independence assumption able us to use this recurrence rate to define a grid of seismic point sources as the input source model to compute the hazard.
 
@@
 
-The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.
-
-All the hazard calculations was performed on OpenQuake \citep{pagani_2014} open source software also from GEM Foundation.
+% The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.
+
+All the hazard computations was performed on OpenQuake \citep{pagani_2014} open source software also from GEM Foundation.
 
@@
 The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region.
-However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.
-
-A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better definition of seismic zones. For this reason an updated hazard map is needed.
+However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GSHAP-based building codes.
+
+A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better empirical and updated definition of the seismicity and hazard levels.
+
+% definition of seismic zones. For this reason an updated hazard map is needed.
 
@@
 
-Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.
+Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology often used to assess seismic hazard is using expert opinions to pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.
 

### 2015-02-10 process/figure_completeness_maps.py

@@ -1,0 +1,464 @@
+# -*- coding: utf-8 -*-
+# Python Numerical and Plotting Libraries
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+# from matplotlib.lines import mlines
+from copy import deepcopy
+
+# HMTK Catalogue Import/Export Libraries
+from hmtk.parsers.catalogue.csv_catalogue_parser import CsvCatalogueParser, CsvCatalogueWriter
+
+
+# HMTK Declustering Tools
+from hmtk.seismicity.declusterer.dec_afteran import Afteran
+from hmtk.seismicity.declusterer.dec_gardner_knopoff import GardnerKnopoffType1
+from hmtk.seismicity.declusterer.distance_time_windows import (GardnerKnopoffWindow,
+                                                               GruenthalWindow,
+                                                               UhrhammerWindow)
+
+# HMTK Completeness Tools
+from hmtk.seismicity.completeness.comp_stepp_1971 import Stepp1971
+
+from hmtk.plotting.seismicity.completeness.cumulative_rate_analysis import SimpleCumulativeRate
+from hmtk.plotting.seismicity.completeness.plot_stepp_1972 import create_stepp_plot
+
+# HMTK Plotting Tools
+from hmtk.plotting.mapping import HMTKBaseMap
+from hmtk.plotting.seismicity.catalogue_plots import (plot_depth_histogram,
+                                                      plot_magnitude_time_scatter,
+                                                      plot_magnitude_time_density,
+                                                      plot_magnitude_depth_density,
+                                                      plot_rate,
+                                                      plot_weekday_histogram,
+                                                      plot_hour_histogram,
+                                                      plot_observed_recurrence)
+
+
+print 'Imports OK!'
+
+project_home = "/Users/pirchiner/dev/pshab_source_models"
+_model = "background"
+
+config = {
+#    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
+    'catalog': project_home + "/data/catalogs/hmtk_bsb2014.11.csv",
+    'declustered_catalog': project_home + "/"+_model+"/catalogs/bsb2014.11_declustered_catalog.csv",
[…]

### 2015-02-10 process/figure_completeness_temporal.py

@@ -2,6 +2,8 @@
 # Python Numerical and Plotting Libraries
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+# from matplotlib.lines import mlines
 from copy import deepcopy
 
 # HMTK Catalogue Import/Export Libraries
@@ -45,17 +47,17 @@ config = {
     #'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
     'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
     # BR_Geral
-    'fixed_completeness1' : np.array([[2014., 3.5],
-                                     [1980., 4.5],
-                                     [1968., 5.0],
-                                     [1962., 6.0],
-                                     [1940., 6.5]]),
+    'fixed_completeness1' : np.array([[2014., 3.16],
+                                      [1980., 4.28],
+                                      [1968., 4.85],
+                                      [1962., 6.0],
+                                      [1940., 6.5]]),
 
     # pirchiner_simplificado
-    'fixed_completeness2' : np.array([[1985., 2.0],
-                                      [1975., 3.0],
-                                      [1960., 4.0],
-                                      [1935., 6.5]]),
+    'fixed_completeness2' : np.array([[2014., 3.0],
+                                      [1975., 3.8],
+                                      [1960., 4.8],
+                                      [1938., 6.5]]),
 
 }
 
@@ -71,7 +73,7 @@ catalogue.sort_catalogue_chronologically()
 print 'Catalogue sorted chronologically!'
 
 # 3.0 - 0.3 uncertainty
-min_mag = 2.0
+min_mag = 3.0
 # filter magnitudes Mw lower than 3.0
 _idx = catalogue.data['magnitude'] >= min_mag
 catalogue.purge_catalogue(_idx)
@@ -119,7 +121,7 @@ catalogue_uh.get_number_events()
 #print config['fixed_completeness']
 # completeness
 _config = {'magnitude_bin': 0.5,
-           'time_bin': 2.0,
+           'time_bin': 5.0,
[…]

### 2015-02-13 manuscript.tex
some figure changes, add histogram (counting) model uniform catalogs...

@@
 
-The seismicity \gls{kde} or the smoothed seismicity grid allow to define a grid of seismic point sources.
+The \gls{kde} (or smoothed) seismicity grid allow to define a grid of seismic point sources.
 The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.
@@
 	\begin{table}[H]
+		\begin{tabular}{ p{0.25\textwidth}
+						 p{0.25\textwidth}
+						 p{0.25\textwidth}
+						 p{0.25\textwidth} }
+		\footnotesize(a) &
+		\footnotesize(b) &
+		\footnotesize(c) &
+		\footnotesize(d)
+		\\
+		\includegraphics[height=4cm]{z_img_2dhist_assump_25}  &
+		\includegraphics[height=4cm]{z_img_f_assump_50_25} &
+		\includegraphics[height=4cm]{z_img_f_assump_150_025} &
+		\includegraphics[height=4cm]{z_img_woo_assump_25} 
+		\\
+		\footnotesize(e) &
+		\footnotesize(f) &
+		\footnotesize(g) &
+		\footnotesize(h)
+		\\
+		\includegraphics[height=4cm]{z_img_2dhist_stepp_25}  &
+		\includegraphics[height=4cm]{z_img_f_stepp_50_025} &
+		\includegraphics[height=4cm]{z_img_f_stepp_150_025} &
+		\includegraphics[height=4cm]{z_img_woo_stepp_25} 
+		\\
+		\footnotesize(i) &
+		\footnotesize(j) &
+		\footnotesize(l) &
+		\footnotesize(m)
+		\\
+		\includegraphics[height=4cm]{z_img_2dhist_assump_50}  &
+		\includegraphics[height=4cm]{z_img_f_assump_50_50} &
+		\includegraphics[height=4cm]{z_img_f_assump_150_050} &
+		\includegraphics[height=4cm]{z_img_woo_assump_50} 
+		\\
+		\footnotesize(n) &
+		\footnotesize(o) &
+		\footnotesize(p) &
+		\footnotesize(q)
+		\\
+		\includegraphics[height=4cm]{z_img_2dhist_stepp_50}  &
+		\includegraphics[height=4cm]{z_img_f_stepp_50_050} &
+		\includegraphics[height=4cm]{z_img_f_stepp_150_050} &
+		\includegraphics[height=4cm]{z_img_woo_stepp_50}
+
+		\end{tabular}
+	\end{table}
+	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
+	\label{fig_rates2}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+	\begin{table}[H]
 		\begin{tabular}{ p{0.33\textwidth}
@@
 		& 0.3 \\
-		Area felt $A_f$
+		Felt area$A_f$
 		&$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 0.00084\sqrt{A_f}$& 0.4 \\

### 2015-02-13 process/figure_completeness_maps.py

@@ -113,8 +113,8 @@ catalogue_uh.purge_catalogue(cluster_flags_uh == 0)
 # cluster_flags == 0: mainshocks
 
 print 'Gardner-Knopoff/UhrhammerWindow\tbefore: ', 
-catalogue.get_number_events(), " after: ",
-catalogue_uh.get_number_events()
+  catalogue.get_number_events(), " after: ",
+  catalogue_uh.get_number_events()
 
 #completeness_table =
 
@@ -135,13 +135,14 @@ print completeness_table
 print config['fixed_completeness2']
 
 
-map_dpi = 300
+map_dpi = 150
 add_geology = False
 add_sourcemodel = False
 savefig=True
 
 #map_title = 'Brazilian Seismic Zones'
 map_title = 'Earthquake epicenters on the BSB2014.11 catalog'
+map_title = 'Completeness spatial implications [BSB 2014.11]  '
 #map_title = 'ISC-GEM Catalogue'
 #map_title = 'South-American Lithology'
 

### 2015-02-13 process/figure_completeness_temporal.py

@@ -47,7 +47,12 @@ config = {
     #'source_model': project_home + "/"+_model+"/raw/"+_model+"_bsb2014.11_geometries.xml",
     'map_config': {'min_lon': -80.0, 'max_lon': -30.0, 'min_lat': -37.0, 'max_lat': 13.0, 'resolution':'l'},
     # BR_Geral
-    'fixed_completeness1' : np.array([[2014., 3.16],
+    'fixed_completeness1' : np.array([
+                                      # [1980., 3.16],
+                                      # [1968., 4.28],
+                                      # [1962., 4.85],
+                                      # [1940., 6.0]]),
+                                      [2014., 3.16],
                                       [1980., 4.28],
                                       [1968., 4.85],
                                       [1962., 6.0],
@@ -135,9 +140,25 @@ print completeness_table
 print config['fixed_completeness2']
 
 
+# _w = CsvCatalogueWriter("hmtk_bsb2014.11_uniform_stepp.csv")
+# _w.write_file(catalogue=catalogue_uh, 
+#   magnitude_table=completeness_table_stepp)
+
+# _w = CsvCatalogueWriter("hmtk_bsb2014.11_uniform_assump.csv")
+# _w.write_file(catalogue=catalogue_uh, 
+#   magnitude_table=config['fixed_completeness1'])
+
+
+# writing uniform catalogs
+# _c = deepcopy(catalogue_uh)
+# _c.catalogue_mt_filter(completeness_table_stepp, reverse=False)
+#_c.write_catalogue("hmtk_bsb2014.11_uniform_stepp.csv")
+
 _c = deepcopy(catalogue_uh)
 # _c.catalogue_mt_filter(config['fixed_completeness2'], reverse=True)
 _c.catalogue_mt_filter(config['fixed_completeness1'], reverse=False)
+#_c.write_catalogue("hmtk_bsb2014.11_uniform_assump.csv")
+
 
 plot_magnitude_time_scatter(_c,
   # completeness_table=config['fixed_completeness2'],

### 2015-02-13 process/figure_histogram2d_maps.py

@@ -1,0 +1,245 @@
+# -*- coding: utf-8 -*-
+# Python Numerical and Plotting Libraries
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+# from matplotlib.lines import mlines
+from copy import deepcopy
+
+# HMTK Catalogue Import/Export Libraries
+from hmtk.parsers.catalogue.csv_catalogue_parser import CsvCatalogueParser, CsvCatalogueWriter
+
+
+# HMTK Declustering Tools
+from hmtk.seismicity.declusterer.dec_afteran import Afteran
+from hmtk.seismicity.declusterer.dec_gardner_knopoff import GardnerKnopoffType1
+from hmtk.seismicity.declusterer.distance_time_windows import (GardnerKnopoffWindow,
+                                                               GruenthalWindow,
+                                                               UhrhammerWindow)
+
+# HMTK Completeness Tools
+from hmtk.seismicity.completeness.comp_stepp_1971 import Stepp1971
+
+from hmtk.plotting.seismicity.completeness.cumulative_rate_analysis import SimpleCumulativeRate
+from hmtk.plotting.seismicity.completeness.plot_stepp_1972 import create_stepp_plot
+
+# HMTK Plotting Tools
+from hmtk.plotting.mapping import HMTKBaseMap
+from hmtk.plotting.seismicity.catalogue_plots import (plot_depth_histogram,
+                                                      plot_magnitude_time_scatter,
+                                                      plot_magnitude_time_density,
+                                                      plot_magnitude_depth_density,
+                                                      plot_rate,
+                                                      plot_weekday_histogram,
+                                                      plot_hour_histogram,
+                                                      plot_observed_recurrence)
+
+
+print 'Imports OK!'
+
+project_home = "."
+_model = "background"
+
+config = {
+#    'catalog': project_home + "/data/catalogs/hmtk_bsb2013.csv",
+    'catalog': project_home + "/hmtk_bsb2014.11_uniform_assump.csv",
[…]

### 2015-02-19 manuscript.tex
add figures, change tables, pre-round03

@@
 
-% \documentclass[draft, grl]{agutex}
-\documentclass[grl]{agutex}
+\documentclass[draft, grl]{agutex}
+% \documentclass[grl]{agutex}
 
@@
 		\includegraphics[height=4cm]{z_img_f_assump_50_25} &
-		\includegraphics[height=4cm]{z_img_f_assump_150_025} &
+		\includegraphics[height=4cm]{z_img_f_assump_150_25} &
 		\includegraphics[height=4cm]{z_img_woo_assump_25} 
@@
 		\includegraphics[height=4cm]{z_img_2dhist_stepp_25}  &
-		\includegraphics[height=4cm]{z_img_f_stepp_50_025} &
-		\includegraphics[height=4cm]{z_img_f_stepp_150_025} &
+		\includegraphics[height=4cm]{z_img_f_stepp_50_25} &
+		\includegraphics[height=4cm]{z_img_f_stepp_150_25} &
 		\includegraphics[height=4cm]{z_img_woo_stepp_25} 
@@
 		\includegraphics[height=4cm]{z_img_f_assump_50_50} &
-		\includegraphics[height=4cm]{z_img_f_assump_150_050} &
+		\includegraphics[height=4cm]{z_img_f_assump_150_50} &
 		\includegraphics[height=4cm]{z_img_woo_assump_50} 
@@
 		\includegraphics[height=4cm]{z_img_2dhist_stepp_50}  &
-		\includegraphics[height=4cm]{z_img_f_stepp_50_050} &
-		\includegraphics[height=4cm]{z_img_f_stepp_150_050} &
+		\includegraphics[height=4cm]{z_img_f_stepp_50_50} &
+		\includegraphics[height=4cm]{z_img_f_stepp_150_50} &
 		\includegraphics[height=4cm]{z_img_woo_stepp_50}
@@
 \begin{table}[H]
-	\caption{Temporal completeness comparision. For each magnitude interval,
-	the completess was computed by the \citet{stepp_1972} method and the respective assigned value.}
-	\begin{tabular}{c c c}
+	\caption{Temporal completeness. For each magnitude interval, the completess was computed by the \citet{stepp_1972} method.}
+	\begin{tabular}{c c }
 		\hline
-	\textbf{magnitude} & \textbf{year (Stepp)} & \textbf{year (simplest)} \\
+	\textbf{magnitude} & \textbf{year (Stepp)} \\
 		\hline
-		2.0 & 1982 	& 1885 \\
-		2.5 & 1980 	& 1965 \\
-		3.0 & 1976 	&  \\
-		3.5 & 1969 	& 1960 \\
-		4.0 & 1955 	&  \\
-		4.5 &  		& 1935 \\
-		5.0 & 1952 	&  \\
-		5.5 & 1927 	&  \\
-		6.0 &  		&  \\
-		6.5 &  		&  \\
+		3.0 & 1970 \\
+		% 3.5 &  	   \\
+		4.0 & 1959 \\
+		4.5 & 1951 \\
+		% 5.0 &      \\
+		% 5.5 &  	   \\
+		6.0 & 1933 \\
+		\hline
+	\end{tabular}
+	\label{tab:completeness_stepp}
+\end{table}
+
+\begin{table}[H]
+	\caption{Temporal completeness. For each magnitude interval, the completess comes from \citet{assumpcao_et_al_2014} converted to $M_W$ from $m_b$ following table \ref{tab:guidelines}.}
+	\begin{tabular}{c c}
+		\hline
+	\textbf{magnitude} & \textbf{year (Assumpção)} \\
+		\hline
+		3.1 & 1980 \\
+		4.3 & 1968 \\
+		4.8 & 1962 \\
+		% 5.0 &      \\
+		% 5.5 &  	   \\
+		6.0 & 1940 \\
 		\hline

### 2015-05-13 manuscript.tex
rebuilding from intro adding marcelo's suggestion (to be reviewed)

@@
 
+% package used by \citep and \citet
+% \usepackage[sort&compress,square,comma,authoryear]{natbib}
+
+% makes color citations
+% \usepackage[dvips,dvipdfm,colorlinks=true,urlcolor=blue,citecolor=red,linkcolor=red,bookmarks=true]{hyperref}
+
 % To create numbered lines:
@@
 
-COMENTAR CAPITULO DO LIVRO
+Brazil occupies a large stable continental region with seismicity rate lower than Central and Eastern North America, Australia, India or Northern Europe \citep{giardini_1999,johnston_1996a}.
+Spanning most of the central region of the South American plate, Brazil has earthquakes above magnitude 6 mb with a return period of 50 years or so.
+Historical records are very short with almost no data before the XX century.
+Historical evidence of magnitude a 7 has been reported in the Amazon \citep{berrocal_1984,veloso_2014}.
+However, as in other intraplate settings, seismic hazard is not negligible due to the low seismic attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear plants \citep{almeida_etal_2013}.
+
+Very few studies related to seismic hazard, such as \citet{berrocal_1996} for SE Brazil, have been carried out so far and the first Brazilian seismic building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}.
+Brazilian seismicity is concentrated in several seismic zones \citep{berrocal_1984, assumpcao_et_al_2014}.
+Except for a peak of activity in Northeast Brazil, both the GSHAP model and the ABNT zoning have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 mb in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}.
+For this reason, it is important to review and update the hazard estimates for Brazil, especially since the increase in the number of recordings of Brazilian earthquakes in the last three decades.
+In addition, the results presented here are part of the \acrfull{gem} supported SARA project (South America Risk Assessment).
+
+Despite attempts to explain the origin of some of the seismic zones, such as stress concentrations from lithospheric thinning \citep{assumpcao_2004, azevedo_2015} or flexural effects \citep{zoback_1996, assumpcao_sacek_2013}, most of them have ill-defined boundaries.
+Attempts to produce seismic hazard maps \citep{dourado_2014} based on seismogenic zones are heavily dependent on subjective expert opinion, even more so in low seismic areas such as Brazil.
+As an important contribution to the new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.
 

### 2015-05-13 manuscript.tex
bibliography and abstract changes

@@
 \begin{abstract}
-Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+Intraplate seismicity is a hard problem to solve in tectonics.
+Brazil, as others intraplate environments, mainly during the lack of observed seismicity the seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
 % Intraplate seismicity is a hard problem to solve in tectonics.
@@
 % Other possible approach to deal this is develop zoneless or smoothed  models for the seismicity, conside
-
-Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assessment are discussed.
-
-All these smoothing methods are based on a \gls{kde}. The main difference between them is the bandwidth selection process.
-
-The method of \citet{frankel_1995} uses a \emph{fixed} bandwidth in space.
+Three smoothed seismicity techniques are reviewed and their application on construct models of Brazilian seismicity and its applications on long-term seismic hazard assessment in Brazil are also discussed.
+All these smoothing methods are based on a \gls{kde}.
+The main difference between them is the bandwidth selection process.
+
+The method proposed by \citet{frankel_1995} uses a \emph{fixed} bandwidth in space.
 \citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}.
-\citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.
-
-This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast which time-independence assumption able us to use this recurrence rate to define a grid of seismic point sources as the input source model to compute the hazard.
-
-The results shown that all methods clearly define some known seismic regions and could be used in the Brazilian seismicity.
-All methods improve the spatial resolution of the GSHAP model, the only previously available for the region.
-
-% The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.
-
-All the hazard computations was performed on OpenQuake \citep{pagani_2014} open source software also from GEM Foundation.
+\citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event as background (time-independent) model on their forecast experiments.
+
+All these methods compute the seismicity rate surface in a binned space, time and magnitude, which able us to use this occurrence rate to define a grid of seismogenic point sources as the input source model to the hazard computation.
+
+%The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.
+
+In addition, we used the information score and the error diagrams \citep{kagan_2009} to evaluate the models, optimize the smoothing distance on the \citet{frankel_1995} method, and finally give a weighted source models logic-tree to hazard computation.
+
+All the hazard computations was performed on \gls{oq} \citep{pagani_2014} open source software from \acrfull{gem} Foundation.
+Until the present moment, no focal mechanism information has been included or considered on these smoothed seismicity models.
+All hazard computations has been done using a single \gls{GMPE}.
+A more detailed study on \glspl{GMPE} selection should be developed further and is out of this paper scope.
+
+The results shown that all methods clearly most of the well known seismogenic regions and could be used in the Brazilian seismicity modeling.
+All methods improve the spatial resolution of the GSHAP hazard model, the only previously available for the region.
+These results should be considered on future reviews of Brazilian design building codes and procedures.
+%Tectonic connections between these seismicity models and stress release could be developed in the future.
 

### 2015-05-13 d03-acronyms.tex
fixing typos and cleaning for now
adding completeness biblio, text and figures
completeness and introduction almost finished... figure and bibliography update
start of introduction review

@@
 \newacronym{gr}{GR}{Gutenberg-Richter}
-\newacronym{cnn}{CNN}{Coupled-Nearest-Neighbour}
+\newacronym{cnn}{CNN}{Coupled-Nearest-Neighbor}
 \newacronym{va}{r.v.}{random variable}
-\newacronym{iid}{i.i.d.}{independent and identicaly distributed}
+\newacronym{iid}{i.i.d.}{independent and identically distributed}
 \newacronym{pdf}{density}{probability density function}
@@
 \newacronym{kde}{KDE}{Kernel Density Estimation}
+\newacronym{pde}{PDE}{Probability of Detecting an Earthquake}
+\newacronym{tcef}{TCEF}{temporal course of earthquake frequency}
 
@@
 \newacronym{unb}{UnB}{Brasilia University}
-\newacronym{iag}{IAG}{Institute of Astronomy, Geophysics and Atmosferical Sciences}
+\newacronym{iag}{IAG}{Institute of Astronomy, Geophysics and Atmospherical Sciences}
 \newacronym{usp}{USP}{São Paulo University}
@@
 \newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{ISC-GEM catalog for South America}
+
+\newacronym{relm}{RELM}{Regional Earthquake Likelihood Models}
+\newacronym{ed}{ED}{error diagrams}

### 2015-05-13 manuscript.tex
start of introduction review

@@
 Brazil, as others intraplate environments, mainly during the lack of observed seismicity the seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
-% Intraplate seismicity is a hard problem to solve in tectonics.
-In general, the earthquake occurrence is lower than others ``active'' regions.
-In the Brazilian case, there is a lack of observed seismicity due the short (in geological scale) observation timeframe.
-In addition, the populational density decreases towards to interior of the country and changes the reporting capabilities historicaly.
-
-% This diffused intraplate seismicity pattern is often modeled as polygons representing different seismic zones or area sources for \gls{psha} purposes.
-Their geometries are defined by expert opinions.
-
-% Other possible approach to deal this is develop zoneless or smoothed  models for the seismicity, conside
+
 Three smoothed seismicity techniques are reviewed and their application on construct models of Brazilian seismicity and its applications on long-term seismic hazard assessment in Brazil are also discussed.
@@
 All the hazard computations was performed on \gls{oq} \citep{pagani_2014} open source software from \acrfull{gem} Foundation.
-Until the present moment, no focal mechanism information has been included or considered on these smoothed seismicity models.
+No focal mechanism information has been included or considered on these smoothed seismicity models.
 All hazard computations has been done using a single \gls{GMPE}.
-A more detailed study on \glspl{GMPE} selection should be developed further and is out of this paper scope.
+A more detailed study on \glspl{GMPE} selection should be developed further and it is beyond the scope of this work.
 
@@
 These results should be considered on future reviews of Brazilian design building codes and procedures.
-%Tectonic connections between these seismicity models and stress release could be developed in the future.
 
@@
 
-\subsection{Intraplate seismicity}
+\subsection{Intraplate seismicity and previous studies in Brazil}
 
@@
 Historical evidence of magnitude a 7 has been reported in the Amazon \citep{berrocal_1984,veloso_2014}.
-However, as in other intraplate settings, seismic hazard is not negligible due to the low seismic attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear plants \citep{almeida_etal_2013}.
+However, as in other intraplate settings, seismic hazard is not negligible due to the low seismic attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.
 
@@
 For this reason, it is important to review and update the hazard estimates for Brazil, especially since the increase in the number of recordings of Brazilian earthquakes in the last three decades.
-In addition, the results presented here are part of the \acrfull{gem} supported SARA project (South America Risk Assessment).
+%In addition, the results presented here are part of the \acrfull{gem} supported SARA project (South America Risk Assessment).
 
@@
 Attempts to produce seismic hazard maps \citep{dourado_2014} based on seismogenic zones are heavily dependent on subjective expert opinion, even more so in low seismic areas such as Brazil.
-As an important contribution to the new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.
-
-\subsection{Previous studies in Brazil}
-
-Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984, berrocal_1996, assumpcao_et_al_2014}, Brazil has no seismic hazard map at national scale.
-
-The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region.
+As a contribution to the new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.
+
+%Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984, berrocal_1996, assumpcao_et_al_2014}, Brazil has no seismic hazard map at national scale.
+
+%The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region.
 However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GSHAP-based building codes.
 
-A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better empirical and updated definition of the seismicity and hazard levels.
+%A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better empirical and updated definition of the seismicity and hazard levels.
 
@@
 
-\subsection{Smoothed seismicity}
-
-Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology often used to assess seismic hazard is using expert opinions to pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.
-
-The \gls{kde} (or smoothed) seismicity grid allow to define a grid of seismic point sources.
-The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.
+\subsection{Smoothed seismicity models}
+
+Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to pre-defined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
+
+The (smoothed) seismicity \gls{kde} techniques allow us to define another grid of seismic point sources.
+The seismicity rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods.
 
 To perform the \gls{kde}, both its shape and bandwidth needs to be defined.
-Most kernels are Gaussian or power-laws or one of their variants.
+Most kernels are Gaussian, power-laws or one of their variants like Fisher kernels.
 The main distinction between the methods described here is the bandwidth selection process.
 
-\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate smoothing the 2D histogram counting.
-This version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1}.
-There is also a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which sum the contribution of each earthquake kernel on given cell.
+\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} kernel smooth technique with a \emph{fixed smoothing-distance} bandwidth to estimate the seismicity rate by smoothing the 2D histogram from earthquakes counting.
+The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which sum the contribution of each earthquake kernel on given cell.
+
 \citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins.
-\citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth, based on kernel estimations for the background seismicity rate in long-term forecasts on both space and time dimensions.
-
-Using the background seismic rate computed to the forecast and assuming time-independence of that, the same assumption of PSHA, it is possible, preserving their differences \citep{marzocchi_2011}, to use these forecasted seismicity values on seismic sources characterization \citep{weatherill_pagani_2014}.
+
+\citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth on both space and time dimensions for the background (time-independent) seismicity rate on their long-term forecasts experiments.
+
+Using the long-term background seismic rate computed for forecast and assuming its time-independence, the same \gls{psha} assumption, it is possible, preserving their purposes differences \citep{marzocchi_2011}, to use these background seismicity rate values on seismic sources characterization \citep{weatherill_pagani_2014}.
+
+\subsection{Information scores and \gls{ed}}
+
+From these three smoothing methods, only \cite{helmstetter_2012} since its formalization use optimizations to define their model parameters.
+All the other methods are concept to use the whole catalog and some fixed parameter (e.g.
+Frankel's smoothing distance).
+
+To compare the performance of these three models, the information scores and \gls{ed} \citep{molchan_1992, kagan_2007, kagan_2009} where computed as alternative to the tests suggested to the \gls{relm} forecasts experiments by \citet{schorlemmer_2007}.
+
+On the evaluation phase all models learned from earthquakes into 1970-2004 observation time and were tested against the target realizations occurred between 2004 and 2014-01-01.
+
+Scores from some Frankel-based models (from different smoothing distances) was computed and the best scores should give us an reasonable choice for this parameter.
+
+HYBRID DECISION APPROACH BIRD/JORDAN[2013?]
 
@@
 
-To perform the hazard computation, the OpenQuake suite \citep{pagani_2014} was chosen.
-The modelling tool was the \gls{hmtk} also open available by \gls{gem} scientific board.
-The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.
+To perform the hazard computation, the \acrfull{oq} suite \citep{pagani_2014} was chosen.
+The modeling tool was the \gls{hmtk} also open available by \gls{gem} scientific board.
+The Frankel method was already implemented on the toolkit and two others was implemented in the context of this paper and they are free available on a public repository.
+
+The other parameters, like \emph{b}-values, minimum and maximum magnitudes for the frequency-magnitude distributions, strike, dip and rate probabilities to the rupture plane, its top and bottom limits on the subsurface, magnitude-area scale relationships and also the depth distributions must be assigned in other ways to proceed the hazard computation and take advantages of the new developed \glspl{gmpe}.
+
+In this work, to keep our focus on seismicity models as smoothed seismic sources, we perform hazard defining a single \gls{gmpe}  [EXPOSE WHICH ONE]  for a stable continental crust tectonic environment.
 
@@
 
-The specific purposes of this paper are: (i) review 3 distincts smoothing/zoneless methods to describe spatially the seismicity rate, (ii) update the seismic hazard studies at national scale by proposing a new hazard model,
-(iii) use the GEM-OpenQuake tools to perform hazard calculation and (iv) add and share with the community two more new smoothing methods into the \gls{hmtk}.
+The specific purposes of this paper are:
+
+(i) review three distinct smoothing (zoneless) methods to spatially model the seismic rate,
+(ii) update the seismic hazard studies at national scale by proposing a new simple hazard model,
+(iii) use the \gls{gem}-\glsdesc{oq} tools to perform hazard calculation and 
+(iv) add and share with the community two more new smoothing methods into the \gls{hmtk}.
 
 To aim this, the Brazilian seismicity data is presented including some overview plots.
-Next the method's theory is presented follow by the specific decisions and optimizations performed on each method modelling.
+Next the method's theory is given by the specific decisions and optimizations performed on each  method.
 And last, the results, conclusions and further considerations are discussed.
@@
 The other way could be called \emph{clean}, where all non-seismic and low quality events were removed.
-This last one was used in this paper and the total number of entries is [CALCULAR O NUMERO DE EVENTOS NO CATALOGO].
-
-This catalog only cover crustal events. Andean subduction earthquakes deeper than 50km in the Brazil-Peru border was discarded. The epicenters of the catalog are shown in figure \ref{fig_seismicity}.
+This last one was used in this paper.
+There is 1941 earthquakes recorded from 1724 to the end of 2013.
+
+This catalog only cover crustal events.
+Andean subduction earthquakes deeper than 50 km in the Brazil-Peru border was discarded as well earthquakes from the Mid-Atlantic ridge.
+The epicenters of the catalog are shown in figure \ref{fig_seismicity}.
 

### 2015-05-14 manuscript.tex
+bibliography and final introduccion and abstract review including infoscores and ED

@@
 Intraplate seismicity is a hard problem to solve in tectonics.
-Brazil, as others intraplate environments, mainly during the lack of observed seismicity the seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
-
-Three smoothed seismicity techniques are reviewed and their application on construct models of Brazilian seismicity and its applications on long-term seismic hazard assessment in Brazil are also discussed.
+In Brazil, as other intraplate tectonic settings, mainly during the lack of observed seismicity, the seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
+
+Three smoothed seismicity techniques are reviewed and their application on construct models of Brazilian seismicity and its applications to long-term seismic hazard assessment in Brazil are also discussed.
 All these smoothing methods are based on a \gls{kde}.
@@
 Historical evidence of magnitude a 7 has been reported in the Amazon \citep{berrocal_1984,veloso_2014}.
-However, as in other intraplate settings, seismic hazard is not negligible due to the low seismic attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.
-
-Very few studies related to seismic hazard, such as \citet{berrocal_1996} for SE Brazil, have been carried out so far and the first Brazilian seismic building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}.
+However, as in other intraplate settings, seismic hazard is not negligible due to the low attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.
+
+Few studies related to seismic hazard, such as \citet{berrocal_1996,santos_2010} for SE and NE Brazil, have been carried out so far and the first Brazilian seismic building code \citep{nbr_15421_2006,santos_lima_2008} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}.
 Brazilian seismicity is concentrated in several seismic zones \citep{berrocal_1984, assumpcao_et_al_2014}.
-Except for a peak of activity in Northeast Brazil, both the GSHAP model and the ABNT zoning have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 mb in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}.
-For this reason, it is important to review and update the hazard estimates for Brazil, especially since the increase in the number of recordings of Brazilian earthquakes in the last three decades.
+Except for a peak of activity in Northeast Brazil, both the GSHAP model and the \citet{nbr_15421_2006} zoning (even when the Brazilian recommended levels are in line with those of other nations with higher occurrence of earthquakes \citep{santos_2013}) have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 mb in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}.
+For this reason, it is important to review and update the hazard estimates for Brazil, especially since the number of Brazilian earthquake records has been increased in the last three decades.
+%the increasing number of Brazilian earthquake records in the last three decades.
 %In addition, the results presented here are part of the \acrfull{gem} supported SARA project (South America Risk Assessment).
 
-Despite attempts to explain the origin of some of the seismic zones, such as stress concentrations from lithospheric thinning \citep{assumpcao_2004, azevedo_2015} or flexural effects \citep{zoback_1996, assumpcao_sacek_2013}, most of them have ill-defined boundaries.
+Despite attempts to explain the origin of some of the seismic zones, such as stress concentrations from lithospheric thinning \citep{assumpcao_2004, assumpcao_et_al_2014, azevedo_2015} or flexural effects \citep{zoback_1996, assumpcao_sacek_2013}, most of them have ill-defined boundaries.
 Attempts to produce seismic hazard maps \citep{dourado_2014} based on seismogenic zones are heavily dependent on subjective expert opinion, even more so in low seismic areas such as Brazil.
-As a contribution to the new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.
+Like a contribution to some new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.
 
@@
 
-Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to pre-defined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
-
-The (smoothed) seismicity \gls{kde} techniques allow us to define another grid of seismic point sources.
+Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
+
+The (smoothed) seismicity \gls{kde} techniques allow us to define another grid of seismic point sources given an earthquake catalog.
 The seismicity rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods.
@@
 \citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} kernel smooth technique with a \emph{fixed smoothing-distance} bandwidth to estimate the seismicity rate by smoothing the 2D histogram from earthquakes counting.
-The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which sum the contribution of each earthquake kernel on given cell.
-
-\citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins.
-
-\citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth on both space and time dimensions for the background (time-independent) seismicity rate on their long-term forecasts experiments.
+The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which adds the contribution of each earthquake kernel on given cell.
+
+\citet{woo_1996} proposed a \emph{magnitude-dependent} kernel bandwidth based on the regression for nearest neighbor mean distance between earthquakes binned by magnitude.
+
+\citet{helmstetter_2012} proposed a \emph{locally-adaptive} bandwidth on both space and time dimensions for the background (time-independent) seismicity rate on their long-term forecasts experiments.
 
@@
 
-\subsection{Information scores and \gls{ed}}
-
-From these three smoothing methods, only \cite{helmstetter_2012} since its formalization use optimizations to define their model parameters.
+\subsection{Information scores and \glspl{ed}}
+
+From these three smoothing methods, only \cite{helmstetter_2012} since its formalization cross-validate their model parameters.
 All the other methods are concept to use the whole catalog and some fixed parameter (e.g.
@@
 
-To compare the performance of these three models, the information scores and \gls{ed} \citep{molchan_1992, kagan_2007, kagan_2009} where computed as alternative to the tests suggested to the \gls{relm} forecasts experiments by \citet{schorlemmer_2007}.
+To compare the performance of these three models, the information scores and error diagrams (\glspl{ed}) \citep{molchan_1992, kagan_2007, kagan_2009} were computed as alternative to the tests suggested to the \gls{relm} forecasts experiments \citep{schorlemmer_2007}.
 
@@
 
-Scores from some Frankel-based models (from different smoothing distances) was computed and the best scores should give us an reasonable choice for this parameter.
+Scores from some Frankel-based models (from different smoothing distances) was computed and the best scores gave us an reasonable choice for this parameter.
 
@@
 The modeling tool was the \gls{hmtk} also open available by \gls{gem} scientific board.
-The Frankel method was already implemented on the toolkit and two others was implemented in the context of this paper and they are free available on a public repository.
-
-The other parameters, like \emph{b}-values, minimum and maximum magnitudes for the frequency-magnitude distributions, strike, dip and rate probabilities to the rupture plane, its top and bottom limits on the subsurface, magnitude-area scale relationships and also the depth distributions must be assigned in other ways to proceed the hazard computation and take advantages of the new developed \glspl{gmpe}.
-
-In this work, to keep our focus on seismicity models as smoothed seismic sources, we perform hazard defining a single \gls{gmpe}  [EXPOSE WHICH ONE]  for a stable continental crust tectonic environment.
-
-The earthquake catalog data comes from the Brazilian seismological research authorities.
+The Frankel method was already implemented on the toolkit and two others was rewritten in the context of this paper based on sources given by the authors and their papers.
+% and should be opened on a public repository.
+
+The other seismic source parameters, like \emph{b}-values, minimum and maximum magnitudes for the frequency-magnitude distributions, strike, dip and rate probabilities to the rupture planes on each source, its top and bottom limits on the subsurface, magnitude-area scale relationships and also the depth distributions must be assigned in other ways to proceed the hazard computation and take advantages of the latest generation \glspl{gmpe}.
+
+In this work, to keep our focus on compute the hazard given a grid of punctual seismic sources computed from smoothed seismicity models, we compute the hazard probabilities using a single \gls{gmpe} [EXPOSE WHICH ONE] for a stable continental crust tectonic environment.
+
+The earthquake catalog data comes from the Brazilian seismological research authorities \citep{bsb_2014}.
 
@@
 The specific purposes of this paper are: 
-
-(i) review three distinct smoothing (zoneless) methods to spatially model the seismic rate,
-(ii) update the seismic hazard studies at national scale by proposing a new simple hazard model,
+(i) to present the \citet{bsb_2014} earthquake catalog,
+(ii) review three distinct smoothing (zoneless) methods to spatially model the seismic rate,
+(iii) evaluate each model by its information scores and error diagrams,
+(iv) add and share into the \gls{hmtk} two more smoothing methods to be used by the hazard modeling community,
 (iii) use the \gls{gem}-\glsdesc{oq} tools to perform hazard calculation and
-(iv) add and share with the community two more new smoothing methods into the \gls{hmtk}.
-
-To aim this, the Brazilian seismicity data is presented including some overview plots.
-Next the method's theory is given by the specific decisions and optimizations performed on each  method.
-And last, the results, conclusions and further considerations are discussed.
+(v) update the hazard studies at national scale by proposing a simple hazard map based on these smoothed seismicity models.
+
+Is not on our purposes at this time: (i) make an exhaustive evaluation of completeness magnitude in dependence of space and time, (ii) go deep on the \gls{gmpe} selection methodology and neither (iii) create the definitive Brazilian hazard map which we handle as a continuous process.
+
+To aim this, the Brazilian seismicity data is presented and some preprocessing steps are described.
+Next the theory behind this smoothed seismicity methodologies are presented as the results are discussed in terms of its scores.
+Then the hazard is computed from these source models in a logic-tree and presented to allow us take some lessons, try to conclude or even suggest something.
 

### 2015-05-14 manuscript.tex
data section almost reviewed. waiting mfd new figure

@@
 Brazil occupies a large stable continental region with seismicity rate lower than Central and Eastern North America, Australia, India or Northern Europe \citep{giardini_1999,johnston_1996a}.
-Spanning most of the central region of the South American plate, Brazil has earthquakes above magnitude 6 mb with a return period of 50 years or so.
+Spanning most of the central region of the South American plate, Brazil has earthquakes above magnitude 6 $mb$ with a return period of 50 years or so.
 Historical records are very short with almost no data before the XX century.
 Historical evidence of magnitude a 7 has been reported in the Amazon \citep{berrocal_1984,veloso_2014}.
-However, as in other intraplate settings, seismic hazard is not negligible due to the low attenuation in the cold and thick lithosphere (magnitude 5 mb can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.
+However, as in other intraplate settings, seismic hazard is not negligible due to the low attenuation in the cold and thick lithosphere (magnitude 5 $mb$ can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.
 
@@
 Brazilian seismicity is concentrated in several seismic zones \citep{berrocal_1984, assumpcao_et_al_2014}.
-Except for a peak of activity in Northeast Brazil, both the GSHAP model and the \citet{nbr_15421_2006} zoning (even when the Brazilian recommended levels are in line with those of other nations with higher occurrence of earthquakes \citep{santos_2013}) have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 mb in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}.
+Except for a peak of activity in Northeast Brazil, both the GSHAP model and the \citet{nbr_15421_2006} zoning (even when the Brazilian recommended levels are in line with those of other nations with higher occurrence of earthquakes \citep{santos_2013}) have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 $mb$ in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}.
 For this reason, it is important to review and update the hazard estimates for Brazil, especially since the number of Brazilian earthquake records has been increased in the last three decades.
@@
 
-Data on intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo \citep{bsb_2014}.
-This catalog is a joint compilation from different sources.
+Data on intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo.
+This catalog is a joint compilation from different sources \citep{bsb_2014}.
 
@@
 
-For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation of Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
+For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation from Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
 From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica.
@@
 Andean subduction earthquakes deeper than 50 km in the Brazil-Peru border was discarded as well earthquakes from the Mid-Atlantic ridge.
-The epicenters of the catalog are shown in figure \ref{fig_seismicity}.
+Epicenters into the catalog are shown in figure \ref{fig_seismicity}.
 
@@
 Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude.
-For historical events magnitudes were estimated from macroseismic data (felt area and maximum intensity) \citep{berrocal_1984}.
-Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
+For historical events magnitudes were estimated from macroseismic data (\glsdesc{sym:Af}, \gls{sym:Af} and \glsdesc{sym:I0}, \gls{sym:I0}) \citep{berrocal_1984}.
+
+Since most \glspl{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
 REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS.
 (Johnston)
-% Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.
+
+Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst most magnitudes on \citet{bsb_2014} catalog had values lower than this.
 
@@
 
-Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night.
-This could suggest some influence of daily noise level.
+Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night hours.
+This could suggest some influence of daily noise level on the station records.
 
@@
 In the case of Brazilian catalog the methods did not present large differences at the final number of earthquakes (fig.
-\ref{fig_decluster}a) and the Gardner-Knopoff/Uhrhammer combination was chosen to maximize the number of events.
-
-The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where long earthquake series tend to occur more frequently.
+\ref{fig_decluster}a) and the Gardner-Knopoff/Uhrhammer combination was chosen to preserve the maximum number of events.
+
+The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where  earthquake sequences tend to occur more frequently.
 
@@
 
-Brazil just recently was installed its permanent seismic network \citep{pirchiner_2011}.
+Brazil just recently installed its permanent seismographic network \citep{pirchiner_2011}.
 The number and quality of stations increased over time.
-The population density, previously concentrated along the coast increased in other regions changing the historical detectability.
-
-There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, stucchi_2004, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveforms \citep{schorlemmer_2008}, but the time and the effort to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance, was not intended to be done on the scope of this paper.
-
-The waveforms from the earliest years was not in digital form.
+The population density, previously concentrated along the coast increased also on other regions changing the historical network coverage.
+%Trace back this history is a huge work with the current available station metadata, mainly from the analog stations.
+There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, stucchi_2004, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveform records \citep{schorlemmer_2008}, but the time and efforts to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance is beyond the scope of this work.
+The waveforms from earliest years was not in digital form.
 The histories of most regional (temporary) stations are not easy available.
-In addition, the low number of earthquakes in the catalog also difficults some high resolution statistical methods.
+In addition, the low number of earthquakes in the catalog also difficult some high resolution statistical methods.
 All of this hinders the direct application of most that methodologies.
-
-For this reason, the spatial completeness was not considered and the temporal completeness parameters for the whole country was defined by an easy cumulative rate criteria using historical information about the network evolution and the \citet{stepp_1972} plot (fig.
+For this reason, the spatial completeness was not considered and just simple temporal \gls{sym:Mc} parameters for the whole country was defined by an easy cumulative rate criteria using historical information about the records evolution and the \citet{stepp_1972} plot (fig.
 \ref{fig_completeness}a).
-TALK ABOUT BRASILIA STATION???
-
-Figure \ref{fig_completeness}b shows the magnitude time distribution in a scatter plot as the cumulative number of events which is important to perform the \gls{tcef} \citep{nasir_2013}.
+
+Figure \ref{fig_completeness}b shows the magnitude by time distribution in a scatter plot as well the cumulative number of events which is important to perform the \gls{tcef} \citep{nasir_2013}.
 The linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend.
-Figure \ref{fig_completeness}b also show three (from many) possible completeness tables to allow quick check their distinctions.
-We chosen the ``simplest'' one.
-
-IMPROVE THIS DECISION EXPLANATION
+For magnitudes greater than 3, this threshold occurs about 1970.
+For the magnitudes greater than 4 and 5, these limits could be extrapolated to 1950 and 1940, respective and roughly.
+Figure \ref{fig_completeness}b also show the completeness tables from the \citet{stepp_1972} test using $\Delta_{mag}=0.5$ and $\Delta_{time}=2$ as magnitude and time bins, and from the literature \citep{assumpcao_et_al_2014}.
+Tables \ref{tab:completeness_stepp} and \ref{tab:completeness} show each one of these completeness tables.
 
@@
 \section{Smoothing methods}
-Three smoothing methods are now briefly explained.
+The smoothing methods considered in this article are now briefly explained.
 
@@
 	\includegraphics[width=0.99\linewidth]{z_img_seismicity}
-	\caption{Brazilian seismicity from 1767 to 2014. Source: \citep{bsb_2014}.}
+	\caption{Brazilian seismicity from 1767 to 2014-01-01. Source: \citep{bsb_2014}.}
 	\label{fig_seismicity}
@@
 There is no significant differences between final results.
-The map (b) of clusters from selected method shows places where earthquake sequences tends to occur more frequently.}
+The map (b) of clusters from selected method shows places where earthquake sequences tends to occur.}
 	\label{fig_decluster}

### 2015-05-16 manuscript.tex
reccurrence plot and legend

@@
 Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
-A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.
+For the minimum magnitude the threshold was 2.9.
+The blue x and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity.
+Stepp completeness table and \citet{bsb_2014} declustered catalogue was used.
+Also a fixed b-value (b=1) incremental distribution is shown in blue continuous line, and the fitted by maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} (orange, $0.76\pm0.04$) are shown as continuous lines.
+The respective a-values was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.
+The minimum magnitude on that completeness table was 2.9.
+
+COMMENT
 
@@
 \section{Smoothing methods}
-The smoothing methods considered in this article are now briefly explained.
+The smoothing methods considered in this article are now briefly explained and their results are discussed.
 
@@
 \begin{figure}
-    \includegraphics[width=.99\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
-    \caption{Annual earthquake recurrence.}
+    \includegraphics[width=.99\linewidth]{z_img_bsb2014_recurrence}
+    \caption{Annual earthquake recurrence.
+The blue x and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity.
+Stepp completeness table and \citet{bsb_2014} declustered catalogue was used.
+Also a fixed b-value (b=1) incremental distribution is shown in blue continuous line, and the fitted by maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} (orange, $0.76\pm0.04$) are shown as continuous lines.
+The respective $10^a$ was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.
+    }
     \label{fig_occurrence}

### 2015-05-16 manuscript.tex
increasing biblio and finish comments + minor changes on the sec 2, data presentation.

@@
 
-Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
+Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods \citep{riznichenko_1959} based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
 
@@
 The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where  earthquake sequences tend to occur more frequently.
-
-\subsection{Frequency-magnitude distribution}
-
-Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
-For the minimum magnitude the threshold was 2.9.
-The blue x and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity.
-Stepp completeness table and \citet{bsb_2014} declustered catalogue was used.
-Also a fixed b-value (b=1) incremental distribution is shown in blue continuous line, and the fitted by maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} (orange, $0.76\pm0.04$) are shown as continuous lines.
-The respective a-values was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.
-The minimum magnitude on that completeness table was 2.9.
-
-COMMENT
 
@@
 Tables \ref{tab:completeness_stepp} and \ref{tab:completeness} show each one of these completeness tables.
+
+\subsection{Frequency-magnitude distribution}
+
+Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
+For the minimum magnitude threshold was 2.9.
+The blue exes and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity.
+Stepp completeness (table \ref{tab:completeness_stepp}) and \citet{bsb_2014} declustered catalogue was used.
+
+Simple \citet{gutenberg_richter_1944} incremental \glspl{mfd},
+\begin{equation}
+\log{N}(m|a,b) = a - b \times m,
+\end{equation}
+were fitted using a fixed $b$-value (b=1), showed as the blue continuous line (fig.
+\ref{fig_occurrence}), by the maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} methods (orange, $b=0.76\pm0.04$) are shown as continuous lines.
+The respective $a$-values was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.
+
+From these distributions over the incremental catalog data, its possible to note just few records with magnitudes greater than 5, even considering different magnitude completeness on time, due probably the small observation time (few than 100 years) on Brazil stable crustal set.
+
+Other important feature extracted from figure \ref{fig_occurrence}, is the low values of the $b$-values relative to the unity.
+Incremental b=1 \gls{mfd} underestimate the earthquake occurrences, specially under great magnitudes.
+From now on, we will assume the $b$-value of 0.9 as better representative for the whole region.
+On the future more investigations on the spatial $b$-value variations should be undertaken.
 

### 2015-05-17 d01-terms.tex
start of introduction review
bibliography, more comments in the text using the figures and improoving legends (section 2)

@@
 
+\newglossaryentry{ED}
+{
+	name={\gls{ed}},
+	description={error diagram},
+	plural={error diagrams}
+}
+
 \newglossaryentry{hypocenter}
@@
 	name={epicenter},
-	description={orthogonal projecton over Earth surface of \gls{hypocenter}},
+	description={orthogonal projection over Earth surface of \gls{hypocenter}},
 	plural={epicentros}

### 2015-05-17 manuscript.tex
bibliography, more comments in the text using the figures and improoving legends (section 2)

@@
 
-Although the standard \citet{cornell_1968}-\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods \citep{riznichenko_1959} based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
+Although the standard \citet{cornell_1968}--\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods \citep{riznichenko_1959} based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).
 
@@
 \citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} kernel smooth technique with a \emph{fixed smoothing-distance} bandwidth to estimate the seismicity rate by smoothing the 2D histogram from earthquakes counting.
-The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which adds the contribution of each earthquake kernel on given cell.
-
-\citet{woo_1996} proposed a \emph{magnitude-dependent} kernel bandwidth based on the regression for nearest neighbor mean distance between earthquakes binned by magnitude.
+This algorithm is already available on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1}.
+%\citet{zechar_jordan_2010} integrate the contribution of each earthquake kernel on given cell grid and then evaluate each cell.
+%The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which integrates the contribution of each earthquake kernel on given cell grid and then evaluate each cell instead smooth the counts \citep{frankel_1995}.
+
+\citet{woo_1996} proposed a \emph{magnitude-dependent} kernel bandwidth function based on the regression for nearest neighbor mean distance between earthquakes binned by magnitude.
 
@@
 
-Using the long-term background seismic rate computed for forecast and assuming its time-independence, the same \gls{psha} assumption, it is possible, preserving their purposes differences \citep{marzocchi_2011}, to use these background seismicity rate values on seismic sources characterization \citep{weatherill_pagani_2014}.
+Is possible to use the long-term background seismic rate computed for forecast by its time-independence, the same \gls{psha}, assumption, even we preserve their different purposes \citep{marzocchi_2011}, and to apply this background seismicity models on seismogenic sources characterization \citep{weatherill_pagani_2014}.
 
@@
 
-From these three smoothing methods, only \cite{helmstetter_2012} since its formalization cross-validate their model parameters.
+From these three smoothing methods, only \cite{helmstetter_2012}, since its formalization, cross-validate their model parameters.
 All the other methods are concept to use the whole catalog and some fixed parameter (e.g.
@@
 (iii) evaluate each model by its information scores and error diagrams,
-(iv) add and share into the \gls{hmtk} two more smoothing methods to be used by the hazard modeling community,
-(iii) use the \gls{gem}-\glsdesc{oq} tools to perform hazard calculation and
+(iv) add into and share through \gls{hmtk} two more smooth methods to be used by the hazard modeling community,
+(iii) use the open source \gls{gem}-\glsdesc{oq} tools to perform hazard calculations and
 (v) update the hazard studies at national scale by proposing a simple hazard map based on these smoothed seismicity models.
@@
 
-To aim this, the Brazilian seismicity data is presented and some preprocessing steps are described.
-Next the theory behind this smoothed seismicity methodologies are presented as the results are discussed in terms of its scores.
-Then the hazard is computed from these source models in a logic-tree and presented to allow us take some lessons, try to conclude or even suggest something.
+To go there, the Brazilian seismicity data will be presented and some preprocessing steps described.
+Next we will present the theory behind these smoothed seismicity methodologies.
+Its results will discussed in terms of its information scores.
+Then we will present the hazard computation formulating a the logic-tree as input for these source models.
+All this should allow us to take some lessons, trying to conclude or even suggest something which could still be developed in the future.
 
@@
 
-Data on intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo.
+Data on crustal intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo.
 This catalog is a joint compilation from different sources \citep{bsb_2014}.
 
-\subsection{Hypocenters}
-
-For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation from Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
+\subsection{Epicenters and magnitudes}
+
+For events up to 1981, epicenters in the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with the participation from Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT).
 From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica.
-Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP).
+Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP) to compute the epicenters.
 Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.
@@
 
-\subsection{Magnitudes}
-
-Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude.
+Most magnitudes were computed as $mb$ type for teleseismic distances or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude for short (300 to 1200 km) distances.
 For historical events magnitudes were estimated from macroseismic data (\glsdesc{sym:Af}, \gls{sym:Af} and \glsdesc{sym:I0}, \gls{sym:I0}) \citep{berrocal_1984}.
 
-Since most \glspl{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
+Since most \glspl{gmpe} are based on $M_W$ moment magnitude scale, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}.
 REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS.
@@
 
-Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst most magnitudes on \citet{bsb_2014} catalog had values lower than this.
+Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst most magnitudes in \citet{bsb_2014} catalog had values lower than this.
 
@@
 
-Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour following \citet{gulia_2012}.
-Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10km.
-
-Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night hours.
+Figure \ref{fig_overview} shows the distribution of depths, day-of-the-week and origin hour \citep{gulia_2012}.
+Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10 km.
+
+Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day-of-the-week, and a slightly trend to record earthquakes during the night hours.
 This could suggest some influence of daily noise level on the station records.
@@
 
-To preserve the Poissonian assumption about the independence of the events, several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a.
+To preserve the Poisson assumption about the independence between distinct events, several decluster algorithms were tried and the results are shown in the figure \ref{fig_decluster}a.
 The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uenthal \citep{van_stiphout_2012}.
-In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window.
+In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window. All this decluster methods are available in the \gls{hmtk}.
 
@@
 
-The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where  earthquake sequences tend to occur more frequently.
+The map on figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where earthquake sequences tend to occur more frequently.
 
@@
 
-The completeness of Brazilian seismicity data has evident time and space dependence.
-
+The completeness of Brazilian seismicity data has evident space and time dependence.
+
+The number and quality of stations increased over time.
+The population density, previously concentrated along the coast increased also on other regions changing the historical instruments coverage.
 Brazil just recently installed its permanent seismographic network \citep{pirchiner_2011}.
-The number and quality of stations increased over time.
-The population density, previously concentrated along the coast increased also on other regions changing the historical network coverage.
 %Trace back this history is a huge work with the current available station metadata, mainly from the analog stations.
-There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, stucchi_2004, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveform records \citep{schorlemmer_2008}, but the time and efforts to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance is beyond the scope of this work.
+
+There are many methodologies to assessment of the spatial and temporal completeness \citep{mignan_2014, vorobieva_2013, nasir_2013, mignan_2013, mignan_2012,  mignan_woessner_2012, mignan_2011, woessner_2005, stucchi_2004, cao_gao_2002, wiemer_wyss_2000, ogata_1993, stepp_1972}, or explaining how to compute the real \gls{pde} from the waveform records \citep{schorlemmer_2008}, but the time and efforts to proceed this exhaustive study of the \glsdesc{sym:Mc}, \gls{sym:Mc}, despite its relevance, is beyond the scope of this work.
 The waveforms from earliest years was not in digital form.
@@
 All of this hinders the direct application of most that methodologies.
-For this reason, the spatial completeness was not considered and just simple temporal \gls{sym:Mc} parameters for the whole country was defined by an easy cumulative rate criteria using historical information about the records evolution and the \citet{stepp_1972} plot (fig.
-\ref{fig_completeness}a).
-
-Figure \ref{fig_completeness}b shows the magnitude by time distribution in a scatter plot as well the cumulative number of events which is important to perform the \gls{tcef} \citep{nasir_2013}.
-The linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend.
+For this reason, the spatial completeness was not considered in this present work.
+
+On the temporal dimension, for the whole country, two simple \gls{sym:Mc} methods were applied: the more subjective cumulative rate criteria or \gls{tcef} \citep{nasir_2013} which uses historical information about the records evolution and by the \citet{stepp_1972} method, an analytical approach, which uses the same kind information.
+
+For the Stepp algorithm we use 0.5 and 2 as magnitude and time bins.
+The plot can be seen on the figure \ref{fig_completeness}a.
+The Stepp plot shows how each complete interval was set up by the point where, going back on time for each magnitude bin  --- increasing the number of samples, $T$ times 1 year samples --- the earthquakes by year ratio, $\lambda$ (expected to be time-independent, and always the same value), the standard error instead decrease as $1/\sqrt{T}$ samples, diverges.
+We also enforce the condition whilst greater magnitudes should have bigger observation times.
+
+Figure \ref{fig_completeness}b shows the magnitude by time distribution in a scatter plot as well the cumulative number of events which is important to perform the \gls{tcef}.
+A linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend.
 For magnitudes greater than 3, this threshold occurs about 1970.
 For the magnitudes greater than 4 and 5, these limits could be extrapolated to 1950 and 1940, respective and roughly.
-Figure \ref{fig_completeness}b also show the completeness tables from the \citet{stepp_1972} test using $\Delta_{mag}=0.5$ and $\Delta_{time}=2$ as magnitude and time bins, and from the literature \citep{assumpcao_et_al_2014}.
+Figure \ref{fig_completeness}b also show the completeness tables from the \citet{stepp_1972} test and from the literature \citep{assumpcao_et_al_2014}.
 Tables \ref{tab:completeness_stepp} and \ref{tab:completeness} show each one of these completeness tables.
@@
 	\includegraphics[width=0.99\linewidth]{z_img_seismicity}
-	\caption{Brazilian seismicity from 1767 to 2014-01-01. Source: \citep{bsb_2014}.}
+	\caption{Brazilian seismicity from 1767 to 2014-01-01.
+Source: \citep{bsb_2014}.
+The lack of information about the earthquakes' depths on this figure is due to almost none precise computation for them.
+When there is, the depths are around 10-15 km or shallower.}
 	\label{fig_seismicity}
@@
 	\end{table}
-	\caption{Catalogue overview.
+	\caption{Catalog overview.
 The histogram (a) shows the depth distribution.
-The histograms (b) and (c) represent the distributions of weekday and hour which earthquakes occur respectly.
+Histograms (b) and (c) represent respectively the distributions of day-of-the-week and day hour which earthquakes occur.
 Dashed lines represent the mean value.}
@@
 	\end{table}
-	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
+	\caption{Earthquakes by year (a) and discriminated by magnitude (b).
+The minimum magnitude is 2.8.
+The time bin is 1 year for (a) and 2 years for (b).
+The magnitude bin on (b) is 0.5 unity.
+CHECK MAGNITUDE 6 MISSED EVENTS ON (b).
+}
 	\label{fig_records}
@@
 	\end{table}
-	\caption{Decluster evaluation (a) using distinct algorithms and windows.
+	\caption{Declustering evaluation (a) using distinct algorithms and windows.
 There is no significant differences between final results.
@@
 \begin{table}[H]
-	\caption{BSB 2014.11 $M_W$ proxies \emph{ad-hoc} guidelines.}
+	\caption{BSB 2014.11 $M_W$ proxies \emph{ad-hoc} guidelines, sorted by priority. \gls{sym:Af} is the \glsdesc{sym:Af}, \gls{sym:I0} is the \glsdesc{sym:I0} in \gls{mmi} \citep{wood_1931}.}
 	\begin{tabular}{l l l}
@@
 		\hline
+		$mb$ and $A_f$
+		& $M_W(m, A_f) = 0.7 M_W(m) + 0.3 M_W(A_f)$
+		& 0.33 \\
 		% mb or equival & $$ M_W = 1.21m - 0.76 $$  & 0.3 \\
@@
 		& 0.3 \\
-		Felt area$A_f$
+		Felt area $A_f$
 		&$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 0.00084\sqrt{A_f}$& 0.4 \\
@@
 		& 0.6 \\
-		$mb$ and $A_f$
-		& $M_W(m, A_f) = 0.7 M_W(m) + 0.3 M_W(A_f)$
-		& 0.33 \\
 		\hline
@@
 \begin{table}[H]
-	\caption{Temporal completeness. For each magnitude interval, the completess was computed by the \citet{stepp_1972} method.}
+	\caption{Temporal completeness. For each magnitude interval, the completeness was computed by the \citet{stepp_1972} method with time-step of 2 years and magnitude bin of 0.5 unity.}
 	\begin{tabular}{c c }
@@
 \begin{table}[H]
-	\caption{Temporal completeness. For each magnitude interval, the completess comes from \citet{assumpcao_et_al_2014} converted to $M_W$ from $m_b$ following table \ref{tab:guidelines}.}
+	\caption{Temporal completeness. For each magnitude interval, the completeness comes from \citet{assumpcao_et_al_2014} converted to $M_W$ from $mb$ following table \ref{tab:guidelines}.}
 	\begin{tabular}{c c}

### 2015-05-18 manuscript.tex
better explaining completeness figure on text

@@
 For the magnitudes greater than 4 and 5, these limits could be extrapolated to 1950 and 1940, respective and roughly.
-Figure \ref{fig_completeness}b also show the completeness tables from the \citet{stepp_1972} test and from the literature \citep{assumpcao_et_al_2014}.
-Tables \ref{tab:completeness_stepp} and \ref{tab:completeness} show each one of these completeness tables.
+Figure \ref{fig_completeness}b also show the completeness times and magnitudes computed from the \citet{stepp_1972} procedure and from \citet{assumpcao_et_al_2014}, converting its magnitudes to \gls{sym:MW}.
+\ref{tab:completeness_stepp} and \ref{tab:completeness} present their values.
 
